### 2016-07-27 manusctipt.tex
Name changes

\documentclass[9pt,twocolumn,twoside]{gsajnl}
% Use the documentclass option 'lineno' to view line numbers

\articletype{gos} % article type
% {inv} Investigation 
% {gs} Genomic Selection
% {goi} Genetics of Immunity 
% {gos} Genetics of Sex 
% {mp} Multiparental Populations

\title{Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression}

\author[$\ast$,$\dagger$,1]{Josh Hough}
\author[$\dagger$]{Felix Baudry}
\author[$\dagger$]{Wei Wang}
\author[$\dagger$]{Spencer C.H. Barrett}
\author[$\dagger$]{Stephen I. Wright}

\affil[$\ast$]{Department of Plant Sciences, University of California, Davis}
\affil[$\dagger$]{Department of Ecology and Evolutionary Biology, University of Toronto}

%\affil[$\dagger$]{Author two affiliation}
%\affil[$\ddagger$]{Author three affiliation}
%\affil[$\S$]{Author four affiliation}
%\affil[$\ast\ast$]{Author five affiliation}

\keywords{Sex Chromosome Evolution; Nucleotide Diversity; Recombination; Deleterious Mutations}

\runningtitle{Plant sex chromosome evolution} % For use in the footer

\correspondingauthor{Josh Hough}

\begin{abstract}
X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome. On Y chromosomes with suppressed recombination, both positive and negative selection can reduce the frequencies of genetically linked neutral variants, resulting in a reduction in $N_{e}$ and the level of neutral polymorphism on Y compared to X chromosomes or autosomes. However, non-selective factors including biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity. Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae). Strikingly, our analysis revealed that nucleotide diversity on the Y is approximately 12.5 fold lower than predicted from models of neutral evolution. We demonstrate that the magnitude of this reduction is unlikely to be explained by the observed occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness. Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of background selection arising from suppressed recombination. Given the recent origin of \textit{R. hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with background selection playing an important role during the early stages of their evolution.

%this is 254 words and cannot be any longer than 250. Make edits accordingly.

%Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization}

\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\marginmark
\firstpagefootnote
\correspondingauthoraffiliation{jhough@ucdavis.edu}
\vspace{-11pt}%

\section*{Introduction}

\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plant and animal kingdoms \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}. Despite numerous biological differences between the kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes, suggesting that general evolutionary mechanisms might drive sex chromosome divergence \citep{charlesworth1978,charlesworth1996CB,charlesworth2000degeneration}. Investigating the shared features of plant and animal sex chromosomes, including the loss of recombination \citep{bergero2009}, the degeneration of Y chromosomes \citep{hough2014,bergero2015}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}, thus provides an informative context for investigating the evolutionary forces driving sex chromosome evolution.

One fundamental difference between sex chromosomes and autosomes is their respective copy number in a population; whereas autosomes in diploid species are present in two copies in each sex, the X chromosome is present in two copies in females and only one copy in males. As a consequence, the effective population size ($N_{e}$) of the X chromosome is predicted to be 3/4 that of autosomes, whereas the $N_{e}$ of the male-limited Y chromosome is expected to equal 1/4 that of autosomes (assuming an equal number of reproducing females and males). Such differences in $N_{e}$ are predicted to directly affect relative levels of neutral polymorphism maintained on these chromosomes, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ \citep{Kimura1984}. This variation in neutral polymorphism is in turn expected to have important consequences on patterns of DNA sequence evolution, including the effectiveness of both positive and negative selection \citep{charlesworth1987}.

Several demographic factors are expected to modulate or accentuate differences in polymorphism levels maintained on sex chromosomes \citep{ellegren2009}. These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}. For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in current estimates of the X/A ratio in humans, and other demographic processes have also been suggested \citep{hammer2010,bustamante2009}. Theoretical work also indicates that population bottlenecks can lead to disproportionate decreases in sex-linked compared to autosomal variation \citep{pool2007}, a prediction that has been supported in a diverse array of taxa, including chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, Drosophila \citep{andolfatto2001}, and humans \citep{keinan2009}.

In addition to demographic factors, evolutionary theory predicts that genetic variability on sex chromosomes should also be affected by the evolution of suppressed X-Y recombination, which lowers the coalescent $N_{e}$ for Y-linked genes and increases their vulnerability to the effects of selective sweeps \citep{smith1974hitch,aquadro1994} and background selection \citep{charlesworth1996background,charlesworth1994effect}. These processes can be viewed as examples of the Hill–Robertson (HR) effect, in which the effective population size for a given chromosomal region depends strongly on the rate of recombination, such that selection at a focal site is not independent of selection at nearby sites \citep{hill1966HReffect}. Moreover, both the background selection and selective sweep models predict a reduction in $N_{e}$ and, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}. Although the HR effect can occur throughout the genome \citep{mcvean2000}, large genomic regions in which recombination is suppressed, such as the Y chromosome, are expected to experience the most severe reductions in neutral diversity due to chromosome-wide linkage \citep{charlesworth1996CB,charlesworth2000degeneration,bachtrog2013NRG}. Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have found that Y-linked diversity is considerably lower than predicted under models of neutral evolution, suggesting that the HR effect plays a key role determining patterns of nucleotide diversity during sex chromosome evolution across a broad range of animal taxa.

Despite widespread interest in determining the effects of recombination on patterns of genetic diversity and the forces contributing to sex chromosome divergence \citep{ellegren2011,otto2011PAR,bachtrog2013NRG}, we still know little about the influence of the HR effect on patterns nucleotide polymorphism in young sex chromosome systems, where \textit{de novo} recombination suppression has evolved relatively recently \citep{charlesworth2016plant}. The timescales over which selection affects neutral polymorphism on sex chromosomes is therefore not well understood. Recent studies of young sex chromosome system in plants have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}. Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in young plant sex chromosomes. Given that both background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost a large proportion of their genes \citep{bachtrog2008temporal}, studying recently evolved plant sex chromosomes provides an important test of the temporal dynamics of the HR effect.

To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed single nucleotide polymorphisms (SNPs) on sex chromosomes and autosomes in the plant \textit{R. hastatulus }(Polygonaceae). This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than those in mammals \citep{lahn1999,ross2005dna}. \textit{Rumex hastatulus} has received particular attention because of the unique occurrence of an intraspecific polymorphism for sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct chromosomal races \citep{smith1963mechanism}. The $XY_{1}XY_{2}$ sex chromosome system (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}. Despite the recent origin of the sex chromosome systems in this species, there is evidence that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration, though less degeneration was inferred for younger sex-linked genes unique to the North Carolina race \citep{hough2014}.

Of particular relevance to the present study is that \textit{R. hastatulus} populations throughout their geographical range (southern USA) consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}. This is important because female-biased sex ratios are predicted to affect levels of sex chromosome diversity \citep{ellegren2009}, though few studies have attempted to empirically distinguish between the effects of biased sex ratios and selective interference. Here, focusing on the $Y_{1}$ chromosome that is shared between the chromosomal races of \textit{R. hastatulus}, we tested whether levels of nucleotide diversity on X, Y, and autosomal chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success. In addition, we use population genetic parameters estimated from our data to conduct forward-time simulations of both purifying and positive selection to determine the relative support for models of background selection and selective sweeps in explaining the levels of diversity we observed.

\section*{Materials and Methods}
\subsection*{Population Samples and Sex-Linked Genes} 
We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations, 6 populations of each sex chromosome race). Samples of both sex chromosomal races were collected in 2010 from throughout the native range of \textit{R. hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females. We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich). The isolation of mRNA and cDNA synthesis was conducted according to standard Illumina RNAseq procedures, with sequencing conducted on two Illumina HiSeq lanes with 150-bp end reads at the Genome Quebec Innovation Center. Reads from these 24 samples were mapped to the \textit{R. hastatulus} reference transcriptome \citep{hough2014}, and raw sequences are available from the Gen-Bank Short Read Archive under accession no. SRP041588. Reads were mapped using the Burrows–Wheeler Aligner (release 0.6.2-r126 \citep{li2010fast}), followed by Stampy (release 1.0.20; \citep{lunter2011stampy}). We used Picard tools (release 1.78, http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit (GATK version 2.1-11; \citep{mckenna2010genome}) variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores (<20). The population samples analyzed here were previously reported in Hough et al. (2014), where they were used to validate the ascertainment of sex-linked contigs identified through segregation analysis. From the list of sex-linked genes identified in Hough et al. (2014) that were shared between the Texas (XY) and North Carolina (XY1Y2) sex chromosome races, we identified a total of 460 sex-linked genes for analysis.

\subsection*{Autosomal Genes} 
In evaluating evidence for nucleotide diversity differences between X and Y chromosomes, it is important to distinguish between reduced Y-linked diversity, and the possibility that X-linked diversity is elevated above the level predicted from a neutral model. To do this, we normalized sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from a neutral model (described below). Because the criteria for identifying autosomal loci in Hough et al. (2014) were based on the occurrence of four segregating SNPs per locus, and this set is therefore probably higher in diversity than the average autosome, here we incorporated a broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data. We filtered this set of non-sex linked genes to remove those that may have been sex-linked but were not identified as such by the ascertainment criteria in Hough et al. (2014). In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes with SNPs showing fixed heterozygosity in males and fixed homozygosity in females, (iii) any genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicate genes or those with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females. Finally, we removed genes with fewer than 100 synonymous sites following filtering to avoid biasing results toward genes that may have been particularly short due to assembly problems. This filtering resulted in a set of 12,356 and 11,350 autosomal genes in the Texas and North Carolina races, respectively.

\subsection*{Phasing X and Y alleles} 
To estimate polymorphism for the X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males. In previous work, phasing alleles on \textit{R. hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross. Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes in a diploid individual to output a list of phased haplotype blocks containing the SNP variants on each chromosome. Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences with each haplotype block (i.e., sites in which all females were homozygous, and all males were heterozygous). Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output. This was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites - i.e., when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched. SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data. Finally, we filtered SNPs for coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels. This procedure was conducted using a combination of Perl and Bash and this resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes from the 24 individuals in our study.

We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted using data from a genetic cross in Hough et al. (2014). To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data). In particular, we looked for cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was zero, since Y-linked  sequences from a single family should all be identical, as we had confirmed through phasing by segregation. We identified 7 percent of sex-linked genes that had phasing errors of this kind, or that we determined to have genotyping errors resulting in a false SNP calls on the Y chromosome, leading to a SNP error rate estimate of 1.7 x 10-4. This rate is very low relative to the population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism. However, because this rate is high relative to the expected level of true polymorphism on the Y-chromosome, we further filtered genes in which we found evidence for false-positive Y-polymorphisms arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression. This filtering was done by manually inspecting sequences in IGV \citep{robinson2011integrative} and identifying each individual putative polymorphism on the Y chromosome.

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes} 
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1). ). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate. To address this issue, we used a maximum likelihood approach implemented in the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio. Here, [explain more]. Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males. As no significant difference was observed, we report only results from females.

\subsection*{Neutral predictions and the effect of sex ratio bias on diversity} 
Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number

\begin{equation}
S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i
\label{eq:refname1}
\end{equation}

\subsection*{Simulations of Background Selection and Selective Sweeps}

\section*{Results and Discussion}

\subsection*{Evolutionary relationships of sex chromosome races in \textit{R. hastatulus}} 
\subsection*{Y-chromosome diversity is significantly lower than predicted by neutrality}

\begin{table*}[htbp]
\centering
\caption{\bf Students and their grades}
\begin{tableminipage}{\textwidth}
\begin{tabularx}{\textwidth}{XXXX}
\hline
Student & Grade\footnote{This is an example of a footnote in a table. Lowercase, superscript italic letters (a, b, c, etc.) are used by default. You can also use *, **, and *** to indicate conventional levels of statistical significance, explained below the table.} & Rank & Notes \\
\hline
Alice & 82\% & 1 & Performed very well.\\
Bob & 65\% & 3 & Not up to his usual standard.\\
Charlie & 73\% & 2 & A good attempt.\\
\hline
\end{tabularx}
  \label{tab:shape-functions}
\end{tableminipage}
\end{table*}

\subsection*{Male-to-female sex ratio of $N_{e}$}

\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{Figure1.png}
\caption{Example figure from \url{10.1534/genetics.114.173807}. Please include your figures in the manuscript for the review process. You can upload figures to Overleaf via the Project menu. Upon acceptance, we'll ask for your figure files to be uploaded in any of the following formats: TIFF (.tiff), JPEG (.jpg), Microsoft PowerPoint (.ppt), EPS (.eps), or Adobe Illustrator (.ai).  Images should be a minimum of 300 dpi in resolution and 500 dpi minimum if line art images.  RGB, CMYK, and Grayscale are all acceptable. Halftones should be high contrast with sharp detail, because some loss of detail and contrast is inevitable in the production process. Figures should be 10-20 cm in width and 1-25 cm in height. Graph axes must be exactly perpendicular and all lines of equal density.
Label multiple figure parts with A, B, etc. in bolded type, and use Arrows and numbers to draw attention to areas you want to highlight. Legends should start with a brief title and should be a self-contained description of the content of the figure that provides enough detail to fully understand the data presented. All conventional symbols used to indicate figure data points are available for typesetting; unconventional symbols should not be used. Italicize all mathematical variables (both in the figure legend and figure) , genotypes, and additional symbols that are normally italicized.  
}%
\label{fig:spectrum}
\end{figure}

\subsection*{Background Selection and Selective Sweeps}

\section*{Conclusions}
\section*{Acknowledgments}

\bibliography{example-bibliography}
\end{document}

### 2016-07-27 manusctipt.tex
bib fix
readme edits, minor ms changes
neutral predictions section and bib

@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number
+Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number.
+For a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes relative to autosomes is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
+In a neutral model with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$.
+For populations with female-biased sex ratios, however, $N_{m}$ becomes small relative to $N_{f}$ and this ratio can become larger than 1, approaching 1.125 at the limit \citep{caballero1995}.
+For the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.
+Under neutrality, the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
+Assuming equal neutral mutation rates for all genes and an equal number of reproducing males and females, Y-linked genes should therefore have 1/3 the polymorphism of X-linked genes, and 1/4 that of autosomal genes.
+Figure 1 shows the predicted relative effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\linewidth]{Figure1.png}
+\caption{\textbf{A:} The relation between relative effective population size and sex ratio bias for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y).
+The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})=0$, ranging from 0 to 1, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively.
+Calculations assume a constant population size, non-overlapping generations, and a Poisson distribution of offspring number \citep{wright1931evolution,charlesworth2009effective}.
+The $N_{e}$ for each chromosome type is calculated relative to the predicted $N_{e}$ for genes on autosomes with an equal sex ratio.\textbf{B:} The corresponding expected relation between sex ratio bias and the sex-chromosome-to-autosome diversity ratio, assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates for all genes.
+}%
+\label{fig:spectrum}
+\end{figure}
+
+\subsection*{Simulations of Background Selection and Selective Sweeps}
 
@@
 \end{equation}
-
-\subsection*{Simulations of Background Selection and Selective Sweeps}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=\linewidth]{Figure1.png}
-\caption{Example figure from \url{10.1534/genetics.114.173807}.
-Please include your figures in the manuscript for the review process.
-You can upload figures to Overleaf via the Project menu.
-Upon acceptance, we'll ask for your figure files to be uploaded in any of the following formats: TIFF (.tiff), JPEG (.jpg), Microsoft PowerPoint (.ppt), EPS (.eps), or Adobe Illustrator (.ai).
-Images should be a minimum of 300 dpi in resolution and 500 dpi minimum if line art images.
-RGB, CMYK, and Grayscale are all acceptable.
-Halftones should be high contrast with sharp detail, because some loss of detail and contrast is inevitable in the production process.
-Figures should be 10-20 cm in width and 1-25 cm in height.
-Graph axes must be exactly perpendicular and all lines of equal density.
-Label multiple figure parts with A, B, etc.
-in bolded type, and use Arrows and numbers to draw attention to areas you want to highlight.
-Legends should start with a brief title and should be a self-contained description of the content of the figure that provides enough detail to fully understand the data presented.
-All conventional symbols used to indicate figure data points are available for typesetting; unconventional symbols should not be used.
-Italicize all mathematical variables (both in the figure legend and figure) , genotypes, and additional symbols that are normally italicized.
-}%
-\label{fig:spectrum}
-\end{figure}
-
 \subsection*{Background Selection and Selective Sweeps}
@@
 
-\bibliography{example-bibliography}
+\bibliography{bibliography}
 \end{document}

### 2016-07-27 manuscript.tex
renamed files, updated README

--- 2016-07-27 manusctipt.tex
@@
 \subsection*{Simulations of Background Selection and Selective Sweeps}
+To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a %using the software packages fwdpp and fwdpy (Thornton, 2014), and analyzed the output using libsequence (Thornton 2003) and standard Python.
+To match the sample size of our empirical data, we simulated sex chromosomes and autosomes from 24 individuals , and collected diversity statistics after evolving populations under background selection for 10N generations with 1000 replicates each.
+To consider only the effects of selection against deleterious alleles, we ignored adaptive mutations and sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient s = -0.01, neutral mutation rate μ = 0.005, deleterious mutation rate U = 0.01 (diploid), and dominance term h = 1 .
+Our simulations assumed that recombination did not occur on Y-chromosomes, whereas we allowed X chromosomes and autosomes to experience free recombination uniformly throughout their length.
+Initial effective population sizes for simulated sex chromosomes were set such that NeX/NeA = 3/4 and NeY/NeA = 1/4, with an initial NeA of 1000.
+To compare the simulation output with our empirical data, we calculated the average θ over all simulations, and calculated normalized X/A and Y/A ratios of neutral diversity (Figure 2).
+Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
+
+%We modeled both purifying and positive selection using the forward simulation software SFS_CODE [64].
+Y chromosome per generation per base pair mutation rate, and sampling 8 chromosomes per simulation to match the sample size of our observed data.
+SFS_CODE starts with the ancestral number of chromosomes and simulates a haploid population forward in time.
+Thus, when rescaling the effective population size from the autosomal estimates, for SFS_CODE we used the same diploid autosomal ancestral effective population size for both populations (N=10,000).
+The Y chromosome effective size was then found using the same process described above for the neutral coalescent simulations.
+
+%We ran 1000 replicates for each set of selection parameters in each population.
+For each replicate we calculated p*, the simulated per site nucleotide diversity (average number of pairwise differences).
+The similarly calculated observed Y diversity is denoted pobs.
+For each set of parameter values we then calculated P1, the proportion of simulation replicates with p*.pobs was used to calculate a 2-sided P-value by P2 = 1226|P220.5|.
+Models with could not be rejected and were considered to fit the observed data.
 
@@
 \subsection*{Evolutionary relationships of sex chromosome races in \textit{R. hastatulus}}
-\subsection*{Y-chromosome diversity is significantly lower than predicted by neutrality}
-
-\begin{table*}[htbp]
-\centering
-\caption{\bf Students and their grades}
-\begin{tableminipage}{\textwidth}
-\begin{tabularx}{\textwidth}{XXXX}
-\hline
-Student & Grade\footnote{This is an example of a footnote in a table.
-Lowercase, superscript italic letters (a, b, c, etc.) are used by default.
-You can also use *, **, and *** to indicate conventional levels of statistical significance, explained below the table.} & Rank & Notes \\
-\hline
-Alice & 82\% & 1 & Performed very well.\\
-Bob & 65\% & 3 & Not up to his usual standard.\\
-Charlie & 73\% & 2 & A good attempt.\\
-\hline
-\end{tabularx}
-  \label{tab:shape-functions}
-\end{tableminipage}
-\end{table*}
 

### 2016-07-27 manuscript.tex
minor edits, added simulation directories

--- 2016-07-27 manusctipt.tex
@@
 % Use the documentclass option 'lineno' to view line numbers
+
+%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
+%\setlength{\marginparwidth}{1.25cm}
+\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
 
@@
 \subsection*{Simulations of Background Selection and Selective Sweeps}
+To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a
+
+\josh{
+\textit{to do}
+\begin{itemize}
+\item \textit{describe parameters used and justify}
+\item \textit{do both purifying and positive selection sims}
+\end{itemize}
+}
+
+We modeled both purifying and positive selection using the forward simulation software SFS_CODE \citep{hernandez2008flexible}.
+To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data ...
+A with our empirical sampling, the simulations sampled 8 chromosomes per simulation.
+
+%using the software packages fwdpp and fwdpy (Thornton, 2014), and analyzed the output using libsequence (Thornton 2003) and standard Python.
+To match the sample size of our empirical data, we simulated sex chromosomes and autosomes from 24 individuals , and collected diversity statistics after evolving populations under background selection for 10N generations with 1000 replicates each.
+To consider only the effects of selection against deleterious alleles, we ignored adaptive mutations and sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient s = -0.01, neutral mutation rate μ = 0.005, deleterious mutation rate U = 0.01 (diploid), and dominance term h = 1 .
+Our simulations assumed that recombination did not occur on Y-chromosomes, whereas we allowed X chromosomes and autosomes to experience free recombination uniformly throughout their length.
+Initial effective population sizes for simulated sex chromosomes were set such that NeX/NeA = 3/4 and NeY/NeA = 1/4, with an initial NeA of 1000.
+To compare the simulation output with our empirical data, we calculated the average θ over all simulations, and calculated normalized X/A and Y/A ratios of neutral diversity (Figure 2).
+Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
 

### 2016-07-28 SFS_code/convert_SFS_CODE.sh
added SFS scripts

@@ -1,0 +1,21 @@
+#!/bin/bash -l
+#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
+#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/convert-stdout_%j.txt
+#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/convert_stderr_%j.txt
+#SBATCH -J convert_SFS
+#SBATCH --mem-per-cpu=5G
+#SBATCH --cpus-per-task=1
+
+set -e
+set -u
+set -o pipefail
+
+##Josh Hough, UC Davis, July 08, 2016
+
+convertSFS_CODE [commands] | msstats/src/msstats >
+
+
+#Then, using R,
+#x=read.table("output",header=T)
+#mean(x$hsm01[x$pop==0])
+#[1] 0.3511204

### 2016-07-28 SFS_code/sfs_code.sh

@@ -1,0 +1,22 @@
+#!/bin/bash -l
+#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
+#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code-stdout_%j.txt
+#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code_stderr_%j.txt
+#SBATCH -J sfs
+#SBATCH --cpus-per-task=10
+
+set -e
+set -u
+set -o pipefail
+
+##Josh Hough, UC Davis, July 19, 2016
+
+echo "Starting Job: test1"
+date
+
+sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A >test1 \
+-o test1.out --errfile test1.err
+
+echo "Ending Job: "
+date

### 2016-07-29 manuscript.tex
minor edits
simulation section edits

--- 2016-07-27 manusctipt.tex
@@
 % Use the documentclass option 'lineno' to view line numbers
+
+%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
+%\setlength{\marginparwidth}{1.25cm}
+\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
+\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
+\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
+\newcommand{\X}{\textcolor{red}{\bf X\,}}
 
@@
 Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
-To address this issue, we used a maximum likelihood approach implemented in the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio.
-Here, [explain more].
+To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp.
+Figures).
+Here, \X{...explain more}.
 Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males.
-As no significant difference was observed, we report only results from females.
+As no significant difference was observed (\X Supp.
+fig.), we report only results from females.
 
@@
 \subsection*{Simulations of Background Selection and Selective Sweeps}
+To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a
+
+\josh{
+\textit{to do}
+\begin{itemize}
+\item \textit{describe parameters used and justify}
+\item \textit{do both purifying and positive selection sims}
+\end{itemize}
+}
+
+We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}.
+To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data.
+To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$.
+We then calculated the average \theta over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure).
+Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
 

### 2016-07-29 SFS_code/sfs_code.sh
sfs script

@@ -9,9 +9,7 @@ set -e
 set -u
 set -o pipefail
 
-##Josh Hough, UC Davis, July 19, 2016
-
-echo "Starting Job: test1"
+echo "Starting Job: test"
 date
 
 sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \

### 2016-07-29 manuscript.tex
moved phylogeny section

@@
 To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1).
-).
 We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
+Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated in our , since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
 To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp.
@@
 Here, \X{...explain more}.
+To test for population substructure between the two sex chromosome races, which could affect estimates diversity, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+The analysis was conducted on an alignment of X- and Y-linked genes from \texit{R.
+hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \texit{R.
+bucephalophorus} used to root the tree.
+Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987).
+The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al.
+2004).
 Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males.
@@
 To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$.
-We then calculated the average \theta over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure).
+We then calculated the average $\theta$ over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure).
 Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
@@
 
+%In contrast to the Texas sex chromosome race, our analysis revealed that Y-chromosome diversity was considerably higher in the North Carolina race (see Supplementary Material; Figure S1).
+This difference could have arisen due to population substructure, resulting in an elevated overall Y diversity in the pooled sample.
+To test this hypothesis, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+The resulting sequence phylogeny (Figure 4) revealed strong support for the Y-linked genes from the XY1Y2 race being paraphyletic (98% bootstrap support).
+In particular, our analysis revealed that samples from two populations, both from South Carolina (hereafter the SC sub-clade), formed a monophyletic group, whereas samples from Florida and Georgia (hereafter the FL sub-clade) were also monophyletic, but were more closely related to the XY Texas race.
+Although our sampling here is limited, these results suggest that introgression may have occurred between the ancestral Texas (XY race) and the derived North Carolina (XY1Y2) race, leading to a second, derived, XY1Y2 sub-clade and an elevated level of Y-diversity in our pooled sample.
+As these races are inter-fertile (Smith reference about hybrid zone), we suggest that the SC sub-clade likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY race.
+Notably, when we separated samples by the two inferred sub-clades of Y–linked sequences, the estimated level of Y diversity was consistent across all three clades, suggesting that the extensive reduction in diversity on this chromosome is not a population or race-specific phenomenon (see Supplementary Material, Figure S1).
+
+%figure
+
+%Figure 4.
+Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987).
+The percentage of replicate trees in which the associated sequences clustered together in the bootstrap test (1000 replicates) is shown next to branches (Felsenstein 1985).
+The phylogeny is drawn to scale, with branch lengths in the same units as those of the evolutionary distances used to infer the tree.
+The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al.
+2004) and are in the units of the number of base substitutions per site.
+The analysis was conducted on an alignment of X- and Y-linked genes from Rumex hastatulus, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species Rumex bucephalophorus used to root the tree.
+‘N’ designates the North Carolina race, and ‘T’ the Texas race, with numbers 1-6 corresponding to each of the 6 populations sampled.
+The inferred SC (South Carolina), and FL (Florida) sub-clades are indicated.
+
 \subsection*{Male-to-female sex ratio of $N_{e}$}

### 2016-07-30 manuscript.tex
minor ms edits

@@
 \author[$\ast$,$\dagger$,1]{Josh Hough}
-\author[$\dagger$]{Felix Baudry}
 \author[$\dagger$]{Wei Wang}
@@
 To test for population substructure between the two sex chromosome races, which could affect estimates diversity, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
-The analysis was conducted on an alignment of X- and Y-linked genes from \texit{R.
-hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \texit{R.
+The analysis was conducted on an alignment of X- and Y-linked genes from \textit{R.
+hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R.
 bucephalophorus} used to root the tree.
@@
 Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males.
-As no significant difference was observed (\X Supp.
-fig.), we report only results from females.
+As no significant difference was observed, we report only results from females.
 

### 2016-07-30 SFS_code/convert_SFS_CODE.sh
sfs scripts

@@ -10,12 +10,11 @@ set -e
 set -u
 set -o pipefail
 
-##Josh Hough, UC Davis, July 08, 2016
+#command structure: convertSFS_CODE [commands] | msstats/src/msstats >out
+# used: - for ms style output T 0 says to output sysnonymous mutations
 
-convertSFS_CODE [commands] | msstats/src/msstats >
+convertSFS_CODE -m T 0 | msstats/src/.msstats > output.txt
 
-
-#Then, using R,
-#x=read.table("output",header=T)
-#mean(x$hsm01[x$pop==0])
-#[1] 0.3511204
+# Then, using R,
+# x=read.table("output",header=T)
+# mean(x$hsm01[x$pop==0])

### 2016-07-30 SFS_code/sfs_code.sh

@@ -9,12 +9,29 @@ set -e
 set -u
 set -o pipefail
 
-echo "Starting Job: test"
+
+#command structure: sfs_code <Npops> <Niter> [<options> [arguments]]
+
+# Parameters used here
+# Npops:        1
+# Niter:        50
+# theta(-t):    0.001375
+# rho(-r):      0
+# ploidy(-P):   1
+# tau(-TE):     0.3 (scaled in units of P*N generations)
+# loci(-L):     2 regions, lengths: 500000, 800000 (<R> <L1> [<L2>..] [R]): R regions with lengths <L1><L2>...
+# annotate(-a): N noncoding
+# fitness(-W):  L 1 2 0 1 1 0.184 0.003 (here, gamma shape = 0.184, and s=0.01. i.e., 0.184/5000=0.003)
+# -A tells simualtion to end
+# sampsize(-n): 8 chromosomes per sample
+# popsize:(-N)  500
+
+echo "Starting Job: coding regions only"
 date
 
 sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
--a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A >test1 \
--o test1.out --errfile test1.err
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A \
+-o SFS.out --errfile SFS.err
 
-echo "Ending Job: "
+echo "Ending Job: coding regions "
 date

### 2016-07-30 SFS_code/convert_SFS_CODE.sh
covert_SFS_CODE

@@ -3,18 +3,18 @@
 #SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/convert-stdout_%j.txt
 #SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/convert_stderr_%j.txt
 #SBATCH -J convert_SFS
-#SBATCH --mem-per-cpu=5G
-#SBATCH --cpus-per-task=1
+#SBATCH --cpus-per-task=5
 
 set -e
 set -u
 set -o pipefail
 
-#command structure: convertSFS_CODE [commands] | msstats/src/msstats >out
-# used: - for ms style output T 0 says to output sysnonymous mutations
+#see https://github.com/molpopgen/msstats
+#command structure: convertSFS CODE <input file> <option [args]> | msstats/src/msstats >out
+# used: - for ms style output; T 0 says to output only sysnonymous mutations
 
-convertSFS_CODE -m T 0 | msstats/src/.msstats > output.txt
+convertSFS_CODE SFS.out -m T 0 | msstats/src/.msstats > stats.txt
 
 # Then, using R,
-# x=read.table("output",header=T)
+# x=read.table("stats",header=T)
 # mean(x$hsm01[x$pop==0])

### 2016-07-31 SFS_code/convert_SFS_CODE.sh
minor edit

@@ -13,7 +13,7 @@ set -o pipefail
 #command structure: convertSFS CODE <input file> <option [args]> | msstats/src/msstats >out
 # used: - for ms style output; T 0 says to output only sysnonymous mutations
 
-convertSFS_CODE SFS.out -m T 0 | msstats/src/.msstats > stats.txt
+convertSFS_CODE SFS.out -m T 0 | ./msstats/src/msstats > stats.txt
 
 # Then, using R,
 # x=read.table("stats",header=T)

### 2016-07-31 manuscript.tex
plots, sampling, minor edits
ms edit, added supp info

@@
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated in our , since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
-To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp.
+Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
+To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to obtain independent estimates of the Y/A ratio (\X Supp.
 Figures).
 Here, \X{...explain more}.
+
 To test for population substructure between the two sex chromosome races, which could affect estimates diversity, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
@@
 2004).
+
 Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males.

### 2016-07-31 supporting-information/supporting_info.tex
ms edit, added supp info

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Remove the twocolumn option to create a single column SI file if required. 
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnassupportinginfo}

\title{Supporting Information}
\author{Hough et al.}

\doi{genetics.XXXXXXXXXX}

\begin{document}

\maketitle

\section*{Supporting Information (SI)}

This supporting information may also be accessed at gihub at https://github.com/houghjosh/XYdiversity

\subsection*{SI Figures}

\subsection*{SI Tables}
Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

\begin{table}[tbhp!]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{SI Datasets}

Supply Excel (.xls), RTF, CSV, TXT, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsection*{SI software}

\end{document}

### 2016-08-01 manuscript.tex
supp info, tables, edits

@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, both positive and negative selection can reduce the frequencies of genetically linked neutral variants, resulting in a reduction in $N_{e}$ and the level of neutral polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in a reduction in $N_{e}$ and the level of neutral polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
@@
 
-Several demographic factors are expected to modulate or accentuate differences in polymorphism levels maintained on sex chromosomes \citep{ellegren2009}.
+Several demographic factors can modulate or accentuate differences in polymorphism levels maintained on sex chromosomes \citep{ellegren2009}.
 These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
-For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in current estimates of the X/A ratio in humans, and other demographic processes have also been suggested \citep{hammer2010,bustamante2009}.
+For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in estimates of the X/A ratio in humans, and other demographic processes have also been suggested \citep{hammer2010,bustamante2009}.
 Theoretical work also indicates that population bottlenecks can lead to disproportionate decreases in sex-linked compared to autosomal variation \citep{pool2007}, a prediction that has been supported in a diverse array of taxa, including chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, Drosophila \citep{andolfatto2001}, and humans \citep{keinan2009}.
@@
 These processes can be viewed as examples of the Hill–Robertson (HR) effect, in which the effective population size for a given chromosomal region depends strongly on the rate of recombination, such that selection at a focal site is not independent of selection at nearby sites \citep{hill1966HReffect}.
-Moreover, both the background selection and selective sweep models predict a reduction in $N_{e}$ and, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
+Moreover, both background selection and selective sweeps predict a reduction in $N_{e}$ and, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
 Although the HR effect can occur throughout the genome \citep{mcvean2000}, large genomic regions in which recombination is suppressed, such as the Y chromosome, are expected to experience the most severe reductions in neutral diversity due to chromosome-wide linkage \citep{charlesworth1996CB,charlesworth2000degeneration,bachtrog2013NRG}.
@@
 
-Despite widespread interest in determining the effects of recombination on patterns of genetic diversity and the forces contributing to sex chromosome divergence \citep{ellegren2011,otto2011PAR,bachtrog2013NRG}, we still know little about the influence of the HR effect on patterns nucleotide polymorphism in young sex chromosome systems, where \textit{de novo} recombination suppression has evolved relatively recently \citep{charlesworth2016plant}.
+Despite widespread interest in determining the effects of recombination on patterns of genetic diversity and the forces contributing to sex chromosome divergence \citep{ellegren2011,otto2011PAR,bachtrog2013NRG}, we still know little about the influence of the HR effect on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
 The timescales over which selection affects neutral polymorphism on sex chromosomes is therefore not well understood.
+
 Recent studies of young sex chromosome system in plants have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}.
 Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in young plant sex chromosomes.
-Given that both background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost a large proportion of their genes \citep{bachtrog2008temporal}, studying recently evolved plant sex chromosomes provides an important test of the temporal dynamics of the HR effect.
+Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved plant sex chromosomes provides an important test of the temporal dynamics of the HR effect.
 
@@
 hastatulus }(Polygonaceae).
-This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than those in mammals \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has received particular attention because of the unique occurrence of an intraspecific polymorphism for sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct chromosomal races \citep{smith1963mechanism}.
-The $XY_{1}XY_{2}$ sex chromosome system (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of the sex chromosome systems in this species, there is evidence that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration, though less degeneration was inferred for younger sex-linked genes unique to the North Carolina race \citep{hough2014}.
+This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than the well-studied mammalian sex chromosomes \citep{lahn1999,ross2005dna}.
+\textit{R.
+hastatulus} has also received particular attention because of the unique occurrence of an intraspecific polymorphism for sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct chromosomal races \citep{smith1963mechanism}.
+The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
+Despite the recent origin of sex chromosomes in both races, there is evidence that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration\citep{hough2014}.
 
-Of particular relevance to the present study is that \textit{R.
-hastatulus} populations throughout their geographical range (southern USA) consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
-This is important because female-biased sex ratios are predicted to affect levels of sex chromosome diversity \citep{ellegren2009}, though few studies have attempted to empirically distinguish between the effects of biased sex ratios and selective interference.
-Here, focusing on the $Y_{1}$ chromosome that is shared between the chromosomal races of \textit{R.
-hastatulus}, we tested whether levels of nucleotide diversity on X, Y, and autosomal chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success.
-In addition, we use population genetic parameters estimated from our data to conduct forward-time simulations of both purifying and positive selection to determine the relative support for models of background selection and selective sweeps in explaining the levels of diversity we observed.
+Of particular relevance to the present study, \textit{R.
+hastatulus} populations throughout their geographical range (southern USA) have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
+This is important because female-biased sex ratios are predicted to affect levels of sex chromosome diversity \citep{ellegren2009}.
+Here, we focus on the $Y_{1}$ chromosome that is shared between the chromosomal races of \textit{R.
+hastatulus}, and investigated patterns of nucleotide diversity on X, Y, and autosomal chromosomes.
+Our goal was to determine whether the levels of neutral diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection (background selection and/or selective sweeps) has likely had a greater impact on polymorphism levels.
+To do this, we used empirically-estimated population genetic parameters to conduct forward-time simulations of purifying and positive selection and evaluated the relative support for these models in explaining the levels of diversity observed in our data.
 
@@
 Samples of both sex chromosomal races were collected in 2010 from throughout the native range of \textit{R.
-hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
+hastatulus} (locations in Table S1\X), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
 We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich).

### 2016-08-01 supporting-information/supporting_info.tex
supp infor edits
supp info, tables, edits

@@
 \begin{document}
+\onecolumn
 
@@
 
-\subsection*{SI Figures}
+\subsection*{SI Tables}
 
-\subsection*{SI Tables}
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
-
-\begin{table}[tbhp!]
+\begin{table}[h!]
 \centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
+\caption{Population identities (ID) and location information for \textit{R. hastatulus} samples from Texas and North Carolina}
 \begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
+Population ID & Location & Altitude & Latitude & Longitude \\
 \midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+TX-MTP & Mount Pleasant, Texas & 130	 & 33.17453 & 94.98799 \\
+OK-RAT & Rattan, Oklahoma & 138 & 34.15755 & 95.41325 \\
+TX-LIV & Livingston, Texas & 83 & 30.69947 & 94.79981 \\ 
+LA-DER & De Ridder, Lousiana & 67 & 30.8941 & 93.3143 \\
+TX-ATH & Athens, Texas & 145 & 32.18471 & 95.8032 \\
+OK-WIL & Willis, Oklahoma & 211 & 33.89663 & 96.83533 \\
 \bottomrule
@@
 
+North Carolina				
+Population ID	Location	Altitude	Latitude	Long
+SC-PRO	Prosperity, South Carolina	126	34.10792	81.43711
+GA-BEL	Belfast, Georgia	15	31.84293	81.28405
+GA-STA	Statesboro, Georgia	78	32.45237	81.84849
+SC-BRA	Branchville, South Carolina	34	33.25082	80.80761
+FL-HAM	Hammock, Florida	3	29.06816	82.64664
+GA-GLA	Gladys, Georgia	97	31.48198	83.23783
+
+\subsection*{SI Figures}
+
 \subsection*{SI Datasets}
-
-Supply Excel (.xls), RTF, CSV, TXT, or PDF files. This file type will be published in raw format and will not be edited or composed.
-
-\subsection*{SI software}
+links:
 

### 2016-08-01 supporting-information/supporting_info.tex
supp info

--- 2016-07-31 supporting-information/supporting_info.tex
@@
-\documentclass[9pt,twocolumn,twoside]{pnas-new}
+\documentclass[9pt,onecolumn,twoside]{pnas-new}
 % Remove the twocolumn option to create a single column SI file if required. 
@@
 
-This supporting information may also be accessed at gihub at https://github.com/houghjosh/XYdiversity
+All supporting information can be obtained from https://github.com/houghjosh/XYdiversity.
 
-\subsection*{SI Figures}
-
-\subsection*{SI Tables}
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
+\section*{SI Tables}
 
@@
 \centering
-\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-\begin{tabular}{lrrr}
-Species & CBS & CV & G3 \\
+\caption{Population identities (ID) and location information for \textit{R. hastatulus} samples from Texas and North Carolina}
+\begin{tabular}{lllll}
+Population ID & Location & Altitude & Latitude & Longitude \\
 \midrule
-1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+\textbf{Texas} &  &  &  &  \\
+TX-MTP & Mount Pleasant, Texas & 130	 & 33.17453 & 94.98799 \\
+OK-RAT & Rattan, Oklahoma & 138 & 34.15755 & 95.41325 \\
+TX-LIV & Livingston, Texas & 83 & 30.69947 & 94.79981 \\ 
+LA-DER & De Ridder, Lousiana & 67 & 30.8941 & 93.3143 \\
+TX-ATH & Athens, Texas & 145 & 32.18471 & 95.8032 \\
+OK-WIL & Willis, Oklahoma & 211 & 33.89663 & 96.83533 \\
+\textbf{North Carolina} &  &  &  &  \\
+SC-PRO & Prosperity, South Carolina & 126 & 34.10792 & 81.43711 \\
+GA-BEL & Belfast, Georgia & 15 & 31.84293 & 81.28405 \\
+GA-STA & Statesboro, Georgia & 78 & 32.45237 & 81.84849 \\
+SC-BRA & Branchville, South Carolina & 	34 & 33.25082 & 80.80761 \\
+FL-HAM & Hammock, Florida & 3 & 29.06816 & 82.64664 \\
+GA-GLA & Gladys, Georgia	 & 97 & 31.48198 & 83.23783 \\
 \bottomrule
 \end{tabular}
-
-\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
 \end{table}
 
-\subsection*{SI Datasets}
+\section*{SI Figures}
 
-Supply Excel (.xls), RTF, CSV, TXT, or PDF files. This file type will be published in raw format and will not be edited or composed.
-
-\subsection*{SI software}
+\section*{SI Datasets}
 

### 2016-08-01 manuscript.tex
fixed bibliography issue
minor edits

@@
 %\setlength{\marginparwidth}{1.25cm}
-\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
-\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
-\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
-\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
+%\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+%\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
+%\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
+%\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
 \newcommand{\X}{\textcolor{red}{\bf X\,}}
@@
 hastatulus} populations throughout their geographical range (southern USA) have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
-This is important because female-biased sex ratios are predicted to affect levels of sex chromosome diversity \citep{ellegren2009}.
-Here, we focus on the $Y_{1}$ chromosome that is shared between the chromosomal races of \textit{R.
-hastatulus}, and investigated patterns of nucleotide diversity on X, Y, and autosomal chromosomes.
-Our goal was to determine whether the levels of neutral diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection (background selection and/or selective sweeps) has likely had a greater impact on polymorphism levels.
-To do this, we used empirically-estimated population genetic parameters to conduct forward-time simulations of purifying and positive selection and evaluated the relative support for these models in explaining the levels of diversity observed in our data.
+This is important because female-biased sex ratios are predicted to affect genetic diversity on sex chromosomes and autosomes \citep{ellegren2009}.
+Here, we investigated the factors affecting nucleotide diversity on X, Y, and autosomal chromosomes.
+Our goal was to determine whether levels of diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection has been the dominant factor affecting patterns of neutral polymorphism on sex chromosomes.
 
@@
 \subsection*{Simulations of Background Selection and Selective Sweeps}
-To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a
-
-\josh{
-\textit{to do}
-\begin{itemize}
-\item \textit{describe parameters used and justify}
-\item \textit{do both purifying and positive selection sims}
-\end{itemize}
-}
-
-We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}.
+To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations.....We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}.
 To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data.

### 2016-08-01 manuscript.tex
ms edits, FigureS1

@@
 We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations, 6 populations of each sex chromosome race).
-Samples of both sex chromosomal races were collected in 2010 from throughout the native range of \textit{R.
-hastatulus} (locations in Table S1\X), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
+Samples of both sex chromosomal race were collected in 2010 from throughout the native range of \textit{R.
+hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
 We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich).
@@
 SRP041588.
-Reads were mapped using the Burrows–Wheeler Aligner (release 0.6.2-r126 \citep{li2010fast}), followed by Stampy (release 1.0.20; \citep{lunter2011stampy}).
-We used Picard tools (release 1.78, http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit (GATK version 2.1-11; \citep{mckenna2010genome}) variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores (<20).
+Reads were mapped using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy\citep{lunter2011stampy}.
+We used Picard tools (http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores <20.
 The population samples analyzed here were previously reported in Hough et al.
@@
 Because the criteria for identifying autosomal loci in Hough et al.
-(2014) were based on the occurrence of four segregating SNPs per locus, and this set is therefore probably higher in diversity than the average autosome, here we incorporated a broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
-We filtered this set of non-sex linked genes to remove those that may have been sex-linked but were not identified as such by the ascertainment criteria in Hough et al.
+(2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosome.
+Here, we incorporated the broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
+We filtered this set to remove those that may have been sex-linked but were not identified as such by the ascertainment criteria in Hough et al.
 (2014).
 In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes with SNPs showing fixed heterozygosity in males and fixed homozygosity in females, (iii) any genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicate genes or those with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
-Finally, we removed genes with fewer than 100 synonymous sites following filtering to avoid biasing results toward genes that may have been particularly short due to assembly problems.
+Finally, we removed genes with fewer than 100 synonymous sites following this filtering to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
 This filtering resulted in a set of 12,356 and 11,350 autosomal genes in the Texas and North Carolina races, respectively.
@@
 hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross.
-Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes in a diploid individual to output a list of phased haplotype blocks containing the SNP variants on each chromosome.
+Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome.
 Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences with each haplotype block (i.e., sites in which all females were homozygous, and all males were heterozygous).
@@
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
-Finally, we filtered SNPs for coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels.
-This procedure was conducted using a combination of Perl and Bash and this resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes from the 24 individuals in our study.
+Finally, we filtered out SNPs with coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels.
+This procedure was conducted using a combination of Perl and Bash scripts, and resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes from the 24 individuals in our study.
 
@@
 To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data).
-In particular, we looked for cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was zero, since Y-linked  sequences from a single family should all be identical, as we had confirmed through phasing by segregation.
-We identified 7 percent of sex-linked genes that had phasing errors of this kind, or that we determined to have genotyping errors resulting in a false SNP calls on the Y chromosome, leading to a SNP error rate estimate of 1.7 x 10-4.
+In particular, we looked for cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was inferred to be zero by phasing-by-segregation.
+We identified 7 percent of sex-linked genes with phasing errors of this kind, or that were determined to have genotyping errors resulting in a false SNP calls on the Y chromosome, with a SNP error rate estimate of 1.7 x 10-4.
 This rate is very low relative to the population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism.
@@
 For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
-To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1).
+To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1).
 We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate.
-To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to obtain independent estimates of the Y/A ratio (\X Supp.
-Figures).
-Here, \X{...explain more}.
+Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated such that the true uncertainty in our Y/A diversity estimate obtained by this method is likely underestimated.
+To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (see Figure S1).
+We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
+As no significant difference was observed (\X Figure S2), we report only results from females.
 
-To test for population substructure between the two sex chromosome races, which could affect estimates diversity, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+Finally, we tested for population substructure between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations might have affected estimates of diversity.
+To do this we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
 The analysis was conducted on an alignment of X- and Y-linked genes from \textit{R.
@@
 2004).
-
-Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males.
-As no significant difference was observed, we report only results from females.
 

### 2016-08-01 manuscript.tex
ms edits

@@
 We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
-As no significant difference was observed (\X Figure S2), we report only results from females.
+As no significant difference was observed (\X Figure?
+Table?), we report only results from females.
 
 Finally, we tested for population substructure between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations might have affected estimates of diversity.
-To do this we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+To do this, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
 The analysis was conducted on an alignment of X- and Y-linked genes from \textit{R.
 hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R.
-bucephalophorus} used to root the tree.
-Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987).
-The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al.
-2004).
+bucephalophorus} used to root the tree (\X Figure S2).
+The phylogeny was inferred using the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
+The inferred sequence phylogeny revealed strong support for the Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, which is more closely related to the XY (Texas) race.
+In light of this evidence for substructure among our samples, we used the inferred phylogeny in our estimation of X and Y diversity.
 

### 2016-08-01 manuscript.tex
ms edits, FigureS2, renaming

@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number.
+To test whether our estimates of neutral diversity on sex chromosomes and autosomes could be explained by biased sex ratios or variance in reproductive success, we evaluated the fit of our data to predictions from a standard neutral model.
+Here, for a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
+Similarly, for the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.
+Thus, with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
+For populations with female-biased sex ratios, however, $N_{m}$ becomes small relative to $N_{f}$ and this ratio can become larger than 1, approaching 1.125 at the limit \citep{caballero1995}.
+With the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
+Figure 1 shows the predicted effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
+
+%with the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$
+
+%In particular, Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number.
 For a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes relative to autosomes is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
-In a neutral model with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$.
+With equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$.
 For populations with female-biased sex ratios, however, $N_{m}$ becomes small relative to $N_{f}$ and this ratio can become larger than 1, approaching 1.125 at the limit \citep{caballero1995}.
@@
 Assuming equal neutral mutation rates for all genes and an equal number of reproducing males and females, Y-linked genes should therefore have 1/3 the polymorphism of X-linked genes, and 1/4 that of autosomal genes.
-Figure 1 shows the predicted relative effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
+Figure 1 shows the predicted effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
 

### 2016-08-01 supporting-information/supporting_info.tex
ms edits, FigureS1
ms edits, FigureS2, renaming

@@
 
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=.5\linewidth]{FigureS1.png}
+\caption{MLHKA likelihood estimation of Y/A ratio. Explain more.}
+\label{figure:FigureS1}
+\end{figure}
+
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=.9\linewidth]{FigureS2.png}
+\caption{Evolutionary relationships of sex chromosome races in \textit{Rumex hastatulus}, inferred using the Neighbor-Joining method (Saitou and Nei 1987).
+The percentage of replicate trees in which the associated sequences clustered together in the bootstrap test (1000 replicates) is shown next to branches (Felsenstein 1985).
+The phylogeny is drawn to scale, with branch lengths in the same units as those of the evolutionary distances used to infer the tree.
+The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al.
+2004) and are in the units of the number of base substitutions per site.
+The analysis was conducted on an alignment of X- and Y-linked genes from \textit{R.
+hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{Rumex bucephalophorus} used to root the tree.
+‘N’ designates the North Carolina race, and ‘T’ the Texas race, with numbers 1-6 corresponding to each of the 6 populations sampled.
+The inferred SC (South Carolina), and FL (Florida) sub-clades are indicated.}
+\label{figure:FigureS2}
+\end{figure}
+
 \section*{SI Datasets}

### 2016-08-02 manuscript.tex
edits to neutral predictions section

@@
 
-Finally, we tested for population substructure between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations might have affected estimates of diversity.
+Finally, we tested for population substructure between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
 To do this, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
@@
 hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R.
-bucephalophorus} used to root the tree (\X Figure S2).
+bucephalophorus} used to root the tree (Figure S2).
 The phylogeny was inferred using the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
-The inferred sequence phylogeny revealed strong support for the Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, which is more closely related to the XY (Texas) race.
-In light of this evidence for substructure among our samples, we used the inferred phylogeny in our estimation of X and Y diversity.
+The inferred sequence phylogeny revealed strong support for the Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, but more closely related to the XY (Texas) race.
+In light of this evidence for substructure among our samples, we used the inferred phylogeny (Figure S2) as the basis for our estimates of X and Y diversity.
 
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimates of neutral diversity on sex chromosomes and autosomes could be explained by biased sex ratios or variance in reproductive success, we evaluated the fit of our data to predictions from a standard neutral model.
-Here, for a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
+To test whether the levels of neutral diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical estimates of diversity to predictions from a neutral model that jointly considers these two effects.
+Here, following Nomura (2002), the effective population sizes for autosomal loci are given by:
+
+\begin{equation}
+N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{d{f}})+N_{f}(1+C^2_{d{m}})}\label{eq:NeA}
+\end{equation}
+
+where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{d{i}}$ is the coefficient of variation in in mating success for the $i$th sex.
+Similarly, for X loci,
+
+\begin{equation}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{d{f}})+2N_{f}(1+C^2_{d{m}})}\label{eq:NeX}
+\end{equation}
+
+and
+
+\begin{equation}
+N_{e{Y}} = \frac{N_{m}(1+C^2_{d{f}})}{2}\label{eq:NeY}
+\end{equation}
+
+Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3 reduces to the familiar expressions:
+
+\begin{equation}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}},\label{eq:NeX}
+\end{equation}
+
+and
+
+\begin{equation}
+N_{e{Y}} = \frac{N_{m}}{2}\label{eq:NeY}
+\end{equation}
+
+respectively.
+To consider the joint effects of biased sex ratios and variance in male reproductive success (which is more likely in a wind-pollinated annual), our neutrally-expected X/A and Y/A levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{d{f}}=0$) and we considered values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
+With the level of neutral polymorphism in a population equalling the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success (assuming equal mutation rates among genes).
+
+%The effect of biased reproductive sex ratios on the $N_{e}$ for an autosomal gene can be estimated using the equation \citep{wright1931evolution}:
+
+%\begin{equation}
+%N_{e} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:Ne}
+%\end{equation}
+
+%where $N_{m}$ and $N_{f}$ are the numbers of reproducing males and females, respectively.
+Similarly, for an X-linked gene, where females contain two-thirds and males one-third of the alleles, the effective population size is given by:
+
+%\begin{equation}
+%N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}}\label{eq:NeX}
+%\end{equation}
+
+%And, for the male-limited Y chromosome:
+
+%\begin{equation}
+%N_{e{Y}} = \frac{N_{m}}{2}\label{eq:NeY}
+%\end{equation}
+
+%\begin{equation}
+%\frac{N_{e_{X}}}{N_{e_{A}}} = \frac{9(N_{f}+N_{m})}{8(2N_{f}+N_{m})}\label{eq:X/A}
+%\end{equation}
+
+%and,
+
+%\begin{equation}
+%\frac{N_{e_{Y}}}{N_{e_{A}}} = \frac{9(N_{f}+N_{m})}{8(2N_{f}+N_{m})}\label{eq:X/A}
+%\end{equation}
+
+ Given a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
 Similarly, for the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.

### 2016-08-02 Rnotebook/ne.Rmd
Rnotebook, major edits on Ne section

---
title: "Nucleotide diversity predictions"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

```{r}
Nf=100 #number of females
Nm=100 #number of males
N=Nf+Nm #total number of individuals
x=Nm/(Nf+Nm) #proportion of males

op <- par(mfrow = c(1, 2))
m=0
f=0

curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m))))/N, from=0, to=1, lty=2,
      xlab="proportion of males", ylab="NeA/N", main="Autosomes", ylim=c(0,1), xlim=c(0,1))

mh=0.5
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh))))/N, from=0, to=1, add=T)

m1=1
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1))))/N, from=0, to=1, add=T)

m2=2
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2))))/N, from=0, to=1, add=T)

m3=3
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3))))/N, from=0, to=1, add=T)

#for sex-linked genes
curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/N, from=0, to=1, lty=2, xlab="proportion of males", ylab="Ne/N", main="Sex chromosomes", ylim=c(0,1),xlim=c(0,1))

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+mh))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m1))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m2))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m3))))/N, from=0, to=1, add=T)

curve((((N*(x)*(1+f))/2))/N, from=0, to=1, col="red", add=T)

par(op)

```

```{r}

ol <- par(mfrow = c(1, 2))

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, lty=2,
      xlab="proportion of males", ylab="Relative Ne", main="X/A")

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+mh))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m1))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m2))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m3))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T)

#unsure if this is right; substituting in the modified Nm's here based on crow model where Nm*(km-1))/((km-1)+(x/km)) is the effect of male rep variance on male Ne.

curve(((N*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, col="red", lty=2, ylim=c(0,1), xlab="proportion of males", ylab="Relative Ne", main="Y/A")

curve(((80*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T, col="red")

curve(((66.66667*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T, col="red")

curve(((50*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T, col="red")

curve(((40*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T, col="red")

par(ol)
```

```{r}
curve(((N*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, col="red", lty=2, ylim=c(0,1), xlab="proportion of males", ylab="Relative Ne", main="Y/A")

curve(((80*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T, col="red")

curve(((66.66667*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T, col="red")

curve(((50*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T, col="red")

curve(((40*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T, col="red")

```

### 2016-08-02 manuscript.tex
Rnotebook, major edits on Ne section

@@
 
-\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
+\subsection*{Estimating polymorphism levels on sex chromosomes and autosomes}
 For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
@@
 
-Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3 reduces to the familiar expressions:
+Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3, respectively, reduce to the familiar expressions from \citep{wright1931evolution}:
 
@@
 
-respectively.
 To consider the joint effects of biased sex ratios and variance in male reproductive success (which is more likely in a wind-pollinated annual), our neutrally-expected X/A and Y/A levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{d{f}}=0$) and we considered values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).

### 2016-08-02 Rnotebook/crap.R

@@ -1,0 +1,15 @@
+kf=2 #mean number of offspring of females
+km=2 #mean number of offspring of males
+vm=2
+
+#effect of male reproductive variance on male Ne
+curve((Nm*(km-1))/((km-1)+(x/km))/N, from = 0, to = 1)
+
+#1:1
+mvalue=0.25
+mhvalue=0.125
+m1value=0.124875
+m2value=0.125
+m3value=0.125
+
+barplot(c(mvalue,mhvalue,0.14375,0.141525,0.14))

### 2016-08-02 Rnotebook/ne.Rmd
manuscript and Figure1 edits

@@
 curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m))))/N, from=0, to=1, lty=2,
-      xlab="proportion of males", ylab="NeA/N", main="Autosomes", ylim=c(0,1), xlim=c(0,1))
+      xlab=expression(italic('N'[m]/'N'[m]+'N'[f])),
+      ylab=expression(italic('N'[e]/N)), 
+      main="Autosomes", ylim=c(0,1), xlim=c(0,1))
 
@@
 #for sex-linked genes
-curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/N, from=0, to=1, lty=2, xlab="proportion of males", ylab="Ne/N", main="Sex chromosomes", ylim=c(0,1),xlim=c(0,1))
+curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[m]/'N'[m]+'N'[f])), 
+      ylab=expression(italic('N'[e]/N)), 
+      main="Sex chromosomes", ylim=c(0,1),xlim=c(0,1))
 

### 2016-08-02 manuscript.tex
manuscript and Figure1 edits

@@
 The phylogeny was inferred using the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
-The inferred sequence phylogeny revealed strong support for the Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, but more closely related to the XY (Texas) race.
-In light of this evidence for substructure among our samples, we used the inferred phylogeny (Figure S2) as the basis for our estimates of X and Y diversity.
+The inferred sequence phylogeny revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, but more closely related to the XY (Texas) race (Figure S2).
+In light of this evidence for substructure, we used the inferred phylogeny as the basis for estimating X and Y diversity.
 
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether the levels of neutral diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical estimates of diversity to predictions from a neutral model that jointly considers these two effects.
+To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical estimates of diversity to predictions from a neutral model that jointly considers these two effects.
 Here, following Nomura (2002), the effective population sizes for autosomal loci are given by:
@@
 \begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{d{f}})+N_{f}(1+C^2_{d{m}})}\label{eq:NeA}
+N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{f})+N_{f}(1+C^2_{m})}\label{eq:NeA}
 \end{equation}
 
-where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{d{i}}$ is the coefficient of variation in in mating success for the $i$th sex.
-Similarly, for X loci,
+where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, given by the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
+Similarly, for X and Y linked loci,
 
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{d{f}})+2N_{f}(1+C^2_{d{m}})}\label{eq:NeX}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{f})+2N_{f}(1+C^2_{m})}\label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e{Y}} = \frac{N_{m}(1+C^2_{d{f}})}{2}\label{eq:NeY}
+N_{e{Y}} = \frac{N_{m}(1+C^2_{f})}{2}\label{eq:NeY}
 \end{equation}
 
-Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3, respectively, reduce to the familiar expressions from \citep{wright1931evolution}:
+With with no variation in mating success in either sex ($C^2_{m}=C^2_{f}=0$), equations 1-3 reduce to the familiar expressions from \citep{wright1931evolution}:
 
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}},\label{eq:NeX}
+N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:NeA}
 \end{equation}
 
-and
+\begin{equation}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}}\label{eq:NeX}
+\end{equation}
 
@@
 
-To consider the joint effects of biased sex ratios and variance in male reproductive success (which is more likely in a wind-pollinated annual), our neutrally-expected X/A and Y/A levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{d{f}}=0$) and we considered values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-With the level of neutral polymorphism in a population equalling the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success (assuming equal mutation rates among genes).
+To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{f}=0$) and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
+Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success.
+Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R.
+hastatulus}, where males produce copious amounts of pollen and compete to fertilize the uniovulate females.
 
@@
 
- Given a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
+ %Given a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
 Similarly, for the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.

### 2016-08-02 Rnotebook/Figure1.R

@@ -1,0 +1,45 @@
+Nf=100 #number of females
+Nm=100 #number of males
+N=Nf+Nm #total number of individuals
+x=Nm/(Nf+Nm) #proportion of males
+
+op <- par(mfrow = c(1, 2))
+m=0
+f=0
+
+curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[m]/'N'[m]+'N'[f])),
+      ylab=expression(italic('N'[e]/N)), 
+      main="Autosomes", ylim=c(0,1), xlim=c(0,1))
+
+mh=0.5
+curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh))))/N, from=0, to=1, add=T)
+
+m1=1
+curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1))))/N, from=0, to=1, add=T)
+
+m2=2
+curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2))))/N, from=0, to=1, add=T)
+
+m3=3
+curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3))))/N, from=0, to=1, add=T)
+
+#for sex-linked genes
+curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[m]/'N'[m]+'N'[f])), 
+      ylab=expression(italic('N'[e]/N)), 
+      main="Sex chromosomes", ylim=c(0,1),xlim=c(0,1))
+
+curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+mh))))/N, from=0, to=1, add=T)
+
+curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m1))))/N, from=0, to=1, add=T)
+
+curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m2))))/N, from=0, to=1, add=T)
+
[…]

### 2016-08-02 manuscript.tex
finished section on Ne predictions

@@
 Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R.
-hastatulus}, where males produce copious amounts of pollen and compete to fertilize the uniovulate females.
+hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females.
 
@@
 \includegraphics[width=\linewidth]{Figure1.png}
-\caption{\textbf{A:} The relation between relative effective population size and sex ratio bias for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y).
-The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})=0$, ranging from 0 to 1, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively.
-Calculations assume a constant population size, non-overlapping generations, and a Poisson distribution of offspring number \citep{wright1931evolution,charlesworth2009effective}.
-The $N_{e}$ for each chromosome type is calculated relative to the predicted $N_{e}$ for genes on autosomes with an equal sex ratio.\textbf{B:} The corresponding expected relation between sex ratio bias and the sex-chromosome-to-autosome diversity ratio, assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates for all genes.
+\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}; Y chromosome in red).
+The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$.
+Dotted lines show predictions where both males and females produce Poisson-distributed offspring numbers.
+In that case, the $N_{e}$'s are given by $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}.
+The other curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
+Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate our null expectations for the X/A and Y/A ratios of diversity.
 }%

### 2016-08-02 manuscript.tex
minor edits

@@
 \includegraphics[width=\linewidth]{Figure1.png}
-\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}; Y chromosome in red).
+\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B} (Y chromosome in red).
 The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$.
@@
 In that case, the $N_{e}$'s are given by $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}.
-The other curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
+The solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
 Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate our null expectations for the X/A and Y/A ratios of diversity.
@@
 
-\subsection*{Simulations of Background Selection and Selective Sweeps}
-To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations.....We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}.
-To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data.
-To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$.
-We then calculated the average $\theta$ over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure).
-Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
-
-\begin{equation}
-S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
-      = \frac{1}{n}\sum_{i}^{n} X_i
-\label{eq:refname1}
-\end{equation}
+\subsection*{Simulations of positive and purifying selection}
+To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations, modeling both purifying and positive selection on the Y chromosome using the software SFSCODE \citep{hernandez2008flexible}.
+To make the simulation output comparable to our data, we initialized simulations with an empirically estimated value of $\theta$ (from autosomes), and sampled 8 Y chromosomes per simulation from 24 populations.
+To consider the effects of selection against deleterious alleles (purifying selection), we sampled selective coefficients from a gamma distribution with shape parameter $k=0.1$, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid).
+For positive selection simulations, we used \X [parameters here].
+We ran 1000 replicates for each model and we calculated the average $\theta_{sim}$ over all simulation output to obtain the simulated Y/A ratio of neutral diversity under the both models (purifying and positive selection), taking into account unequal male and female population sizes.
+We compared simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+Commands for running the simulations are provided in the Supplementary Material.
 

### 2016-08-03 manuscript.tex
results section edits

@@
 
-\subsection*{Evolutionary relationships of sex chromosome races in \textit{R. hastatulus}}
+\subsection*{Y-chromosome diversity in \textit{R. hastatulus} is very low}
+Our analysis of polymorphism in \textit{R.
+hastatulus} revealed that diversity on the Y chromosome is significantly lower than neutral predictions, with estimates indicating $Y/A=0.02$, a 12.5 fold reduction compared to the neutrally expected $Y/A = 0.25 (P<0.0001)$.
+This corresponds to a 50-fold reduction compared to the mean diversity estimated on autosomes (\X), and a 40-fold reduction compared to X chromosome diversity (Table 1).
+By normalizing X- and Y-chromosome diversity estimates by autosomal diversity, these results suggest that the X-Y difference we observe (an \X fold difference) is not due to some process specifically elevating diversity on the X chromosome.
+In addition, a recent analysis of sex-linked substitution rates in this species found that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, were not significantly different \citep{hough2014}.
+This further suggests that the low diversity on the Y is not likely due to a lower mutation rate.
+
+ A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection), ...and several cases....
+but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity.
+Moreover, a further reduction in Y diversity could occure because of high variance in male reproductive success (Figure 1).
+This is not unusual in annual species such as \textit{R.
+hastatulus}s, which commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+Conceivably, such plastic differences among plants, owing to spatial heterogeneity in resources, could influence the intensity of male-male competition, and thus affect variance in male reproductive success.
+Given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, one might expect strong competition among males.
+
+ In common with most flowering plants we do not have marker-based estimates of the variance in male reproductive success in \textit{R.
+hastatulus}.
+However, ...
+by modeling...we evaluated whether such an effect could result in a level of Y/A diversity similar to what we observed by comparing our results to predictions from a  model for the joint effects of sex ratio bias and male reproductive variance (see Methods).
+Estimates of sex ratio bias in \textit{R.
+hastatulus} have been estimated, and they are consistently female-biased, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ (with a mean of 0.62; N=46 \citep{pickup2013influence}.The Y/A diversity ratio predicted from this sex ratio, assuming a Poisson distribution of offspring numbers, is approximately 0.2 (Table1, Figure 3).
+Our estimated mean Y/A ratio, however, is 0.02, which is significantly lower ($\textit{P}<0.0001$).
+Indeed, even for higher values for the variance in male reproductive success (\X), the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
+Finally, neutral models that could not be rejected, i.e., in which a highly female-biased sex ratio (\X) and a high variance in male reproductive success predicted an Y/A ratio that fit our data (\X), simultaneously predicts X/A ratios that are significantly different from what we observed (\X Figure).
+Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X and Y diversity in \textit{R.
+hastatulus}.
+
+ %e.g., X/A diversity ratio that is 30% larger than what we found for the Texas race (X/ATX=0.85), and 80% larger than  the X/A ratio in the North Carolina race (X/ANC=0.35, further discussed below).
 
@@
 
-%figure
-
-%Figure 4.
-Evolutionary relationships of sex chromosome races in Rumex hastatulus, inferred using the Neighbor-Joining method (Saitou and Nei 1987).
-The percentage of replicate trees in which the associated sequences clustered together in the bootstrap test (1000 replicates) is shown next to branches (Felsenstein 1985).
-The phylogeny is drawn to scale, with branch lengths in the same units as those of the evolutionary distances used to infer the tree.
-The evolutionary distances were computed using the Maximum Composite Likelihood method (Tamura et al.
-2004) and are in the units of the number of base substitutions per site.
-The analysis was conducted on an alignment of X- and Y-linked genes from Rumex hastatulus, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species Rumex bucephalophorus used to root the tree.
-‘N’ designates the North Carolina race, and ‘T’ the Texas race, with numbers 1-6 corresponding to each of the 6 populations sampled.
-The inferred SC (South Carolina), and FL (Florida) sub-clades are indicated.
-
 \subsection*{Male-to-female sex ratio of $N_{e}$}

### 2016-08-07 manuscript.tex
manuscript edits, sim code

@@
 The inferred sequence phylogeny revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, but more closely related to the XY (Texas) race (Figure S2).
-In light of this evidence for substructure, we used the inferred phylogeny as the basis for estimating X and Y diversity.
+In light of this evidence for population substructure, we used the inferred phylogeny as the basis for estimating X and Y diversity.
 
@@
 To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{f}=0$) and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success.
+Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
+From these neutral predictions we calculated the normalized X/A and Y/A ratios of diversity as the basis for comparison with our empirical data.
 Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R.
@@
 \subsection*{Simulations of positive and purifying selection}
-To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations, modeling both purifying and positive selection on the Y chromosome using the software SFSCODE \citep{hernandez2008flexible}.
-To make the simulation output comparable to our data, we initialized simulations with an empirically estimated value of $\theta$ (from autosomes), and sampled 8 Y chromosomes per simulation from 24 populations.
+To investigate whether our observed level of Y-chromosome diversity could be explained by selection, we conducted two sets of forward-time simulations using the software SFSCODE \citep{hernandez2008flexible}: (i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection on beneficial mutations reducing variation by selective sweeps (refs \X).
+For background selection simulations, we sampled selection coefficients from a gamma distributions, and investigated a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X).
+mutations rate $\mu=0.01$ as estimated in Hough et al.
+2014,
+
+%The effect of background selection is a function of the distribution of selection coefficients for new, deleterious mutations, and can be modeled by varying the mutation rate, the number of sites affected by selection (L), and the selection coefficient acting on new mutations (s) [21].
+When evaluating models with different strengths of purifying selection, we assumed that selection coefficients for the nonsynonymous sites were drawn from a gamma distribution.
+Previous studies found this distribution to fit the observed autosomal frequency spectrum well [37,38,68,69], and there is little reason to believe that the shape of the gamma distribution varies across chromosomes.
+
+To make the simulation output comparable to our data, we initialized simulations with our empirically estimated value of $\theta$ (from autosomes), and sampled 8 Y chromosomes per simulation from 24 populations.
 To consider the effects of selection against deleterious alleles (purifying selection), we sampled selective coefficients from a gamma distribution with shape parameter $k=0.1$, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid).

### 2016-08-08 manuscript.tex
edits, table1

@@
 % Use the documentclass option 'lineno' to view line numbers
-
+\usepackage{booktabs}
+\usepackage{siunitx}
+\usepackage{geometry} % just not to bother with the table width
 %\usepackage[colorinlistoftodos]{todonotes} % comments in margins
@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in a reduction in $N_{e}$ and the level of neutral polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in a reduction in $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, our analysis revealed that nucleotide diversity on the Y is approximately 12.5 fold lower than predicted from models of neutral evolution.
-We demonstrate that the magnitude of this reduction is unlikely to be explained by the observed occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of background selection arising from suppressed recombination.
+Strikingly, we find that nucleotide diversity on the Y chromosome is 40 fold lower than on the X chromosome.
+We demonstrate that the magnitude of this reduction is not consistent with a purely neutral model based on the observed occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of purifying selection removing deleterious mutations and linked neutral variation.
 Given the recent origin of \textit{R.
 hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with background selection playing an important role during the early stages of their evolution.
-
-%this is 254 words and cannot be any longer than 250. Make edits accordingly.
-
-%Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization}
-
 \end{abstract}
@@
 \correspondingauthoraffiliation{jhough@ucdavis.edu}
-\vspace{-11pt}%
+\vspace{-11pt}
 
@@
 
-In addition to demographic factors, evolutionary theory predicts that genetic variability on sex chromosomes should also be affected by the evolution of suppressed X-Y recombination, which lowers the coalescent $N_{e}$ for Y-linked genes and increases their vulnerability to the effects of selective sweeps \citep{smith1974hitch,aquadro1994} and background selection \citep{charlesworth1996background,charlesworth1994effect}.
+In addition to demographic factors, evolutionary theory predicts that genetic variability on sex chromosomes should be affected by the evolution of suppressed X-Y recombination, which lowers the coalescent $N_{e}$ for Y-linked genes and increases their vulnerability to the effects of selective sweeps \citep{smith1974hitch,aquadro1994} and background selection \citep{charlesworth1996background,charlesworth1994effect}.
 These processes can be viewed as examples of the Hill–Robertson (HR) effect, in which the effective population size for a given chromosomal region depends strongly on the rate of recombination, such that selection at a focal site is not independent of selection at nearby sites \citep{hill1966HReffect}.
@@
 
-Despite widespread interest in determining the effects of recombination on patterns of genetic diversity and the forces contributing to sex chromosome divergence \citep{ellegren2011,otto2011PAR,bachtrog2013NRG}, we still know little about the influence of the HR effect on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
-The timescales over which selection affects neutral polymorphism on sex chromosomes is therefore not well understood.
-
-Recent studies of young sex chromosome system in plants have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}.
-Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in young plant sex chromosomes.
-Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved plant sex chromosomes provides an important test of the temporal dynamics of the HR effect.
+Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of the HR effect on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
+Recent studies of young plant sex chromosome systems have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}.
+Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in plant sex chromosomes.
+Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR effect.
 
@@
 This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than the well-studied mammalian sex chromosomes \citep{lahn1999,ross2005dna}.
-\textit{R.
-hastatulus} has also received particular attention because of the unique occurrence of an intraspecific polymorphism for sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct chromosomal races \citep{smith1963mechanism}.
+\textit{Rumex hastatulus} has also received particular attention because of the unique occurrence of an intraspecific sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct 'chromosomal races' \citep{smith1963mechanism}.
 The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, there is evidence that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration\citep{hough2014}.
+Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration\citep{hough2014}.
 
 Of particular relevance to the present study, \textit{R.
-hastatulus} populations throughout their geographical range (southern USA) have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
-This is important because female-biased sex ratios are predicted to affect genetic diversity on sex chromosomes and autosomes \citep{ellegren2009}.
-Here, we investigated the factors affecting nucleotide diversity on X, Y, and autosomal chromosomes.
-Our goal was to determine whether levels of diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection has been the dominant factor affecting patterns of neutral polymorphism on sex chromosomes.
+hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
+This is important because female-biased sex ratios are predicted to affect chromosomal $N_{e}$ and therefor the level of genetic diversity maintained on sex chromosomes and autosomes \citep{ellegren2009}.
+Here, we investigated whether the patterns of diversity we observed on sex chromosomes and autosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection (background selection and selective sweeps) has played a greater role.
 
@@
 \subsection*{Population Samples and Sex-Linked Genes}
-We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations, 6 populations of each sex chromosome race).
-Samples of both sex chromosomal race were collected in 2010 from throughout the native range of \textit{R.
+We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations; 6 populations of each sex chromosome race).
+Samples were collected in 2010 from throughout the native range of \textit{R.
 hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
@@
 SRP041588.
-Reads were mapped using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy\citep{lunter2011stampy}.
+Reads were mapped using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}.
 We used Picard tools (http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores <20.
 The population samples analyzed here were previously reported in Hough et al.
-(2014), where they were used to validate the ascertainment of sex-linked contigs identified through segregation analysis.
-From the list of sex-linked genes identified in Hough et al.
-(2014) that were shared between the Texas (XY) and North Carolina (XY1Y2) sex chromosome races, we identified a total of 460 sex-linked genes for analysis.
+(2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis.
+Here, we focused on the set of 460 sex-linked genes for which the Y homolog was inferred to be on the $y_{1}$ chromosome (i.e., sex-linked genes that are shared between the Texas and North Carolina races).
 
@@
 (2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosome.
-Here, we incorporated the broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
-We filtered this set to remove those that may have been sex-linked but were not identified as such by the ascertainment criteria in Hough et al.
+Therefore, for the present analysis we incorporated the broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
+We also filtered this set to remove genes that may have been sex-linked but were not identified as such by the conservative ascertainment criteria in Hough et al.
 (2014).
@@
 
+\begin{table}
+\centering
+\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
+\begin{tabular}{ccccccccc}
+\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
+chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
+\midrule
+A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\    
+X & 0.0047 & 0.85 & 0.002 & 0.33 & 0.0047 & 0.37 \\ 
+Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\ 
+\addlinespace
+
+\bottomrule
+\end{tabular}
+
+\end{table}
+
  A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection), ...and several cases....

### 2016-08-10 manuscript.tex
supp info edits, ms edits

@@
 Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated such that the true uncertainty in our Y/A diversity estimate obtained by this method is likely underestimated.
-To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (see Figure S1).
+To address this issue, we used a maximum likelihood approach, implemented a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
@@
 To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical estimates of diversity to predictions from a neutral model that jointly considers these two effects.
-Here, following Nomura (2002), the effective population sizes for autosomal loci are given by:
+Here, following Nomura (2002), the effective population sizes for autosomal loci is given by:
 
@@
 
-where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, given by the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
+where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, defined as the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
 Similarly, for X and Y linked loci,
@@
 Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
-From these neutral predictions we calculated the normalized X/A and Y/A ratios of diversity as the basis for comparison with our empirical data.
-Note that we have only considered variance in male reproductive success because this is most likely to occur in wind pollinated plant like \textit{R.
-hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females.
+Using these neutral predictions, we compared our empirically estimated diversity levels to normalized X/A and Y/A ratios of diversity.
+Note that we have only considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
+hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females (\X ref).
 
+%old stuff
 %The effect of biased reproductive sex ratios on the $N_{e}$ for an autosomal gene can be estimated using the equation \citep{wright1931evolution}:
@@
 
-%with the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$
-
 %In particular, Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number.
@@
 \includegraphics[width=\linewidth]{Figure1.png}
-\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B} (Y chromosome in red).
-The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$.
-Dotted lines show predictions where both males and females produce Poisson-distributed offspring numbers.
-In that case, the $N_{e}$'s are given by $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}.
+\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).
+The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$ for sex chromosomes and autosomes is given in equations 1-3.
+Dotted lines show predictions in the standard neutral model, where both males and females produce Poisson-distributed offspring numbers, where $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}.
 The solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
-Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate our null expectations for the X/A and Y/A ratios of diversity.
+Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate null predictions for X/A and Y/A ratios of diversity.
 }%
@@
 \subsection*{Simulations of positive and purifying selection}
-To investigate whether our observed level of Y-chromosome diversity could be explained by selection, we conducted two sets of forward-time simulations using the software SFSCODE \citep{hernandez2008flexible}: (i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection on beneficial mutations reducing variation by selective sweeps (refs \X).
-For background selection simulations, we sampled selection coefficients from a gamma distributions, and investigated a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X).
-mutations rate $\mu=0.01$ as estimated in Hough et al.
-2014,
-
-%The effect of background selection is a function of the distribution of selection coefficients for new, deleterious mutations, and can be modeled by varying the mutation rate, the number of sites affected by selection (L), and the selection coefficient acting on new mutations (s) [21].
-When evaluating models with different strengths of purifying selection, we assumed that selection coefficients for the nonsynonymous sites were drawn from a gamma distribution.
-Previous studies found this distribution to fit the observed autosomal frequency spectrum well [37,38,68,69], and there is little reason to believe that the shape of the gamma distribution varies across chromosomes.
-
+To test whether our observed level of Y-chromosome diversity could be explained by selection, we conducted two sets of forward-time simulations using the software SFSCODE \citep{hernandez2008flexible}: (i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection acting on beneficial mutations and reducing variation through selective sweeps (refs \X).
+For background selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X) from (\X) to \X (see SI text for simulation code).
 To make the simulation output comparable to our data, we initialized simulations with our empirically estimated value of $\theta$ (from autosomes), and sampled 8 Y chromosomes per simulation from 24 populations.
-To consider the effects of selection against deleterious alleles (purifying selection), we sampled selective coefficients from a gamma distribution with shape parameter $k=0.1$, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid).
-For positive selection simulations, we used \X [parameters here].
-We ran 1000 replicates for each model and we calculated the average $\theta_{sim}$ over all simulation output to obtain the simulated Y/A ratio of neutral diversity under the both models (purifying and positive selection), taking into account unequal male and female population sizes.
-We compared simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
-Commands for running the simulations are provided in the Supplementary Material.
+To consider the effects of selection against deleterious alleles (purifying selection), we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ from \X to \X.
+For simulating positive selection, we used \X [parameters here].
+We ran 1000 replicates for each model and we calculated the average $\theta_{sim}$ to obtain the simulated Y/A ratio of neutral diversity under the both sets of models separately (purifying and positive selection).
+We tested the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+Commands for running the simulations are provided in the Supporting Information.
 

### 2016-08-10 manuscript.tex
ms edits

@@
 \usepackage{siunitx}
+\usepackage{adjustbox}
 \usepackage{geometry} % just not to bother with the table width
@@
 \subsection*{Y-chromosome diversity in \textit{R. hastatulus} is very low}
-Our analysis of polymorphism in \textit{R.
-hastatulus} revealed that diversity on the Y chromosome is significantly lower than neutral predictions, with estimates indicating $Y/A=0.02$, a 12.5 fold reduction compared to the neutrally expected $Y/A = 0.25 (P<0.0001)$.
-This corresponds to a 50-fold reduction compared to the mean diversity estimated on autosomes (\X), and a 40-fold reduction compared to X chromosome diversity (Table 1).
-By normalizing X- and Y-chromosome diversity estimates by autosomal diversity, these results suggest that the X-Y difference we observe (an \X fold difference) is not due to some process specifically elevating diversity on the X chromosome.
-In addition, a recent analysis of sex-linked substitution rates in this species found that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, were not significantly different \citep{hough2014}.
-This further suggests that the low diversity on the Y is not likely due to a lower mutation rate.
+Our analysis revealed that diversity on the \textit{R.
+hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating $Y/A=0.02$, which is 12.5 fold lower than the standard neutral prediction of $Y/A = 0.25 (P<0.0001)$, and a 40-fold reduction compared to mean diversity on the X chromosome (Table 1).
+Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but a Y-specific diversity reduction.
+Conceivably, low diversity on the Y could arise from a low Y-chromosome mutation rate or a lower mutation rate in males compared to females.
+However, recent evidence indicates that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, were not significantly different \citep{hough2014}.
 
-\begin{table}
+% The pattern of reduced diversity on the Y chromosome is observed in both Africans and Europeans, suggesting that the effect is not population-specific, and holds regardless of whether the neutral sequence analyzed is near or far from genes (Table 1).
+Previous
+
+%Diversity on the Y chromosome is likely not being under-estimated due to the inability to call variants in haploid regions of the genome because diversity on the X measured in females, where the X is diploid, is nearly identical to diversity on the X measured in males, where the X is haploid (Figure S2).
+
+\begin{table}[t!]
 \centering
@@
 \end{tabular}
+\end{table}
 
-\end{table}
+%In contrast to the Texas sex chromosome race, our analysis revealed that Y-chromosome diversity was considerably higher in the North Carolina race (see Supplementary Material Figure S1).
+This difference could have arisen due to population substructure, resulting in an elevated overall Y diversity in the pooled sample.
+To test this hypothesis, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+The resulting sequence phylogeny (Figure 4) revealed strong support for the Y-linked genes from the XY1Y2 race being paraphyletic (98% bootstrap support).
+In particular, our analysis revealed that samples from two populations, both from South Carolina (hereafter the SC sub-clade), formed a monophyletic group, whereas samples from Florida and Georgia (hereafter the FL sub-clade) were also monophyletic, but were more closely related to the XY Texas race.
+Although our sampling here is limited, these results suggest that introgression may have occurred between the ancestral Texas (XY race) and the derived North Carolina (XY1Y2) race, leading to a second, derived, XY1Y2 sub-clade and an elevated level of Y-diversity in our pooled sample.
+As these races are inter-fertile (Smith reference about hybrid zone), we suggest that the SC sub-clade likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY race.
+Notably, when we separated samples by the two inferred sub-clades of Y–linked sequences, the estimated level of Y diversity was consistent across all three clades, suggesting that the extensive reduction in diversity on this chromosome is not a population or race-specific phenomenon (see Supplementary Material, Figure S1).
+
+%Our data indicate a significant reduction in X/A polymorphism in the derived SC and FL sub-clades of the North Carolina race (X/AFL=0.33 and X/ASC=0.37) compared to the Texas race (X/ATX=0.85).
+This reduction in diversity may be associated with the recent origin of the XY1Y2 sex chromosome system that occurs in these two sub-clades.
+The XY1Y2 sex chromosome system is thought to have originated through an X-autosome fusion (Smith 1964) involving the ancestral 3rd chromosome in the Texas race.
+Evidence supporting this autosomal origin was recently obtained by Grabowska-Joachimiak et al.
+(2014), who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the Y2 sex chromosomes in the derived North Carolina race.
+That this fusion is likely to have been recent is suggested by C-banding/DAPI experiments showing that while the ancestral Y chromosome is highly heterochromatic, the translocated neo-Y is still euchromatic (Grabowska-Joachimiak et al.
+2014), and also shows less genetic degeneration than the ancestral Y chromosome (Hough et al.
+2014).
+Therefore, if recent positive selection was responsible for driving the evolution of this fusion, then the formerly autosomal segment on the X chromosome in the XY1Y2 sub-clades (the neo-X) may have experienced a selective sweep.
+This process may explain the reduction in diversity that we observed in the derived XY1Y2 sub-clades, though further work is needed to determine the factors driving the establishment of the X-autosome fusion.
+
+%We next consider several possible models that could explain this unexpectedly low amount of diversity found on the Y chromosome relative to other genomic regions.
+Such models include differences in the variance in reproductive success between males and females, purifying selection on the Y chromosome, and positive selection on the Y chromosome.
+
+\subsection*{Female biased sex ratios and variance in male fitness}
 
@@
 
-\subsection*{Male-to-female sex ratio of $N_{e}$}
-
 \subsection*{Background Selection and Selective Sweeps}

### 2016-08-15 manuscript.tex
ms edits

@@
 
-To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{f}=0$) and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-Assuming the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
+To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated using equations 1-3, assuming no variance in female mating (e.g., $C^2_{f}=0$), and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
+Assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
 Using these neutral predictions, we compared our empirically estimated diversity levels to normalized X/A and Y/A ratios of diversity.
@@
 The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$ for sex chromosomes and autosomes is given in equations 1-3.
-Dotted lines show predictions in the standard neutral model, where both males and females produce Poisson-distributed offspring numbers, where $N_{e{A}}=4N_{m}N_{f}/(N_{m}+N_{f})$ for autosomes, $N_{e{X}}=9N_{m}N_{f}/(4N_{m}+2N_{f})$ for the X, and $N_{e{Y}}=N_{m}/2$ for the Y \citep{wright1931evolution}.
-The solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
+Dotted curves show predictions in the standard neutral model, where both males and females produce Poisson-distributed offspring numbers and the chromosomal $N_{e}$'s are given by equations 4-6 \citep{wright1931evolution}.
+Solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
 Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate null predictions for X/A and Y/A ratios of diversity.
@@
 To test whether our observed level of Y-chromosome diversity could be explained by selection, we conducted two sets of forward-time simulations using the software SFSCODE \citep{hernandez2008flexible}: (i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection acting on beneficial mutations and reducing variation through selective sweeps (refs \X).
-For background selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X) from (\X) to \X (see SI text for simulation code).
-To make the simulation output comparable to our data, we initialized simulations with our empirically estimated value of $\theta$ (from autosomes), and sampled 8 Y chromosomes per simulation from 24 populations.
-To consider the effects of selection against deleterious alleles (purifying selection), we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ from \X to \X.
+For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X) from (\X) to \X and keeping the shape parameter, \X, constant (see Supporting Information for simulation code).
+To make the simulation output comparable to our data, we initialized simulations with our empirically estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
 For simulating positive selection, we used \X [parameters here].
-We ran 1000 replicates for each model and we calculated the average $\theta_{sim}$ to obtain the simulated Y/A ratio of neutral diversity under the both sets of models separately (purifying and positive selection).
-We tested the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of models separately (purifying and positive selection).
+We then tested the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
 Commands for running the simulations are provided in the Supporting Information.
@@
 hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating $Y/A=0.02$, which is 12.5 fold lower than the standard neutral prediction of $Y/A = 0.25 (P<0.0001)$, and a 40-fold reduction compared to mean diversity on the X chromosome (Table 1).
-Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but a Y-specific diversity reduction.
-Conceivably, low diversity on the Y could arise from a low Y-chromosome mutation rate or a lower mutation rate in males compared to females.
-However, recent evidence indicates that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, were not significantly different \citep{hough2014}.
-
-% The pattern of reduced diversity on the Y chromosome is observed in both Africans and Europeans, suggesting that the effect is not population-specific, and holds regardless of whether the neutral sequence analyzed is near or far from genes (Table 1).
-Previous
-
-%Diversity on the Y chromosome is likely not being under-estimated due to the inability to call variants in haploid regions of the genome because diversity on the X measured in females, where the X is diploid, is nearly identical to diversity on the X measured in males, where the X is haploid (Figure S2).
+Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
+Conceivably, such low diversity on the Y could arise at least in part from a low mutation rate on the Y, or a lower mutation rate in males compared to females.
+However, these possibilities can be rejected because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
+In addition, the pattern of reduced Y chromosome diversity occurs in each of the \textit{R.
+hastatulus} sub-clades that were inferred to be monophyletic (\X Figure 2; Table 1; Figure S1,S2), suggesting that the effect is not population-specific.
 

### 2016-08-15 manuscript.tex
ms edits

@@
 Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
-Conceivably, such low diversity on the Y could arise at least in part from a low mutation rate on the Y, or a lower mutation rate in males compared to females.
+Conceivably, such low diversity on the Y could arise from a low mutation rate on the Y chromosome, or a lower mutation rate in males compared to females.
 However, these possibilities can be rejected because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
-In addition, the pattern of reduced Y chromosome diversity occurs in each of the \textit{R.
-hastatulus} sub-clades that were inferred to be monophyletic (\X Figure 2; Table 1; Figure S1,S2), suggesting that the effect is not population-specific.
+It is worth noting that our results only apply to the $Y_{1}$ chromosome, however, as estimates of diversity were calculated using the set of sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome races.
+Previous work investigating patterns of genetic degeneration in both systems found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes have undergone classic signs of degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations, but we cannot say at present whether the neo-Y has also undergone a significant reduction in diversity.
+Nonetheless, our results indicate that the pattern of reduced diversity on the $Y_{1}$ chromosome occurs in each of the \textit{R.
+hastatulus} sub-clades that were inferred to be monophyletic (\X Figure 2; Table 1; Figure S1,S2), suggesting that this effect is not population-specific.
 
@@
 
-%In contrast to the Texas sex chromosome race, our analysis revealed that Y-chromosome diversity was considerably higher in the North Carolina race (see Supplementary Material Figure S1).
-This difference could have arisen due to population substructure, resulting in an elevated overall Y diversity in the pooled sample.
-To test this hypothesis, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
-The resulting sequence phylogeny (Figure 4) revealed strong support for the Y-linked genes from the XY1Y2 race being paraphyletic (98% bootstrap support).
-In particular, our analysis revealed that samples from two populations, both from South Carolina (hereafter the SC sub-clade), formed a monophyletic group, whereas samples from Florida and Georgia (hereafter the FL sub-clade) were also monophyletic, but were more closely related to the XY Texas race.
-Although our sampling here is limited, these results suggest that introgression may have occurred between the ancestral Texas (XY race) and the derived North Carolina (XY1Y2) race, leading to a second, derived, XY1Y2 sub-clade and an elevated level of Y-diversity in our pooled sample.
-As these races are inter-fertile (Smith reference about hybrid zone), we suggest that the SC sub-clade likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY race.
-Notably, when we separated samples by the two inferred sub-clades of Y–linked sequences, the estimated level of Y diversity was consistent across all three clades, suggesting that the extensive reduction in diversity on this chromosome is not a population or race-specific phenomenon (see Supplementary Material, Figure S1).
+Although our sampling of each \textit{R.
+hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting as it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a derived $XY_{1}Y_{2}$ sub-clade.
+As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the SC sub-clade inferred here likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY Texas race.
+Notably, when we pooled the samples from each sub-clade together, the estimated level of Y diversity was significantly higher in the pooled data, highlighting the presence of strong substructure between these groups (Figure S2).
 

### 2016-08-15 manuscript.tex
ms edits, figure S3

@@
 To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated using equations 1-3, assuming no variance in female mating (e.g., $C^2_{f}=0$), and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-Assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
-Using these neutral predictions, we compared our empirically estimated diversity levels to normalized X/A and Y/A ratios of diversity.
+Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
+Using these neutral predictions, we compared our empirically estimated diversity levels to normalized X/A and Y/A ratios of diversity, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
 Note that we have only considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
@@
 For simulating positive selection, we used \X [parameters here].
-We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of models separately (purifying and positive selection).
-We then tested the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of models separately.
+We then tested the significant of the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
 Commands for running the simulations are provided in the Supporting Information.
@@
 Our analysis revealed that diversity on the \textit{R.
-hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating $Y/A=0.02$, which is 12.5 fold lower than the standard neutral prediction of $Y/A = 0.25 (P<0.0001)$, and a 40-fold reduction compared to mean diversity on the X chromosome (Table 1).
+hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001), and 40-fold lower compared to mean diversity on the X chromosome (Table 1).
 Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
@@
 However, these possibilities can be rejected because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
-It is worth noting that our results only apply to the $Y_{1}$ chromosome, however, as estimates of diversity were calculated using the set of sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome races.
-Previous work investigating patterns of genetic degeneration in both systems found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes have undergone classic signs of degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations, but we cannot say at present whether the neo-Y has also undergone a significant reduction in diversity.
-Nonetheless, our results indicate that the pattern of reduced diversity on the $Y_{1}$ chromosome occurs in each of the \textit{R.
-hastatulus} sub-clades that were inferred to be monophyletic (\X Figure 2; Table 1; Figure S1,S2), suggesting that this effect is not population-specific.
+
+It is worth noting that our results only apply to the $Y_{1}$ chromosome, however, as our estimates of diversity were calculated using the set of sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
+Previous work investigating patterns of genetic degeneration found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibit signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a significant reduction in diversity.
+However, our results do indicate that the pattern of reduced diversity on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
+hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
 
@@
 Although our sampling of each \textit{R.
-hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting as it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a derived $XY_{1}Y_{2}$ sub-clade.
+hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a derived $XY_{1}Y_{2}$ sub-clade.
 As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the SC sub-clade inferred here likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY Texas race.
@@
 
-%Our data indicate a significant reduction in X/A polymorphism in the derived SC and FL sub-clades of the North Carolina race (X/AFL=0.33 and X/ASC=0.37) compared to the Texas race (X/ATX=0.85).
-This reduction in diversity may be associated with the recent origin of the XY1Y2 sex chromosome system that occurs in these two sub-clades.
+Interestingly, our results indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
+Although not expected, this reduction in diversity may be associated with the recent origin of the XY1Y2 sex chromosome system.
 The XY1Y2 sex chromosome system is thought to have originated through an X-autosome fusion (Smith 1964) involving the ancestral 3rd chromosome in the Texas race.
@@
 (2014), who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the Y2 sex chromosomes in the derived North Carolina race.
-That this fusion is likely to have been recent is suggested by C-banding/DAPI experiments showing that while the ancestral Y chromosome is highly heterochromatic, the translocated neo-Y is still euchromatic (Grabowska-Joachimiak et al.
-2014), and also shows less genetic degeneration than the ancestral Y chromosome (Hough et al.
-2014).
-Therefore, if recent positive selection was responsible for driving the evolution of this fusion, then the formerly autosomal segment on the X chromosome in the XY1Y2 sub-clades (the neo-X) may have experienced a selective sweep.
-This process may explain the reduction in diversity that we observed in the derived XY1Y2 sub-clades, though further work is needed to determine the factors driving the establishment of the X-autosome fusion.
+If recent positive selection was responsible for driving the evolution of this fusion, then the formerly autosomal segment on the X chromosome in the XY1Y2 sub-clades (the neo-X) may have experienced a selective sweep, resulting in reduced X-linked diversity in the derived XY1Y2 sub-clades.
+It will be important for future work to investigate the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
 
-%We next consider several possible models that could explain this unexpectedly low amount of diversity found on the Y chromosome relative to other genomic regions.
-Such models include differences in the variance in reproductive success between males and females, purifying selection on the Y chromosome, and positive selection on the Y chromosome.
+In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now test several possible models that might explain this diversity reduction.
 
@@
 
- A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection), ...and several cases....
-but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity.
-Moreover, a further reduction in Y diversity could occure because of high variance in male reproductive success (Figure 1).
-This is not unusual in annual species such as \textit{R.
-hastatulus}s, which commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+ A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection) (\X refs), ...and several cases....(\X refs), but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity in the absence of such selective interference.
+Moreover, such a reduction in Y-linked diversity could be accented due to high variance in male reproductive success (Figure 1).
+This is not unusual in annual plants such as \textit{R.
+hastatulus}, which commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
 Conceivably, such plastic differences among plants, owing to spatial heterogeneity in resources, could influence the intensity of male-male competition, and thus affect variance in male reproductive success.
-Given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, one might expect strong competition among males.
+Given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, strong competition among males is expected.
 
@@
 
- %e.g., X/A diversity ratio that is 30% larger than what we found for the Texas race (X/ATX=0.85), and 80% larger than  the X/A ratio in the North Carolina race (X/ANC=0.35, further discussed below).
-
-%In contrast to the Texas sex chromosome race, our analysis revealed that Y-chromosome diversity was considerably higher in the North Carolina race (see Supplementary Material; Figure S1).
-This difference could have arisen due to population substructure, resulting in an elevated overall Y diversity in the pooled sample.
-To test this hypothesis, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
-The resulting sequence phylogeny (Figure 4) revealed strong support for the Y-linked genes from the XY1Y2 race being paraphyletic (98% bootstrap support).
-In particular, our analysis revealed that samples from two populations, both from South Carolina (hereafter the SC sub-clade), formed a monophyletic group, whereas samples from Florida and Georgia (hereafter the FL sub-clade) were also monophyletic, but were more closely related to the XY Texas race.
-Although our sampling here is limited, these results suggest that introgression may have occurred between the ancestral Texas (XY race) and the derived North Carolina (XY1Y2) race, leading to a second, derived, XY1Y2 sub-clade and an elevated level of Y-diversity in our pooled sample.
-As these races are inter-fertile (Smith reference about hybrid zone), we suggest that the SC sub-clade likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY race.
-Notably, when we separated samples by the two inferred sub-clades of Y–linked sequences, the estimated level of Y diversity was consistent across all three clades, suggesting that the extensive reduction in diversity on this chromosome is not a population or race-specific phenomenon (see Supplementary Material, Figure S1).
-
 \subsection*{Background Selection and Selective Sweeps}
@@
 \section*{Conclusions}
+
 \section*{Acknowledgments}

### 2016-08-15 supporting-information/supporting_info.tex
supp info edits, ms edits
ms edits, figure S3

@@
 \templatetype{pnassupportinginfo}
-
+\usepackage{listings}
+\lstset{basicstyle=\ttfamily,
+  showstringspaces=false,
+  commentstyle=\color{red},
+  keywordstyle=\color{blue}
+}
 \title{Supporting Information}
@@
 
-\section*{SI Datasets}
+\begin{figure}[tbhp!]
+\centering
+\includegraphics[width=1\linewidth]{FigureS3.png}
+\caption{Normalized X/A and Y/A nucleotide diversity ratios for XY and $XY_{1}Y_{2}$ races of \textit{Rumex hastatulus}.
+In \textbf{A}, the South Carolina and Florida sub-clades are collapsed into a single race (‘NC’), and in \textbf{B}, they are separated into the inferred ‘SC’ and ‘FL’ sub-clades.}
+\label{figure:FigureS3}
+\end{figure}
+
+%Figure A4.1.2: Autosomal nucleotide diversity for the TX, SC, and FL clades of Rumex hastatulus, with confidence intervals calculated from bootstrapping 20000 replicates using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
+
+\newpage
+\section*{Simulation code}
+
+\begin{lstlisting}
+#!/bin/bash
+
+#purifying selection
+sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A
+
+#positive selection
+sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A
+\end{lstlisting}
 

### 2016-08-16 manuscript.tex
minor edits

@@
 
-It is worth noting that our results only apply to the $Y_{1}$ chromosome, however, as our estimates of diversity were calculated using the set of sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
-Previous work investigating patterns of genetic degeneration found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibit signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a significant reduction in diversity.
-However, our results do indicate that the pattern of reduced diversity on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
+It is worth noting that our results only apply to the $Y_{1}$ chromosome as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
+Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
+However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
 hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
@@
 
-Interestingly, our results indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
-Although not expected, this reduction in diversity may be associated with the recent origin of the XY1Y2 sex chromosome system.
-The XY1Y2 sex chromosome system is thought to have originated through an X-autosome fusion (Smith 1964) involving the ancestral 3rd chromosome in the Texas race.
-Evidence supporting this autosomal origin was recently obtained by Grabowska-Joachimiak et al.
-(2014), who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the Y2 sex chromosomes in the derived North Carolina race.
-If recent positive selection was responsible for driving the evolution of this fusion, then the formerly autosomal segment on the X chromosome in the XY1Y2 sub-clades (the neo-X) may have experienced a selective sweep, resulting in reduced X-linked diversity in the derived XY1Y2 sub-clades.
-It will be important for future work to investigate the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
+Interestingly, our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
+Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}.
+Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race.
+If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests as likely \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
+It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
 
-In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now test several possible models that might explain this diversity reduction.
+In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now consider several models that might explain reduction.
 
-\subsection*{Female biased sex ratios and variance in male fitness}
+\subsection*{Female biased sex ratio and high variance in male fitness}
 

### 2016-08-16 manuscript.tex
minor edits

@@
 On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in a reduction in $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
+However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
 Strikingly, we find that nucleotide diversity on the Y chromosome is 40 fold lower than on the X chromosome.
-We demonstrate that the magnitude of this reduction is not consistent with a purely neutral model based on the observed occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+We demonstrate that the magnitude of this reduction is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
 Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of purifying selection removing deleterious mutations and linked neutral variation.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with background selection playing an important role during the early stages of their evolution.
+hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes can evolve soon after sex chromosomes originate, with background selection playing an important role during the early stages of their evolution.
 \end{abstract}
@@
 
-In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now consider several models that might explain reduction.
+In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now consider several models that might explain this reduction.
 
@@
 
- A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection) (\X refs), ...and several cases....(\X refs), but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity in the absence of such selective interference.
-Moreover, such a reduction in Y-linked diversity could be accented due to high variance in male reproductive success (Figure 1).
+ A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection), but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity in the absence of such selective interference (\X ref).
+Moreover, reduced Y-linked diversity could be accented by high variance in male reproductive success (Figure 1).
 This is not unusual in annual plants such as \textit{R.
@@
 hastatulus}.
-However, ...
-by modeling...we evaluated whether such an effect could result in a level of Y/A diversity similar to what we observed by comparing our results to predictions from a  model for the joint effects of sex ratio bias and male reproductive variance (see Methods).
-Estimates of sex ratio bias in \textit{R.
-hastatulus} have been estimated, and they are consistently female-biased, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ (with a mean of 0.62; N=46 \citep{pickup2013influence}.The Y/A diversity ratio predicted from this sex ratio, assuming a Poisson distribution of offspring numbers, is approximately 0.2 (Table1, Figure 3).
-Our estimated mean Y/A ratio, however, is 0.02, which is significantly lower ($\textit{P}<0.0001$).
-Indeed, even for higher values for the variance in male reproductive success (\X), the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
+However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could explain the level of Y/A diversity that we observed (see Methods).
+Conditioning on estimates of sex ratio bias in \textit{R.
+hastatulus} that have been estimated, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  under a Poisson distribution of offspring numbers, is approximately 0.2 (Table1, Figure 3).
+Our estimated mean Y/A ratio, however, is 0.02, which is significantly lower ($\textit{P}<0.0001$), suggesting that the sex ratio effect alone is insufficient to explain our data.
+For larger values for the variance in male reproductive success (\X), the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
  Finally, neutral models that could not be rejected, i.e., in which a highly female-biased sex ratio (\X) and a high variance in male reproductive success predicted an Y/A ratio that fit our data (\X), simultaneously predicts X/A ratios that are significantly different from what we observed (\X Figure).

### 2016-08-20 manuscript.tex
minor edits

@@
 
-\subsection*{Estimating polymorphism levels on sex chromosomes and autosomes}
+\subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
 For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
@@
 
-Interestingly, our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
+Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
 Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}.
@@
 Our estimated mean Y/A ratio, however, is 0.02, which is significantly lower ($\textit{P}<0.0001$), suggesting that the sex ratio effect alone is insufficient to explain our data.
-For larger values for the variance in male reproductive success (\X), the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
- Finally, neutral models that could not be rejected, i.e., in which a highly female-biased sex ratio (\X) and a high variance in male reproductive success predicted an Y/A ratio that fit our data (\X), simultaneously predicts X/A ratios that are significantly different from what we observed (\X Figure).
-Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X and Y diversity in \textit{R.
+For larger values for the variance in male reproductive success (\X) given this sex ratio, the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
+ Finally, neutral models that could not be rejected, i.e., in which a highly female-biased sex ratio (\X) and a high variance in male reproductive success predict a Y/A ratio that fit out data (\X), can be excluded because they simultaneously predicted X/A ratios that were significantly different from what we observed (\X Figure).
+Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity in \textit{R.
 hastatulus}.

### 2016-08-20 manuscript.tex
edits up to bgs and rhh sims

@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in a reduction in $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
@@
 Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of purifying selection removing deleterious mutations and linked neutral variation.
+We do not exclude the possibility that positive selection on beneficial mutations has also reduced Y chromosome variability, and suggest that both forms of linked selection have likely been important during plant sex chromosome evolution.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes can evolve soon after sex chromosomes originate, with background selection playing an important role during the early stages of their evolution.
+hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with linked selection playing an important role during the early stages of their evolution.
 \end{abstract}
@@
 The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant genetic degeneration\citep{hough2014}.
+Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant gene loss and functional deterioration \citep{hough2014}.
 
@@
 hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
-This is important because female-biased sex ratios are predicted to affect chromosomal $N_{e}$ and therefor the level of genetic diversity maintained on sex chromosomes and autosomes \citep{ellegren2009}.
-Here, we investigated whether the patterns of diversity we observed on sex chromosomes and autosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection (background selection and selective sweeps) has played a greater role.
+This is important because female-biased sex ratios are predicted to affect chromosomal $N_{e}$ and therefore the level of neutral polymorphism maintained on sex chromosomes and autosomes \citep{ellegren2009}.
+Here, we test whether patterns of nucleotide diversity on sex chromosomes and autosomes in \textit{R.
+hastatulus} are jointly consistent with neutral evolutionary processes - including the effects of ratio bias and variance in reproductive success - or whether the removal of linked neutral variation caused by positive and purifying selection has played a more significant role.
 
@@
 (2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis.
-Here, we focused on the set of 460 sex-linked genes for which the Y homolog was inferred to be on the $y_{1}$ chromosome (i.e., sex-linked genes that are shared between the Texas and North Carolina races).
+Here, to consider sex linked genes that were identified in each of our sequenced population samples, we focused on the set of 460 X/Y genes for which the Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).
 
@@
 In evaluating evidence for nucleotide diversity differences between X and Y chromosomes, it is important to distinguish between reduced Y-linked diversity, and the possibility that X-linked diversity is elevated above the level predicted from a neutral model.
-To do this, we normalized sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from a neutral model (described below).
+To do this, we normalized sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below).
 Because the criteria for identifying autosomal loci in Hough et al.
@@
 Therefore, for the present analysis we incorporated the broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
-We also filtered this set to remove genes that may have been sex-linked but were not identified as such by the conservative ascertainment criteria in Hough et al.
-(2014).
+We also filtered this set to remove genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
 In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes with SNPs showing fixed heterozygosity in males and fixed homozygosity in females, (iii) any genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicate genes or those with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
@@
 Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output.
-This was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites - i.e., when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
+This was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
@@
 
-We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted using data from a genetic cross in Hough et al.
+We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in Hough et al.
 (2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data).
-In particular, we looked for cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was inferred to be zero by phasing-by-segregation.
-We identified 7 percent of sex-linked genes with phasing errors of this kind, or that were determined to have genotyping errors resulting in a false SNP calls on the Y chromosome, with a SNP error rate estimate of 1.7 x 10-4.
-This rate is very low relative to the population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism.
-However, because this rate is high relative to the expected level of true polymorphism on the Y-chromosome, we further filtered genes in which we found evidence for false-positive Y-polymorphisms arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was inferred to be zero by phasing-by-segregation.
+We identified 7 percent of sex-linked genes with phasing errors of this kind, or which were otherwise determined to have genotyping errors resulting in a false SNP calls.
+This corresponds to SNP error rate estimate of 1.7 x 10-4.
+Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism.
+However, because the rate is high relative to the expected level of true polymorphism on the Y-chromosome, we further filtered genes in which we found evidence for false-positive Y-polymorphisms arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
 This filtering was done by manually inspecting sequences in IGV \citep{robinson2011integrative} and identifying each individual putative polymorphism on the Y chromosome.
@@
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated such that the true uncertainty in our Y/A diversity estimate obtained by this method is likely underestimated.
-To address this issue, we used a maximum likelihood approach, implemented a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate is likely an underestimate.
+To address this issue, we used a maximum likelihood approach, implemented a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
@@
 To do this, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
-The analysis was conducted on an alignment of X- and Y-linked genes from \textit{R.
+The analysis was conducted on an alignment of the X- and Y-linked genes from \textit{R.
 hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R.
@@
 The phylogeny was inferred using the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
-The inferred sequence phylogeny revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98 percent bootstrap support), with samples from two populations (hereafter the SC sub-clade) forming a monophyletic group, and samples from Florida and Georgia (hereafter the FL sub-clade) also forming a monophyletic group, but more closely related to the XY (Texas) race (Figure S2).
-In light of this evidence for population substructure, we used the inferred phylogeny as the basis for estimating X and Y diversity.
+The inferred sequence phylogeny revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2).
+In light of this evidence for significant population substructure, and to consider between population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
 
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical estimates of diversity to predictions from a neutral model that jointly considers these two effects.
-Here, following Nomura (2002), the effective population sizes for autosomal loci is given by:
+To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that jointly considers these two effects.
+Here, following Nomura (2002), the neutrally-expected effective population size for autosomal loci is given by:
 
@@
 
-To consider the joint effects of biased sex ratios and variance in male reproductive success, our neutrally-expected levels of diversity were calculated using equations 1-3, assuming no variance in female mating (e.g., $C^2_{f}=0$), and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio and variance in male reproductive success.
-Using these neutral predictions, we compared our empirically estimated diversity levels to normalized X/A and Y/A ratios of diversity, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
-Note that we have only considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
+To consider the joint effects of biased sex ratios and variance in male reproductive success, neutrally-expected levels of diversity were calculated using equations 1-3, assuming no variance in female mating (e.g., $C^2_{f}=0$), and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
+Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio under this model.
+Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
+hastatulus} sex ratio $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
+Note that we considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
 hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females (\X ref).
@@
 \subsection*{Simulations of positive and purifying selection}
-To test whether our observed level of Y-chromosome diversity could be explained by selection, we conducted two sets of forward-time simulations using the software SFSCODE \citep{hernandez2008flexible}: (i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection acting on beneficial mutations and reducing variation through selective sweeps (refs \X).
-For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient ($s$) by varying the scale parameter, $k$ (range \X) from (\X) to \X and keeping the shape parameter, \X, constant (see Supporting Information for simulation code).
-To make the simulation output comparable to our data, we initialized simulations with our empirically estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
-For simulating positive selection, we used \X [parameters here].
+To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates with those from simulations.
+This approach was preferred over using analytical predictions primarily because the BGS model over-predicts the reduction in diversity when there are many linked sites under selection, which is expected to be the case for the Y chromosome where there is no crossing over.
+Moreover .....[\X explain here]
+
+%We chose to use forward simulations, rather than using standard analytical background selection models, which assume the effect of background selection is a simple reduction in effective population size, for several reasons.
+First, the standard formulas were derived for equilibrium demographic models, but human populations have a more complex demographic history with unknown effects on the process of background selection.
+
+%Second, many mutations have been shown to be weakly deleterious and may persist in the population due to genetic drift [37,38].
+The standard theory does not allow for this.
+Finally, simulations studies suggest that the standard theory can over-predict the reduction in genetic diversity due to background selection if there are many weakly selected linked mutations [26].
+The forward simulations that we conducted address all of these concerns.
+
+Because we were interested primarily in contrasting the
+
+(i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection acting on beneficial mutations and reducing variation through selective sweeps (refs \X).
+For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations ($s$) by varying the scale parameter, $k$ from \X to \X while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
+To make the simulation output comparable to our data, we initialized the simulations with our estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
+
+For simulating positive selection on the Y, we included a neutral locus of length Ln = 103 bp that is flanked by loci on either side that are subject to selection and are each of length Lf = 105 bp.
+
+In a forward simulation, you can only specify the mutation rate, not the substitution rate.
+So, for a given substitution rate  , we must divide by the fixation probability to obtain the corresponding mutation rate of selected mutations.
+Let ⌫ be the per base pair mutation rate in the flanking loci:
+
+Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity to asses the effects of selection in each model.
 We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of models separately.

### 2016-08-20 manuscript.tex
minor edits

@@
 Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
-hastatulus} sex ratio $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
+hastatulus} sex ratio $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}.
 Note that we considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
 hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females (\X ref).
-
-%old stuff
-%The effect of biased reproductive sex ratios on the $N_{e}$ for an autosomal gene can be estimated using the equation \citep{wright1931evolution}:
-
-%\begin{equation}
-%N_{e} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:Ne}
-%\end{equation}
-
-%where $N_{m}$ and $N_{f}$ are the numbers of reproducing males and females, respectively.
-Similarly, for an X-linked gene, where females contain two-thirds and males one-third of the alleles, the effective population size is given by:
-
-%\begin{equation}
-%N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}}\label{eq:NeX}
-%\end{equation}
-
-%And, for the male-limited Y chromosome:
-
-%\begin{equation}
-%N_{e{Y}} = \frac{N_{m}}{2}\label{eq:NeY}
-%\end{equation}
-
-%\begin{equation}
-%\frac{N_{e_{X}}}{N_{e_{A}}} = \frac{9(N_{f}+N_{m})}{8(2N_{f}+N_{m})}\label{eq:X/A}
-%\end{equation}
-
-%and,
-
-%\begin{equation}
-%\frac{N_{e_{Y}}}{N_{e_{A}}} = \frac{9(N_{f}+N_{m})}{8(2N_{f}+N_{m})}\label{eq:X/A}
-%\end{equation}
-
- %Given a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes, relative to autosomes, is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
-Similarly, for the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.
-Thus, with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-For populations with female-biased sex ratios, however, $N_{m}$ becomes small relative to $N_{f}$ and this ratio can become larger than 1, approaching 1.125 at the limit \citep{caballero1995}.
-With the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
-Figure 1 shows the predicted effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
-
-%In particular, Wright's (1931) formula $N_{e}=4N_{f}N_{m}/(N_{f} + N_{m})$ gives the effective population size of a randomly mating population with separate sexes and a Poisson distribution of offspring number.
-For a female effective population size $N_{f}$, and a male effective population size $N_{m}$, the predicted effective population size for X-linked genes relative to autosomes is given by $N_{e_{X}}/N_{e_{A}} = 9(N_{f}+N_{m})/8(2N_{f}+N_{m})$ (Wright 1931).
-With equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$.
-For populations with female-biased sex ratios, however, $N_{m}$ becomes small relative to $N_{f}$ and this ratio can become larger than 1, approaching 1.125 at the limit \citep{caballero1995}.
-For the Y chromosome, the effective population size is given by 1/4 the total $N_{e}$ or 1/2 the $N_{e}$ of males, $N_{m}$/2.
-Under neutrality, the level of nucleotide polymorphism maintained in a population is proportional to the product of the mutation rate and the effective population size: $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}.
-Assuming equal neutral mutation rates for all genes and an equal number of reproducing males and females, Y-linked genes should therefore have 1/3 the polymorphism of X-linked genes, and 1/4 that of autosomal genes.
-Figure 1 shows the predicted effective population size ratios for autosomal, X-linked, and Y-linked genes (left) and the corresponding X/A and Y/A ratios of diversity (right) for a sex ratio bias ranging from $N_{f}/(N_{f}+N_{m})=0$ to $N_{f}/(N_{f}+N_{m})=1$.
 
@@
 Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate null predictions for X/A and Y/A ratios of diversity.
-}%
+}
 \label{fig:spectrum}
@@
 To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates with those from simulations.
-This approach was preferred over using analytical predictions primarily because the BGS model over-predicts the reduction in diversity when there are many linked sites under selection, which is expected to be the case for the Y chromosome where there is no crossing over.
-Moreover .....[\X explain here]
+This approach was preferred over using analytical predictions primarily because the background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for Y chromosomes that lack crossing over.
 
-%We chose to use forward simulations, rather than using standard analytical background selection models, which assume the effect of background selection is a simple reduction in effective population size, for several reasons.
-First, the standard formulas were derived for equilibrium demographic models, but human populations have a more complex demographic history with unknown effects on the process of background selection.
-
-%Second, many mutations have been shown to be weakly deleterious and may persist in the population due to genetic drift [37,38].
-The standard theory does not allow for this.
-Finally, simulations studies suggest that the standard theory can over-predict the reduction in genetic diversity due to background selection if there are many weakly selected linked mutations [26].
-The forward simulations that we conducted address all of these concerns.
-
-Because we were interested primarily in contrasting the
-
-(i) Purifying selection acting to remove deleterious mutations and linked neutral variation (i.e., background selection), and (ii) Positive selection acting on beneficial mutations and reducing variation through selective sweeps (refs \X).
 For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations ($s$) by varying the scale parameter, $k$ from \X to \X while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
 To make the simulation output comparable to our data, we initialized the simulations with our estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
-
 For simulating positive selection on the Y, we included a neutral locus of length Ln = 103 bp that is flanked by loci on either side that are subject to selection and are each of length Lf = 105 bp.
-
-In a forward simulation, you can only specify the mutation rate, not the substitution rate.
-So, for a given substitution rate  , we must divide by the fixation probability to obtain the corresponding mutation rate of selected mutations.
-Let ⌫ be the per base pair mutation rate in the flanking loci:
-
-Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity to asses the effects of selection in each model.
-We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of models separately.
-We then tested the significant of the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity for each model.
+We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of selection models separately.
+Finally, we tested the significance of the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
 Commands for running the simulations are provided in the Supporting Information.

### 2016-08-21 manuscript.tex
minor edits

@@
 
-For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations ($s$) by varying the scale parameter, $k$ from \X to \X while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
-To make the simulation output comparable to our data, we initialized the simulations with our estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
-For simulating positive selection on the Y, we included a neutral locus of length Ln = 103 bp that is flanked by loci on either side that are subject to selection and are each of length Lf = 105 bp.
-Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity for each model.
-We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of selection models separately.
-Finally, we tested the significance of the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations, $s$, by varying the scale parameter, $k$ from 0.001 to 0.1 while keeping the distribution shape parameter, $\lambda$, constant (see Supporting Information for simulation commands).
+To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, scaled to the  $\theta_{Y}=\theta_{A}/4$ and adjusted for a sex ratio of $N_{m}/N_{f}=0.4$, and sampled 8 haploid chromosomes per simulation from 24 populations.
+Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity.
+For simulating positive selection, we included a neutral locus of length $Ln = 103$ bp (\X might change), which was flanked by loci on either side that were subject to positive selection.
+We ran 1000 replicates for each model and calculated the average $\theta_{Y{sim}}$ and the corresponding Y/A ratio of neutral diversity under the each set of selection models.
+Finally, we tested the significance of the fit of our simulated diversity output to our empirical Y/A ratio (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{{Y}sim}$ was not significantly different from $\theta_{{Y}observed}$, using a two-sided P value.
 Commands for running the simulations are provided in the Supporting Information.

### 2016-08-21 manuscript.tex
results edits

--- 2016-08-20 manuscript.tex
@@
 
-For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations ($s$) by varying the scale parameter, $k$ from \X to \X while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
-To make the simulation output comparable to our data, we initialized the simulations with our estimated value of $\theta$ (estimated from autosomes), and sampled 8 chromosomes per simulation from 24 populations.
-For simulating positive selection on the Y, we included a neutral locus of length Ln = 103 bp that is flanked by loci on either side that are subject to selection and are each of length Lf = 105 bp.
-Simulated sequences also contained 500 kb of linked neutral sequence from which we calculated diversity for each model.
-We ran 1000 replicates for each model and calculated the average $\theta_{sim}$ and the corresponding Y/A ratio of neutral diversity under the both sets of selection models separately.
-Finally, we tested the significance of the fit of our simulated diversity output to our empirical data (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{sim}$ was not significantly different from $\theta_{observed}$, using a two-sided P value.
+For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations, $s$, by varying the scale parameter from 0.0001 to 0.1 while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
+To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, scaled to the Y chromosome as $\theta_{Y}=\theta_{A}/4$ and adjusted for a sex ratio of $N_{m}/N_{f}=0.4$, and we sampled 8 haploid chromosomes per simulation.
+Simulated sequences contained 500 kb of linked neutral sequence from which we calculated diversity.
+
+For positive selection with recurrent hitchhiking, we used sfscoder to implement the method of \citep{uricchio2014robust}, and simulated a 500kb (\X) neutral locus that was flanked by 100kb loci on either side that were subject to positive selection.
+We ran 1000 replicates for each model and calculated the average $\theta_{Y{sim}}$ values and the corresponding Y/A ratios for each model seperately.
+Finally, we tested the significance of the fit of the simulated diversity to our empirical Y/A ratio (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{{Y}sim}$ was not significantly different from $\theta_{{Y}observed}$, using a two-sided P value.
 Commands for running the simulations are provided in the Supporting Information.
@@
 Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race.
-If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests as likely \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
+If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can driving the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
 It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
-
-In contrast to the situation on the X chromosome, our data indicate a strong and consistent diversity reduction on the Y chromosome, and we now consider several models that might explain this reduction.
+In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome - an approximately 40 fold reduction compared to the mean $theta_{X}$.
+We next consider several possible models - neutral and selective - that might explain this reduction.
 
@@
 
- A highly reduced level of Y chromosome variability is a predicted outcome of models Y-chromosome evolution based on the removal of deleterious mutations and linked neutral variants (e.g., background selection), but the occurrence of female-biased sex ratios in this species has also been predicted to lower Y diversity in the absence of such selective interference (\X ref).
-Moreover, reduced Y-linked diversity could be accented by high variance in male reproductive success (Figure 1).
-This is not unusual in annual plants such as \textit{R.
-hastatulus}, which commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Conceivably, such plastic differences among plants, owing to spatial heterogeneity in resources, could influence the intensity of male-male competition, and thus affect variance in male reproductive success.
-Given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, strong competition among males is expected.
+ The occurrence of female-biased sex ratios in this species has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome (\X ref).
+This reduction $N_{e}$ on the Y is also expected to be accented if there is high variance in male reproductive success (Figure 1), which is not unusual in annual plants such as \textit{R.
+hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+Conceivably, such plastic differences, owing to spatial heterogeneity in resources, could influence the intensity of male-male competition and thus affect variance in male reproductive success.
+Moreover, given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, strong competition among males is expected.
 
@@
 Conditioning on estimates of sex ratio bias in \textit{R.
-hastatulus} that have been estimated, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  under a Poisson distribution of offspring numbers, is approximately 0.2 (Table1, Figure 3).
-Our estimated mean Y/A ratio, however, is 0.02, which is significantly lower ($\textit{P}<0.0001$), suggesting that the sex ratio effect alone is insufficient to explain our data.
-For larger values for the variance in male reproductive success (\X) given this sex ratio, the corresponding predictions for Y/A are all significantly higher than our observed ratio (\X table).
- Finally, neutral models that could not be rejected, i.e., in which a highly female-biased sex ratio (\X) and a high variance in male reproductive success predict a Y/A ratio that fit out data (\X), can be excluded because they simultaneously predicted X/A ratios that were significantly different from what we observed (\X Figure).
-Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity in \textit{R.
-hastatulus}.
+hastatulus} that have been estimated, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
+This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), even if we consider the upper bound from the bootstrapped or likelihood based estimated of the confidence interval.
+This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.
+
+Assuming that there is extensive variance in male reproductive success (\X), the predicted ratios of Y/A diversity (\X supp) were also significantly higher than our estimates for the empirically estimated sex ratio (\X table) .
+We also find, however, that purely neutral models that assumed a highly female-biased sex ratio (\X) and a level of variance in male reproductive success on the order of (\X) predicted a Y/A ratio that fit out data (\X).
+However, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure).
+Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity.
 

### 2016-08-22 manuscript.tex
minor edits

@@
 
-It is worth noting that our results only apply to the $Y_{1}$ chromosome as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
+It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
 Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
@@
 It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
-In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome - an approximately 40 fold reduction compared to the mean $theta_{X}$.
+
+In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 40 fold reduction compared to the mean $\theta_{X}$.
 We next consider several possible models - neutral and selective - that might explain this reduction.
@@
 
- The occurrence of female-biased sex ratios in this species has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome (\X ref).
-This reduction $N_{e}$ on the Y is also expected to be accented if there is high variance in male reproductive success (Figure 1), which is not unusual in annual plants such as \textit{R.
+ The occurrence of female-biased sex ratios in this species has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the neutrally expected $N_{e}$ of the Y chromosome (\X ref).
+This reduction $N_{e}$ on the Y is expected to be further accented if there is high variance in male reproductive success (Figure 1), which is not unusual in annual plants such as \textit{R.
 hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Conceivably, such plastic differences, owing to spatial heterogeneity in resources, could influence the intensity of male-male competition and thus affect variance in male reproductive success.
-Moreover, given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, strong competition among males is expected.
+Moreover, given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, we expected that there is strong competition among males to fertilize females.
 
@@
 hastatulus} that have been estimated, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
-This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), even if we consider the upper bound from the bootstrapped or likelihood based estimated of the confidence interval.
+This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), and remains significant even if we consider an upper bound estimate obtained from our likelihood based estimated of the confidence interval.
 This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.
+(\X I think some of this can be written better...)
 
 Assuming that there is extensive variance in male reproductive success (\X), the predicted ratios of Y/A diversity (\X supp) were also significantly higher than our estimates for the empirically estimated sex ratio (\X table) .
-We also find, however, that purely neutral models that assumed a highly female-biased sex ratio (\X) and a level of variance in male reproductive success on the order of (\X) predicted a Y/A ratio that fit out data (\X).
-However, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure).
+We also find, however, that purely neutral models in which the sex ratio was highly female-biased (\X), with level of variance in male reproductive success on the order of (\X), predicted a Y/A ratio that could not be rejected (\X).
+However, because a highly females biased sex ratio is expected to increase the X/A ratio as well, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure).
 Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity.

### 2016-09-21 manuscript.tex
stephen's changes + minor edits

@@
 In addition to demographic factors, evolutionary theory predicts that genetic variability on sex chromosomes should be affected by the evolution of suppressed X-Y recombination, which lowers the coalescent $N_{e}$ for Y-linked genes and increases their vulnerability to the effects of selective sweeps \citep{smith1974hitch,aquadro1994} and background selection \citep{charlesworth1996background,charlesworth1994effect}.
-These processes can be viewed as examples of the Hill–Robertson (HR) effect, in which the effective population size for a given chromosomal region depends strongly on the rate of recombination, such that selection at a focal site is not independent of selection at nearby sites \citep{hill1966HReffect}.
 Moreover, both background selection and selective sweeps predict a reduction in $N_{e}$ and, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
-Although the HR effect can occur throughout the genome \citep{mcvean2000}, large genomic regions in which recombination is suppressed, such as the Y chromosome, are expected to experience the most severe reductions in neutral diversity due to chromosome-wide linkage \citep{charlesworth1996CB,charlesworth2000degeneration,bachtrog2013NRG}.
-Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have found that Y-linked diversity is considerably lower than predicted under models of neutral evolution, suggesting that the HR effect plays a key role determining patterns of nucleotide diversity during sex chromosome evolution across a broad range of animal taxa.
+Although the such linked selective effects can occur throughout the genome \citep{mcvean2000}, large genomic regions in which recombination is suppressed, such as the Y chromosome, are expected to experience the most severe reductions in neutral diversity due to chromosome-wide linkage \citep{charlesworth1996CB,charlesworth2000degeneration,bachtrog2013NRG}.
+Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have found that Y-linked diversity is considerably lower than predicted under models of neutral evolution, suggesting that the linked selection plays a key role determining patterns of nucleotide diversity during sex chromosome evolution across a broad range of animal taxa.
 
-Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of the HR effect on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
+Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
 Recent studies of young plant sex chromosome systems have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}.
 Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in plant sex chromosomes.
-Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR effect.
+Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the linked selection.
 
@@
 
-Finally, we tested for population substructure between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
+Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
 To do this, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
@@
 For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations, $s$, by varying the scale parameter from 0.0001 to 0.1 while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
-To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, scaled to the Y chromosome as $\theta_{Y}=\theta_{A}/4$ and adjusted for a sex ratio of $N_{m}/N_{f}=0.4$, and we sampled 8 haploid chromosomes per simulation.
+To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for a sex ratio of $N_{m}/N_{f}=0.4$, and we sampled 8 haploid chromosomes per simulation.
 Simulated sequences contained 500 kb of linked neutral sequence from which we calculated diversity.
@@
 However, these possibilities can be rejected because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
-
-It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
-Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
-However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
-hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
 
@@
 
+It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
+Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
+However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
+hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
+
 \section*{Conclusions}

### 2016-10-03 manuscript.tex
organized folders, added temp

@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y chromosome is 40 fold lower than on the X chromosome.
-We demonstrate that the magnitude of this reduction is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of purifying selection removing deleterious mutations and linked neutral variation.
-We do not exclude the possibility that positive selection on beneficial mutations has also reduced Y chromosome variability, and suggest that both forms of linked selection have likely been important during plant sex chromosome evolution.
+We find that nucleotide diversity on the Y chromosome is approximately 40 fold lower than on the X chromosome, suggesting an extensive and relatively rapid loss of diversity.
+We demonstrate that the magnitude of this diversit reduction is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the removal of deleterious mutations and genetically-linked neutral variation.
+In particular, we find that ....
+hr realm...
+
+Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with linked selection playing an important role during the early stages of their evolution.
+hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.
 \end{abstract}

### 2016-10-03 Rnotebook/Figure1.R

@@ -1,45 +1,97 @@
 Nf=100 #number of females
 Nm=100 #number of males
 N=Nf+Nm #total number of individuals
-x=Nm/(Nf+Nm) #proportion of males
+#x=Nf/(Nf+Nm) #proportion of males
 
-op <- par(mfrow = c(1, 2))
+op <- par(mfrow = c(1, 3))
 m=0
 f=0
 
-curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m))))/N, from=0, to=1, lty=2,
-      xlab=expression(italic('N'[m]/'N'[m]+'N'[f])),
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[f]/'N'[m]+'N'[f])),
       ylab=expression(italic('N'[e]/N)), 
       main="Autosomes", ylim=c(0,1), xlim=c(0,1))
 
 mh=0.5
-curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh))))/N, from=0, to=1, add=T)
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+mh))))/N, from=0, to=1, add=T)
 
 m1=1
-curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1))))/N, from=0, to=1, add=T)
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T, col="blue")
 
 m2=2
-curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2))))/N, from=0, to=1, add=T)
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m2))))/N, from=0, to=1, add=T)
 
 m3=3
-curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3))))/N, from=0, to=1, add=T)
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m3))))/N, from=0, to=1, add=T)
 
 #for sex-linked genes
[…]

### 2016-10-03 SFS_code/sfs_code.sh

@@ -9,29 +9,49 @@ set -e
 set -u
 set -o pipefail
 
-
-#command structure: sfs_code <Npops> <Niter> [<options> [arguments]]
-
-# Parameters used here
-# Npops:        1
-# Niter:        50
-# theta(-t):    0.001375
+# Npops:        1 #number of pops
+# Niter:        50 #iterations
+# theta(-t):    0.001375 #=1/4 autosomal theta: 0.0055/4
 # rho(-r):      0
 # ploidy(-P):   1
-# tau(-TE):     0.3 (scaled in units of P*N generations)
-# loci(-L):     2 regions, lengths: 500000, 800000 (<R> <L1> [<L2>..] [R]): R regions with lengths <L1><L2>...
-# annotate(-a): N noncoding
-# fitness(-W):  L 1 2 0 1 1 0.184 0.003 (here, gamma shape = 0.184, and s=0.01. i.e., 0.184/5000=0.003)
-# -A tells simualtion to end
-# sampsize(-n): 8 chromosomes per sample
-# popsize:(-N)  500
+# tau(-TE):     0.3 #generations scaled in units of P*N (this seems too small?)
+# loci(-L):     2 <500000> <800000> #2 regions with lengths <L1><L2>...
+# annotate(-a): N #noncoding
+# fitness(-W):  L 1 2 0 1 1 0.258 0.005 # Gamma T model. probability positive s = 0, gamma shape = 0.258 from the DFE for x-linked genes. For a selection coefficient = 0.01, the scale parameter is 0.258/(0.01 * 5000)= 0.005.
+# sampsize(-n): 8 #chromosomes per sample
+# popsize:(-N)  5000
+# -A tells simulation to end
 
-echo "Starting Job: coding regions only"
+./sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A
+# set paramters as bash variables
+aN=0.2579 #shape parameter from X-linked DFE
+s=
+lN=${aN}/${s}*5000 # scale parameter. Hold aN constant, test a range for s. 5000=5kb
+
+#############################################################################
+echo "Starting Job: Non-coding regions"
[…]

### 2016-10-03 temp/TOP.R

@@ -1,0 +1,144 @@
+vm <- c(0:100)
+for (val in vm) {
+  Nf=100000 #numner of females
+  Nm=100000 #number of males
+  N=Nf+Nf
+  vf=0 #variance in female rep success
+  kf=2 #mean number of offspring of females
+  km=2 #mean number of offspring of males
+  Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
+  Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
+  Ne=Nem+Nef
+  Ne_06=4*Ne*(1-0.6)*0.6
+  Ne_07=4*Ne*(1-0.7)*0.7
+  Ne_08=4*Ne*(1-0.8)*0.8
+  Ne_09=4*Ne*(1-0.9)*0.9
+  Ne_ratio=Nef/Nem
+  A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
+  X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
+  Y=Nem/2
+  XA=X/A
+  YA=Y/A
+}
+d <- data.frame(vm,Nem,Nef,Ne_ratio,A,X,Y,XA,YA, Ne, Ne_06,Ne_07,Ne_08,Ne_09)
+
+plot(vm,(Nem/2)/N, type="l", xlim = c(0,100))
+plot(vm,(Ne)/N, type="l", xlim = c(0,20))
+plot(vm,YA,type="l", xlim = c(0,100))
+plot(vm,XA,type="l", xlim = c(0,100))
+
+
+require(ggplot2)
+ggplot(d, aes(vm)) +                   
+  geom_line(aes(y=(Ne/4)/N), colour="red") +
+  geom_line(aes(y=(Ne_06/4)/N), colour="blue") +
+  geom_line(aes(y=(Ne_07/4)/N), colour="green") +
+  geom_line(aes(y=(Ne_08/4)/N), colour="red") +
+  geom_line(aes(y=(Ne_09/4)/N), colour="brown")
+  
+Nf=100000 #numner of females
+Nm=100000 #number of males
+N=Nf+Nf
+ne_A_poisson=(4*N*(1-r)*r)/N
+ne_X_poisson=(9*N*r*(r-1)/(2*(r-2)))/N
+  
[…]

### 2016-10-10 manuscript.tex
minor edits, changes to abstract which disappeared

--- 2016-09-21 manuscript.tex
@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, both positive and negative selection can remove genetically linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral varia tion, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y chromosome is 40 fold lower than on the X chromosome.
-We demonstrate that the magnitude of this reduction is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations, we show that the diversity reduction on the Y is consistent with the predicted effects of purifying selection removing deleterious mutations and linked neutral variation.
-We do not exclude the possibility that positive selection on beneficial mutations has also reduced Y chromosome variability, and suggest that both forms of linked selection have likely been important during plant sex chromosome evolution.
+Strikingly, we find that nucleotide diversity on the Y chromosome is approximately 40 fold lower than on the X chromosome, suggesting an extensive and rapid loss of diversity.
+We demonstrate that the magnitude of this diversity loss is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations, we show that observed Y diversity is consistent selection on deleterious mutations.
+
+\end{abstract}
+%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+
+%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
+
+%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
+
+%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
+
+%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
+
+%The agreement is reasonably good given the uncertainties involved.
+
+%silent site diversity for 22 genes on the D. miranda neo-Y is
+%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
+
+% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
+So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
+
+%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results indicate that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve soon after sex chromosomes originate, with linked selection playing an important role during the early stages of their evolution.
-\end{abstract}
+hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}
 
@@
 
+%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+
+%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
+
+%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
+
+%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
+
+%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
+
+%The agreement is reasonably good given the uncertainties involved.
+
+%silent site diversity for 22 genes on the D. miranda neo-Y is
+%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
+
+% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
+So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
+
 It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.

### 2016-10-10 manuscript.tex
fixed lost text, minor edits

@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral varia tion, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y chromosome is approximately 40 fold lower than on the X chromosome, suggesting an extensive and rapid loss of diversity.
-We demonstrate that the magnitude of this diversity loss is not consistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations, we show that observed Y diversity is consistent selection on deleterious mutations.
+Strikingly, we find that nucleotide diversity on the Y is 40 fold lower than on the X, suggesting an extensive and rapid loss of diversity.
+We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations, we show that Y diversity can be explained by purifying selection.
+Interestingly, whereas standard background selection theory predicted a diversity reduction n times what we observed, our simulations indicate that Y-linked variability can be explained by interference among a large number (>800kb) of loci subject to weak purifying selection (s).
+Given the recent origin of \textit{R.
+hastatulus} sex chromosomes, our study suggests that the low level of sequence diversity commonly observed on nonrecombining Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against weakly deleterious mutations playing an important role during the earliest stages of their evolution.
+
+%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
 
 \end{abstract}
-%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-
-%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
-
-%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
-
-%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
-
-%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
-
-%The agreement is reasonably good given the uncertainties involved.
-
-%silent site diversity for 22 genes on the D. miranda neo-Y is
-%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
-
-% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
-So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
-
-%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}
 
@@
 
-One fundamental difference between sex chromosomes and autosomes is their respective copy number in a population; whereas autosomes in diploid species are present in two copies in each sex, the X chromosome is present in two copies in females and only one copy in males.
+One fundamental difference between sex chromosomes and autosomes is their copy number in a population; whereas autosomes in diploid species are present in two copies in each sex, the X chromosome is present in two copies in females and only one copy in males.
 As a consequence, the effective population size ($N_{e}$) of the X chromosome is predicted to be 3/4 that of autosomes, whereas the $N_{e}$ of the male-limited Y chromosome is expected to equal 1/4 that of autosomes (assuming an equal number of reproducing females and males).
 Such differences in $N_{e}$ are predicted to directly affect relative levels of neutral polymorphism maintained on these chromosomes, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ \citep{Kimura1984}.
-This variation in neutral polymorphism is in turn expected to have important consequences on patterns of DNA sequence evolution, including the effectiveness of both positive and negative selection \citep{charlesworth1987}.
+Such variation in neutral polymorphism within the genome can have important consequences on patterns of DNA sequence evolution, including the effectiveness of both positive and negative selection \citep{charlesworth1987}.
 
@@
 These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
-For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in estimates of the X/A ratio in humans, and other demographic processes have also been suggested \citep{hammer2010,bustamante2009}.
-Theoretical work also indicates that population bottlenecks can lead to disproportionate decreases in sex-linked compared to autosomal variation \citep{pool2007}, a prediction that has been supported in a diverse array of taxa, including chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, Drosophila \citep{andolfatto2001}, and humans \citep{keinan2009}.
+For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in estimates of the X/A ratio in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}.
+Theoretical work also indicates that population bottlenecks should lead to a disproportionate decrease in sex-linked compared to autosomal variation \citep{pool2007}, as discovered in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, Drosophila \citep{andolfatto2001}, and humans \citep{keinan2009}.
 
-In addition to demographic factors, evolutionary theory predicts that genetic variability on sex chromosomes should be affected by the evolution of suppressed X-Y recombination, which lowers the coalescent $N_{e}$ for Y-linked genes and increases their vulnerability to the effects of selective sweeps \citep{smith1974hitch,aquadro1994} and background selection \citep{charlesworth1996background,charlesworth1994effect}.
-Moreover, both background selection and selective sweeps predict a reduction in $N_{e}$ and, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
-Although the such linked selective effects can occur throughout the genome \citep{mcvean2000}, large genomic regions in which recombination is suppressed, such as the Y chromosome, are expected to experience the most severe reductions in neutral diversity due to chromosome-wide linkage \citep{charlesworth1996CB,charlesworth2000degeneration,bachtrog2013NRG}.
-Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have found that Y-linked diversity is considerably lower than predicted under models of neutral evolution, suggesting that the linked selection plays a key role determining patterns of nucleotide diversity during sex chromosome evolution across a broad range of animal taxa.
+In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by linked selection, including selective sweeps of benficial mutations \citep{smith1974hitch,aquadro1994} and the removal of strongly deleterious mutations \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}.
+Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution.
+
+Although...
+associations among selected alleles in regions of suppressed recombination is expected to lower the coalescent $N_{e}$ for Y-linked genes, so that they behave as if they were subject to weaker selection \citep{hill1966HReffect, mcvean2000}...
+thereby reducing their effects on linked neutral or nearly neutral variants.
+This reduction in $N_{e}$ and the effecacy of selection, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
 
 Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
-Recent studies of young plant sex chromosome systems have revealed that several canonical features of X and Y chromosome evolution are shared between plants and animals, including the genetic deterioration of the Y chromosome \citep{bergero2015,hough2014}, the occurrence of ‘evolutionary strata’ \citep{bergero2009}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}.
-Nonetheless, except for a small number of genes in \textit{Silene latifolia} \citep{filatov2001diversity,qiu2010nucleotide}, the associated changes in genetic diversity that are predicted from evolutionary models of X- and Y-chromosome evolution have not been widely studied in plant sex chromosomes.
-Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the linked selection.
+Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR-effect.
 
@@
 So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
+%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
+
+%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
+
+%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
+
+%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
+
+%The agreement is reasonably good given the uncertainties involved.
+
+%silent site diversity for 22 genes on the D. miranda neo-Y is
+%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
+
+% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
+So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
+
+%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+Given the recent origin of \textit{R.
+hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}
 

### 2016-10-14 manuscript.tex
minor edits, pdf

@@
 
-Several demographic factors can modulate or accentuate differences in polymorphism levels maintained on sex chromosomes \citep{ellegren2009}.
+Several demographic factors are expected to modulate or accentuate differences in neutral polymorphism maintained on sex chromosomes \citep{ellegren2009}.
 These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
-For example, studies comparing levels of variability on the X chromosome and autosomes in humans have found evidence that sex-biased dispersal has reduced levels of X-to-autosome diversity below the neutrally expected level of 0.75 \citep{keinan2009}, though there is considerable variation in estimates of the X/A ratio in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}.
-Theoretical work also indicates that population bottlenecks should lead to a disproportionate decrease in sex-linked compared to autosomal variation \citep{pool2007}, as discovered in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, Drosophila \citep{andolfatto2001}, and humans \citep{keinan2009}.
+In theory, these processes all result in asymmetries in female vs male $N_{e}$, that could lead to different levels of neutral polymorphism on the sex chromosomes than are expected based on the autosomes (\citep{pool2007}).
+As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-africa dispersal result in levels of X/A neutral diversity simular to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
+In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferreed in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
 
-In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by linked selection, including selective sweeps of benficial mutations \citep{smith1974hitch,aquadro1994} and the removal of strongly deleterious mutations \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}.
-Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution.
+In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which refers to the reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
+Two types of HRI include (i) selection on stronlgy benficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} and (ii) selection against strongly deleterious mutations
+) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}.
+In both processes, linkage  disequalibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absense of selection.
 
-Although...
-associations among selected alleles in regions of suppressed recombination is expected to lower the coalescent $N_{e}$ for Y-linked genes, so that they behave as if they were subject to weaker selection \citep{hill1966HReffect, mcvean2000}...
-thereby reducing their effects on linked neutral or nearly neutral variants.
-This reduction in $N_{e}$ and the effecacy of selection, as a result, reduced levels of neutral polymorphism, selection efficacy, and rate of adaptation \citep{comeron2008}.
+Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with purifying selection playing a key role.
 
-Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
-Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}, studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR-effect.
+In addition to reducing neutral diversity due to reduced Ne, high linkage disequilibrum (LD) can also occur between the sites experiencing selection such that the net effect that selection has on reducing neutral (or nearly neutral) variation at linked sites is reduced \citep{hill1966HReffect, mcvean2000,comeron2008}.
+This "weak Hill-Robertson Interference" (weak-HRI) effect is predicted to be greatest in large genomic regions where recombination is rare or absent (\X cite kaiser paper), making Y chromosomes important for empirically investigating this processes.
+
+%discussion point?:
+%Moreover, because recombination suppression was relatively recent on plant Y chromosomes, it is an open question whether weak-HRI has had enough time to have a large effect.
+Although selective interference (indlucing BGS and Selective Sweeps) is expected to have the greatest effects during the early stages of Y-chromosome evolution \citep{bachtrog2008temporal} the R.
+hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.
+
+%Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
+Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes , studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR-effect.
 

### 2016-10-25 manuscript.tex

@@
 On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult infer the causes of low Y-chromosome diversity.
+However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y is 40 fold lower than on the X, suggesting an extensive and rapid loss of diversity.
-We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model based on the occurrence of female biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations, we show that Y diversity can be explained by purifying selection.
-Interestingly, whereas standard background selection theory predicted a diversity reduction n times what we observed, our simulations indicate that Y-linked variability can be explained by interference among a large number (>800kb) of loci subject to weak purifying selection (s).
+Strikingly, we find that nucleotide diversity on the Y is 40-fold lower than on the X, suggesting an extensive reduction in diversity.
+We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations and population genetic estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity levels can be explained by the action of background selection.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that the low level of sequence diversity commonly observed on nonrecombining Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against weakly deleterious mutations playing an important role during the earliest stages of their evolution.
-
-%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromsome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites.
+Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
 
@@
 Several demographic factors are expected to modulate or accentuate differences in neutral polymorphism maintained on sex chromosomes \citep{ellegren2009}.
-These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
+These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high sex-specific variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
 In theory, these processes all result in asymmetries in female vs male $N_{e}$, that could lead to different levels of neutral polymorphism on the sex chromosomes than are expected based on the autosomes (\citep{pool2007}).
-As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-africa dispersal result in levels of X/A neutral diversity simular to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
+As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-africa dispersal result in levels of X/A neutral diversity similar to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
 In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferreed in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
@@
 In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which refers to the reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
-Two types of HRI include (i) selection on stronlgy benficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} and (ii) selection against strongly deleterious mutations
+Two types of HRI include (i) selection on strongly benficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} and (ii) selection against strongly deleterious mutations
 ) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}.
-In both processes, linkage  disequalibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absense of selection.
+In both processes, linkage  disequalibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
 
@@
 
-%Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
+Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
 Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes , studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR-effect.

### 2016-10-26 manuscript.tex
edits to intro and methods

@@
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y is 40-fold lower than on the X, suggesting an extensive reduction in diversity.
+Strikingly, we find that nucleotide diversity on the Y is more than 40-fold lower than on the X, suggesting an extensive reduction in diversity.
 We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
@@
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromsome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites.
+hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites.
 Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
@@
 As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-africa dispersal result in levels of X/A neutral diversity similar to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
-In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferreed in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
+In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferred in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
 
-In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which refers to the reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
-Two types of HRI include (i) selection on strongly benficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} and (ii) selection against strongly deleterious mutations
-) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}.
-In both processes, linkage  disequalibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
+In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
+The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations
+) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference).
+In all of these processes, linkage  disequilibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
+In a large nonrecombining region experiencing sufficiently high deleterious mutation rates, the distinction between the effects of strongly- and weakly- selected mutations begins to break down, and have been explored primarily using forward computer simulations (refs).
 
-Consistent with this, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with purifying selection playing a key role.
-
-In addition to reducing neutral diversity due to reduced Ne, high linkage disequilibrum (LD) can also occur between the sites experiencing selection such that the net effect that selection has on reducing neutral (or nearly neutral) variation at linked sites is reduced \citep{hill1966HReffect, mcvean2000,comeron2008}.
-This "weak Hill-Robertson Interference" (weak-HRI) effect is predicted to be greatest in large genomic regions where recombination is rare or absent (\X cite kaiser paper), making Y chromosomes important for empirically investigating this processes.
+Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with purifying selection playing a key role.
 
@@
 Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
-Given that background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes , studying recently evolved sex chromosomes provides an important test of the temporal dynamics of the HR-effect.
+Background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes.
+On the other hand, if gene silencing and Y degeneration occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimal even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes.
 
@@
 Here, we test whether patterns of nucleotide diversity on sex chromosomes and autosomes in \textit{R.
-hastatulus} are jointly consistent with neutral evolutionary processes - including the effects of ratio bias and variance in reproductive success - or whether the removal of linked neutral variation caused by positive and purifying selection has played a more significant role.
+hastatulus} are jointly consistent with neutral evolutionary processes - including the effects of sex ratio bias and variance in reproductive success - or whether the removal of linked neutral variation caused by positive and purifying selection has played a more significant role.
 
@@
 (2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosome.
-Therefore, for the present analysis we incorporated the broader set of all non-sex linked (punitively autosomal) genes from our transcriptome data.
+Therefore, for the present analysis we incorporated the broader set of all non-sex linked (putatively autosomal) genes from our transcriptome data.
 We also filtered this set to remove genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
@@
 (2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was inferred to be zero by phasing-by-segregation.
-We identified 7 percent of sex-linked genes with phasing errors of this kind, or which were otherwise determined to have genotyping errors resulting in a false SNP calls.
-This corresponds to SNP error rate estimate of 1.7 x 10-4.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was known to be zero.
+We identified 7 percent of sex-linked genes with phasing errors of this kind, or which were otherwise determined to have genotyping errors resulting in false SNP calls.
+This corresponds to a SNP error rate estimate of 1.7 x 10-4.
 Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism.
@@
 Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate is likely an underestimate.
-To address this issue, we used a maximum likelihood approach, implemented a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+To address this issue, we used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this would involve maximizing the likelihood estimate of the mutation rate for each locus independently.
+Therefore, we applied a method that assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
+Our model therefore had two parameters; theta for autosomal genes and the Y/A ratio.
+We varied both parameters across a grid of points and evaluated the likelihood, where the Y/A ratio varied from 0.001 to 1, and Watterson's theta on the autosomes varied from 0.001 to 0.01.
 We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
@@
 Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
-To do this, we conducted a phylogenetic analysis of sex-linked sequences from all populations in our study.
+To do this, we constructed neighbor-joining trees of sex-linked sequences from all populations in our study.
 The analysis was conducted on an alignment of the X- and Y-linked genes from \textit{R.
@@
 bucephalophorus} used to root the tree (Figure S2).
-The phylogeny was inferred using the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
-The inferred sequence phylogeny revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2).
+We used the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
+The inferred trees revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2).
 In light of this evidence for significant population substructure, and to consider between population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
@@
 Note that we considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
-hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate females (\X ref).
+hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate female flowers (\X ref).
 

### 2016-10-26 manuscript.tex
edits

@@
 We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations and population genetic estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity levels can be explained by the action of background selection.
+Rather, using forward simulations and population genetic estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity levels can be explained by the action of linked selection against deleterious mutations, although selective sweeps may also be playing a role.
 Given the recent origin of \textit{R.
 hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites.
-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-
+These results suggest that selective interference is playing a major role in the ongoing degeneration of the Y chromosome, rather than simply neutral drift following gene silencing.
 \end{abstract}
@@
 Background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes.
-On the other hand, if gene silencing and Y degeneration occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimal even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes.
+On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimal even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes.
 
@@
 
-With with no variation in mating success in either sex ($C^2_{m}=C^2_{f}=0$), equations 1-3 reduce to the familiar expressions from \citep{wright1931evolution}:
+With no variation in mating success in either sex ($C^2_{m}=C^2_{f}=0$), equations 1-3 reduce to the familiar expressions from \citep{wright1931evolution}:
 

### 2016-10-27 manuscript.tex
edits

@@
 On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
@@
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites.
-These results suggest that selective interference is playing a major role in the ongoing degeneration of the Y chromosome, rather than simply neutral drift following gene silencing.
+hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites, rather than simply neutral drift following gene silencing.
 \end{abstract}
@@
 In theory, these processes all result in asymmetries in female vs male $N_{e}$, that could lead to different levels of neutral polymorphism on the sex chromosomes than are expected based on the autosomes (\citep{pool2007}).
-As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-africa dispersal result in levels of X/A neutral diversity similar to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
+As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-Africa dispersal result in levels of X/A neutral diversity similar to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
 In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferred in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
@@
 In all of these processes, linkage  disequilibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
-In a large nonrecombining region experiencing sufficiently high deleterious mutation rates, the distinction between the effects of strongly- and weakly- selected mutations begins to break down, and have been explored primarily using forward computer simulations (refs).
 
-Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with purifying selection playing a key role.
+Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with linked selection playing a key role.
 
@@
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate is likely an underestimate.
-To address this issue, we used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes.
+To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this would involve maximizing the likelihood estimate of the mutation rate for each locus independently.
@@
 The inferred trees revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2).
-In light of this evidence for significant population substructure, and to consider between population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
+In light of this evidence for significant population substructure, and to consider between-population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
 

### 2016-10-28 manuscript.tex
edits_SW

@@
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y is more than 40-fold lower than on the X, suggesting an extensive reduction in diversity.
+Strikingly, we find that nucleotide diversity on the Y is more than 50-fold lower than on the autosomes, suggesting an extensive reduction in diversity.
 We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
@@
 
-Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with linked selection playing a key role.
+Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with evidence for linked selection playing a key role.
 
@@
 Background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes.
-On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimal even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes.
+On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimized rapidly even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes (cite backtrog NRG review).
 
@@
 Of particular relevance to the present study, \textit{R.
-hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}.
+hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
 This is important because female-biased sex ratios are predicted to affect chromosomal $N_{e}$ and therefore the level of neutral polymorphism maintained on sex chromosomes and autosomes \citep{ellegren2009}.
@@
 Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes.
-To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio using coalescent theory (Figure S1).
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this would involve maximizing the likelihood estimate of the mutation rate for each locus independently.
@@
 To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that jointly considers these two effects.
-Here, following Nomura (2002), the neutrally-expected effective population size for autosomal loci is given by:
 
-\begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{f})+N_{f}(1+C^2_{m})}\label{eq:NeA}
-\end{equation}
-
-where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, defined as the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
-Similarly, for X and Y linked loci,
-
-\begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{f})+2N_{f}(1+C^2_{m})}\label{eq:NeX}
-\end{equation}
-
-and
-
-\begin{equation}
-N_{e{Y}} = \frac{N_{m}(1+C^2_{f})}{2}\label{eq:NeY}
-\end{equation}
-
-With no variation in mating success in either sex ($C^2_{m}=C^2_{f}=0$), equations 1-3 reduce to the familiar expressions from \citep{wright1931evolution}:
+In particular, the effective population size for genes on each chromosome can be given by  \citep{wright1931evolution}:
 
@@
 
-To consider the joint effects of biased sex ratios and variance in male reproductive success, neutrally-expected levels of diversity were calculated using equations 1-3, assuming no variance in female mating (e.g., $C^2_{f}=0$), and we considered a range of values for $\sigma_{m}$: 0 (no variance), 1/2 (Poisson/2) 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-Figure 1 shows the predicted effective population size of autosomes and sex chromosomes as a function of the sex ratio under this model.
  Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
-hastatulus} sex ratio $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}.
-Note that we considered variance in male reproductive success here because this is most likely to occur in wind pollinated plant like \textit{R.
-hastatulus}, where males produce copious amounts of pollen and compete to fertilize uniovulate female flowers (\X ref).
+hastatulus} sex ratio (N_{f}/$N_{m}+N_{f})=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}.
+Sex-specific variance in reproductive success can lead to departures from this prediction, and so we also evaluated whether any ratio of male- to female- effective population sizes could explain our observed combination of X/A and Y/A diversity.
 
@@
 
-\subsection*{Simulations of positive and purifying selection}
-To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates with those from simulations.
-This approach was preferred over using analytical predictions primarily because the background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for Y chromosomes that lack crossing over.
+\subsection*{Simulations of purifying selection}
+To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a nonrecombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
+This approach was preferred over using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
 
-For purifying selection simulations, we assumed that selection coefficients followed a gamma distribution, and we considered a range of values for the mean selection coefficient against deleterious mutations, $s$, by varying the scale parameter from 0.0001 to 0.1 while keeping the distribution shape parameter constant (see Supporting Information for simulation commands).
-To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for a sex ratio of $N_{m}/N_{f}=0.4$, and we sampled 8 haploid chromosomes per simulation.
-Simulated sequences contained 500 kb of linked neutral sequence from which we calculated diversity.
-
-For positive selection with recurrent hitchhiking, we used sfscoder to implement the method of \citep{uricchio2014robust}, and simulated a 500kb (\X) neutral locus that was flanked by 100kb loci on either side that were subject to positive selection.
-We ran 1000 replicates for each model and calculated the average $\theta_{Y{sim}}$ values and the corresponding Y/A ratios for each model seperately.
-Finally, we tested the significance of the fit of the simulated diversity to our empirical Y/A ratio (\X new figure) by calculating the proportion of simulation replicates for which $\theta_{{Y}sim}$ was not significantly different from $\theta_{{Y}observed}$, using a two-sided P value.
-Commands for running the simulations are provided in the Supporting Information.
+To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (Keightley and Eyre-Walker ref).
+This program fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism from the population genetic analyses above.
+We used this estimated gamma distribution in our forward simulations.
+To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of pi_synonymous .
+To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of Nf/(Nf+Nm)=0.6, and we sampled 6 haploid chromosomes per simulation.
+Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information for simulation commands).
 
@@
 \subsection*{Y-chromosome diversity in \textit{R. hastatulus} is very low}
-Our analysis revealed that diversity on the \textit{R.
-hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001), and 40-fold lower compared to mean diversity on the X chromosome (Table 1).
+Our analysis revealed that, for all sex chromosome races, diversity on the \textit{R.
+hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
+In the XY chromosome race, we also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
 Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
 Conceivably, such low diversity on the Y could arise from a low mutation rate on the Y chromosome, or a lower mutation rate in males compared to females.
-However, these possibilities can be rejected because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
+However, these possibilities are unlikely because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
+Similarly, our estimates of weighted average synonymous substitution rate between R.
+hastatulus and an outgroup species R.
+bucephalophorus from Hough et al(ref) reveal comparable levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219).
 
@@
 Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race.
-If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can driving the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
+If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can drive the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
+However, this would require extensive recombination suppression between the fused and unfused X chromosomes, and it is also possible that sex-specific demographic history has driven these patterns.
 It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
 
-In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 40 fold reduction compared to the mean $\theta_{X}$.
+In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 50 fold reduction compared to the mean $\theta_{Aut}$.
 We next consider several possible models - neutral and selective - that might explain this reduction.

### 2016-10-31 manuscript.tex
edits to methods
minor edits
updated abstract, added latest pdf

@@
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that nucleotide diversity on the Y is more than 50-fold lower than on the autosomes, suggesting an extensive reduction in diversity.
+Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the recombining X chromosome, indicating a significant departure from neutrality and a mean chromosome-wide reduction of $\approx 93\%$.
 We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations and population genetic estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity levels can be explained by the action of linked selection against deleterious mutations, although selective sweeps may also be playing a role.
+Rather, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that levels of neutral polymorphism on the Y can be explained by purifying selection against deleterious mutations, although selective sweeps may have also contributed.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong effects of selective interference experienced by a large number of selected sites, rather than simply neutral drift following gene silencing.
+hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites (at least \textasciitilde 800kb), rather than simply occuring through neutral drift following gene silencing.
 \end{abstract}
@@
 In light of this evidence for significant population substructure, and to consider between-population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
+For model-based analyses, we focused on the population sample from the Texas XY race, given that this reflects the presumed ancestral karyotype and has the largest sample size.
 

### 2016-10-31 manuscript.tex
minor edit to abstract, new pdf

@@
 Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the recombining X chromosome, indicating a significant departure from neutrality and a mean chromosome-wide reduction of $\approx 93\%$.
+Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the recombining X chromosome, indicating a significant departure from neutrality and a mean chromosome-wide reduction of \textasciitilde 93\%.
 We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.

### 2016-10-31 manuscript.tex
fixed sw latex fail in methods

@@
 We used this estimated gamma distribution in our forward simulations.
-To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of pi_synonymous .
+To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of $\pi_{synonymous}$ .
 To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of Nf/(Nf+Nm)=0.6, and we sampled 6 haploid chromosomes per simulation.

### 2016-11-01 manuscript.tex
abstract edits

@@
 On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
-Here, we investigate the factors affecting X- and Y-linked polymorphism during the early stages of plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the recombining X chromosome, indicating a significant departure from neutrality and a mean chromosome-wide reduction of \textasciitilde 93\%.
-We demonstrate that the magnitude of diversity loss is inconsistent with a purely neutral model that accounts for the occurrence of female-biased sex ratios in this species, or by reduced male $N_{e}$ caused by high variance in male fitness.
-Rather, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that levels of neutral polymorphism on the Y can be explained by purifying selection against deleterious mutations, although selective sweeps may have also contributed.
+However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
+Here, we investigate the factors affecting X- and Y-linked polymorphism during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
+Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
+We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male fitness.
+Rather, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that neutral polymorphism on the Y can be explained by the effects of linked purifying selection, although selective sweeps may have also contributed.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study is consistent with the prediction that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites (at least \textasciitilde 800kb), rather than simply occuring through neutral drift following gene silencing.
+hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than occuring through gene silencing follwed by neutral genetic drift.
+%abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.
 \end{abstract}

### 2016-11-01 manuscript.tex
changed title
intro edits, more in progress

@@
 
-\title{Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression}
+\title{Nucleotide Diversity Loss on a Plant Y Chromosome Following Recent Recombination Suppression}
 
@@
 \lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plant and animal kingdoms \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}.
-Despite numerous biological differences between the kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes, suggesting that general evolutionary mechanisms might drive sex chromosome divergence \citep{charlesworth1978,charlesworth1996CB,charlesworth2000degeneration}.
-Investigating the shared features of plant and animal sex chromosomes, including the loss of recombination \citep{bergero2009}, the degeneration of Y chromosomes \citep{hough2014,bergero2015}, and the evolution of dosage compensation \citep{muyle2012,papadopulos2015}, thus provides an informative context for investigating the evolutionary forces driving sex chromosome evolution.
+Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
+For example, in both groups the loss of recombination between X and Y chromosomes is associated with genetic degeneration and loss of genes on the Y chromosome \citep{hough2014,bergero2015}.
+In some species this has also led to the evoltion dosage compensation \citep{charlesworth1996CB,muyle2012,papadopulos2015}, and changes to levels of X and Y genetic diversity \citep{ellegren2011,bachtrog2013NRG}.
+The independent evolution of these phenomena in a broad range of species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in regions that lack recombination is a difficult challenge for both theoreticians and experimentalists \citep{charlesworth1978,charlesworth1996CB,charlesworth2000degeneration}.
 
-One fundamental difference between sex chromosomes and autosomes is their copy number in a population; whereas autosomes in diploid species are present in two copies in each sex, the X chromosome is present in two copies in females and only one copy in males.
-As a consequence, the effective population size ($N_{e}$) of the X chromosome is predicted to be 3/4 that of autosomes, whereas the $N_{e}$ of the male-limited Y chromosome is expected to equal 1/4 that of autosomes (assuming an equal number of reproducing females and males).
+One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
+As a consequence, the neutrally-expectied effective population size ($N_{e}$) of the Y chromosome is predicted to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
 Such differences in $N_{e}$ are predicted to directly affect relative levels of neutral polymorphism maintained on these chromosomes, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ \citep{Kimura1984}.
@@
 
+%In species with male heterogamety, the high variance in male reproductive success associated with sexual selection could further reduce the e¡ective population size of the Y chromosome and (to a smaller degree) of the autosomes, relative to that of the X chromosome (Caballero 1995).
+Ne-values can be estimated from data on variation in silent DNA sequences, if mutation rates for neutral or nearly neutral changes are accurately known (Kimura 1983).
+Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
+
 In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
@@
 
+%distinct but related, process is the `weak selec- tion Hill^Robertson’ e¡ect (McVean & Charlesworth 2000).
+Closely linked selected alleles interfere with each other, inhibiting the spread of favourable alleles and the elimination of deleterious ones (Fisher 1930, chapter5; Muller 1932; Hill & Robertson 1966; Felsenstein 1974; Birky & Walsh 1988).
+
 %discussion point?:
@@
 hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.
+
+% Because the mean size of a eukaryote gene is about 500 codons, and a Drosophila or plant chromosome might have about 2000^3000 genes, the number of third coding p ositions at which silent muta- tions can occur on a proto-Ychromosome is of the order of 106, ten times the number for the most extreme example shownin¢gure1.
 
@@
 
+%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
 \subsection*{Background Selection and Selective Sweeps}
@@
 
+%A predicted consequence of a reduction in Ne is an accelerated rate of ¢xation of slightly deleterious substitu- tions, and a reduction in the rate of ¢xation of slightly advantageous mutations (Kimura 1983).
+Thus we might expect an increased rate of amino-acid replacement rela- tive to silent substitutions on evolving Y chromosomes (provided that su¤cient time has elapsed since the origin of the sex-chromosome system), and more ¢xations of non-preferred versus preferred synonymous substitutions at silent sites in exons (Charlesworth 1996a).
+
 It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
@@
 
+% A further interesting aspect of the results from the S.
+latifolia Y chromosome is that they do not readily ¢t a recent selective sweep.
+Such an event would eliminate all diversity, so that existing SLY1 polymorphisms must have accumulated since the selective sweep and should thus mostly be at low frequencies (see table 1).
+However, all six SLY1 polymorphic sites found are non-singletons, and no signi¢cant deviations from neutrality are detectable by standard test statistics (Fu & Li 1993).
+However, this does not take into account the probable population subdivision in this species.
+Although a selective sweep eliminating variability throughout the worldwide population of this species seems unlikely, e¡ects of selective sweeps within local populations cannot be excluded.
+
+%because the lack of recombination means that all genes on the Yof any given species share a common genealogy, which limits the amount of information that can be obtained from data on within-species variability.
+
+%This raises the further issue of whether Y-chromosome degeneration is driven mainly by the accumulation of deleterious mutations on the Y or is due to processes turning down the expression of genes on an evolving Y.
+%We have focused on the ¢rst of these possibilities, but it might well not completely explain Y degeneration.
+If mutations of relatively minor ¢tness e¡ects, with Nes not much bigger than 1, become ¢xed more rapidly on a degenerating Ychromosome than on the X, selection pres- sure would develop to reduce the transcription of Y genes, provided that this was compensated for in the hetero- gametic sex by an increased activity of their counterparts on the X.
+
 \section*{Conclusions}
+
+The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8).
+The accumulation of deleterious mutations and extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the reduced ability of selection to eliminate deleterious alleles from large genomic regions that lack recombination.
 

### 2016-11-01 manuscript.tex
minor abstract edit

@@
 However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
-Here, we investigate the factors affecting X- and Y-linked polymorphism during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism for genes on the Y chromosome is \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
+Here, we investigate the factors affecting X- and Y-linked diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
+Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
 We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male fitness.
@@
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than occuring through gene silencing follwed by neutral genetic drift.
+hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than gene silencing followed by neutral genetic drift.
 %abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.

### 2016-11-01 manuscript.tex
extensive edits to intro. Still in progress. Changes to methods (variance in rep success) are staged.

@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-For example, in both groups the loss of recombination between X and Y chromosomes is associated with genetic degeneration and loss of genes on the Y chromosome \citep{hough2014,bergero2015}.
-In some species this has also led to the evoltion dosage compensation \citep{charlesworth1996CB,muyle2012,papadopulos2015}, and changes to levels of X and Y genetic diversity \citep{ellegren2011,bachtrog2013NRG}.
-The independent evolution of these phenomena in a broad range of species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in regions that lack recombination is a difficult challenge for both theoreticians and experimentalists \citep{charlesworth1978,charlesworth1996CB,charlesworth2000degeneration}.
+For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species this genetic degeneration has led to the evolution dosage compensation \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}, and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.
+The independent evolution of these phenomena in a broad range of species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a difficult challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
 
 One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-As a consequence, the neutrally-expectied effective population size ($N_{e}$) of the Y chromosome is predicted to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-Such differences in $N_{e}$ are predicted to directly affect relative levels of neutral polymorphism maintained on these chromosomes, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ \citep{Kimura1984}.
-Such variation in neutral polymorphism within the genome can have important consequences on patterns of DNA sequence evolution, including the effectiveness of both positive and negative selection \citep{charlesworth1987}.
+Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that delterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by selective interference among genetically linked sites (the "Hill-Robertson Effect").
+These effects both predict a reduction in the level of neutral polymorphism maintained on the Y chromosome, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
 
-Several demographic factors are expected to modulate or accentuate differences in neutral polymorphism maintained on sex chromosomes \citep{ellegren2009}.
-These include population subdivision and sex-biased dispersal, deviations from a 1:1 breeding sex ratio, and high sex-specific variance in reproductive success \citep{caballero1995,charlesworth2001,laporte2002,pool2007}.
-In theory, these processes all result in asymmetries in female vs male $N_{e}$, that could lead to different levels of neutral polymorphism on the sex chromosomes than are expected based on the autosomes (\citep{pool2007}).
-As evidence for this, coalescent simulations have shown that models of male-biased migration during human out-of-Africa dispersal result in levels of X/A neutral diversity similar to what is observed \citep{keinan2009} (though it is worth noting that there is considerable variation in current estimates of X/A in humans, and other demographic scenarios have been suggested \citep{hammer2010,bustamante2009}).
-In addition, disproportionate losses of sex-linked compared to autosomal variation arising from population bottleneck events have been inferred in chimpanzees and orangutans \citep{kaessmann2001,fischer2006}, the house mouse \textit{Mus musculus} \citep{baines2007}, and Drosophila \citep{andolfatto2001}, highlighting the importance of demography in affecting patterns of sex-linked and autosomal neutral variation.
+In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success could also reduce the $N_{e}$ of the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting the need to distinguish the effects of variance in reproductive success from the effects selective interference just mentioned.
+Because high variance in male reproductive success is expected to reduce both Y-linked $N_{e}$ and and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by examining levels of neutral variability on X and Y chromosomes relative to values for the autosomes.
+In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in X/A.
+Based on this comparison, studies of the X/A ratio in humans have suggested a historical excess of breeding females over males resulting in an inflated X/A ratio (though there is considerable variation in estimates of X/A in humans, and other demographic scenarios have also been suggested \citep{hammer2010,bustamante2009}).
+%as modeled by \citep{kimura1964number} and \citep{nomura2002effective}.
 
-%In species with male heterogamety, the high variance in male reproductive success associated with sexual selection could further reduce the e¡ective population size of the Y chromosome and (to a smaller degree) of the autosomes, relative to that of the X chromosome (Caballero 1995).
-Ne-values can be estimated from data on variation in silent DNA sequences, if mutation rates for neutral or nearly neutral changes are accurately known (Kimura 1983).
-Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
-
-In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
-The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations
-) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference).
+%In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
+The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference).
 In all of these processes, linkage  disequilibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
 
-Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with evidence for linked selection playing a key role.
-
-%distinct but related, process is the `weak selec- tion Hill^Robertson’ e¡ect (McVean & Charlesworth 2000).
-Closely linked selected alleles interfere with each other, inhibiting the spread of favourable alleles and the elimination of deleterious ones (Fisher 1930, chapter5; Muller 1932; Hill & Robertson 1966; Felsenstein 1974; Birky & Walsh 1988).
+%Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with evidence for linked selection playing a key role.
 
@@
 
-% Because the mean size of a eukaryote gene is about 500 codons, and a Drosophila or plant chromosome might have about 2000^3000 genes, the number of third coding p ositions at which silent muta- tions can occur on a proto-Ychromosome is of the order of 106, ten times the number for the most extreme example shownin¢gure1.
-
-Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
+%Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
 Background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes.
@@
 
-To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed single nucleotide polymorphisms (SNPs) on sex chromosomes and autosomes in the plant \textit{R.
+To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R.
 hastatulus }(Polygonaceae).
@@
 hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-This is important because female-biased sex ratios are predicted to affect chromosomal $N_{e}$ and therefore the level of neutral polymorphism maintained on sex chromosomes and autosomes \citep{ellegren2009}.
-Here, we test whether patterns of nucleotide diversity on sex chromosomes and autosomes in \textit{R.
-hastatulus} are jointly consistent with neutral evolutionary processes - including the effects of sex ratio bias and variance in reproductive success - or whether the removal of linked neutral variation caused by positive and purifying selection has played a more significant role.
+Using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and variance in reproductive success, or whether the removal of linked neutral variation caused by purifying selection has played a more significant role.
 

### 2016-11-01 manuscript.tex
Update manuscript.tex

@@
 Here, we investigate the factors affecting X- and Y-linked diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
+Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value for autosomes, indicating a significant departure from neutrality.
 We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male fitness.
@@
 Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that delterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by selective interference among genetically linked sites (the "Hill-Robertson Effect").
+The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by selective interference among genetically linked sites (the "Hill-Robertson Effect").
 These effects both predict a reduction in the level of neutral polymorphism maintained on the Y chromosome, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
@@
 In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success could also reduce the $N_{e}$ of the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting the need to distinguish the effects of variance in reproductive success from the effects selective interference just mentioned.
-Because high variance in male reproductive success is expected to reduce both Y-linked $N_{e}$ and and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by examining levels of neutral variability on X and Y chromosomes relative to values for the autosomes.
+Because high variance in male reproductive success is expected to reduce both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by examining levels of neutral variability on X and Y chromosomes relative to values for the autosomes.
 In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in X/A.
@@
 hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and variance in reproductive success, or whether the removal of linked neutral variation caused by purifying selection has played a more significant role.
+Using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and variance in reproductive success, or whether the removal of linked neutral variation caused by purifying selection is also playing a significant role.
 
@@
 Reads from these 24 samples were mapped to the \textit{R.
-hastatulus} reference transcriptome \citep{hough2014}, and raw sequences are available from the Gen-Bank Short Read Archive under accession no.
+hastatulus} reference transcriptome \citep{hough2014}, and raw sequences are available from the GenBank Short Read Archive under accession no.
 SRP041588.
@@
 Because the criteria for identifying autosomal loci in Hough et al.
-(2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosome.
+(2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosomal gene.
 Therefore, for the present analysis we incorporated the broader set of all non-sex linked (putatively autosomal) genes from our transcriptome data.
@@
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are likely violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes.
+Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes.
 To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio using coalescent theory (Figure S1).
-Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this would involve maximizing the likelihood estimate of the mutation rate for each locus independently.
+Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this involves maximizing the likelihood estimate of the mutation rate for each locus independently.
 Therefore, we applied a method that assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
@@
 To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a nonrecombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
-This approach was preferred over using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
+This approach was implemented rather than using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
 
@@
 A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
-X & 0.0047 & 0.85 & 0.002 & 0.33 & 0.0047 & 0.37 \\
+X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
 Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\

### 2016-11-01 manuscript.tex
changes abstract back. See comments

@@
 Here, we investigate the factors affecting X- and Y-linked diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value for autosomes, indicating a significant departure from neutrality.
+Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
 We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male fitness.

### 2016-11-02 manuscript.tex
Update manuscript.tex

@@
 On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male fitness can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
 Here, we investigate the factors affecting X- and Y-linked diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
 Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
-We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male fitness.
+We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male reproductive success.
 Rather, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that neutral polymorphism on the Y can be explained by the effects of linked purifying selection, although selective sweeps may have also contributed.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than primarily gene silencing followed by neutral genetic drift.
 %abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.

### 2016-11-02 manuscript.tex
Update manuscript.tex

@@
 \subsection*{Y-chromosome diversity in \textit{R. hastatulus} is very low}
-Our analysis revealed that, for all sex chromosome races, diversity on the \textit{R.
+Our analysis reveals that, for the Texas sex chromosome race, diversity on the \textit{R.
 hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
-In the XY chromosome race, we also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
+We also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
 Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
@@
 bucephalophorus from Hough et al(ref) reveal comparable levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219).
+
+After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig).
+Surprisingly, this analysis revealed that the Y chromosome from this race appears to be paraphyletic, with XYY samples from Florida being more closely related to the TX race than either is to samples from South Carolina (suppfig).
+Once we examine diversity for these two subclades separately, a strong loss of diversity on the Y chromosome is once again apparent (Table 1).
+
+Although our sampling of each \textit{R.
+hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a secondarily-derived $XY_{1}Y_{2}$ sub-clade.
+As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the Florida sub-clade inferred here likely originated through hybridization between a female from the SC clade harboring the X-autosome fusion, and a male from the XY Texas race.
 
@@
 \end{table}
-
-Although our sampling of each \textit{R.
-hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a derived $XY_{1}Y_{2}$ sub-clade.
-As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the SC sub-clade inferred here likely originated through hybridization between a female from the FL clade harboring the X-autosome fusion, and a male from the XY Texas race.
-Notably, when we pooled the samples from each sub-clade together, the estimated level of Y diversity was significantly higher in the pooled data, highlighting the presence of strong substructure between these groups (Figure S2).
 

### 2016-11-02 manuscript.tex
new changes to abstract

@@
 \usepackage{geometry} % just not to bother with the table width
-%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
-%\setlength{\marginparwidth}{1.25cm}
-%\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
-%\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
-%\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
-%\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
-\newcommand{\X}{\textcolor{red}{\bf X\,}}
+
+\usepackage{todonotes}
 
@@
 On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome diversity.
-Here, we investigate the factors affecting X- and Y-linked diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that average neutral polymorphism levels for genes on the Y chromosome are \textasciitilde 2.1\% the value of their homologues on the X chromosome, indicating a significant departure from neutrality and a chromosome-wide reduction of \textasciitilde 93\%.
-We demonstrate that the magnitude of this diversity loss is inconsistent with neutral models that account for the occurrence of female-biased sex ratios, or by reduced male $N_{e}$ caused by high variance in male reproductive success.
-Rather, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that neutral polymorphism on the Y can be explained by the effects of linked purifying selection, although selective sweeps may have also contributed.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome polymorphism.
+Here, we investigate the factors affecting X- and Y-linked nucleotide diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
+Strikingly, we find that neutral polymorphism levles for genes on the Y are on average \textasciitilde 2.1\% the value of their homologues on the X, indicating a significant chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
+We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
+Rather, using forward population genetic simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that neutral polymorphism on the Y can be explained by the effect of purifying selection elminating linked neutral variation, although selective sweeps may have also contributed.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that the early stages of Y chromosome degeneration are characterized by strong selective interference among many linked sites rather than primarily gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes, our study suggests that Y chromosome degeneration in the early stages occurs through strong selective interference effects among many linked sites rather than primarily through Y-inactivation followed by neutral genetic drift.
 %abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.
+
 \end{abstract}

### 2016-11-03 manuscript.tex
cleanup of .tex metadata
abstract edits

@@
 \usepackage{adjustbox}
-\usepackage{geometry} % just not to bother with the table width
-
-\usepackage{todonotes}
-
+\usepackage{geometry}
 \articletype{gos} % article type
-% {inv} Investigation
-% {gs} Genomic Selection
-% {goi} Genetics of Immunity
-% {gos} Genetics of Sex
-% {mp} Multiparental Populations
 
@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can significantly affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, selection can remove genetically-linked neutral variation, resulting in reduced $N_{e}$ and the level of polymorphism on Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-chromosome polymorphism.
-Here, we investigate the factors affecting X- and Y-linked nucleotide diversity during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
-Strikingly, we find that neutral polymorphism levles for genes on the Y are on average \textasciitilde 2.1\% the value of their homologues on the X, indicating a significant chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which affect levels of genetic variability within the genome.
+On Y chromosomes with suppressed recombination, selection is expected to remove linked neutral variation and reduce the $N_{e}$ of the Y compared to X chromosomes or autosomes.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the underlying causes of low Y-linked nucleotide diversity.
+Here, we investigate the factors affecting X- and Y-linked neutral polymorphsm during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
+Strikingly, we find that neutral polymorphsm for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corrsponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Rather, using forward population genetic simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that neutral polymorphism on the Y can be explained by the effect of purifying selection elminating linked neutral variation, although selective sweeps may have also contributed.
+Rather, using forward population genetic simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity can be explained by purifying selection removing linked neutral variation, although selective sweeps may have also contributed.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that Y chromosome degeneration in the early stages occurs through strong selective interference effects among many linked sites rather than primarily through Y-inactivation followed by neutral genetic drift.
+hastatulus} sex chromosomes, our study suggests that Y chromosome degeneration in the early stages occurs through strong selective interference among many linked sites rather than primarily through Y-inactivation followed by neutral genetic drift.
 %abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.
-
 \end{abstract}

### 2016-11-07 manuscript.tex
edits up to methods, removal of NC from study

@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, selection is expected to remove linked neutral variation and reduce the $N_{e}$ of the Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the underlying causes of low Y-linked nucleotide diversity.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability within the genome.
+On Y chromosomes with suppressed recombination, selection is expected to remove linked neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y diversity.
 Here, we investigate the factors affecting X- and Y-linked neutral polymorphsm during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
-Strikingly, we find that neutral polymorphsm for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corrsponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
+Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corrsponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species this genetic degeneration has led to the evolution dosage compensation \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}, and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.
-The independent evolution of these phenomena in a broad range of species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a difficult challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
+For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, this genetic degeneration of the Y has led to the evolution dosage compensation of the X \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+The independent evolution of these phenomena across a broad range of distantly-related species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
+% and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.
 
@@
 
-In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success could also reduce the $N_{e}$ of the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting the need to distinguish the effects of variance in reproductive success from the effects selective interference just mentioned.
-Because high variance in male reproductive success is expected to reduce both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by examining levels of neutral variability on X and Y chromosomes relative to values for the autosomes.
-In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in X/A.
-Based on this comparison, studies of the X/A ratio in humans have suggested a historical excess of breeding females over males resulting in an inflated X/A ratio (though there is considerable variation in estimates of X/A in humans, and other demographic scenarios have also been suggested \citep{hammer2010,bustamante2009}).
+In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of linked selection on the Y need to be distinguished from neutral models that include sex ratio variation as well as male-female differences in reproductive success.
+High variance in male reproductive success, for example, is predicted to reduce both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, and therefore evidence for this effect can be obtained by comparing levels of neutral variability on X and Y chromosomes relative to the autosomes.
+In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
+Based on such comparisons, studies in humans have found examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, and the inflated X/A ratio has been suggested to reflect a historical excess of breeding females over males \citep{hammer2008sex}, though there is considerable variation in X/A estimates in human and other demographic scenarios have also been suggested \citep{bustamante2009,hammer2010}.
+
+Despite widespread interest in determining the effects of selection patterns of sex-linked variability diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about how linked selection has affected patterns nucleotide polymorphism.
+In humans and Drosophila \citep{mcallister1999,bachtrog2000}, where large population genomic data sets are readily available, studies examining Y chromosome variability have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014}.
+However, it remains unclear whether such inferences about the effects of linked selection on the Y should also apply to systems in which sex chromosomes evolved \textit{do novo} from autosomes much more recently.
+The effects of diversity loss due to selection might be expeted to be lower in such systems because of a shorter history of recombination suppression, but contrary to this, simulations reveal that background selection and selective sweeps both have strongest effects during the early stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
+Indeed, studies of Y-degeneration in the young plant sex chromosomes \textit{R.
+hastatulus} and \textit{Silene latifolia} have both revealved signs of rapid Y chromosome degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
+In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing) \citep{orr1998adaptive}, then it is possible that the strength of linked selection would then be minimized such that gene loss was mostly driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG} rather than selective interference due to widespread linkage disequilibrium.
+
+%The strength of this effect is dependent on the rate of recombination between a selected locus and a neighbouring region; essentially, the effect becomes more pronounced with less recombination because of linkage disequilibrium between adjacent loci.
+
+%On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimized rapidly even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes (cite backtrog NRG review).
+
+%because recombination suppression was relatively recent on plant Y chromosomes, it is an open question whether weak-HRI has had enough time to have a large effect.
+Although selective interference (indlucing BGS and Selective Sweeps) is expected to have the greatest effects during the early stages of Y-chromosome evolution \citep{bachtrog2008temporal} the R.
+hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.
+
+%Background selection (purifying selection against deleter- ious alleles) [27] and selective sweeps (positive selection for beneficial variants) [28] both have the effect of reducing nucleotide diversity at linked sites.
 %as modeled by \citep{kimura1964number} and \citep{nomura2002effective}.
 
-%In addition to demographic factors, evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
+%evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
 The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference).
@@
 
-%Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, studies examining Y chromosome variability in mammals \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014} and Drosophila \citep{mcallister1999,bachtrog2000} have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution, with evidence for linked selection playing a key role.
+%Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, s
 
 %discussion point?:
-%Moreover, because recombination suppression was relatively recent on plant Y chromosomes, it is an open question whether weak-HRI has had enough time to have a large effect.
-Although selective interference (indlucing BGS and Selective Sweeps) is expected to have the greatest effects during the early stages of Y-chromosome evolution \citep{bachtrog2008temporal} the R.
-hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.
-
-%Despite widespread interest in determining the effects of recombination on patterns of genetic diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of linked selection on patterns nucleotide polymorphism in younger sex chromosome systems, where \textit{de novo} recombination suppression evolved relatively recently \citep{charlesworth2016plant}.
-Background selection and selective sweeps are predicted to have the strongest effects during the earliest stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes.
-On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimized rapidly even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes (cite backtrog NRG review).
+%Moreover,
 
@@
 Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant gene loss and functional deterioration \citep{hough2014}.
+In order to simplify our comparison of polymorphism levels within the genome, we focus on the ancestral Y chromosome that occurs in both sex chromosome races.
 
 Of particular relevance to the present study, \textit{R.
-hastatulus} populations have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and variance in reproductive success, or whether the removal of linked neutral variation caused by purifying selection is also playing a significant role.
+hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
+Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{Rumex} provides an excellent opportunity to study both the demographic and selective effects contributing to sex-linked variability.
+Here, using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and high variance in reproductive success, or whether the removal of linked neutral variation caused by selection on the Y has led to a loss of diversity greater than expected under neutrality.
 

### 2016-11-07 manuscript.tex
minor edit on last push

@@
 Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{Rumex} provides an excellent opportunity to study both the demographic and selective effects contributing to sex-linked variability.
-Here, using genome-wide estimates of neutral polymorphism together with forward population genetic simulations, we test whether levels of genetic diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and high variance in reproductive success, or whether the removal of linked neutral variation caused by selection on the Y has led to a loss of diversity greater than expected under neutrality.
+Here, using genome-wide estimates of polymorphism together with forward population simulations, we test whether levels of neutral diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and high variance in reproductive success, or whether the removal of linked neutral variation caused by selection on the Y has led to greater loss of diversity than expected under neutrality.
 

### 2016-11-07 manuscript.tex
minor edit on last push

@@
 (2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosomal gene.
-Therefore, for the present analysis we incorporated the broader set of all non-sex linked (putatively autosomal) genes from our transcriptome data.
+Therefore, for the present analysis we incorporated the larger set of all non-sex linked (putatively autosomal) genes from our transcriptome data.
 We also filtered this set to remove genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
@@
 To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1).
-We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
+We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
@@
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this involves maximizing the likelihood estimate of the mutation rate for each locus independently.
-Therefore, we applied a method that assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
-Our model therefore had two parameters; theta for autosomal genes and the Y/A ratio.
-We varied both parameters across a grid of points and evaluated the likelihood, where the Y/A ratio varied from 0.001 to 1, and Watterson's theta on the autosomes varied from 0.001 to 0.01.
-We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
+Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
+Our model therefore had two parameters; $\theta_{autosomal}$ and the Y/A ratio.
+We varied both parameters across a grid of points and evaluated the likelihood, where the Y/A ratio varied from 0.001 to 1, and $\theta_{autosomal}$ varied from 0.001 to 0.01.
+%We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
 As no significant difference was observed (\X Figure?
@@
 
-Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
+%Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
 To do this, we constructed neighbor-joining trees of sex-linked sequences from all populations in our study.
@@
 To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of Nf/(Nf+Nm)=0.6, and we sampled 6 haploid chromosomes per simulation.
-Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information for simulation commands).
+Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information).
 
@@
 
-\subsection*{Y-chromosome diversity in \textit{R. hastatulus} is very low}
-Our analysis reveals that, for the Texas sex chromosome race, diversity on the \textit{R.
-hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
+\subsection*{Extensive loss of Y-chromosome diversity}
+Our analysis reveals that diversity on the \textit{R.
+hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, ....which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
 We also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
 Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
-Conceivably, such low diversity on the Y could arise from a low mutation rate on the Y chromosome, or a lower mutation rate in males compared to females.
-However, these possibilities are unlikely because there is no evidence that the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are significantly different \citep{hough2014}.
-Similarly, our estimates of weighted average synonymous substitution rate between R.
-hastatulus and an outgroup species R.
-bucephalophorus from Hough et al(ref) reveal comparable levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219).
+Conceivably, such low diversity on the Y could arise from a low Y-linked mutation rate, or a lower mutation rate in males compared to females.
+However, these possibilities are unlikely because the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are not significantly different \citep{hough2014}.
+Similarly, our estimates of weighted average synonymous substitution rate between \textit{R.
+hastatulus} and the outgroup \textit{R.
+bucephalophorus} reveal similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
 
-After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig).
+%After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig).
 Surprisingly, this analysis revealed that the Y chromosome from this race appears to be paraphyletic, with XYY samples from Florida being more closely related to the TX race than either is to samples from South Carolina (suppfig).
@@
 
-Although our sampling of each \textit{R.
+%Although our sampling of each \textit{R.
 hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a secondarily-derived $XY_{1}Y_{2}$ sub-clade.
@@
 
-Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
+%Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
 Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}.
@@
 
- The occurrence of female-biased sex ratios in this species has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the neutrally expected $N_{e}$ of the Y chromosome (\X ref).
-This reduction $N_{e}$ on the Y is expected to be further accented if there is high variance in male reproductive success (Figure 1), which is not unusual in annual plants such as \textit{R.
-hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Moreover, given that male plants in this wind-pollinated species produce large amounts of pollen, and that female flowers are uniovulate, we expected that there is strong competition among males to fertilize females.
+The occurrence of female-biased sex ratios in \textit{R.
+hastatulus} has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the neutrally expected $N_{e}$ of the Y chromosome (\X ref).
+This reduction $N_{e}$ on the Y is expected to be further accenuated with high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
+hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+Moreover, given that male plants in this species produce large amounts of pollen, and that female flowers are uniovulate, we expected strong competition among males to fertilize females.
+%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
 
@@
 hastatulus}.
-However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could explain the level of Y/A diversity that we observed (see Methods).
+However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could jointly explain the level of Y/A diversity that we observed (see Methods).
 Conditioning on estimates of sex ratio bias in \textit{R.
-hastatulus} that have been estimated, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
+hastatulus}, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
 This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), and remains significant even if we consider an upper bound estimate obtained from our likelihood based estimated of the confidence interval.
 This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.
-(\X I think some of this can be written better...)
 
@@
 
-%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
 \subsection*{Background Selection and Selective Sweeps}
@@
 
-It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
+%It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
 Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
@@
 %This raises the further issue of whether Y-chromosome degeneration is driven mainly by the accumulation of deleterious mutations on the Y or is due to processes turning down the expression of genes on an evolving Y.
+
 %We have focused on the ¢rst of these possibilities, but it might well not completely explain Y degeneration.
@@
 The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8).
-The accumulation of deleterious mutations and extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the reduced ability of selection to eliminate deleterious alleles from large genomic regions that lack recombination.
+The extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the ....
+a large genomic regions that lack recombination.
 

### 2016-11-07 manuscript.tex
edits to variance in rep success section

@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that jointly considers these two effects.
 
-In particular, the effective population size for genes on each chromosome can be given by  \citep{wright1931evolution}:
+To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that considers these two effects (crow and kimura).
+In particular, assuming a sex ratio of 0.6, we calculated the effect of varaince in reproductive success on male and female $N_{e}$, as:
+
+by Ne = ((N(k -1))/((k -1)+(Vk / k)), where k and Vk are the mean and variance in offspring number, respectively (Kimura and Crow 1964).
+Using sex-specific values of k and Vk to calculate male and female Ne (Nm and Nf , respectively)
+
+respectively, where Vk is the variance in the number of progeny (Kimura and Crow, 1963; Crow and Denniston, 1988).
+Assuming a constant population size, the Ne for autosomes and sex chromosomes then be calculated using the expression from \citep{wright1931evolution}:
 
 \begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:NeA}
+N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:Ne}
 \end{equation}
@@
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}}\label{eq:NeX}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}},\label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e{Y}} = \frac{N_{m}}{2}\label{eq:NeY}
+N_{e{Y}} = \frac{N_{m}}{2},\label{eq:NeY}
 \end{equation}
+
+and the X/A and Y/A ratios as:
+
+\begin{equation}
+\frac{N_{e{X}}}{N_{e{A}} } = \frac{9(N_{m} + N_{f}}{8(2N_{m} + N_{f})} \label{eq:NeXNeA}
+\end{equation}
+
+and
+
+respectively.
+Thus, with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
+However, as $Ne_{m}$ is reduced relative to $Ne_{f}$, the X/A ratio increases amd may become greater than 1, approaching 1.125 at the limit \citep{caballero1995}.
+With the level of neutral polymorphism at equilibrium given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1C shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio across a range of values for the variance in male reproductive success.
+
+%To consider the combined effects on diversity of our empirically estimated female-biased sex ratio ($(N_{f}$/$N_{m}+N_{f})$=$0.6$ together with high variance in male reproductive success, we calculated the nuetrally-expected X/A and Y/A ratios under a 0.6 sex ratio across a range of values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
+%\begin{equation}
+%N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{f})+N_{f}(1+C^2_{m})}\label{eq:NeA}
+%\end{equation}
+
+%where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, defined as the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
+Similarly, for X-linked loci, where females contain two-thirds and males one-third of the alleles:
+
+%\begin{equation}
+%N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{f})+2N_{f}(1+C^2_{m})}\label{eq:NeX}
+%\end{equation}
+
+%With no variation in mating success in either sex ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 1 and 2 reduced to the familiar expressions from \citep{wright1931evolution}:
+
+%\begin{equation}
+%\label{eq:NeY}
+%\end{equation}
+
+%Because the $N_{e}$ for the Y chromosome depends on the male effective population size, $N_{m}$, the effect of variance in reproductive success on $N_{Y}$ can be modeled using the approach of Crow and Kimura (refs; and see ) and $N_{m}$ is given by:
+
+%\begin{equation}
+%N_{e{X}} = \frac{N(k-1)}{k-1}
+%\end{equation}
+
+%\begin{equation}
+%\label{eq:NeY}
+%\end{equation}
 
@@
 \centering
-\includegraphics[width=\linewidth]{Figure1.png}
+\noindent
+\includegraphics[width=\linewidth]{figure1.jpg}
 \caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).

### 2016-11-08 manuscript.tex
finished new edits up to methods

@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, this genetic degeneration of the Y has led to the evolution dosage compensation of the X \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the genetic degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
 The independent evolution of these phenomena across a broad range of distantly-related species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
@@
 Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by selective interference among genetically linked sites (the "Hill-Robertson Effect").
-These effects both predict a reduction in the level of neutral polymorphism maintained on the Y chromosome, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
+The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by interference among selected but genetically linked sites (the "Hill-Robertson Effect").
+Thus, a reduction in the level of neutral polymorphism maintained on the Y chromosome is expected, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
 
-In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of linked selection on the Y need to be distinguished from neutral models that include sex ratio variation as well as male-female differences in reproductive success.
-High variance in male reproductive success, for example, is predicted to reduce both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, and therefore evidence for this effect can be obtained by comparing levels of neutral variability on X and Y chromosomes relative to the autosomes.
+In species with female-biased reproductive sex ratios or extensive male-male competiton, high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of linked selection on the Y need to be distinguished from neutral models that include sex ratio variation as well as male-female differences in reproductive success.
+Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by comparing levels of neutral variability on X and Y chromosomes, relative to values obtained for the autosomes.
 In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans have found examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, and the inflated X/A ratio has been suggested to reflect a historical excess of breeding females over males \citep{hammer2008sex}, though there is considerable variation in X/A estimates in human and other demographic scenarios have also been suggested \citep{bustamante2009,hammer2010}.
+Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though it the inflated X/A ratio in humans has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic})
 
-Despite widespread interest in determining the effects of selection patterns of sex-linked variability diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about how linked selection has affected patterns nucleotide polymorphism.
-In humans and Drosophila \citep{mcallister1999,bachtrog2000}, where large population genomic data sets are readily available, studies examining Y chromosome variability have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014}.
-However, it remains unclear whether such inferences about the effects of linked selection on the Y should also apply to systems in which sex chromosomes evolved \textit{do novo} from autosomes much more recently.
-The effects of diversity loss due to selection might be expeted to be lower in such systems because of a shorter history of recombination suppression, but contrary to this, simulations reveal that background selection and selective sweeps both have strongest effects during the early stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
+Despite widespread interest in determining the evolutionary factors affecting sex-linked variability \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of demography or selection on patterns of nucleotide diversity in more recently evolved sex chromosomes, such as in dioecious plants.
+We therefore don't yet know the time-scales over which processes such as linked selection on the Y chromosome are likely to be important.
+In humans, for example, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, and simulations have revealed that levels of Y-diversity are consistent with a combination of very low $N_{m}$ and strong purifying selection eliminiting linked neutral variation \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is an open question whether such strong effects of purifying selection should also apply to systems where sex chromosomes evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA as in the case of diecious plants \citep{charlesworth2015plant}).
+The effects of diversity loss due to purifying selection might on the one hand be expeted to be lower in such systems because of a shorter history of recombination suppression.
+Contrary to this intuition, however, simulations have shown that background selection and selective sweeps both have the strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
 Indeed, studies of Y-degeneration in the young plant sex chromosomes \textit{R.
-hastatulus} and \textit{Silene latifolia} have both revealved signs of rapid Y chromosome degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing) \citep{orr1998adaptive}, then it is possible that the strength of linked selection would then be minimized such that gene loss was mostly driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG} rather than selective interference due to widespread linkage disequilibrium.
+hastatulus} and \textit{Silene latifolia} have found signs of rapid Y chromosome degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
+In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then it is possible that the strength of linked selection would then be minimized such that the inferred gene loss was primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
 
@@
 
-To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R.
+To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R.
 hastatulus }(Polygonaceae).
 This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than the well-studied mammalian sex chromosomes \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the unique occurrence of an intraspecific sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct 'chromosomal races' \citep{smith1963mechanism}.
+\textit{Rumex hastatulus} has also received particular attention because of the unique occurrence of an intraspecific polymorphism in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ('chromosomal races') \citep{smith1963mechanism}.
 The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant gene loss and functional deterioration \citep{hough2014}.
-In order to simplify our comparison of polymorphism levels within the genome, we focus on the ancestral Y chromosome that occurs in both sex chromosome races.
+Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone significant gene loss and functional deterioration \citep{hough2014}.
+Here, in order to simplify our comparison of polymorphism levels between X, Y, and autosomes, we focused on the ancestral Y chromosome, which occurs in both sex chromosome races.
 
@@
 hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{Rumex} provides an excellent opportunity to study both the demographic and selective effects contributing to sex-linked variability.
-Here, using genome-wide estimates of polymorphism together with forward population simulations, we test whether levels of neutral diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and high variance in reproductive success, or whether the removal of linked neutral variation caused by selection on the Y has led to greater loss of diversity than expected under neutrality.
+Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{R.
+hastatulus} provides an excellent opportunity to study both the demographic and selective factors contributing to sex-linked variability.
+Using estimates of polymorphism estimated on X, Y, and autosomals, together with forward population simulations, our study aims to test whether levels of sex-linked neutral diversity in this species are consistent with models of neutral evolution, including the effects of sex ratio bias and high variance in reproductive success, or whether linked selection has had a significant impact on patterns of nucleotide variability during the early stages of sex chromosome evolution.
 

### 2016-11-08 manuscript.tex
methods: finished edits up to simulations

@@
 \subsection*{Population Samples and Sex-Linked Genes}
-We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations; 6 populations of each sex chromosome race).
+We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 12 males and 12 females (1 male and 1 female from each of 6 populations).
 Samples were collected in 2010 from throughout the native range of \textit{R.
-hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females.
+hastatulus} (locations in Table S1), and plants were grown in the glasshouse at the University of Toronto from seeds collected from open-pollinated females.
 We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich).
-The isolation of mRNA and cDNA synthesis was conducted according to standard Illumina RNAseq procedures, with sequencing conducted on two Illumina HiSeq lanes with 150-bp end reads at the Genome Quebec Innovation Center.
-Reads from these 24 samples were mapped to the \textit{R.
-hastatulus} reference transcriptome \citep{hough2014}, and raw sequences are available from the GenBank Short Read Archive under accession no.
+Isolation of mRNA and cDNA synthesis was conducted according to standard Illumina RNAseq procedures, and sequencing was conducted on two Illumina HiSeq lanes with 150-bp end reads at the Genome Quebec Innovation Center.
+Reads from these samples were mapped to the \textit{R.
+hastatulus} reference transcriptome \citep{hough2014} using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}.
+We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
+The population samples analyzed here were previously reported in Hough et al.
+(2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
 SRP041588.
-Reads were mapped using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}.
-We used Picard tools (http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores <20.
-The population samples analyzed here were previously reported in Hough et al.
-(2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis.
-Here, to consider sex linked genes that were identified in each of our sequenced population samples, we focused on the set of 460 X/Y genes for which the Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).
+Here, to consider sex linked genes that were identified in all of our sequenced population samples, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).
 
@@
 In evaluating evidence for nucleotide diversity differences between X and Y chromosomes, it is important to distinguish between reduced Y-linked diversity, and the possibility that X-linked diversity is elevated above the level predicted from a neutral model.
-To do this, we normalized sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below).
-Because the criteria for identifying autosomal loci in Hough et al.
-(2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosomal gene.
-Therefore, for the present analysis we incorporated the larger set of all non-sex linked (putatively autosomal) genes from our transcriptome data.
-We also filtered this set to remove genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
-In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes with SNPs showing fixed heterozygosity in males and fixed homozygosity in females, (iii) any genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicate genes or those with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
-Finally, we removed genes with fewer than 100 synonymous sites following this filtering to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
-This filtering resulted in a set of 12,356 and 11,350 autosomal genes in the Texas and North Carolina races, respectively.
+To do this, we normalized our sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below).
+Because the criteria for ascertaining autosomal loci in Hough et al.
+(2014) were based on identifying four segregating SNPs per locus, and since this set of genes is likely to be higher in diversity than the average autosomal gene, here we instead used the larger set of all non-sex linked (putatively autosomal) genes as our autosomal reference.
+We filtered genes in this set to remove any genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
+In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicates or genes with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
+Finally, we removed genes with fewer than 100 synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
+This filtering resulted in a final set of 12,356 autosomal genes.
 
 \subsection*{Phasing X and Y alleles}
-To estimate polymorphism for the X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
+To estimate polymorphism for X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
 In previous work, phasing alleles on \textit{R.
@@
 Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome.
-Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences with each haplotype block (i.e., sites in which all females were homozygous, and all males were heterozygous).
+Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences within each haplotype block (i.e., sites where all females were homozygous, and all males were heterozygous).
 Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output.
-This was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
+In particular, this was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
 Finally, we filtered out SNPs with coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels.
-This procedure was conducted using a combination of Perl and Bash scripts, and resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes from the 24 individuals in our study.
+This filtering procedure resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes.
 
@@
 (2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was known to be zero.
-We identified 7 percent of sex-linked genes with phasing errors of this kind, or which were otherwise determined to have genotyping errors resulting in false SNP calls.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level polymorphism, obtained from the genetic cross, was zero.
+We identified 7 \% of sex-linked genes that either had  phasing errors of this kind genotyping errors.
 This corresponds to a SNP error rate estimate of 1.7 x 10-4.
-Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism.
-However, because the rate is high relative to the expected level of true polymorphism on the Y-chromosome, we  further filtered genes in which we found evidence for false-positive Y-polymorphisms arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
-This filtering was done by manually inspecting sequences in IGV \citep{robinson2011integrative} and identifying each individual putative polymorphism on the Y chromosome.
+Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of the X/A ratio.
+However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we further filtered genes in which we found evidence for false-positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
+This final filtering was conducted by manually checking each individual putative polymorphism on the Y chromosome using IGV \citep{robinson2011integrative}.
 
@@
 We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
-Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes.
-To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio using coalescent theory (Figure S1).
-Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this involves maximizing the likelihood estimate of the mutation rate for each locus independently.
+Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes.
+Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping.
+To address this, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate was not feasible, as this would require maximizing the likelihood estimate of the mutation rate for each locus independently.
 Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
-Our model therefore had two parameters; $\theta_{autosomal}$ and the Y/A ratio.
-We varied both parameters across a grid of points and evaluated the likelihood, where the Y/A ratio varied from 0.001 to 1, and $\theta_{autosomal}$ varied from 0.001 to 0.01.
+Our model thus had two parameters: $\theta_{autosomal}$ and Y/A.
+We varied both parameters and evaluated the likelihood for Y/A from 0.001 to 1, and $\theta_{autosomal}$ from 0.001 to 0.01.
+
+%from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
 %We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-
-To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that considers these two effects (crow and kimura).
-In particular, assuming a sex ratio of 0.6, we calculated the effect of varaince in reproductive success on male and female $N_{e}$, as:
-
-by Ne = ((N(k -1))/((k -1)+(Vk / k)), where k and Vk are the mean and variance in offspring number, respectively (Kimura and Crow 1964).
-Using sex-specific values of k and Vk to calculate male and female Ne (Nm and Nf , respectively)
-
-respectively, where Vk is the variance in the number of progeny (Kimura and Crow, 1963; Crow and Denniston, 1988).
-Assuming a constant population size, the Ne for autosomes and sex chromosomes then be calculated using the expression from \citep{wright1931evolution}:
+To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that considers these two effects \citep{kimura1964number}.
+In particular, we calculated the effect of varaince in reproductive success on $N_{e}$ as:
 
 \begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:Ne}
+N_{e}_{i} = \frac{N_{i}k_{i} - 1}{k_{i} - 1 + \frac{V_{k{i}}}{k}} \label{eq:Vk}
+\end{equation}
+
+where $k$ and $V_{k{i}}$ are the mean and variance in offspring number, respectively, for the $i$th sex (Kimura and Crow 1964).
+Using sex-specific values of $k$ and $V_{k{i}}$ to calculate $N_{m}$ and $N_{f}$, the expected $N_{e}$ for autosomes and sex chromosomes is then obtained using equations from \citep{wright1931evolution}:
+
+\begin{equation}
+N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}} \label{eq:Ne}
 \end{equation}
@@
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}},\label{eq:NeX}
+N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}} \label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e{Y}} = \frac{N_{m}}{2},\label{eq:NeY}
+N_{e{Y}} = \frac{N_{m}}{2} \label{eq:NeY}
 \end{equation}
 
-and the X/A and Y/A ratios as:
-
-\begin{equation}
-\frac{N_{e{X}}}{N_{e{A}} } = \frac{9(N_{m} + N_{f}}{8(2N_{m} + N_{f})} \label{eq:NeXNeA}
-\end{equation}
-
-and
-
-respectively.
-Thus, with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-However, as $Ne_{m}$ is reduced relative to $Ne_{f}$, the X/A ratio increases amd may become greater than 1, approaching 1.125 at the limit \citep{caballero1995}.
-With the level of neutral polymorphism at equilibrium given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1C shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio across a range of values for the variance in male reproductive success.
+Note that with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
+Figure 1 shows the expected X/A and Y/A ratios as a function of the reproductive sex ratio across a range of values for the variance in male reproductive success, and we thus compared our empirical diversity ratios to these predictions (Figure 2) using a sex ratio of 0.6 as estimated in \textit{R.
+hastatulus}
 
@@
 
- Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
+ %Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
 hastatulus} sex ratio (N_{f}/$N_{m}+N_{f})=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}.

### 2016-11-08 manuscript.tex
minor edits, added figure2

@@
 Here, we investigate the factors affecting X- and Y-linked neutral polymorphsm during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
-Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corrsponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
+Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
@@
 We therefore don't yet know the time-scales over which processes such as linked selection on the Y chromosome are likely to be important.
-In humans, for example, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, and simulations have revealed that levels of Y-diversity are consistent with a combination of very low $N_{m}$ and strong purifying selection eliminiting linked neutral variation \citep{Wilsonsayres2014}.
+In humans, for example, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, and simulations have revealed that levels of Y-diversity are consistent with a combination of very low $N_{m}$ and strong purifying selection eliminating linked neutral variation \citep{Wilsonsayres2014}.
 However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is an open question whether such strong effects of purifying selection should also apply to systems where sex chromosomes evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA as in the case of diecious plants \citep{charlesworth2015plant}).
@@
 To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that considers these two effects \citep{kimura1964number}.
-In particular, we calculated the effect of varaince in reproductive success on $N_{e}$ as:
+In particular, we calculated the effect of variance in reproductive success on $N_{e}$ as:
 
 \begin{equation}
-N_{e}_{i} = \frac{N_{i}k_{i} - 1}{k_{i} - 1 + \frac{V_{k{i}}}{k}} \label{eq:Vk}
+N_{e_{i}}=\frac{N_{i}k_{i}-1}{k_{i}-1+\frac{V_{k_{i}}}{\overline{k}}} \label{eq:Vk}
 \end{equation}
 
-where $k$ and $V_{k{i}}$ are the mean and variance in offspring number, respectively, for the $i$th sex (Kimura and Crow 1964).
+where $k_{i}$ and $V_{k_{i}}$ are the mean and variance in offspring number, respectively, for the $i$th sex and $\overline{k}$ is the total mean number of progeny (Kimura and Crow 1964).
 Using sex-specific values of $k$ and $V_{k{i}}$ to calculate $N_{m}$ and $N_{f}$, the expected $N_{e}$ for autosomes and sex chromosomes is then obtained using equations from \citep{wright1931evolution}:
@@
 Note that with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-Figure 1 shows the expected X/A and Y/A ratios as a function of the reproductive sex ratio across a range of values for the variance in male reproductive success, and we thus compared our empirical diversity ratios to these predictions (Figure 2) using a sex ratio of 0.6 as estimated in \textit{R.
+Figure 1 shows the expected X/A and Y/A ratios as a function of the reproductive sex ratio across a range of values for the variance in male reproductive success.
+We therefore compared our empirical diversity ratios to these predictions (Figure 2) using a sex ratio of 0.6 as estimated in \textit{R.
 hastatulus}
@@
 \subsection*{Simulations of purifying selection}
-To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a nonrecombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
+To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a non-recombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
 This approach was implemented rather than using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
@@
 
-\begin{table}[t!]
-\centering
-\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
-\begin{tabular}{ccccccccc}
-\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
-chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
-\midrule
-A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
-X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
-Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
-\addlinespace
+%\begin{table}[htb]
+%\centering
+%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
+%\begin{tabular}{ccccccccc}
+%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
+%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
+%\midrule
+%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
+%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
+%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
+%\addlinespace
 
-\bottomrule
-\end{tabular}
-\end{table}
+%\bottomrule
+%\end{tabular}
+%\end{table}
 
@@
 The occurrence of female-biased sex ratios in \textit{R.
-hastatulus} has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the neutrally expected $N_{e}$ of the Y chromosome (\X ref).
-This reduction $N_{e}$ on the Y is expected to be further accenuated with high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
+hastatulus} is expected to lower Y diversity through lowering male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome.
+In addition, the $N_{e}_{Y}$ might be further reduced due to high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
 hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Moreover, given that male plants in this species produce large amounts of pollen, and that female flowers are uniovulate, we expected strong competition among males to fertilize females.
+Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there is reason to believe that there could be strong competition among males.
+
 %Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
+
+\begin{figure}[htbp]
+\centering
+\noindent
+\includegraphics[width=\linewidth]{figure2.png}
+\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).
+}
+\label{fig:spectrum}
+\end{figure}
 

### 2016-11-09 temp/fwdpy_sims/fwdpp.Rmd
major reorganization

---
output:
  pdf_document: default
  html_notebook: default
---

###Can background selection explain low Y diversity in _R. hastatulus_?

We have done similations of background selection on the Y chromosome to test whether our observed level of Y/A diversity, $\theta_Y$/$\theta_A = 0.02$, is consistent with the effects of background selection on this chromosome.

#### Given our simulation parameters, what does the theory predict?

Here are the input parameters:

* $Ne_Y = 1/4 Ne_A$, where the inital $Ne_A = 1000$
* $\mu = 0.005$ (neutral mutation rate, pre gamete, per generation)
* $U = 0.02$ (deleterious mutation rate, per _diploid_; so, twice the per gamete rate)
* $r = 0$ (recombination rate)
* $h = 1$ (dominance)
* $s =-0.01$ (mean s; sampled from a gamma dist)

Under the BGS model, the expected $\pi$ is $E[\pi]=\pi_0e^{-\frac{U}{2sh+r}},$ $U$ is the mutation rate to strongly-deleterious alleles, $\pi_0$ is the value expected in the absence of BGS (_i.e._ $\pi_0 = \theta = 4N_e\mu$), $s$ and $h$ are the selection and dominance coefficients, and $r$ is the recombination rate.

For our parameters, we have:
$\pi_0 = \theta = 4N_e\mu = 5$, so

$$E[\pi_{Y}] = 5e^{-\frac{0.02}{0.1}} = 1.839397$$

For the autosomes, with $r = 0.5$, we have

$$E[\pi_{A}] = 20e^{-\frac{0.02}{0.1+0.5}} = 19.24537$$

and for the X chromosome

$$E[\pi_{X}] = 15e^{-\frac{0.02}{0.1+0.5}} = 14.43403$$

#### Expected Y/A and X/A diversity

OK, so then the expected ratios of diversity under background selection are:

$$\frac{E[\pi]_Y}{E[\pi]_A} = \frac{1.839397}{19.24537} = 0.095$$

and

$$\frac{E[\pi]_X}{E[\pi]_A} = \frac{14.43403}{19.24537} = 0.75$$

#### Simulated Y/A and X/A diversity

```{r}
#autosomes
a <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/sims/a_fixed.csv", header = T)
#X-chromosome
x <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/sims/x_fixed.csv", header = T)
#Y-chromsoome
y <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/sims/y_fixed.csv", header = T)
```

Simple box plot of simulated $\pi_Y$, $\pi_X$, and $\pi_A$

```{r}
boxplot(a$thetapi, x$thetapi, y$thetapi, names = c("A","X","Y"),
        ylab = expression(paste("Nucleotide Diversity (", pi, " ) ")))
```

The following $\pi$ values are the means from all of the simulations, for each chromosome type. You can see for the autosomes and the X chromosome, they are quite close to the theoretically expectd  values calculated above.

$$\pi_{A_{sim}} = 19.5$$
$$\pi_{X_{sim}} = 14.9$$
$$\pi_{Y_{sim}} = 4.2$$

However, the estimate for the Y chromosome is much higher than predicted the by model. Moreover, ratios for the simulated data:

$$ \large X/A = \frac{\pi_{X_{sim}}}{\pi_{A_{sim}}} = 0.76$$

and

$$ \large Y/A = \frac{\pi_{Y_{sim}}}{\pi_{A_{sim}}} = 0.21$$

show the same effect: the X/A diversity ratio is not very different from our expectations, but the Y is much higher than expected.

---
Here is the code I used in fwdpy. And I did 1000 replicates for the 24 populations.

`nregions = [fp.Region(beg=0,end=1,weight=1)]`

`sregions = [fp.GammaS(1,2,1,-0.01,0.1,1)]`

`recregions = [fp.Region(beg=-1,end=2,weight=1)]`

`N=250`

`pops = fp.evolve_regions(rng,24,N,nlist[0:],0.005, 0.01,0,nregions,sregions,recregions)`

### 2016-11-09 temp/fwdpy_sims/pwdpp2.Rmd
major reorganization

---
output:
  html_notebook: default
  pdf_document: default
---

###Can background selection explain low Y diversity in _R. hastatulus_?

We have done similations of background selection on the Y chromosome to test whether our observed level of Y/A diversity, $\theta_Y$/$\theta_A$, is consistent with the effects of background selection on this chromosome.

####Let's ask: given our simulation parameters, what does the theory predict?

Here are the input parameters:

* $Ne_Y = 1/4 Ne_A$, where the inital $Ne_A = 1000$
* $\mu = 0.005$ (neutral mutation rate, pre gamete, per generation)
* $U = 0.02$ (deleterious mutation rate, per _diploid_; so, twice the per gamete rate)
* $r = 0$ (recombination rate)
* $h = 1$ (dominance)

Under the BGS model, the expected $\pi$ is $E[\pi]=\pi_0e^{-\frac{U}{2sh+r}},$ $U$ is the mutation rate to strongly-deleterious alleles, $\pi_0$ is the value expected in the absence of BGS (_i.e._ $\pi_0 = \theta = 4N_e\mu$), $s$ and $h$ are the selection and dominance coefficients, and $r$ is the recombination rate.

For our parameters, we have:
$\pi_0 = \theta = 4N_e\mu = 5$, so

$$E[\pi] = 5e^{-\frac{0.02}{0.1}} = 1.839397$$

```{r}
5*(exp(-(0.02/((2*0.01)))))
```

Analagously for the autosomes, with $r = 0.5$, we have

$$\pi_0 = \theta = 4N_e\mu = 20$$

and

$$E[\pi] = 20e^{-\frac{0.02}{0.1+0.5}} = 19.24537$$

```{r}
20*(exp(-(0.02/((2*0.01)+0.5))))
```

and for the X chromosome

$$\pi_0 = \theta = 4N_e\mu = 15$$

and

$$E[\pi] = 15e^{-\frac{0.02}{0.1+0.5}} = 14.43403$$

```{r}
15*(exp(-(0.02/((2*0.01)+0.5))))
```

#### Expected Y/A and X/A diversity

OK, so then the expected ratios of diversity under background selection are:

$$\frac{E[\pi]_Y}{E[\pi]_A} = \frac{1.839397}{19.24537} = 0.095$$

and

$$\frac{E[\pi]_X}{E[\pi]_A} = \frac{14.43403}{19.24537} = 0.75$$

#### Simulated Y/A and X/A diversity

How does this compare with the simulation output?

import the simulation data

```{r}
#autosomes
a <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/a_fixed.csv", header = T)
#X-chromosome
x <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/x_fixed.csv", header = T)
#Y-chromsoome
y <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/y_fixed.csv", header = T)
```

Let's see some plots of the simulated $\pi_Y$, $\pi_X$, and $\pi_X$

```{r}
boxplot(a$thetapi, x$thetapi, y$thetapi, names = c("A","X","Y"),
        ylab = expression(paste("Nucleotide Diversity (", pi, " ) ")))
```

The following $\pi$ values are the means from all of the simulations, for each chromosome type. You can see for the autosomes and the X chromosome, they are quite close to the theoretically expectd  values calculated above.

$$\pi_{A_{sim}} = 19.5$$
$$\pi_{X_{sim}} = 14.9$$
$$\pi_{Y_{sim}} = 4.2$$

```{r}
mean(a$thetapi)
mean(x$thetapi)
mean(y$thetapi)
```
However, the estimate for the Y chromosome is much higher than predicted the by model.

Moreover, ratios for the simulated data:

$$ \large X/A = \frac{\pi_{X_{sim}}}{\pi_{A_{sim}}} = 0.76$$

and

$$ \large Y/A = \frac{\pi_{Y_{sim}}}{\pi_{A_{sim}}} = 0.21$$

are not different from neutral expectations

```{r}
mean(x$thetapi)/mean(a$thetapi)
mean(y$thetapi)/mean(a$thetapi)
```

This is basically unchanged (well, even higher for the Y) if one weights these means by the number of polymorphic sites (called "npoly" in the data)

---
_Stephen: I don't know why this is? It might be some issue with the simulation somehow still treating sites on the Y as independent (this is my suspicion), but I have been investigating all day, and not quite sure anymore..._

### 2016-11-09 Rnotebook/Figure1.R

@@ -1,97 +1,0 @@
-Nf=100 #number of females
-Nm=100 #number of males
-N=Nf+Nm #total number of individuals
-#x=Nf/(Nf+Nm) #proportion of males
-
-op <- par(mfrow = c(1, 3))
-m=0
-f=0
-
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m))))/N, from=0, to=1, lty=2,
-      xlab=expression(italic('N'[f]/'N'[m]+'N'[f])),
-      ylab=expression(italic('N'[e]/N)), 
-      main="Autosomes", ylim=c(0,1), xlim=c(0,1))
-
-mh=0.5
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+mh))))/N, from=0, to=1, add=T)
-
-m1=1
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T, col="blue")
-
-m2=2
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m2))))/N, from=0, to=1, add=T)
-
-m3=3
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m3))))/N, from=0, to=1, add=T)
-
-#for sex-linked genes
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m))))/N, from=0, to=1, lty=2,
-      xlab=expression(italic('N'[f]/'N'[f]+'N'[m])), 
-      ylab=expression(italic('N'[e]/N)), 
-      main="X chromosomes", ylim=c(0,1),xlim=c(0,1))
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+mh))))/N, from=0, to=1, add=T)
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m1))))/N, from=0, to=1, add=T, col="blue")
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m2))))/N, from=0, to=1, add=T)
-
[…]

### 2016-11-09 SFS_code/sfs_code.sh

@@ -1,57 +1,0 @@
-#!/bin/bash -l
-#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
-#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code-stdout_%j.txt
-#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code_stderr_%j.txt
-#SBATCH -J sfs
-#SBATCH --cpus-per-task=10
-
-set -e
-set -u
-set -o pipefail
-
-# Npops:        1 #number of pops
-# Niter:        50 #iterations
-# theta(-t):    0.001375 #=1/4 autosomal theta: 0.0055/4
-# rho(-r):      0
-# ploidy(-P):   1
-# tau(-TE):     0.3 #generations scaled in units of P*N (this seems too small?)
-# loci(-L):     2 <500000> <800000> #2 regions with lengths <L1><L2>...
-# annotate(-a): N #noncoding
-# fitness(-W):  L 1 2 0 1 1 0.258 0.005 # Gamma T model. probability positive s = 0, gamma shape = 0.258 from the DFE for x-linked genes. For a selection coefficient = 0.01, the scale parameter is 0.258/(0.01 * 5000)= 0.005.
-# sampsize(-n): 8 #chromosomes per sample
-# popsize:(-N)  5000
-# -A tells simulation to end
-
-./sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
--a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A
-# set paramters as bash variables
-aN=0.2579 #shape parameter from X-linked DFE
-s=
-lN=${aN}/${s}*5000 # scale parameter. Hold aN constant, test a range for s. 5000=5kb
-
-#############################################################################
-echo "Starting Job: Non-coding regions"
-date
-
-cmd="./sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
--a N -W L 1 2 0 1 1 0.258 0.005 -n 8 -N 5000 -A \
--o SFS.out --errfile SFS.err"
-
-echo $cmd
-eval $cmd
-echo "Ending Job: Non-coding regions "
-date
-
-# fitness model in SFS_code
[…]

### 2016-11-09 code/R/Figure1.R

@@ -1,0 +1,97 @@
+Nf=100 #number of females
+Nm=100 #number of males
+N=Nf+Nm #total number of individuals
+#x=Nf/(Nf+Nm) #proportion of males
+
+op <- par(mfrow = c(1, 3))
+m=0
+f=0
+
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[f]/'N'[m]+'N'[f])),
+      ylab=expression(italic('N'[e]/N)), 
+      main="Autosomes", ylim=c(0,1), xlim=c(0,1))
+
+mh=0.5
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+mh))))/N, from=0, to=1, add=T)
+
+m1=1
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T, col="blue")
+
+m2=2
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m2))))/N, from=0, to=1, add=T)
+
+m3=3
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m3))))/N, from=0, to=1, add=T)
+
+#for sex-linked genes
+curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m))))/N, from=0, to=1, lty=2,
+      xlab=expression(italic('N'[f]/'N'[f]+'N'[m])), 
+      ylab=expression(italic('N'[e]/N)), 
+      main="X chromosomes", ylim=c(0,1),xlim=c(0,1))
+
+curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+mh))))/N, from=0, to=1, add=T)
+
+curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m1))))/N, from=0, to=1, add=T, col="blue")
+
+curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m2))))/N, from=0, to=1, add=T)
+
[…]

### 2016-11-09 code/R/bgs.R

@@ -1,0 +1,27 @@
+print(20*math.exp(-0.02/(0.1+0.005)))
+
+curve(0.2083333*(exp(-(0.02/((2*x)+0)))), from=0, to=1,
+      main="Expected Y/A diversity under BGS",
+      xlab="selection coefficient (s)",
+      ylab = expression(paste("E[", pi,"]")), las=1,
+      ylim=c(0,0.3),
+      xlim=c(0,0.2),lwd=3)
+abline(h=0.2,lty=2)
+abline(h=0.02,lty=1, col="red",lwd=3)
+legend(x=0.12,y=0.3, cex=0.8, box.lty=0, c("Nf/(Nf+Nm) = 0.6","U = 0.02","h = 1","r = 0", expression(paste("", pi[0], paste(" = ",N[em]/2)))))
+
+text(0.03,0.029, cex=1, expression(paste(, theta,)))
+text(0.03,0.21,expression(paste(, pi[0],)))
+
+0.2083333*(exp(-0.02/((2*0.02)+0)))
+
+
+
+((9*N*0.6*(0.6-1)/(2*(0.6-2))))/(4*N*(1-0.6)*0.6)
+
+
+((N*(1-0.6)/2)/(4*N*(1-0.6)*0.6)
+)
+
+(0.2083333*(exp(-(0.02/((2*0.1)+0))))
+)

### 2016-11-09 code/SFS_code/convert_SFS_CODE.sh

@@ -1,0 +1,20 @@
+#!/bin/bash -l
+#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
+#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/convert-stdout_%j.txt
+#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/convert_stderr_%j.txt
+#SBATCH -J convert_SFS
+#SBATCH --cpus-per-task=5
+
+set -e
+set -u
+set -o pipefail
+
+#see https://github.com/molpopgen/msstats
+#command structure: convertSFS CODE <input file> <option [args]> | msstats/src/msstats >out
+# used: - for ms style output; T 0 says to output only sysnonymous mutations
+
+convertSFS_CODE SFS.out -m T 0 | ./msstats/src/msstats > stats.txt
+
+# Then, using R,
+# x=read.table("stats",header=T)
+# mean(x$hsm01[x$pop==0])

### 2016-11-09 code/SFS_code/newsfscode.sh

@@ -1,0 +1,42 @@
+#Purifying Selection
+
+./sfs_code 2 1000 -N 740 -t 0.008 -r 0.0008 -n %d-W %s TE 0.40
+
+#Where,%d and %s are substitutions for the sample size and selection coefficients.
+
+#a)Sample size%d = (5, 50, 500, 5000)
+
+#b)Selection coefficients
+  i)Neutral mutation %s:‘0’
+  ii)Purifying selection
+    (%s):
+    -|Υ|:"L 0 2 0 0.0 1.0 0.206 0.0004931"
+    -|Υ|*0.1:"L 0 2 0 0.0 1.0 0.206 0.004931"
+    -|Υ|*0.025:"L 0 2 0 0.0 1.0 0.206 0.0197240"
+    -|Υ|*0.01:"L 0 2 0 0.0 1.0 0.206 0.04931"3
+    -|Υ|*0.005:"L 0 2 0 0.0 1.0 0.206 0.0986200"
+    -|Υ|*0.0025:"L 0 2 0 0.0 1.0 0.206 0.1972400"
+    -|Υ|*0.001:"L 0 2 0 0.0 1.0 0.206 0.4931"
+    -|Υ|*0.0001:"L 0 2 0 0.0 1.0 0.206 4.931"
+
+iii)Positive selection (%s):
+|Υ|:"L 0 2 1.0 0.206 0.0004931 0.206 0.0004931"
+|Υ|*0.1:"L 0 2 1.0 0.206 0.004931 0.206 0.004931"
+|Υ|*0.025:"L 0 2 1.0 0.206 0.0197240 0.206 0.0197240"
+|Υ|*0.01:"L 0 2 1.0 0.206 0.0493100 0.206 0.0493100"
+|Υ|*0.005:"L 0 2 1.0 0.206 0.0986200 0.206 0.0986200"
+|Υ|*0.0025:"L 0 2 1.0 0.206 0.1972400 0.206 0.1972400"
+|Υ|*0.001:"L 0 2 1.0 0.206 0.4931 0.206 0.4931"
+|Υ|*0.0001:"L 0 2 1.0 0.206 4.931 0.206 4.931"
+
+iv)Mixture of negative selection and positive selection (%s)
+-|Υ|*0.1 and|Υ|*0.1: "L 0 1 41.77652 %n %p"
+where, %n = (0.10, 0.20, ..., 1.00) istheproportion of sites underpurifying selection,
+%p= (0.0, 0.02,..., 0.10)istheproportion of sites under positive selection,
[…]

### 2016-11-09 code/SFS_code/sfs_code.sh

@@ -1,0 +1,57 @@
+#!/bin/bash -l
+#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
+#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code-stdout_%j.txt
+#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/sfs_code_stderr_%j.txt
+#SBATCH -J sfs
+#SBATCH --cpus-per-task=10
+
+set -e
+set -u
+set -o pipefail
+
+# Npops:        1 #number of pops
+# Niter:        50 #iterations
+# theta(-t):    0.001375 #=1/4 autosomal theta: 0.0055/4
+# rho(-r):      0
+# ploidy(-P):   1
+# tau(-TE):     0.3 #generations scaled in units of P*N (this seems too small?)
+# loci(-L):     2 <500000> <800000> #2 regions with lengths <L1><L2>...
+# annotate(-a): N #noncoding
+# fitness(-W):  L 1 2 0 1 1 0.258 0.005 # Gamma T model. probability positive s = 0, gamma shape = 0.258 from the DFE for x-linked genes. For a selection coefficient = 0.01, the scale parameter is 0.258/(0.01 * 5000)= 0.005.
+# sampsize(-n): 8 #chromosomes per sample
+# popsize:(-N)  5000
+# -A tells simulation to end
+
+./sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.184 0.003 -n 8 -N 500 -A
+# set paramters as bash variables
+aN=0.2579 #shape parameter from X-linked DFE
+s=
+lN=${aN}/${s}*5000 # scale parameter. Hold aN constant, test a range for s. 5000=5kb
+
+#############################################################################
+echo "Starting Job: Non-coding regions"
+date
+
+cmd="./sfs_code 1 50 -r 0 -t 0.001375 -P 1 -TE 0.3 -L 2 500000 800000 \
+-a N -W L 1 2 0 1 1 0.258 0.005 -n 8 -N 5000 -A \
+-o SFS.out --errfile SFS.err"
+
+echo $cmd
+eval $cmd
+echo "Ending Job: Non-coding regions "
+date
+
+# fitness model in SFS_code
[…]

### 2016-11-09 temp/newsfscode.sh

@@ -1,42 +1,0 @@
-#Purifying Selection
-
-./sfs_code 2 1000 -N 740 -t 0.008 -r 0.0008 -n %d-W %s TE 0.40
-
-#Where,%d and %s are substitutions for the sample size and selection coefficients.
-
-#a)Sample size%d = (5, 50, 500, 5000)
-
-#b)Selection coefficients
-  i)Neutral mutation %s:‘0’
-  ii)Purifying selection
-    (%s):
-    -|Υ|:"L 0 2 0 0.0 1.0 0.206 0.0004931"
-    -|Υ|*0.1:"L 0 2 0 0.0 1.0 0.206 0.004931"
-    -|Υ|*0.025:"L 0 2 0 0.0 1.0 0.206 0.0197240"
-    -|Υ|*0.01:"L 0 2 0 0.0 1.0 0.206 0.04931"3
-    -|Υ|*0.005:"L 0 2 0 0.0 1.0 0.206 0.0986200"
-    -|Υ|*0.0025:"L 0 2 0 0.0 1.0 0.206 0.1972400"
-    -|Υ|*0.001:"L 0 2 0 0.0 1.0 0.206 0.4931"
-    -|Υ|*0.0001:"L 0 2 0 0.0 1.0 0.206 4.931"
-
-iii)Positive selection (%s):
-|Υ|:"L 0 2 1.0 0.206 0.0004931 0.206 0.0004931"
-|Υ|*0.1:"L 0 2 1.0 0.206 0.004931 0.206 0.004931"
-|Υ|*0.025:"L 0 2 1.0 0.206 0.0197240 0.206 0.0197240"
-|Υ|*0.01:"L 0 2 1.0 0.206 0.0493100 0.206 0.0493100"
-|Υ|*0.005:"L 0 2 1.0 0.206 0.0986200 0.206 0.0986200"
-|Υ|*0.0025:"L 0 2 1.0 0.206 0.1972400 0.206 0.1972400"
-|Υ|*0.001:"L 0 2 1.0 0.206 0.4931 0.206 0.4931"
-|Υ|*0.0001:"L 0 2 1.0 0.206 4.931 0.206 4.931"
-
-iv)Mixture of negative selection and positive selection (%s)
--|Υ|*0.1 and|Υ|*0.1: "L 0 1 41.77652 %n %p"
-where, %n = (0.10, 0.20, ..., 1.00) istheproportion of sites underpurifying selection,
-%p= (0.0, 0.02,..., 0.10)istheproportion of sites under positive selection,
[…]

### 2016-11-09 temp/workspace/Nucleotide_diversity_predictions.r

@@ -1,0 +1,71 @@
+Nf=5000 #number of females
+Nm=5000 #number of males
+N=Nf+Nm #total number of individuals
+x=Nf/(Nf+Nm) #proportion of females
+
+#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
+curve(4*N*(1-x)*x, from=0,to=1, #autosomes
+      ylab="Effective population size (Ne)", 
+      xlab= "Nf/(Nf+Nm)", 
+      ylim=c(0,11000))
+curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
+curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
+text(0.5,2800, labels="Y")
+text(0.5,7800, labels="X")
+text(0.5,10300, labels="A")
+abline(v=0.6,lty=2)
+
+#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
+curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
+      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
+      xlab="Nf/(Nf+Nm)",
+      ylim=c(0,2))
+curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
+text(0.8,1.05, labels="X/A")
+text(0.8,0.25, labels="Y/A")
+#abline(v=0.6,lty=2)
+points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
+points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+
+
+#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
+#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
+points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+
+
[…]

### 2016-11-09 temp/workspace/autosomal.R

@@ -1,0 +1,25 @@
+#autosomal diversity
+
+tm_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/tm_all.csv")
+tf_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/tf_all.csv")
+nf_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/nf_all.csv")
+nm_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/nm_all.csv")
+
+#new columns=SynSitestot*theta_syntot (for each cell)
+#calculate average weighted theta
+
+#tf_all
+tf_all$weighted_theta <- tf_all$SynSitestot * tf_all$theta_syntot            
+tf_all_ave_weighted_theta=(sum(tf_all$weighted_theta))/(sum(tf_all$SynSitestot))
+#tm_all
+tm_all$weighted_theta <- tm_all$SynSitestot * tm_all$theta_syntot            
+tm_all_ave_weighted_theta=(sum(tm_all$weighted_theta))/(sum(tm_all$SynSitestot))
+#nf_all
+nf_all$weighted_theta <- nf_all$SynSitestot * nf_all$theta_syntot            
+nf_all_ave_weighted_theta=(sum(nf_all$weighted_theta))/(sum(nf_all$SynSitestot))
+#nm_all
+nm_all$weighted_theta <- nm_all$SynSitestot * nm_all$theta_syntot            
+nm_all_ave_weighted_theta=(sum(nm_all$weighted_theta))/(sum(nm_all$SynSitestot))
+
+#overall average weighted theta for autosomes:
+mean_all=mean(c(tf_all_ave_weighted_theta, tm_all_ave_weighted_theta, nf_all_ave_weighted_theta, nm_all_ave_weighted_theta))

### 2016-11-09 temp/workspace/bootingtheta.r

@@ -1,0 +1,47 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+#x-linked genes
+txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
+
+#data for x linked genes
+a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+abline(v=txx_ave_weighted_theta, col="red",lwd=3)
+abline(v=0.0041,col="black", lwd=3)
+abline(v=0.0053,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+qqnorm(z$t)  # normal quantiles of replicate estimates
+boot.ci(z, type = "bca")  # 95% confidence interval using BCa
+boot.ci(z, type = "perc") # same using percentile method
+
+#implied range for X/A
+lowerXA=0.0041/tf_all_ave_weighted_theta
+upperXA=0.0053/tf_all_ave_weighted_theta
+
+
+#autosomal genes
+#tf_all$weighted_theta <- tf_all$SynSitestot * tf_all$theta_syntot            
+#tf_all_ave_weighted_theta=(sum(tf_all$weighted_theta))/(sum(tf_all$SynSitestot))
+#tx_x_auto=txx_ave_weighted_theta/tf_all_ave_weighted_theta
+
+

### 2016-11-09 temp/workspace/bootingthetaY.r

@@ -1,0 +1,36 @@
+#x-linked genes
+txy$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+txy_ave_weighted_theta=(sum(txy$weighted_theta))/(sum(txy$SynSitestot))
+
+#data for y linked genes
+p=fly$SynSitestot #stores the number of synonymous sites for y-linked genes in vector
+q=fly$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d2 <- cbind.data.frame(p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function2 <- function(d2,i){
+  p <- d2$p[i]
+  q <- d2$q[i]
+  boot.function2 <- sum(p*q)/sum(p) # calculates average weighted theta from d
+}
+
+z2 <- boot(d2, boot.function2, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z2)    # bootstrap calculation of bias and SE
+sd(z2$t)     # another way to get the standard error
+
+#hist(z2$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped Y genes"
+#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+#abline(v=txy_ave_weighted_theta, col="red",lwd=3)
+#abline(v=0.0001,col="black", lwd=3)
+#abline(v=0.0002,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+#qqnorm(z2$t)  # normal quantiles of replicate estimates
+boot.ci(z2, type = "bca")  # 95% confidence interval using BCa
+#boot.ci(z2, type = "perc") # same using percentile method
+
+#implied range for X/A
+#lowerYA=0.0001/tf_all_ave_weighted_theta; lowerYA
+#upperYA=0.0002/tf_all_ave_weighted_theta; upperYA
+

### 2016-11-09 temp/workspace/bootingtheta_FL.R

@@ -1,0 +1,39 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+#x-linked genes
+flx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+flx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
+
+#data for x linked genes
+a=flx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=flx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+#hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+#abline(v=txx_ave_weighted_theta, col="red",lwd=3)
+#abline(v=0.0041,col="black", lwd=3)
+#abline(v=0.0053,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+#qqnorm(z$t)  # normal quantiles of replicate estimates
+boot.ci(z, type = "bca")  # 95% confidence interval using BCa
+
+#implied range for X/A
+lowerXA_FL=0.0014/tf_all_ave_weighted_theta
+upperXA_FL=0.0021/tf_all_ave_weighted_theta
+

### 2016-11-09 temp/workspace/bootingtheta_SC.R

@@ -1,0 +1,37 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+#x-linked genes
+
+#data for x linked genes
+a=scx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=scx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+#hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+#abline(v=txx_ave_weighted_theta, col="red",lwd=3)
+#abline(v=0.0041,col="black", lwd=3)
+#abline(v=0.0053,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+#qqnorm(z$t)  # normal quantiles of replicate estimates
+
+#implied range for X/A
+lowerSC_XA=0.0014/tf_all_ave_weighted_theta
+upperSC_XA=0.0025/tf_all_ave_weighted_theta
+
+

### 2016-11-09 temp/workspace/bootingtheta_TX.R

@@ -1,0 +1,40 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+#x-linked genes
+txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
+
+#data for x linked genes
+a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+abline(v=txx_ave_weighted_theta, col="red",lwd=3)
+abline(v=0.0041,col="black", lwd=3)
+abline(v=0.0053,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+qqnorm(z$t)  # normal quantiles of replicate estimates
+boot.ci(z, type = "bca")  # 95% confidence interval using BCa
+boot.ci(z, type = "perc") # same using percentile method
+
+#implied CI for X/A
+lowerXA=0.0042/tf_all_ave_weighted_theta
+upperXA=0.0054/tf_all_ave_weighted_theta
+

### 2016-11-09 temp/workspace/bootingtheta_TX_auto copy copy.R

@@ -1,0 +1,52 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+#x-linked genes
+txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
+
+#data for x linked genes
+a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+#autosomal genes
+p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
+q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
+
+max.len = max(length(a), length(b), length(p), length(q))
+a = c(a, rep(NA, max.len - length(a)))
+b = c(b, rep(NA, max.len - length(b)))
+p = c(p, rep(NA, max.len - length(p)))
+q = c(q, rep(NA, max.len - length(q)))
+
+d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  p <- d$p[i]
+  q <- d$q[i]
+  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
+abline(v=txx_ave_weighted_theta, col="red",lwd=3)
+abline(v=0.0041,col="black", lwd=3)
+abline(v=0.0053,col="black", lwd=3)
+
+# histogram of boostrap replicate estimates
+qqnorm(z$t)  # normal quantiles of replicate estimates
+boot.ci(z, type = "bca")  # 95% confidence interval using BCa
[…]

### 2016-11-09 temp/workspace/bootingtheta_TX_auto copy.R

@@ -1,0 +1,61 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+
+ncx <- read.csv("ncx.csv")
+ncy <- read.csv("ncy.csv")
+txx <- read.csv("txx.csv")
+txy <- read.csv("txy.csv")
+scx <- read.csv("scx.csv")
+scy <- read.csv("scy.csv")
+flx <- read.csv("flx.csv")
+fly <- read.csv("fly.csv")
+scf_all <- read.csv("scf_autosomes.csv")
+scm_all <- read.csv("scm_autosomes.csv")
+flf_all <- read.csv("flf_autosomes.csv")
+flm_all <- read.csv("flm_autosomes.csv")
+
+#data for x linked genes
+a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+#autosomal genes
+p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
+q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
+
+max.len = max(length(a), length(b), length(p), length(q))
+a = c(a, rep(NA, max.len - length(a)))
+b = c(b, rep(NA, max.len - length(b)))
+p = c(p, rep(NA, max.len - length(p)))
+q = c(q, rep(NA, max.len - length(q)))
+
+d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  p <- d$p[i]
+  q <- d$q[i]
+  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z)    # bootstrap calculation of bias and SE
+sd(z$t)     # another way to get the standard error
+boot.ci(z, type="bca")
+
+hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
+     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
[…]

### 2016-11-09 temp/workspace/bootingtheta_TX_auto.R

@@ -1,0 +1,66 @@
+#bootstrapping average weighted theta from polymorphurama output
+library("boot")
+setwd("/Users/joshhough/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/")
+
+txx <- read.csv("txx.csv")
+txy <- read.csv("txy.csv")
+scx <- read.csv("scx.csv")
+scy <- read.csv("scy.csv")
+flx <- read.csv("flx.csv")
+fly <- read.csv("fly.csv")
+scf_all <- read.csv("scf_autosomes.csv")
+scm_all <- read.csv("scm_autosomes.csv")
+flf_all <- read.csv("flf_autosomes.csv")
+flm_all <- read.csv("flm_autosomes.csv")
+tm_all <- read.csv("tm_all_new.csv")
+tf_all <- read.csv("tf_all_new.csv")
+scf_all <- read.csv("scf_autosomes.csv")
+scm_all <- read.csv("scm_autosomes.csv")
+flf_all <- read.csv("flf_autosomes.csv")
+flm_all <- read.csv("flm_autosomes.csv")
+
+#data for x linked genes
+a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
+b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+#autosomal genes
+p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
+q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
+
+max.len = max(length(a), length(b), length(p), length(q))
+a = c(a, rep(NA, max.len - length(a)))
+b = c(b, rep(NA, max.len - length(b)))
+p = c(p, rep(NA, max.len - length(p)))
+q = c(q, rep(NA, max.len - length(q)))
+
+d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function <- function(d,i){
+  a <- d$a[i]
+  b <- d$b[i]
+  p <- d$p[i]
+  q <- d$q[i]
+  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
+}
+
+z <- boot(d, boot.function, R = 20000) #store 10000 bootstrap estimates of boot.function in z
+
[…]

### 2016-11-09 temp/workspace/expected_pies.R

@@ -1,0 +1,7 @@
+pi_X=4*750*0.005
+pi_Y=4*250*0.005
+pi_A=4*1000*0.005
+
+ex_Y=pi_Y*(exp(-(0.02/((2*0.01)+0))));ex_Y
+ex_X=pi_X*(exp(-(0.02/((2*0.01)+0.5))));ex_X
+ex_A=pi_A*(exp(-(0.02/((2*0.01)+0.5))));ex_A

### 2016-11-09 temp/workspace/fold.R

@@ -1,0 +1,7 @@
+aveX=mean(c(scx_ave_weighted_theta,txx_ave_weighted_theta,flx_ave_weighted_theta));aveX
+aveY=mean(c(scy_ave_weighted_theta,txy_ave_weighted_theta,fly_ave_weighted_theta));aveY
+aveX/aveY
+
+mean(c(tf_all_ave_weighted_theta,scf_all_ave_weighted_theta,flf_all_ave_weighted_theta))/aveY
+
+mean(c(tx_x_auto,sc_x_auto,fl_x_auto))/mean(c(tx_y_auto,sc_y_auto,fl_y_auto))

### 2016-11-09 temp/workspace/malerepvariance.R

@@ -1,0 +1,89 @@
+#the effect of male reproductive variance on sex chromosome to autosome ratios of Ne
+
+#(4*Nf*Nm)/(Nf+Nm) #autosome Ne 
+#(9*Nf*Nm)/(((2*Nf))+4*Nm) #X-chromosome Ne
+#Nm/2 #Y chromosome Ne
+#Nf=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of fmeale reproductive variance on female Ne: 
+#Nm=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
+
+#set numbers of males and females
+#Nft=100000 #Nf for Ntheoretical population of autosomes, equal males and females, no rep variance, poisson offspring
+#Nmt=100000 #Nm for theoretical population of autosomes, equal males and females, no rep variance, poisson offspring
+vf=2 #variance in female rep success. Set to zero. (change to 2 for poisson)
+kf=2 #mean number of offspring of females
+km=2 #mean number of offspring of males
+Nf=100000 #female Ne
+Nm=100000 #male Ne
+
+
+#effect of male rep variance on autosomal Ne
+#(4*Nf*Nm)/(4*Nf+Nm) #autosome Ne
+#((4*Nf*Nm)/((4*Nf)+Nm))/((4*Nf*Nm)/((4*Nf)+Nm)) #autosomes relative to autosomes 
+#replace numerator Ne's with rep variance equations for males and females
+#((4*Nf*Nm)/((4*Nf)+Nm))/((4*Nf*Nm)/((4*Nf)+Nm))
+
+#Nf=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of fmeale reproductive variance on female Ne: 
+#Nm=((Nm*(km-1))/((km-1)+(x/km))) #effect of male reproductive variance on male Ne
+
+curve(((4*((Nf*(kf-1))/((kf-1)+(vf/kf)))*((Nm*(km-1))/((km-1)+(x/km))))/((4*((Nf*(kf-1))/((kf-1)+(vf/kf))))+((Nm*(km-1))/((km-1)+(x/km)))))/((4*Nft*Nmt)/((4*Nft)+Nmt)),
[…]

### 2016-11-09 temp/workspace/mle.R

@@ -1,0 +1,19 @@
+YAlikihoods <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/YAlikihoods.csv")
+
+plot(YAlikihoods$YAratio,YAlikihoods$mle, pch=19,xaxt='n', xlab="Y/A Ne ratio", ylab="Likelihood")
+axis(3)
+abline(h=-2, lty=2, lwd=3)
+segments(x0=0.09775,y0=-0.9,x1=0.09775,y1=-1.1,lwd=2, col="red")
+segments(x0=0.009,y0=-0.9,x1=0.009,y1=-1.1,lwd=2, col="red")
+#segments(x0=0.009,y0=1,x1=0.009,y1=-2,lwd=1,lty=2, col="black")
+#segments(x0=0.09775,y0=1,x1=0.09775,y1=-2,lwd=1,lty=2, col="black")
+legend("topright", c("likelihood credibility interval", "LRT significance cut-off"),
+       cex=0.7, box.lty=0, lty=c(1,2),col=c("red","black"), lwd=c(2,2))
+segments(0.009, -1, 0.09775, -1,lwd=2,code = 3,col="red")
+
+#rect(0.009, -2, 0.09775, 1, density = NULL, angle = 45,
+#     col="grey", border = NULL, lty = par("lty"), lwd = par("lwd"))
+
+#cord.x=c(0.009,0.09775)
+#cord.y=c(-2,-2)
+#polygon(cord.x,cord.y, col=gray(0.8))

### 2016-11-09 temp/workspace/mlhka.R

@@ -1,0 +1,64 @@
+#making MLHKA infile.txt
+
+#rename polymorphurama output file from TX_X
+txx_mlhka=txx
+
+#take names from "locus" column in txx_mlhka file,
+#and remove everything after "T
+test <- as.data.frame(sub("T.*$","",txx_mlhka$locus))
+
+#replaced in excel
+write.xlsx(test, file="test.xlsx")
+write.xlsx(txx_mlhka, file="txx_mlhka.xlsx")
+#load fixed file
+txx_mlhka_fixed <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/txx_mlhka_fixed.csv")
+
+#now do same with dNdSFreeRatioNCshared
+test2 <- as.data.frame(sub("T.*$","",dNdSFreeRatioNCshared$LOC))
+write.xlsx(test2, file="test2.xlsx")
+dNdSFreeRatioNCshared_fixed <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/dNdSFreeRatioNCshared_fixed.csv")
+
+crap <- merge(txx_mlhka_fixed, dNdSFreeRatioNCshared_fixed)
+write.xlsx(crap, file="crap.xlsx")
+
+
+#ok i think file called "crap" is good: it is a merge of txx_mlhla_fixed and dNdSFreeRatioNCshared_fixed
+#make new column in crap for "D" =synsites tot* OUTS for D
+crap$D_mlhka <- crap$SynSitestot * crap$OUTS
+
+#fixing D column
+write.xlsx(crap, file="crap2.xlsx")
+
+#ok crap2 now has all relevant information. Copied this file as crap_infile in order to format for infile.txt for MLHKA
+crap_infile <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/crap_infile.csv")
+crap_infile$SynSitestot_rounded <- round(crap_infile$SynSitestot)
+
+write.xlsx(crap_infile, file="crap_infile_top.xlsx")
+
+
+#OK, things needed for MLHKA input, in order:
+#1. Locus ID. This is locus
+#2. L (size in BP). This is SynSitestot_rounded
[…]

### 2016-11-09 temp/workspace/new_malevariance.R

@@ -1,0 +1,118 @@
+#set numbers of males and females
+#vm=50 #variance in male rep success
+vf=2 #variance in female rep success
+kf=2 #mean number of offspring of females
+km=2 #mean number of offspring of males
+Nf=100000 #numner of females
+Nm=100000 #number of males
+
+Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
+Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
+
+Ne_ratio=Nef/Nem
+N=Nf+Nf
+
+A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
+X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
+Y=Nem/2
+XA=X/A
+YA=Y/A
+
+vm <- c(1:100000)
+for (val in vm) {
+}
+plot(((Nm*(km-1))/((km-1)+(vm/km)))/Nef, ylab="Nem/Nef",xlab="vm", type="l")
+
+
+vm <- c(1:50)
+for (val in vm) {
+Nf=100000 #numner of females
+Nm=100000 #number of males
+N=Nf+Nf
+vf=2 #variance in female rep success
+kf=2 #mean number of offspring of females
+km=2 #mean number of offspring of males
+Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
+Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
+Ne_ratio=Nef/Nem
+A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
+X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
+Y=Nem/2
+XA=X/A
+YA=Y/A
+}
+#plot(YA, ylab="Y/A",xlab="vm", type="l")
+#plot(XA, add=TRUE)
+d <- data.frame(vm,Nem,Nef,Ne_ratio,A,X,Y,XA,YA)
+require(ggplot2)
+ggplot(d, aes(vm)) +                    # basic graphical object
+  geom_line(aes(y=YA), colour="red") +  # first layer
+  geom_line(aes(y=XA), colour="green")  # second layer
+g <- g + ylab("Y") + xlab("X")
+g
+
+
+
[…]

### 2016-11-09 temp/workspace/nov24.R

@@ -1,0 +1,76 @@
+Nf=5000 #number of females
+Nm=5000 #number of males
+N=Nf+Nm #total number of individuals
+x=Nf/(Nf+Nm) #proportion of females
+
+#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
+curve(4*N*(1-x)*x, from=0,to=1, #autosomes
+      ylab="Effective population size (Ne)", 
+      xlab= "Nf/(Nf+Nm)", 
+      ylim=c(0,11000))
+curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
+curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
+text(0.5,2800, labels="Y")
+text(0.5,7800, labels="X")
+text(0.5,10300, labels="A")
+abline(v=0.6,lty=2)
+
+#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
+curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
+      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
+      xlab="Nf/(Nf+Nm)",
+      ylim=c(0,2))
+curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
+text(0.8,1.05, labels="X/A")
+text(0.8,0.25, labels="Y/A")
+#abline(v=0.6,lty=2)
+points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
+points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+
+
+#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
+#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
+points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+
+
[…]

### 2016-11-09 temp/workspace/old/polymorphism.R

@@ -1,0 +1,1 @@
+S_Syntot_NMX=

### 2016-11-09 temp/workspace/polymorphurama.R

@@ -1,0 +1,225 @@
+
+#read polymorphurama output (converted to .csv and parsed)
+setwd("/Users/joshhough/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/")
+ncx <- read.csv("ncx.csv")
+ncy <- read.csv("ncy.csv")
+txx <- read.csv("txx.csv")
+txy <- read.csv("txy.csv")
+scx <- read.csv("scx.csv")
+scy <- read.csv("scy.csv")
+flx <- read.csv("flx.csv")
+fly <- read.csv("fly.csv")
+
+#new columns=SynSitestot*theta_syntot (for each cell)
+#calculate average weighted theta
+#NC_X
+ncx$weighted_theta <- ncx$SynSitestot * ncx$theta_syntot            
+ncx_ave_weighted_theta=(sum(ncx$weighted_theta))/(sum(ncx$SynSitestot))
+#NC_Y
+ncy$weighted_theta <- ncy$SynSitestot * ncy$theta_syntot            
+ncy_ave_weighted_theta=(sum(ncy$weighted_theta))/(sum(ncy$SynSitestot))
+#TX_X
+txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
+txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
+#TX_Y
+txy$weighted_theta <- txy$SynSitestot * txy$theta_syntot            
+txy_ave_weighted_theta=(sum(txy$weighted_theta))/(sum(txy$SynSitestot))
+
+#SC_X
+scx$weighted_theta <- scx$SynSitestot * scx$theta_syntot            
+scx_ave_weighted_theta=(sum(scx$weighted_theta))/(sum(scx$SynSitestot))
+#SC_Y
+scy$weighted_theta <- scy$SynSitestot * scy$theta_syntot            
+scy_ave_weighted_theta=(sum(scy$weighted_theta))/(sum(scy$SynSitestot))
+#Fl_X
+flx$weighted_theta <- flx$SynSitestot * flx$theta_syntot            
+flx_ave_weighted_theta=(sum(flx$weighted_theta))/(sum(flx$SynSitestot))
+#Fl_Y
+fly$weighted_theta <- fly$SynSitestot * fly$theta_syntot            
[…]

### 2016-11-09 temp/workspace/predictions.R

@@ -1,0 +1,62 @@
+Nf=5000 #number of females
+Nm=5000 #number of males
+N=Nf+Nm #total number of individuals
+x=Nf/(Nf+Nm) #proportion of females
+
+#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
+curve(4*N*(1-x)*x, from=0,to=1, #autosomes
+      ylab="Effective population size (Ne)", 
+      xlab= "Nf/(Nf+Nm)", 
+      ylim=c(0,10000))
+curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
+curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
+text(0.5,2800, labels="Y")
+text(0.5,7800, labels="X")
+text(0.5,10300, labels="A")
+abline(v=0.6,lty=2)
+
+#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
+curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
+      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
+      xlab="Nf/(Nf+Nm)",
+      ylim=c(0,2))
+curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
+text(0.8,1.05, labels="X/A")
+text(0.8,0.25, labels="Y/A")
+#abline(v=0.6,lty=2)
+#points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
+#points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+#points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
+
+#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
+#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
+#points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+#points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+#points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
+
[…]

### 2016-11-09 temp/workspace/pvalue.R

@@ -1,0 +1,3 @@
+prop.test(4, 200, p=0.2,
+          alternative="two.sided",
+          conf.level = 0.95, correct=TRUE)

### 2016-11-09 temp/workspace/sim_output.R

@@ -1,0 +1,183 @@
+#new 
+
+#Ychromosome
+Y_npoly=16.49500
+Y_thetapi=4.22218
+Y_thetaw=4.41717
+
+#x-chromosome
+X_npoly=55.366667
+X_thetapi=14.915749
+X_thetaw=14.826552
+
+
+#the A
+A_npoly=73.005000
+A_thetapi=19.566661
+A_thetaw=19.549893
+
+
+
+#A=A_npoly*A_thetaw;A
+#X=X_npoly*X_thetaw;X
+#Y=Y_npoly*Y_thetaw;Y
+
+
+y$weighted_theta <- y$npoly * y$thetaw
+yw=(sum(y$weighted_theta))/(sum(y$thetaw));yw
+
+a$weighted_theta <- a$npoly * a$thetaw
+aw=(sum(a$weighted_theta))/(sum(a$thetaw));aw
+
+x$weighted_theta <- x$npoly * x$thetaw
+xw=(sum(x$weighted_theta))/(sum(x$thetaw));xw
+
+#xa=xw/aw;xa
+#ya=yw/aw;ya
+
+X
+
+yesy=mean(y$weighted_theta)/mean(a$weighted_theta);yesy
+
+
+
+weighted.mean(y$thetaw,y$npoly)/weighted.mean(a$thetaw,a$npoly)
+weighted.mean(a$thetaw,a$npoly)
+
+
+[1] 0.06062793
+yesx=mean(x$weighted_theta)/mean(a$weighted_theta);yesx
+[1] 0.5656107
+
+
+boxplot(y$weighted_theta,x$weighted_theta, a$weighted_theta, col="grey", notch=T, 
+        main="thetas from BGS simulations",
+        ylab="theta*synsites",
+        xlab="Y                  X                  A")
+
+barplot(c(tx_x_auto, sc_x_auto, fl_x_auto, yesx, xA_exp_equal, xA_exp_06, xA_exp_075, xA_exp_1,
+          tx_y_auto, sc_y_auto, fl_y_auto, yesy, yA_exp_equal, yA_exp_06, yA_exp_075, yA_exp_1),
+        ylim=c(0,1.2),
+    col=c("darkgreen","darkgreen","darkgreen","red","black", "gray22","gray73","gray87",
+          "darkgreen","darkgreen","darkgreen", "red","black", "gray22","gray73","gray87"),
+    xlab="X/A                                                                    Y/A",
[…]

### 2016-11-09 temp/workspace/sup2.r

@@ -1,0 +1,44 @@
+barplot(c(
+  tf_all_ave_weighted_theta,
+  scf_all_ave_weighted_theta,
+  flf_all_ave_weighted_theta),
+  ylab=expression(paste("Nucleotide Diversity (average weighted ", theta, " )")),
+  names.arg=c("TX","SC","FL"), xlim=c(0,3),width=0.5,ylim=c(0,0.007))
+
+#autosomes 
+flf_all$weighted_theta <- flf_all$SynSitestot * flf_all$theta_syntot            
+flf_all_ave_weighted_theta=(sum(flf_all$weighted_theta))/(sum(flf_all$SynSitestot))
+
+p=flf_all$SynSitestot #stores the number of synonymous sites for y-linked genes in vector
+q=flf_all$theta_syntot #stores theta synonymous for X-linked genes in vector
+
+d2 <- cbind.data.frame(p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
+
+boot.function2 <- function(d2,i){
+  p <- d2$p[i]
+  q <- d2$q[i]
+  boot.function2 <- sum(p*q)/sum(p) # calculates average weighted theta from d
+}
+
+z2 <- boot(d2, boot.function2, R = 15000) #store 10000 bootstrap estimates of boot.function in z
+
+print(z2)    # bootstrap calculation of bias and SE
+sd(z2$t)     # another way to get the standard error
+
+boot.ci(z2, type = "perc")  # 95% confidence interval using perc
+
+#TX
+lowerTX=0.0053
+upperTX=0.0058 
+segments(x0=0.35,y0=lowerTX,x1=0.35,y1=upperTX,lwd=2, col="black")
+
+#SC
+lowerSC=0.0054
+upperSC=0.0062
+segments(x0=0.95,y0=lowerSC,x1=0.95,y1=upperSC,lwd=2, col="black")
+
+#FL
+lowerFL=0.0045
+upperFL=0.0051
+segments(x0=1.54,y0=lowerFL,x1=1.54,y1=upperFL,lwd=2, col="black")
+

### 2016-11-09 temp/workspace/supfigs.R

@@ -1,0 +1,26 @@
+#sup fig1
+
+par(mfrow=c(1,2))
+barplot(c(tx_x_auto,
+  tx_y_auto, 
+  nc_x_auto,
+  nc_y_auto),
+  ylab=expression(paste("Relative Nucleotide Diversity (average weighted ", theta, " )")),
+  names.arg=c("TX_X","TX_Y","NC_X", "NC_Y"),
+  ylim=c(0,1),
+  xlim=c(0,4),
+  width=0.59,
+  cex.names=0.75)
+
+barplot(c(tx_x_auto,
+  tx_y_auto, 
+  sc_x_auto,
+  sc_y_auto,
+  fl_x_auto,
+  fl_y_auto),
+  #ylab=expression(paste("Nucleotide Diversity (average weighted ", theta, " )")),
+  names.arg=c("TX_X","TX_Y","SC_X", "SC_Y", "FL_X", "FL_Y"),
+  ylim=c(0,1),
+  xlim=c(0,7),
+  width=1,
+  cex.names=0.75)

### 2016-11-09 temp/workspace/theoryfig.R

@@ -1,0 +1,26 @@
+Nf=5000 #number of females
+Nm=5000 #number of males
+N=Nf+Nm #total number of individuals
+x=Nf/(Nf+Nm) #proportion of females
+
+#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
+par(mfrow=c(1,2))
+curve(4*N*(1-x)*x, from=0,to=1, #autosomes
+      ylab="Effective population size (Ne)", 
+      xlab= "Nf/(Nf+Nm)", 
+      ylim=c(0,11000), lwd=2)
+curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, lwd=2,add=T) #Xchromosomes
+curve(N*(1-x)/2,from=0,to=1, lwd=2, add=T) #Ychromosomes
+text(0.5,2800, labels="Y")
+text(0.5,7800, labels="X")
+text(0.5,10300, labels="A")
+#abline(v=0.6,lty=2)
+
+#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
+curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,, lwd=2,
+      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
+      xlab="Nf/(Nf+Nm)",
+      ylim=c(0,2))
+curve((N*(1-x)/2)/(4*N*(1-x)*x), lwd=2, from=0, to=1,add=T)
+text(0.8,1.05, labels="X/A")
+text(0.8,0.25, labels="Y/A")

### 2016-11-09 temp/workspace/tree.R

@@ -1,0 +1,14 @@
+install.packages("ape")
+install.packages("phytools")
+library(phytools)
+library(ape)
+
+mytree <- read.tree(file.choose()) 
+mytree               # prints basic tree information
+mytree$tip.label     # species names
+mytree$edge.length   # lengths of all tree branches
+mytree$edge          # identity of branches (from node, to node)
+
+plot(mytree,cex=0.8)
+is.rooted(mytree)
+

### 2016-11-09 temp/workspace/Nucleotide_diversity_predictions.r
major reorganization 2

@@ -1,71 +1,0 @@
-Nf=5000 #number of females
-Nm=5000 #number of males
-N=Nf+Nm #total number of individuals
-x=Nf/(Nf+Nm) #proportion of females
-
-#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
-curve(4*N*(1-x)*x, from=0,to=1, #autosomes
-      ylab="Effective population size (Ne)", 
-      xlab= "Nf/(Nf+Nm)", 
-      ylim=c(0,11000))
-curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
-curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
-text(0.5,2800, labels="Y")
-text(0.5,7800, labels="X")
-text(0.5,10300, labels="A")
-abline(v=0.6,lty=2)
-
-#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
-curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
-      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-      xlab="Nf/(Nf+Nm)",
-      ylim=c(0,2))
-curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
-text(0.8,1.05, labels="X/A")
-text(0.8,0.25, labels="Y/A")
-#abline(v=0.6,lty=2)
-points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
-points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-
-
-#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
-#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
-points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-
-
[…]

### 2016-11-09 temp/workspace/autosomal.R

@@ -1,25 +1,0 @@
-#autosomal diversity
-
-tm_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/tm_all.csv")
-tf_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/tf_all.csv")
-nf_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/nf_all.csv")
-nm_all <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/nm_all.csv")
-
-#new columns=SynSitestot*theta_syntot (for each cell)
-#calculate average weighted theta
-
-#tf_all
-tf_all$weighted_theta <- tf_all$SynSitestot * tf_all$theta_syntot            
-tf_all_ave_weighted_theta=(sum(tf_all$weighted_theta))/(sum(tf_all$SynSitestot))
-#tm_all
-tm_all$weighted_theta <- tm_all$SynSitestot * tm_all$theta_syntot            
-tm_all_ave_weighted_theta=(sum(tm_all$weighted_theta))/(sum(tm_all$SynSitestot))
-#nf_all
-nf_all$weighted_theta <- nf_all$SynSitestot * nf_all$theta_syntot            
-nf_all_ave_weighted_theta=(sum(nf_all$weighted_theta))/(sum(nf_all$SynSitestot))
-#nm_all
-nm_all$weighted_theta <- nm_all$SynSitestot * nm_all$theta_syntot            
-nm_all_ave_weighted_theta=(sum(nm_all$weighted_theta))/(sum(nm_all$SynSitestot))
-
-#overall average weighted theta for autosomes:
-mean_all=mean(c(tf_all_ave_weighted_theta, tm_all_ave_weighted_theta, nf_all_ave_weighted_theta, nm_all_ave_weighted_theta))

### 2016-11-09 temp/workspace/bootingtheta.r

@@ -1,47 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-#x-linked genes
-txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
-
-#data for x linked genes
-a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-abline(v=txx_ave_weighted_theta, col="red",lwd=3)
-abline(v=0.0041,col="black", lwd=3)
-abline(v=0.0053,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-qqnorm(z$t)  # normal quantiles of replicate estimates
-boot.ci(z, type = "bca")  # 95% confidence interval using BCa
-boot.ci(z, type = "perc") # same using percentile method
-
-#implied range for X/A
-lowerXA=0.0041/tf_all_ave_weighted_theta
-upperXA=0.0053/tf_all_ave_weighted_theta
-
-
-#autosomal genes
-#tf_all$weighted_theta <- tf_all$SynSitestot * tf_all$theta_syntot            
-#tf_all_ave_weighted_theta=(sum(tf_all$weighted_theta))/(sum(tf_all$SynSitestot))
-#tx_x_auto=txx_ave_weighted_theta/tf_all_ave_weighted_theta
-
-

### 2016-11-09 temp/workspace/bootingthetaY.r

@@ -1,36 +1,0 @@
-#x-linked genes
-txy$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-txy_ave_weighted_theta=(sum(txy$weighted_theta))/(sum(txy$SynSitestot))
-
-#data for y linked genes
-p=fly$SynSitestot #stores the number of synonymous sites for y-linked genes in vector
-q=fly$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d2 <- cbind.data.frame(p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function2 <- function(d2,i){
-  p <- d2$p[i]
-  q <- d2$q[i]
-  boot.function2 <- sum(p*q)/sum(p) # calculates average weighted theta from d
-}
-
-z2 <- boot(d2, boot.function2, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z2)    # bootstrap calculation of bias and SE
-sd(z2$t)     # another way to get the standard error
-
-#hist(z2$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped Y genes"
-#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-#abline(v=txy_ave_weighted_theta, col="red",lwd=3)
-#abline(v=0.0001,col="black", lwd=3)
-#abline(v=0.0002,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-#qqnorm(z2$t)  # normal quantiles of replicate estimates
-boot.ci(z2, type = "bca")  # 95% confidence interval using BCa
-#boot.ci(z2, type = "perc") # same using percentile method
-
-#implied range for X/A
-#lowerYA=0.0001/tf_all_ave_weighted_theta; lowerYA
-#upperYA=0.0002/tf_all_ave_weighted_theta; upperYA
-

### 2016-11-09 temp/workspace/bootingtheta_FL.R

@@ -1,39 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-#x-linked genes
-flx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-flx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
-
-#data for x linked genes
-a=flx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=flx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-#hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-#abline(v=txx_ave_weighted_theta, col="red",lwd=3)
-#abline(v=0.0041,col="black", lwd=3)
-#abline(v=0.0053,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-#qqnorm(z$t)  # normal quantiles of replicate estimates
-boot.ci(z, type = "bca")  # 95% confidence interval using BCa
-
-#implied range for X/A
-lowerXA_FL=0.0014/tf_all_ave_weighted_theta
-upperXA_FL=0.0021/tf_all_ave_weighted_theta
-

### 2016-11-09 temp/workspace/bootingtheta_SC.R

@@ -1,37 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-#x-linked genes
-
-#data for x linked genes
-a=scx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=scx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-#hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-#     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-#abline(v=txx_ave_weighted_theta, col="red",lwd=3)
-#abline(v=0.0041,col="black", lwd=3)
-#abline(v=0.0053,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-#qqnorm(z$t)  # normal quantiles of replicate estimates
-
-#implied range for X/A
-lowerSC_XA=0.0014/tf_all_ave_weighted_theta
-upperSC_XA=0.0025/tf_all_ave_weighted_theta
-
-

### 2016-11-09 temp/workspace/bootingtheta_TX.R

@@ -1,40 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-#x-linked genes
-txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
-
-#data for x linked genes
-a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d <- cbind.data.frame(a, b, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  boot.function <- sum(a*b)/sum(a) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-abline(v=txx_ave_weighted_theta, col="red",lwd=3)
-abline(v=0.0041,col="black", lwd=3)
-abline(v=0.0053,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-qqnorm(z$t)  # normal quantiles of replicate estimates
-boot.ci(z, type = "bca")  # 95% confidence interval using BCa
-boot.ci(z, type = "perc") # same using percentile method
-
-#implied CI for X/A
-lowerXA=0.0042/tf_all_ave_weighted_theta
-upperXA=0.0054/tf_all_ave_weighted_theta
-

### 2016-11-09 temp/workspace/bootingtheta_TX_auto copy copy.R

@@ -1,52 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-#x-linked genes
-txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
-
-#data for x linked genes
-a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-#autosomal genes
-p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
-q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
-
-max.len = max(length(a), length(b), length(p), length(q))
-a = c(a, rep(NA, max.len - length(a)))
-b = c(b, rep(NA, max.len - length(b)))
-p = c(p, rep(NA, max.len - length(p)))
-q = c(q, rep(NA, max.len - length(q)))
-
-d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  p <- d$p[i]
-  q <- d$q[i]
-  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
-abline(v=txx_ave_weighted_theta, col="red",lwd=3)
-abline(v=0.0041,col="black", lwd=3)
-abline(v=0.0053,col="black", lwd=3)
-
-# histogram of boostrap replicate estimates
-qqnorm(z$t)  # normal quantiles of replicate estimates
-boot.ci(z, type = "bca")  # 95% confidence interval using BCa
[…]

### 2016-11-09 temp/workspace/bootingtheta_TX_auto copy.R

@@ -1,61 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-
-ncx <- read.csv("ncx.csv")
-ncy <- read.csv("ncy.csv")
-txx <- read.csv("txx.csv")
-txy <- read.csv("txy.csv")
-scx <- read.csv("scx.csv")
-scy <- read.csv("scy.csv")
-flx <- read.csv("flx.csv")
-fly <- read.csv("fly.csv")
-scf_all <- read.csv("scf_autosomes.csv")
-scm_all <- read.csv("scm_autosomes.csv")
-flf_all <- read.csv("flf_autosomes.csv")
-flm_all <- read.csv("flm_autosomes.csv")
-
-#data for x linked genes
-a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-#autosomal genes
-p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
-q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
-
-max.len = max(length(a), length(b), length(p), length(q))
-a = c(a, rep(NA, max.len - length(a)))
-b = c(b, rep(NA, max.len - length(b)))
-p = c(p, rep(NA, max.len - length(p)))
-q = c(q, rep(NA, max.len - length(q)))
-
-d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  p <- d$p[i]
-  q <- d$q[i]
-  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 10000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z)    # bootstrap calculation of bias and SE
-sd(z$t)     # another way to get the standard error
-boot.ci(z, type="bca")
-
-hist(z$t, breaks=50, col="grey",ylim=c(0,700), main="bootstrapped X-genes"
-     ,xlab=expression(paste("boostrap replicate estimates of mean weighted ", theta, " ")))
[…]

### 2016-11-09 temp/workspace/bootingtheta_TX_auto.R

@@ -1,66 +1,0 @@
-#bootstrapping average weighted theta from polymorphurama output
-library("boot")
-setwd("/Users/joshhough/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/")
-
-txx <- read.csv("txx.csv")
-txy <- read.csv("txy.csv")
-scx <- read.csv("scx.csv")
-scy <- read.csv("scy.csv")
-flx <- read.csv("flx.csv")
-fly <- read.csv("fly.csv")
-scf_all <- read.csv("scf_autosomes.csv")
-scm_all <- read.csv("scm_autosomes.csv")
-flf_all <- read.csv("flf_autosomes.csv")
-flm_all <- read.csv("flm_autosomes.csv")
-tm_all <- read.csv("tm_all_new.csv")
-tf_all <- read.csv("tf_all_new.csv")
-scf_all <- read.csv("scf_autosomes.csv")
-scm_all <- read.csv("scm_autosomes.csv")
-flf_all <- read.csv("flf_autosomes.csv")
-flm_all <- read.csv("flm_autosomes.csv")
-
-#data for x linked genes
-a=txx$SynSitestot #stores the number of synonymous sites for X-linked genes in vector
-b=txx$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-#autosomal genes
-p=tf_all$SynSitestot #stores the number of synonymous sites for autosomal genes in vector
-q=tf_all$theta_syntot #stores theta synonymous for autosomal genes in vector
-
-max.len = max(length(a), length(b), length(p), length(q))
-a = c(a, rep(NA, max.len - length(a)))
-b = c(b, rep(NA, max.len - length(b)))
-p = c(p, rep(NA, max.len - length(p)))
-q = c(q, rep(NA, max.len - length(q)))
-
-d <- cbind.data.frame(a, b, p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function <- function(d,i){
-  a <- d$a[i]
-  b <- d$b[i]
-  p <- d$p[i]
-  q <- d$q[i]
-  boot.function <- (sum(a*b, na.rm=T)/sum(a, na.rm=T))/(sum(p*q, na.rm=T)/sum(p,na.rm=T)) # calculates average weighted theta from d
-}
-
-z <- boot(d, boot.function, R = 20000) #store 10000 bootstrap estimates of boot.function in z
-
[…]

### 2016-11-09 temp/workspace/expected_pies.R

@@ -1,7 +1,0 @@
-pi_X=4*750*0.005
-pi_Y=4*250*0.005
-pi_A=4*1000*0.005
-
-ex_Y=pi_Y*(exp(-(0.02/((2*0.01)+0))));ex_Y
-ex_X=pi_X*(exp(-(0.02/((2*0.01)+0.5))));ex_X
-ex_A=pi_A*(exp(-(0.02/((2*0.01)+0.5))));ex_A

### 2016-11-09 temp/workspace/fold.R

@@ -1,7 +1,0 @@
-aveX=mean(c(scx_ave_weighted_theta,txx_ave_weighted_theta,flx_ave_weighted_theta));aveX
-aveY=mean(c(scy_ave_weighted_theta,txy_ave_weighted_theta,fly_ave_weighted_theta));aveY
-aveX/aveY
-
-mean(c(tf_all_ave_weighted_theta,scf_all_ave_weighted_theta,flf_all_ave_weighted_theta))/aveY
-
-mean(c(tx_x_auto,sc_x_auto,fl_x_auto))/mean(c(tx_y_auto,sc_y_auto,fl_y_auto))

### 2016-11-09 temp/workspace/malerepvariance.R

@@ -1,89 +1,0 @@
-#the effect of male reproductive variance on sex chromosome to autosome ratios of Ne
-
-#(4*Nf*Nm)/(Nf+Nm) #autosome Ne 
-#(9*Nf*Nm)/(((2*Nf))+4*Nm) #X-chromosome Ne
-#Nm/2 #Y chromosome Ne
-#Nf=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of fmeale reproductive variance on female Ne: 
-#Nm=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
-
-#set numbers of males and females
-#Nft=100000 #Nf for Ntheoretical population of autosomes, equal males and females, no rep variance, poisson offspring
-#Nmt=100000 #Nm for theoretical population of autosomes, equal males and females, no rep variance, poisson offspring
-vf=2 #variance in female rep success. Set to zero. (change to 2 for poisson)
-kf=2 #mean number of offspring of females
-km=2 #mean number of offspring of males
-Nf=100000 #female Ne
-Nm=100000 #male Ne
-
-
-#effect of male rep variance on autosomal Ne
-#(4*Nf*Nm)/(4*Nf+Nm) #autosome Ne
-#((4*Nf*Nm)/((4*Nf)+Nm))/((4*Nf*Nm)/((4*Nf)+Nm)) #autosomes relative to autosomes 
-#replace numerator Ne's with rep variance equations for males and females
-#((4*Nf*Nm)/((4*Nf)+Nm))/((4*Nf*Nm)/((4*Nf)+Nm))
-
-#Nf=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of fmeale reproductive variance on female Ne: 
-#Nm=((Nm*(km-1))/((km-1)+(x/km))) #effect of male reproductive variance on male Ne
-
-curve(((4*((Nf*(kf-1))/((kf-1)+(vf/kf)))*((Nm*(km-1))/((km-1)+(x/km))))/((4*((Nf*(kf-1))/((kf-1)+(vf/kf))))+((Nm*(km-1))/((km-1)+(x/km)))))/((4*Nft*Nmt)/((4*Nft)+Nmt)),
[…]

### 2016-11-09 temp/workspace/mle.R

@@ -1,19 +1,0 @@
-YAlikihoods <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/YAlikihoods.csv")
-
-plot(YAlikihoods$YAratio,YAlikihoods$mle, pch=19,xaxt='n', xlab="Y/A Ne ratio", ylab="Likelihood")
-axis(3)
-abline(h=-2, lty=2, lwd=3)
-segments(x0=0.09775,y0=-0.9,x1=0.09775,y1=-1.1,lwd=2, col="red")
-segments(x0=0.009,y0=-0.9,x1=0.009,y1=-1.1,lwd=2, col="red")
-#segments(x0=0.009,y0=1,x1=0.009,y1=-2,lwd=1,lty=2, col="black")
-#segments(x0=0.09775,y0=1,x1=0.09775,y1=-2,lwd=1,lty=2, col="black")
-legend("topright", c("likelihood credibility interval", "LRT significance cut-off"),
-       cex=0.7, box.lty=0, lty=c(1,2),col=c("red","black"), lwd=c(2,2))
-segments(0.009, -1, 0.09775, -1,lwd=2,code = 3,col="red")
-
-#rect(0.009, -2, 0.09775, 1, density = NULL, angle = 45,
-#     col="grey", border = NULL, lty = par("lty"), lwd = par("lwd"))
-
-#cord.x=c(0.009,0.09775)
-#cord.y=c(-2,-2)
-#polygon(cord.x,cord.y, col=gray(0.8))

### 2016-11-09 temp/workspace/mlhka.R

@@ -1,64 +1,0 @@
-#making MLHKA infile.txt
-
-#rename polymorphurama output file from TX_X
-txx_mlhka=txx
-
-#take names from "locus" column in txx_mlhka file,
-#and remove everything after "T
-test <- as.data.frame(sub("T.*$","",txx_mlhka$locus))
-
-#replaced in excel
-write.xlsx(test, file="test.xlsx")
-write.xlsx(txx_mlhka, file="txx_mlhka.xlsx")
-#load fixed file
-txx_mlhka_fixed <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/txx_mlhka_fixed.csv")
-
-#now do same with dNdSFreeRatioNCshared
-test2 <- as.data.frame(sub("T.*$","",dNdSFreeRatioNCshared$LOC))
-write.xlsx(test2, file="test2.xlsx")
-dNdSFreeRatioNCshared_fixed <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/dNdSFreeRatioNCshared_fixed.csv")
-
-crap <- merge(txx_mlhka_fixed, dNdSFreeRatioNCshared_fixed)
-write.xlsx(crap, file="crap.xlsx")
-
-
-#ok i think file called "crap" is good: it is a merge of txx_mlhla_fixed and dNdSFreeRatioNCshared_fixed
-#make new column in crap for "D" =synsites tot* OUTS for D
-crap$D_mlhka <- crap$SynSitestot * crap$OUTS
-
-#fixing D column
-write.xlsx(crap, file="crap2.xlsx")
-
-#ok crap2 now has all relevant information. Copied this file as crap_infile in order to format for infile.txt for MLHKA
-crap_infile <- read.csv("~/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/crap_infile.csv")
-crap_infile$SynSitestot_rounded <- round(crap_infile$SynSitestot)
-
-write.xlsx(crap_infile, file="crap_infile_top.xlsx")
-
-
-#OK, things needed for MLHKA input, in order:
-#1. Locus ID. This is locus
-#2. L (size in BP). This is SynSitestot_rounded
[…]

### 2016-11-09 temp/workspace/new_malevariance.R

@@ -1,118 +1,0 @@
-#set numbers of males and females
-#vm=50 #variance in male rep success
-vf=2 #variance in female rep success
-kf=2 #mean number of offspring of females
-km=2 #mean number of offspring of males
-Nf=100000 #numner of females
-Nm=100000 #number of males
-
-Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
-Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
-
-Ne_ratio=Nef/Nem
-N=Nf+Nf
-
-A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
-X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
-Y=Nem/2
-XA=X/A
-YA=Y/A
-
-vm <- c(1:100000)
-for (val in vm) {
-}
-plot(((Nm*(km-1))/((km-1)+(vm/km)))/Nef, ylab="Nem/Nef",xlab="vm", type="l")
-
-
-vm <- c(1:50)
-for (val in vm) {
-Nf=100000 #numner of females
-Nm=100000 #number of males
-N=Nf+Nf
-vf=2 #variance in female rep success
-kf=2 #mean number of offspring of females
-km=2 #mean number of offspring of males
-Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
-Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
-Ne_ratio=Nef/Nem
-A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
-X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
-Y=Nem/2
-XA=X/A
-YA=Y/A
-}
-#plot(YA, ylab="Y/A",xlab="vm", type="l")
-#plot(XA, add=TRUE)
-d <- data.frame(vm,Nem,Nef,Ne_ratio,A,X,Y,XA,YA)
-require(ggplot2)
-ggplot(d, aes(vm)) +                    # basic graphical object
-  geom_line(aes(y=YA), colour="red") +  # first layer
-  geom_line(aes(y=XA), colour="green")  # second layer
-g <- g + ylab("Y") + xlab("X")
-g
-
-
-
[…]

### 2016-11-09 temp/workspace/nov24.R

@@ -1,76 +1,0 @@
-Nf=5000 #number of females
-Nm=5000 #number of males
-N=Nf+Nm #total number of individuals
-x=Nf/(Nf+Nm) #proportion of females
-
-#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
-curve(4*N*(1-x)*x, from=0,to=1, #autosomes
-      ylab="Effective population size (Ne)", 
-      xlab= "Nf/(Nf+Nm)", 
-      ylim=c(0,11000))
-curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
-curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
-text(0.5,2800, labels="Y")
-text(0.5,7800, labels="X")
-text(0.5,10300, labels="A")
-abline(v=0.6,lty=2)
-
-#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
-curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
-      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-      xlab="Nf/(Nf+Nm)",
-      ylim=c(0,2))
-curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
-text(0.8,1.05, labels="X/A")
-text(0.8,0.25, labels="Y/A")
-#abline(v=0.6,lty=2)
-points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
-points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-
-
-#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
-#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
-points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-
-
[…]

### 2016-11-09 temp/workspace/old/polymorphism.R

@@ -1,1 +1,0 @@
-S_Syntot_NMX=

### 2016-11-09 temp/workspace/polymorphurama.R

@@ -1,225 +1,0 @@
-
-#read polymorphurama output (converted to .csv and parsed)
-setwd("/Users/joshhough/Dropbox/JoshDocuments/academic/University/University of Toronto/PhD/manuscripts/Hough et al. Pop gen X and Y/workspace/")
-ncx <- read.csv("ncx.csv")
-ncy <- read.csv("ncy.csv")
-txx <- read.csv("txx.csv")
-txy <- read.csv("txy.csv")
-scx <- read.csv("scx.csv")
-scy <- read.csv("scy.csv")
-flx <- read.csv("flx.csv")
-fly <- read.csv("fly.csv")
-
-#new columns=SynSitestot*theta_syntot (for each cell)
-#calculate average weighted theta
-#NC_X
-ncx$weighted_theta <- ncx$SynSitestot * ncx$theta_syntot            
-ncx_ave_weighted_theta=(sum(ncx$weighted_theta))/(sum(ncx$SynSitestot))
-#NC_Y
-ncy$weighted_theta <- ncy$SynSitestot * ncy$theta_syntot            
-ncy_ave_weighted_theta=(sum(ncy$weighted_theta))/(sum(ncy$SynSitestot))
-#TX_X
-txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
-txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
-#TX_Y
-txy$weighted_theta <- txy$SynSitestot * txy$theta_syntot            
-txy_ave_weighted_theta=(sum(txy$weighted_theta))/(sum(txy$SynSitestot))
-
-#SC_X
-scx$weighted_theta <- scx$SynSitestot * scx$theta_syntot            
-scx_ave_weighted_theta=(sum(scx$weighted_theta))/(sum(scx$SynSitestot))
-#SC_Y
-scy$weighted_theta <- scy$SynSitestot * scy$theta_syntot            
-scy_ave_weighted_theta=(sum(scy$weighted_theta))/(sum(scy$SynSitestot))
-#Fl_X
-flx$weighted_theta <- flx$SynSitestot * flx$theta_syntot            
-flx_ave_weighted_theta=(sum(flx$weighted_theta))/(sum(flx$SynSitestot))
-#Fl_Y
-fly$weighted_theta <- fly$SynSitestot * fly$theta_syntot            
[…]

### 2016-11-09 temp/workspace/predictions.R

@@ -1,62 +1,0 @@
-Nf=5000 #number of females
-Nm=5000 #number of males
-N=Nf+Nm #total number of individuals
-x=Nf/(Nf+Nm) #proportion of females
-
-#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
-curve(4*N*(1-x)*x, from=0,to=1, #autosomes
-      ylab="Effective population size (Ne)", 
-      xlab= "Nf/(Nf+Nm)", 
-      ylim=c(0,10000))
-curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, add=T) #Xchromosomes
-curve(N*(1-x)/2,from=0,to=1, add=T) #Ychromosomes
-text(0.5,2800, labels="Y")
-text(0.5,7800, labels="X")
-text(0.5,10300, labels="A")
-abline(v=0.6,lty=2)
-
-#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
-curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,
-      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-      xlab="Nf/(Nf+Nm)",
-      ylim=c(0,2))
-curve((N*(1-x)/2)/(4*N*(1-x)*x), from=0, to=1,add=T)
-text(0.8,1.05, labels="X/A")
-text(0.8,0.25, labels="Y/A")
-#abline(v=0.6,lty=2)
-#points(0.6,tx_x_auto,pch=20,col="darkgreen") #empirical X/A weighted theta
-#points(0.6,sc_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-#points(0.6,fl_x_auto,pch=20,col="darkred") #empirical X/A weighted theta
-
-#points(0.6,1.0,pch=15,col="darkgreen") # X/A mlhka
-#points(0.6,0.017,pch=15,col="red") # Y/A mlhka
-#points(0.6,tx_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-#points(0.6,sc_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-#points(0.6,fl_y_auto,pch=20, col="black") #empirical Y/A weighted theta
-
[…]

### 2016-11-09 temp/workspace/pvalue.R

@@ -1,3 +1,0 @@
-prop.test(4, 200, p=0.2,
-          alternative="two.sided",
-          conf.level = 0.95, correct=TRUE)

### 2016-11-09 temp/workspace/sim_output.R

@@ -1,183 +1,0 @@
-#new 
-
-#Ychromosome
-Y_npoly=16.49500
-Y_thetapi=4.22218
-Y_thetaw=4.41717
-
-#x-chromosome
-X_npoly=55.366667
-X_thetapi=14.915749
-X_thetaw=14.826552
-
-
-#the A
-A_npoly=73.005000
-A_thetapi=19.566661
-A_thetaw=19.549893
-
-
-
-#A=A_npoly*A_thetaw;A
-#X=X_npoly*X_thetaw;X
-#Y=Y_npoly*Y_thetaw;Y
-
-
-y$weighted_theta <- y$npoly * y$thetaw
-yw=(sum(y$weighted_theta))/(sum(y$thetaw));yw
-
-a$weighted_theta <- a$npoly * a$thetaw
-aw=(sum(a$weighted_theta))/(sum(a$thetaw));aw
-
-x$weighted_theta <- x$npoly * x$thetaw
-xw=(sum(x$weighted_theta))/(sum(x$thetaw));xw
-
-#xa=xw/aw;xa
-#ya=yw/aw;ya
-
-X
-
-yesy=mean(y$weighted_theta)/mean(a$weighted_theta);yesy
-
-
-
-weighted.mean(y$thetaw,y$npoly)/weighted.mean(a$thetaw,a$npoly)
-weighted.mean(a$thetaw,a$npoly)
-
-
-[1] 0.06062793
-yesx=mean(x$weighted_theta)/mean(a$weighted_theta);yesx
-[1] 0.5656107
-
-
-boxplot(y$weighted_theta,x$weighted_theta, a$weighted_theta, col="grey", notch=T, 
-        main="thetas from BGS simulations",
-        ylab="theta*synsites",
-        xlab="Y                  X                  A")
-
-barplot(c(tx_x_auto, sc_x_auto, fl_x_auto, yesx, xA_exp_equal, xA_exp_06, xA_exp_075, xA_exp_1,
-          tx_y_auto, sc_y_auto, fl_y_auto, yesy, yA_exp_equal, yA_exp_06, yA_exp_075, yA_exp_1),
-        ylim=c(0,1.2),
-    col=c("darkgreen","darkgreen","darkgreen","red","black", "gray22","gray73","gray87",
-          "darkgreen","darkgreen","darkgreen", "red","black", "gray22","gray73","gray87"),
-    xlab="X/A                                                                    Y/A",
[…]

### 2016-11-09 temp/workspace/sup2.r

@@ -1,44 +1,0 @@
-barplot(c(
-  tf_all_ave_weighted_theta,
-  scf_all_ave_weighted_theta,
-  flf_all_ave_weighted_theta),
-  ylab=expression(paste("Nucleotide Diversity (average weighted ", theta, " )")),
-  names.arg=c("TX","SC","FL"), xlim=c(0,3),width=0.5,ylim=c(0,0.007))
-
-#autosomes 
-flf_all$weighted_theta <- flf_all$SynSitestot * flf_all$theta_syntot            
-flf_all_ave_weighted_theta=(sum(flf_all$weighted_theta))/(sum(flf_all$SynSitestot))
-
-p=flf_all$SynSitestot #stores the number of synonymous sites for y-linked genes in vector
-q=flf_all$theta_syntot #stores theta synonymous for X-linked genes in vector
-
-d2 <- cbind.data.frame(p, q, stringsAsFactors = FALSE) #stores a and b in data frame d
-
-boot.function2 <- function(d2,i){
-  p <- d2$p[i]
-  q <- d2$q[i]
-  boot.function2 <- sum(p*q)/sum(p) # calculates average weighted theta from d
-}
-
-z2 <- boot(d2, boot.function2, R = 15000) #store 10000 bootstrap estimates of boot.function in z
-
-print(z2)    # bootstrap calculation of bias and SE
-sd(z2$t)     # another way to get the standard error
-
-boot.ci(z2, type = "perc")  # 95% confidence interval using perc
-
-#TX
-lowerTX=0.0053
-upperTX=0.0058 
-segments(x0=0.35,y0=lowerTX,x1=0.35,y1=upperTX,lwd=2, col="black")
-
-#SC
-lowerSC=0.0054
-upperSC=0.0062
-segments(x0=0.95,y0=lowerSC,x1=0.95,y1=upperSC,lwd=2, col="black")
-
-#FL
-lowerFL=0.0045
-upperFL=0.0051
-segments(x0=1.54,y0=lowerFL,x1=1.54,y1=upperFL,lwd=2, col="black")
-

### 2016-11-09 temp/workspace/supfigs.R

@@ -1,26 +1,0 @@
-#sup fig1
-
-par(mfrow=c(1,2))
-barplot(c(tx_x_auto,
-  tx_y_auto, 
-  nc_x_auto,
-  nc_y_auto),
-  ylab=expression(paste("Relative Nucleotide Diversity (average weighted ", theta, " )")),
-  names.arg=c("TX_X","TX_Y","NC_X", "NC_Y"),
-  ylim=c(0,1),
-  xlim=c(0,4),
-  width=0.59,
-  cex.names=0.75)
-
-barplot(c(tx_x_auto,
-  tx_y_auto, 
-  sc_x_auto,
-  sc_y_auto,
-  fl_x_auto,
-  fl_y_auto),
-  #ylab=expression(paste("Nucleotide Diversity (average weighted ", theta, " )")),
-  names.arg=c("TX_X","TX_Y","SC_X", "SC_Y", "FL_X", "FL_Y"),
-  ylim=c(0,1),
-  xlim=c(0,7),
-  width=1,
-  cex.names=0.75)

### 2016-11-09 temp/workspace/theoryfig.R

@@ -1,26 +1,0 @@
-Nf=5000 #number of females
-Nm=5000 #number of males
-N=Nf+Nm #total number of individuals
-x=Nf/(Nf+Nm) #proportion of females
-
-#Expected Ne as a function of the sex ratio for different chromosome types in a neutral model
-par(mfrow=c(1,2))
-curve(4*N*(1-x)*x, from=0,to=1, #autosomes
-      ylab="Effective population size (Ne)", 
-      xlab= "Nf/(Nf+Nm)", 
-      ylim=c(0,11000), lwd=2)
-curve(9*N*x*(x-1)/(2*(x-2)),from=0,to=1, lwd=2,add=T) #Xchromosomes
-curve(N*(1-x)/2,from=0,to=1, lwd=2, add=T) #Ychromosomes
-text(0.5,2800, labels="Y")
-text(0.5,7800, labels="X")
-text(0.5,10300, labels="A")
-#abline(v=0.6,lty=2)
-
-#Nucelotide diversity ratios (for a neutral model, assuming equal mutation rates, and theta=4Nemu)
-curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,, lwd=2,
-      ylab=expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-      xlab="Nf/(Nf+Nm)",
-      ylim=c(0,2))
-curve((N*(1-x)/2)/(4*N*(1-x)*x), lwd=2, from=0, to=1,add=T)
-text(0.8,1.05, labels="X/A")
-text(0.8,0.25, labels="Y/A")

### 2016-11-09 temp/workspace/tree.R

@@ -1,14 +1,0 @@
-install.packages("ape")
-install.packages("phytools")
-library(phytools)
-library(ape)
-
-mytree <- read.tree(file.choose()) 
-mytree               # prints basic tree information
-mytree$tip.label     # species names
-mytree$edge.length   # lengths of all tree branches
-mytree$edge          # identity of branches (from node, to node)
-
-plot(mytree,cex=0.8)
-is.rooted(mytree)
-

### 2016-11-09 manuscript/manuscript.tex
major reorganization
push reorganized files, remove temp/
edits to introction, HRI arguments

--- 2016-11-08 manuscript.tex
@@
 For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the genetic degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
-The independent evolution of these phenomena across a broad range of distantly-related species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
+The independent evolution of these phenomena in such taxonomically distant species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
 % and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.
@@
 Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by interference among selected but genetically linked sites (the "Hill-Robertson Effect").
-Thus, a reduction in the level of neutral polymorphism maintained on the Y chromosome is expected, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
+The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of genetic drift \citep{nei1981genetic}.
+Moreover, this loss of recombination on the Y implies that genes on this chromosome will experience strong linkage disequilibrium, making them vulnerable to "Hill-Robertson Interference", in which the genetic associations between selected alleles and their genetic backgrounds cause a reduction in the efficacy of selection \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
+This suggests that Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, though the extent to which selective interference should affect chromosome-wide levels of nucleotide diversity is not known.
+However, given that a large number of sites are probably under weak selection (refs), the "interference" effect might alone cause a substantial diversity loss  important feature of the evolution of nonrecombining genomes.
+
+(though simulations suggest a reduction of 30 \% relative to neutrality when there are 100,000 completely linked sites.
+
+which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
 
@@
 \noindent
-\includegraphics[width=\linewidth]{figure2.png}
+\includegraphics[width=\linewidth]{figure2.jpg}
 \caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).

### 2016-11-09 manuscript/manuscript.tex
more intro edits

@@
 Moreover, this loss of recombination on the Y implies that genes on this chromosome will experience strong linkage disequilibrium, making them vulnerable to "Hill-Robertson Interference", in which the genetic associations between selected alleles and their genetic backgrounds cause a reduction in the efficacy of selection \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
-This suggests that Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, though the extent to which selective interference should affect chromosome-wide levels of nucleotide diversity is not known.
-However, given that a large number of sites are probably under weak selection (refs), the "interference" effect might alone cause a substantial diversity loss  important feature of the evolution of nonrecombining genomes.
+This suggests that Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population (proportional to the product of $N_{e}$ and the mutation rate, $\mu$, at equilibrium), though the extent to which such interference should affect chromosome-wide levels of nucleotide diversity is not known.
+Given that a large number of sites are probably under weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
+Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
 
-(though simulations suggest a reduction of 30 \% relative to neutrality when there are 100,000 completely linked sites.
+In species with female-biased reproductive sex ratios or extensive male-male competition, high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, which suggests that inferences about the effects of HRI (including the effects of purifying and positive selection) on the Y need to be distinguished from neutral models that include sex ratio variation as well as sex differences in reproductive success.
+Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be therefore obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
+For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
+Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though the inflated X/A ratio has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic})
 
-which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.
-
-In species with female-biased reproductive sex ratios or extensive male-male competiton, high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of linked selection on the Y need to be distinguished from neutral models that include sex ratio variation as well as male-female differences in reproductive success.
-Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be obtained by comparing levels of neutral variability on X and Y chromosomes, relative to values obtained for the autosomes.
-In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though it the inflated X/A ratio in humans has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic})
-
-Despite widespread interest in determining the evolutionary factors affecting sex-linked variability \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of demography or selection on patterns of nucleotide diversity in more recently evolved sex chromosomes, such as in dioecious plants.
-We therefore don't yet know the time-scales over which processes such as linked selection on the Y chromosome are likely to be important.
-In humans, for example, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, and simulations have revealed that levels of Y-diversity are consistent with a combination of very low $N_{m}$ and strong purifying selection eliminating linked neutral variation \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is an open question whether such strong effects of purifying selection should also apply to systems where sex chromosomes evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA as in the case of diecious plants \citep{charlesworth2015plant}).
-The effects of diversity loss due to purifying selection might on the one hand be expeted to be lower in such systems because of a shorter history of recombination suppression.
-Contrary to this intuition, however, simulations have shown that background selection and selective sweeps both have the strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
-Indeed, studies of Y-degeneration in the young plant sex chromosomes \textit{R.
-hastatulus} and \textit{Silene latifolia} have found signs of rapid Y chromosome degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then it is possible that the strength of linked selection would then be minimized such that the inferred gene loss was primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
+Despite widespread interest in determining the evolutionary factors affecting neutral diversity on both X and Y chromosomes \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of demography or selection on patterns of sex-linked diversity in more recently evolved sex chromosomes, such as in dioecious plants.
+This information is important, however, as it highlights the relevant time scales over which effects such as linked selection are likely to be important.
+In humans, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, with simulations showing that current levels of Y-diversity are consistent with a combination of very low $N_{m}$ and a history of strong purifying selection \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether such strong effects of purifying selection on diversity should also apply to sex chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of most dioecious plants \citep{charlesworth2015plant}).
+The diversity loss due to purifying selection might on the one hand be expected to be lower in such systems due recombination having been suppressed on Y-chromosome for a shorter length of time.
+Contrary to this intuition, however, simulations have shown that two of the key types if HRI - background selection and selective sweeps - are both expected to have the strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
+Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosomes that have found signs of rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
+In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y chromosome in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then it is possible that linked selection plays a less important role than expected in the early stages, with gene loss instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
 

### 2016-11-09 manuscript/manuscript.tex
minor intro edit

@@
 However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether such strong effects of purifying selection on diversity should also apply to sex chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of most dioecious plants \citep{charlesworth2015plant}).
-The diversity loss due to purifying selection might on the one hand be expected to be lower in such systems due recombination having been suppressed on Y-chromosome for a shorter length of time.
-Contrary to this intuition, however, simulations have shown that two of the key types if HRI - background selection and selective sweeps - are both expected to have the strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
+Y-linked diversity loss due to purifying selection might on the one hand be expected to be lower in such systems due to recombination having been suppressed for a shorter length of time.
+Contrary to this intuition, however, simulations have shown that two key types of HRI - background selection and selective sweeps - both have strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
 Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosomes that have found signs of rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y chromosome in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then it is possible that linked selection plays a less important role than expected in the early stages, with gene loss instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
+In these cases, if the early stages of Y-chromosome evolution involved widespread Y-inactivation (gene silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then linked selection might play a less important role than expected, with the early stages of Y-degeneration being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
 

### 2016-11-09 manuscript/manuscript.tex
final intro edits today

@@
 For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though the inflated X/A ratio has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic})
+Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though the inflated X/A ratio has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic})
 
@@
 Contrary to this intuition, however, simulations have shown that two key types of HRI - background selection and selective sweeps - both have strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
-Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosomes that have found signs of rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-In these cases, if the early stages of Y-chromosome evolution involved widespread Y-inactivation (gene silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then linked selection might play a less important role than expected, with the early stages of Y-degeneration being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
+Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosome systems, which have found evidence for rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
+In these cases, if Y-degeneration involved widespread Y-linked gene inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then linked selection might in fact play a less important role than theory would predict, with the early stages of Y-degeneration instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
 
@@
 Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone significant gene loss and functional deterioration \citep{hough2014}.
-Here, in order to simplify our comparison of polymorphism levels between X, Y, and autosomes, we focused on the ancestral Y chromosome, which occurs in both sex chromosome races.
+Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, we focus only on the ancestral Y chromosome, which occurs in both sex chromosome races.
 
-Of particular relevance to the present study, \textit{R.
+Of particular relevance to our study, \textit{R.
 hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{R.
+Female-biased sex ratios in dioecious plants with heteromorphic X and Y chromosomes are not uncommon (\X refs) \citep{hough2013evolutionarily}, and their occurrence in \textit{R.
 hastatulus} provides an excellent opportunity to study both the demographic and selective factors contributing to sex-linked variability.
-Using estimates of polymorphism estimated on X, Y, and autosomals, together with forward population simulations, our study aims to test whether levels of sex-linked neutral diversity in this species are consistent with models of neutral evolution, including the effects of sex ratio bias and high variance in reproductive success, or whether linked selection has had a significant impact on patterns of nucleotide variability during the early stages of sex chromosome evolution.
+Using estimates of neutral polymorphism on sex chromosomes and autosomes, together with forward population genetic simulations, our study tests whether sex chromosome evolution in the early stages is driven primarily by neutral processes, or whether linked selection has played a significant role.
 

### 2016-11-12 manuscript/manuscript.tex
edits, added new figure1, R codes

@@
 \begin{equation}
-N_{e_{i}}=\frac{N_{i}k_{i}-1}{k_{i}-1+\frac{V_{k_{i}}}{\overline{k}}} \label{eq:Vk}
+N_{e_{m}}=\frac{N_{m}k_{m}-1}{k_{m}-1+\frac{V_{k_{m}}}{\overline{k}}} \label{eq:Vk}
 \end{equation}
 
-where $k_{i}$ and $V_{k_{i}}$ are the mean and variance in offspring number, respectively, for the $i$th sex and $\overline{k}$ is the total mean number of progeny (Kimura and Crow 1964).
-Using sex-specific values of $k$ and $V_{k{i}}$ to calculate $N_{m}$ and $N_{f}$, the expected $N_{e}$ for autosomes and sex chromosomes is then obtained using equations from \citep{wright1931evolution}:
+where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
+Given the expected $N_{m}$ and $N_{f}$ for varying levels $V_{k_{m}}$, the expected $N_{e}$ for autosomes and sex chromosomes was then calculated using equations from \citep{wright1931evolution}:
 
@@
 Note that with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-Figure 1 shows the expected X/A and Y/A ratios as a function of the reproductive sex ratio across a range of values for the variance in male reproductive success.
-We therefore compared our empirical diversity ratios to these predictions (Figure 2) using a sex ratio of 0.6 as estimated in \textit{R.
-hastatulus}
+Figure 1 shows these expected X/A and Y/A ratios as a function of the sex ratio.
+We tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) using the sex ratio of 0.6 as estimated in \textit{R.
+hastatulus}.
 
@@
 \includegraphics[width=\linewidth]{figure1.jpg}
-\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).
-The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$ for sex chromosomes and autosomes is given in equations 1-3.
-Dotted curves show predictions in the standard neutral model, where both males and females produce Poisson-distributed offspring numbers and the chromosomal $N_{e}$'s are given by equations 4-6 \citep{wright1931evolution}.
-Solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods).
-Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate null predictions for X/A and Y/A ratios of diversity.
+\caption{The relation between relative effective population size at equilibrium and sex ratio bias for genes on autosomes (\textbf{A}), the X chromosome (\textbf{B}), and the normalized X/A and Y/A ratios (\textbf{C}).
+The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})$, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$'s' for sex chromosomes and autosomes are given in equations 2-4.
+Dotted curves show predictions when both males and females produce Poisson-distributed offspring numbers ($\overline{k}=2$).
+Solid curves correspond to increasing levels of variance in male reproductive success, ranging from Poisson/2 (top solid curve in panels A and B) to 3*Poisson (bottom most curve).
+Assuming $\theta=4N_{e}\mu$ and an equal neutral mutation rates among genes, these predicted $N_{e}$'s were used to generate our null neutral predictions for the X/A and Y/A ratios of diversity.
 }
@@
 To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a non-recombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
-This approach was implemented rather than using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
+This approach was implemented rather than using analytical predictions of background selection for a few reasons.
+First,...
+because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
 
-To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (Keightley and Eyre-Walker ref).
+To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (\X Keightley and Eyre-Walker ref).
 This program fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism from the population genetic analyses above.
@@
 To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of $\pi_{synonymous}$ .
-To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of Nf/(Nf+Nm)=0.6, and we sampled 6 haploid chromosomes per simulation.
+To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$, and we sampled 6 haploid chromosomes per simulation.
 Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information).
@@
 \includegraphics[width=\linewidth]{figure2.jpg}
-\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red).
+\caption{new figure coming soon).
 }

### 2016-11-12 code/R/Figure1.R

@@ -1,5 +1,5 @@
-Nf=100 #number of females
-Nm=100 #number of males
+Nf=1000 #number of females
+Nm=1000 #number of males
 N=Nf+Nm #total number of individuals
 #x=Nf/(Nf+Nm) #proportion of males
 
@@ -16,7 +16,7 @@ mh=0.5
 curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+mh))))/N, from=0, to=1, add=T)
 
 m1=1
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T, col="blue")
+curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T)
 
 m2=2
 curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m2))))/N, from=0, to=1, add=T)
@@ -32,7 +32,7 @@ curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m))))/N, from=0, to=1, l
 
 curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+mh))))/N, from=0, to=1, add=T)
 
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m1))))/N, from=0, to=1, add=T, col="blue")
+curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m1))))/N, from=0, to=1, add=T)
 
 curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m2))))/N, from=0, to=1, add=T)
 
@@ -59,7 +59,7 @@ curve(((9*N*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1,lty=2,
       xlab=expression(italic('N'[f]/'N'[m]+'N'[f])),
       ylim=c(0,2))
 #curve(((9*Ne_0*x*(x-1)/(2*(x-2))))/(4*N*(1-x)*x), from=0, to=1, add=T)
[…]

### 2016-11-12 manuscript/manuscript.tex
abstract edits

@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability within the genome.
-On Y chromosomes with suppressed recombination, selection is expected to remove linked neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
+On Y chromosomes with suppressed recombination, selection is expected to eliminate neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y diversity.
-Here, we investigate the factors affecting X- and Y-linked neutral polymorphsm during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes.
+Here, we investigate the factors affecting X- and Y-linked neutral polymorphism during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with recently evolved sex chromosomes.
 Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
-We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Rather, using forward population genetic simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity can be explained by purifying selection removing linked neutral variation, although selective sweeps may have also contributed.
+We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male offspring number.
+Instead, using forward simulations, we show that diversity loss on the Y is consistent with interference among a large number (>/= 800Kb) of weakly selected mutations.
+Our results are in agreement with theory on "interference selection" and provide evidence that purifying selection on a large number of linked sites can substantially reduce neutral diversity.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our study suggests that Y chromosome degeneration in the early stages occurs through strong selective interference among many linked sites rather than primarily through Y-inactivation followed by neutral genetic drift.
-%abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.
+hastatulus} sex chromosomes (<15mya), our results further imply that Y chromosome degeneration in the early stages may be largely driven by such interference effects rather than by selectively-favoured gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 
+...there is no general analytical treatment describing the dynamics of interference among many alleles under selection and drift, with variable degrees of selection and linkage.
+Thus, forward computer simulations are still fundamental to the study of realistic situations.
+- dfe for the y is probably still fairly similar to that for the x -- shared genes
+
+- human Y: genes on y are not on X
+
+- 45331 = synsites in observed data after filtering
+  - summary stats phased texas autospy
+
+- doesnt matter about coding/non-coding in non-recombining regions? all of those neutral sites are linked
+
 To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (\X Keightley and Eyre-Walker ref).
@@
 
-\subsection*{Female biased sex ratio and high variance in male fitness}
+\subsection*{Female biased sex ratio and high variance in male offspring number}
 
@@
 
-%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+discussion points to make:
+- Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+
+BGS predictions:
+-"Traditionally, the term ‘‘background selection’’ is used to refer both to the general effects of purifying selection on linked neutral diversity as well as to the limiting behavior that emerges when Ns-->inf.
+"
+- "existing theory struggles to predict genetic diversity when many sites experience selection at the same time, which limits our ability to interpret variation in DNA sequence data.". In particular,
+
+- "For non-crossover regions, both with and without gene conversion, there is a rapid initial decline of B with L, but B levels off at a value of  0.015 for > 640 000 sites.
+B values predicted by the BGS model decrease log-linearly with L"
+
+Very relevant remarks From Good et al.:
+
+- "Here, we have shown that simple behavior emerges in the limit of widespread interference.
+When fitness variation is composed of many individual mutations, the magnitudes and signs of their fitness effects are relatively unimportant.
+Instead, molecular evolution is controlled by the variance in fitness within the population over some effectively asexual segment of the genome" " In other words, we cannot conclude that interference is negligible just because ‘‘Nes’’, as inferred from data, is larger than one."
+
+- estimates of "Nes" ignore linkage by fiat under the assumption that sites evolve independently.
+But these estimates become unreliable precisely when small- and intermediate-effect mutations are most common
+
+- "Individual fitness effects may play a central role in single-site models, but we have shown that global properties like the variance in fitness and the corresponding linkage scale are more relevant for predicting evolution in interfering populations.
+"
+
+- "We have provided further evidence that even weak purifying selection, when aggregated over a sufficiently large number of sites, can generate strong deviations from neutrality. ""
+
+- "Apart from an overall reduction in polymorphism, the most prominent features of this frequency spectrum include an excess of rare alleles "
+
+- "silent site diversity decays as p=p0*1=Ns, while the shape of the site-frequency spectrum, Qn(i), becomes independent of all underlying parameters."
+
+- " we showed that the reduction in silent site diversity on this chromosome (p=p0*7) is consistent with the parameters Ns<30, NU<300, and NR<0, which fall in the middle of the interference selection regime"
 
@@
 
+%Our results indicate that highly reduced levels of variability in regions of the Drosophila genome that lack crossing over can largely be explained by the effects of selection on nonsynonymous sites subject to purifying selection of the magnitude indicated by recent work.
+It seems likely that similar effects will occur in other genomes or genomic regions with low rates of genetic recombination, such as many bacterial genomes, mitochondrial genomes and true Y or W chromosomes that are not completely degenerate, as well as asexual or highly self-fertilizing species.
+
 The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8).

### 2016-11-12 manuscript/manuscript.tex
intro edits

@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability within the genome.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability.
 On Y chromosomes with suppressed recombination, selection is expected to eliminate neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y diversity.
-Here, we investigate the factors affecting X- and Y-linked neutral polymorphism during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with recently evolved sex chromosomes.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity.
+Here, we investigate the factors affecting relative levels of polymorphism within the genome during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
 Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
-We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male offspring number.
-Instead, using forward simulations, we show that diversity loss on the Y is consistent with interference among a large number (>/= 800Kb) of weakly selected mutations.
-Our results are in agreement with theory on "interference selection" and provide evidence that purifying selection on a large number of linked sites can substantially reduce neutral diversity.
+We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
+Rather, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations, but is over-predicted by standard background selection models.
+Our results are in agreement with theory on "interference selection" and provide evidence that the aggregate effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes (<15mya), our results further imply that Y chromosome degeneration in the early stages may be largely driven by such interference effects rather than by selectively-favoured gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes (\textasciitilde 15MYA), our results imply that Y chromosome degeneration in the early stages may be largely driven by selective interference rather than by selectively-favoured gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the genetic degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
 The independent evolution of these phenomena in such taxonomically distant species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
@@
 One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that  deleterious mutations are more likely to rise to high frequency because of genetic drift \citep{nei1981genetic}.
-Moreover, this loss of recombination on the Y implies that genes on this chromosome will experience strong linkage disequilibrium, making them vulnerable to "Hill-Robertson Interference", in which the genetic associations between selected alleles and their genetic backgrounds cause a reduction in the efficacy of selection \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
-This suggests that Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population (proportional to the product of $N_{e}$ and the mutation rate, $\mu$, at equilibrium), though the extent to which such interference should affect chromosome-wide levels of nucleotide diversity is not known.
-Given that a large number of sites are probably under weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
-Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
+The effective population size ($N_{e}$) of the Y chromosome is therefore predicted to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that  deleterious mutations are more likely to rise to high frequency on this chromosome because of genetic drift \citep{nei1981genetic}, and should.
+In the absence of recombination, genes on the Y chromosome are also predicted to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations.
+Moreover, the build-up of linkage disequilibrium between selected mutations in regions without recombination means that selection at a particular site will "interfere" with selection at linked sites, and previous work has shown that such interference can significantly reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
+These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which linked selection or selective interference should affect chromosome-wide levels of diversity, and the relative importance of these processes in natural populations, is not well-understood.
 
-In species with female-biased reproductive sex ratios or extensive male-male competition, high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, which suggests that inferences about the effects of HRI (including the effects of purifying and positive selection) on the Y need to be distinguished from neutral models that include sex ratio variation as well as sex differences in reproductive success.
+In addition to reduced $N_{e}$ arising from selection, species with female-biased sex ratios or extensive male-male competition, a high variance in male reproductive success is also expected reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection an interference need to be distinguished from these effects.
 Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be therefore obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
@@
 
+- Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
+
+- iven that a large number of sites in the genome likely experience weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
+
+Classical population genetics fails to account for int
+erference between linked mutations, which grows increasingly severe as the density of selected polymorphisms increases.
+
 %This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].

### 2016-11-13 manuscript/manuscript.tex
minor edits

@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic diversity.
 On Y chromosomes with suppressed recombination, selection is expected to eliminate neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
 However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity.
-Here, we investigate the factors affecting relative levels of polymorphism within the genome during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
+Here, we investigate the factors affecting levels of polymorphism within the genome during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
 Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Rather, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations, but is over-predicted by standard background selection models.
+Instead, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations.
 Our results are in agreement with theory on "interference selection" and provide evidence that the aggregate effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
@@
 
-\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plant and animal kingdoms \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}.
+\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plants and animals \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}.
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
@@
 One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-The effective population size ($N_{e}$) of the Y chromosome is therefore predicted to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that  deleterious mutations are more likely to rise to high frequency on this chromosome because of genetic drift \citep{nei1981genetic}, and should.
-In the absence of recombination, genes on the Y chromosome are also predicted to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations.
-Moreover, the build-up of linkage disequilibrium between selected mutations in regions without recombination means that selection at a particular site will "interfere" with selection at linked sites, and previous work has shown that such interference can significantly reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
-These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which linked selection or selective interference should affect chromosome-wide levels of diversity, and the relative importance of these processes in natural populations, is not well-understood.
+Thus, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas genes on the X chromosome should have an $N_{e}$ that is 3/4 that of autosomal genes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that that the equilibrium level of neutral polymorphism, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations (genetic hitchhiking).
+Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome, and previous work has shown that such "selective interference" can significantly reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
+These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which background selection, genetic hitchhiking, or selective interference should affect chromosome-wide levels of diversity, and the relative importance of these processes in natural populations, is not well-understood.
 
-In addition to reduced $N_{e}$ arising from selection, species with female-biased sex ratios or extensive male-male competition, a high variance in male reproductive success is also expected reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection an interference need to be distinguished from these effects.
-Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can be therefore obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
+In addition to reduced $N_{e}$ arising from selection, species with female-biased sex ratios or extensive male-male competition, a high variance in male reproductive success is also predicted reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection and interference need to be distinguished from these effects.
+Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can therefore be obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
 For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans have suggested examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, though the inflated X/A ratio has also been attributed to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic})
+Based on such comparisons, studies in humans have suggested that the inflated X/A ratio is attributable to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic})
 
-Despite widespread interest in determining the evolutionary factors affecting neutral diversity on both X and Y chromosomes \citep{ellegren2011,bachtrog2013NRG}, we still know very little about the influence of demography or selection on patterns of sex-linked diversity in more recently evolved sex chromosomes, such as in dioecious plants.
-This information is important, however, as it highlights the relevant time scales over which effects such as linked selection are likely to be important.
-In humans, estimates of Y-linked diversity are considerably lower than predicted under models of neutral evolution, with simulations showing that current levels of Y-diversity are consistent with a combination of very low $N_{m}$ and a history of strong purifying selection \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether such strong effects of purifying selection on diversity should also apply to sex chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of most dioecious plants \citep{charlesworth2015plant}).
-Y-linked diversity loss due to purifying selection might on the one hand be expected to be lower in such systems due to recombination having been suppressed for a shorter length of time.
-Contrary to this intuition, however, simulations have shown that two key types of HRI - background selection and selective sweeps - both have strongest effects during the earliest stages of sex chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}.
+Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of either reproductive variance or selection in determining level of diversity on more recently evolved sex chromosomes.
+The time scales over which these effects are likely to be important is therefore not well understood.
+In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations have shown that current levels of Y-diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether such strong effects of purifying selection on diversity should also apply to sex chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of plants \citep{charlesworth2015plant}).
+Y-linked diversity loss due to purifying selection might be expected to be lower in such systems due to recombination having been suppressed for a shorter period of time.
+However, simulations have shown that background selection and genetic hitchhiking have the strongest effects during the earliest stages of sex chromosome evolution, before the Y has lost the majority of its genes \citep{bachtrog2008temporal}.
 Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosome systems, which have found evidence for rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-In these cases, if Y-degeneration involved widespread Y-linked gene inactivation (silencing), as predicted by \citep{orr1998adaptive}, and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then linked selection might in fact play a less important role than theory would predict, with the early stages of Y-degeneration instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
+Despite this, the large number of sites between which recombination is suppressed on a recently evolved Y chromosome means that classic predictions from background selection theory break down (refs), and it is ,...
+
+%In these cases, if Y-degeneration involved widespread gene silencing, as predicted by \citep{orr1998adaptive} and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then selection might in fact play a less important role than theory would predict.
+%with the early stages of Y-degeneration instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
 

### 2016-11-13 manuscript/manuscript.tex
full edits up to methods: simulations

@@
 Instead, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations.
-Our results are in agreement with theory on "interference selection" and provide evidence that the aggregate effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
+Our results are in agreement with theory on "interference selection", and provide evidence that the effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes (\textasciitilde 15MYA), our results imply that Y chromosome degeneration in the early stages may be largely driven by selective interference rather than by selectively-favoured gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes (\textasciitilde 15MYA), our results imply that Y chromosome degeneration in the early stages may be largely driven by interference rather than positive selection for gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 
-\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plants and animals \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}.
+\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plants and animals \citep{westergaard1958,ohno1967,bull1983,charlesworth1991,charlesworth2015plant}.
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
@@
 The independent evolution of these phenomena in such taxonomically distant species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
-% and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.
 
 One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-Thus, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas genes on the X chromosome should have an $N_{e}$ that is 3/4 that of autosomal genes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that that the equilibrium level of neutral polymorphism, which is proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+Thus, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas the  $N_{e}$ for genes on the X chromosome should be 3/4 that of autosomal genes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that that the equilibrium level of neutral polymorphism - proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ - should be lower for Y-linked genes than for their X-linked counterparts.
 In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations (genetic hitchhiking).
-Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome, and previous work has shown that such "selective interference" can significantly reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
-These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which background selection, genetic hitchhiking, or selective interference should affect chromosome-wide levels of diversity, and the relative importance of these processes in natural populations, is not well-understood.
+Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome such that selection at a focal site may "interfere" with selection at sites with which it is linked.
+Originally developed by Hill & Robertson (1966), a large body of work has now shown that such "selective interference" can substantially reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
+These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which background selection, genetic hitchhiking, or selective interference have affected chromosome-wide levels of diversity in natural populations, and the relative importance of these processes, is not well-understood.
 
-In addition to reduced $N_{e}$ arising from selection, species with female-biased sex ratios or extensive male-male competition, a high variance in male reproductive success is also predicted reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection and interference need to be distinguished from these effects.
-Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this effect can therefore be obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
+In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects positive of purifying selection need to be distinguished from these neutral processes.
+Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
 For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans have suggested that the inflated X/A ratio is attributable to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic})
+Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is attributable to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic}).
 
-Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of either reproductive variance or selection in determining level of diversity on more recently evolved sex chromosomes.
-The time scales over which these effects are likely to be important is therefore not well understood.
-In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations have shown that current levels of Y-diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether such strong effects of purifying selection on diversity should also apply to sex chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of plants \citep{charlesworth2015plant}).
-Y-linked diversity loss due to purifying selection might be expected to be lower in such systems due to recombination having been suppressed for a shorter period of time.
-However, simulations have shown that background selection and genetic hitchhiking have the strongest effects during the earliest stages of sex chromosome evolution, before the Y has lost the majority of its genes \citep{bachtrog2008temporal}.
-Indeed, this is consistent with studies of Y-degeneration in young plant sex chromosome systems, which have found evidence for rapid genetic degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}.
-Despite this, the large number of sites between which recombination is suppressed on a recently evolved Y chromosome means that classic predictions from background selection theory break down (refs), and it is ,...
-
-%In these cases, if Y-degeneration involved widespread gene silencing, as predicted by \citep{orr1998adaptive} and discovered on the neo-Y in \textit{Drosophila mrianada} \citep{zhou2012chromosome}, then selection might in fact play a less important role than theory would predict.
-%with the early stages of Y-degeneration instead being primarily driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG}.
-
-%The strength of this effect is dependent on the rate of recombination between a selected locus and a neighbouring region; essentially, the effect becomes more pronounced with less recombination because of linkage disequilibrium between adjacent loci.
-
-%On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimized rapidly even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes (cite backtrog NRG review).
-
-%because recombination suppression was relatively recent on plant Y chromosomes, it is an open question whether weak-HRI has had enough time to have a large effect.
-Although selective interference (indlucing BGS and Selective Sweeps) is expected to have the greatest effects during the early stages of Y-chromosome evolution \citep{bachtrog2008temporal} the R.
-hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.
-
-%Background selection (purifying selection against deleter- ious alleles) [27] and selective sweeps (positive selection for beneficial variants) [28] both have the effect of reducing nucleotide diversity at linked sites.
-%as modeled by \citep{kimura1964number} and \citep{nomura2002effective}.
-
-%evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection.
-The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference).
-In all of these processes, linkage  disequilibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.
-
-%Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, s
-
-%discussion point?:
-%Moreover,
+Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of either sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes.
+The time scales over which these different effects are likely to be important is therefore not well understood.
+In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations suggest that levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether purifying selection might have such strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of plants \citep{charlesworth2015plant}).
+Y-linked diversity loss might on the one hand be expected to be lower in younger systems due to a shorter history of recombination suppression.
+On the other hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
+Moreover, work by \citep{KaiserCharlesworth} and \citep{good2014genetic} have provided evidence that even weak purifying selection, if acting non-independently across large number of sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases.
+Given that a large number of sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is a good candidate model for exploring the evolutionary dynamics of young plant sex chromosomes.
 
@@
 hastatulus }(Polygonaceae).
-This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than the well-studied mammalian sex chromosomes \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the unique occurrence of an intraspecific polymorphism in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ('chromosomal races') \citep{smith1963mechanism}.
-The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}.
+This species is a dioecious annual with heteromorphic X and Y chromosomes that evolved approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
+\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an intraspecific polymorphism in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ("chromosomal races") \citep{smith1963mechanism}.
+The $XY_{1}XY_{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
 Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone significant gene loss and functional deterioration \citep{hough2014}.
@@
 
-Of particular relevance to our study, \textit{R.
+Of particular importance for our study, \textit{R.
 hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Female-biased sex ratios in dioecious plants with heteromorphic X and Y chromosomes are not uncommon (\X refs) \citep{hough2013evolutionarily}, and their occurrence in \textit{R.
+Female-biased sex ratios are not uncommon in dioecious plants with heteromorphic X and Y chromosomes \citep{field2013comparative,hough2013evolutionarily}, and their occurrence in \textit{R.
 hastatulus} provides an excellent opportunity to study both the demographic and selective factors contributing to sex-linked variability.
-Using estimates of neutral polymorphism on sex chromosomes and autosomes, together with forward population genetic simulations, our study tests whether sex chromosome evolution in the early stages is driven primarily by neutral processes, or whether linked selection has played a significant role.
 
@@
 For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
-To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1).
-We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each.
+To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2).
+We obtained 95\% confidence intervals for X/A and Y/A ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A on each iteration for 20000 replicates each.
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes.
-Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping.
+Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the estimate Y/A may be wider than implied by bootstrapping results.
 To address this, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
@@
 
-%from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race.
-%We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males.
-As no significant difference was observed (\X Figure?
-Table?), we report only results from females.
-
-%Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity.
-To do this, we constructed neighbor-joining trees of sex-linked sequences from all populations in our study.
-The analysis was conducted on an alignment of the X- and Y-linked genes from \textit{R.
-hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R.
-bucephalophorus} used to root the tree (Figure S2).
-We used the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}.
-The inferred trees revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2).
-In light of this evidence for significant population substructure, and to consider between-population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately.
-For model-based analyses, we focused on the population sample from the Texas XY race, given that this reflects the presumed ancestral karyotype and has the largest sample size.
-
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that considers these two effects \citep{kimura1964number}.
-In particular, we calculated the effect of variance in reproductive success on $N_{e}$ as:
+To test whether the levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our diversity estimates to predictions from a model that incorporates variance of progeny number for males and females, and gives the predicted $N_{e}$ for males and females when this variance is either smaller or larger than Poisson \citep{kimura1964number, hedrick2011genetics}.
+As we were primarily interested in the possible influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which such a model might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring, and calculated the male $N_{e}$ as a function of variance in offspring number as:
 
@@
 
-where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
-Given the expected $N_{m}$ and $N_{f}$ for varying levels $V_{k_{m}}$, the expected $N_{e}$ for autosomes and sex chromosomes was then calculated using equations from \citep{wright1931evolution}:
+where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males and $\overline{k}$ is the total mean number of progeny of males and females \citep{kimura1964number}.
+Given the expected values for $N_e_{m}$ across a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding expectation for the $N_{e}$ of autosomes and sex chromosomes at equilibrium were calculated as:
 
@@
 
-Note that with equal sex ratios, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-Figure 1 shows these expected X/A and Y/A ratios as a function of the sex ratio.
-We tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) using the sex ratio of 0.6 as estimated in \textit{R.
-hastatulus}.
-
-%To consider the combined effects on diversity of our empirically estimated female-biased sex ratio ($(N_{f}$/$N_{m}+N_{f})$=$0.6$ together with high variance in male reproductive success, we calculated the nuetrally-expected X/A and Y/A ratios under a 0.6 sex ratio across a range of values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3).
-%\begin{equation}
-%N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}(1+C^2_{f})+N_{f}(1+C^2_{m})}\label{eq:NeA}
-%\end{equation}
-
-%where $N_{m}$ and $N_{f}$ are effective population sizes for males and females, respectively, and $C^2_{i}$ is the coefficient of variation in mating success for the $i$th sex, defined as the ratio of the standard deviation to the mean number of mates $\sigma_{i}/\mu_{d{i}}$.
-Similarly, for X-linked loci, where females contain two-thirds and males one-third of the alleles:
-
-%\begin{equation}
-%N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}(1+C^2_{f})+2N_{f}(1+C^2_{m})}\label{eq:NeX}
-%\end{equation}
-
-%With no variation in mating success in either sex ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 1 and 2 reduced to the familiar expressions from \citep{wright1931evolution}:
-
-%\begin{equation}
-%\label{eq:NeY}
-%\end{equation}
-
-%Because the $N_{e}$ for the Y chromosome depends on the male effective population size, $N_{m}$, the effect of variance in reproductive success on $N_{Y}$ can be modeled using the approach of Crow and Kimura (refs; and see ) and $N_{m}$ is given by:
-
-%\begin{equation}
-%N_{e{X}} = \frac{N(k-1)}{k-1}
-%\end{equation}
-
-%\begin{equation}
-%\label{eq:NeY}
-%\end{equation}
-
-%Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R.
-hastatulus} sex ratio (N_{f}/$N_{m}+N_{f})=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}.
-Sex-specific variance in reproductive success can lead to departures from this prediction, and so we also evaluated whether any ratio of male- to female- effective population sizes could explain our observed combination of X/A and Y/A diversity.
+following \citep{wright1931evolution}.
+Note that with equal sex ratios, and Poisson distributed offspring numbers ($V_{k_{m}} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
+Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) assuming the estimated \textit{R.
+hastatulus} sex ratio of 0.6 \citep{pickup2013influence}.
 
@@
 \includegraphics[width=\linewidth]{figure1.jpg}
-\caption{The relation between relative effective population size at equilibrium and sex ratio bias for genes on autosomes (\textbf{A}), the X chromosome (\textbf{B}), and the normalized X/A and Y/A ratios (\textbf{C}).
-The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})$, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$'s' for sex chromosomes and autosomes are given in equations 2-4.
-Dotted curves show predictions when both males and females produce Poisson-distributed offspring numbers ($\overline{k}=2$).
-Solid curves correspond to increasing levels of variance in male reproductive success, ranging from Poisson/2 (top solid curve in panels A and B) to 3*Poisson (bottom most curve).
-Assuming $\theta=4N_{e}\mu$ and an equal neutral mutation rates among genes, these predicted $N_{e}$'s were used to generate our null neutral predictions for the X/A and Y/A ratios of diversity.
+\caption{The relation between effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and the X chromosome (\textbf{B}), and corresponding normalized X/A and Y/A ratios (\textbf{C}).
+The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})$, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$.
+Dotted curves show equilibrium predictions when both males and females produce Poisson-distributed offspring numbers and ($V_{k}$=$\overline{k}=2$).
+Solid curves correspond to increasing values for $V_{k_{m}}$, the variance in male offspring number, ranging from Poisson/2 (top solid curve in panels A and B) to 3*Poisson (bottom most curve).
+For (\textbf{C}) we assumed equal neutral mutation rates among genes and calculated the expected neutral diversity as $\theta=4N_{e}\mu$.
 }

### 2016-11-13 manuscript/manuscript.tex
final edits up to results, added new ms version for Spencer and Stephen

@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether the levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our diversity estimates to predictions from a model that incorporates variance of progeny number for males and females, and gives the predicted $N_{e}$ for males and females when this variance is either smaller or larger than Poisson \citep{kimura1964number, hedrick2011genetics}.
-As we were primarily interested in the possible influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which such a model might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring, and calculated the male $N_{e}$ as a function of variance in offspring number as:
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model of the expected $N_{e}$ of males and females when the variance in offpring number is either smaller or larger than Poisson \citep{kimura1964number, hedrick2011genetics}.
+As we were primarily interested in the possible influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which such a model might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring, and calculated the male $N_{e}$ as:
 
@@
 
-where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males and $\overline{k}$ is the total mean number of progeny of males and females \citep{kimura1964number}.
-Given the expected values for $N_e_{m}$ across a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding expectation for the $N_{e}$ of autosomes and sex chromosomes at equilibrium were calculated as:
+where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
+Given expected values of $N_e_{m}$ for a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding expectation for the $N_{e}$ of autosomes and sex chromosomes at equilibrium were calculated as:
 
@@
 following \citep{wright1931evolution}.
-Note that with equal sex ratios, and Poisson distributed offspring numbers ($V_{k_{m}} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
+Note that with equal sex ratios $N_{f} = N_{m}$, and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
 Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) assuming the estimated \textit{R.
@@
 \subsection*{Simulations of purifying selection}
-To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a non-recombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations.
-This approach was implemented rather than using analytical predictions of background selection for a few reasons.
-First,...
-because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
-
-...there is no general analytical treatment describing the dynamics of interference among many alleles under selection and drift, with variable degrees of selection and linkage.
-Thus, forward computer simulations are still fundamental to the study of realistic situations.
-- dfe for the y is probably still fairly similar to that for the x -- shared genes
-
-- human Y: genes on y are not on X
-
-- 45331 = synsites in observed data after filtering
-  - summary stats phased texas autospy
-
-- doesnt matter about coding/non-coding in non-recombining regions? all of those neutral sites are linked
-
-To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (\X Keightley and Eyre-Walker ref).
-This program fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism from the population genetic analyses above.
-We used this estimated gamma distribution in our forward simulations.
-To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of $\pi_{synonymous}$ .
-To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$, and we sampled 6 haploid chromosomes per simulation.
-Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information).
+To study the effects of purifying selection on expected levels of Y-chromosome diversity, we used a similar approach to \citep{Wilsonsayres2014} and conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
+We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism.
+We used this estimated gamma distribution to parameterize the simulations, initializing each with our estimated $\theta$ from autosomal genes, but appropriately adjusted to reflect the expected reduction in Y-linked $N_{e}$ for a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$.
+To match our sample size and the number of synonymous sites sample from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and simulated sequences contained 45,331bp of linked neutral sequence from which we calculated diversity.
+To examine the expected reduction in diversity relative to neutrality ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), and to estimate this parameter from our data, we ran simulations over a range of values of $L$ (up to a maximum of $5x10^{6}$) (Figure 3A).
+For each simulation, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}_{simulated}$, matched our empirical estimate, $\pi_{s}_{observed}$ (Figure 3B).
 
@@
 
+This approach was implemented rather than using analytical predictions of background selection for a few reasons.
+First,...
+because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
+
 discussion points to make:
@@
 erference between linked mutations, which grows increasingly severe as the density of selected polymorphisms increases.
+
+...there is no general analytical treatment describing the dynamics of interference among many alleles under selection and drift, with variable degrees of selection and linkage.
+Thus, forward computer simulations are still fundamental to the study of realistic situations.
+- dfe for the y is probably still fairly similar to that for the x -- shared genes
+
+- human Y: genes on y are not on X
+
+- 45331 = synsites in observed data after filtering
+  - summary stats phased texas autospy
+
+- doesnt matter about coding/non-coding in non-recombining regions? all of those neutral sites are linked
 

### 2016-11-13 manuscript/Ydiversity_nov12.tex
new ms version for spencer and stephen

--- 2016-11-13 manuscript/manuscript.tex
@@
 \subsection*{Extensive loss of Y-chromosome diversity}
-Our analysis reveals that diversity on the \textit{R.
-hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, ....which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
-We also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
-Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
-Conceivably, such low diversity on the Y could arise from a low Y-linked mutation rate, or a lower mutation rate in males compared to females.
-However, these possibilities are unlikely because the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are not significantly different \citep{hough2014}.
-Similarly, our estimates of weighted average synonymous substitution rate between \textit{R.
-hastatulus} and the outgroup \textit{R.
-bucephalophorus} reveal similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
 
-%After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig).
-Surprisingly, this analysis revealed that the Y chromosome from this race appears to be paraphyletic, with XYY samples from Florida being more closely related to the TX race than either is to samples from South Carolina (suppfig).
-Once we examine diversity for these two subclades separately, a strong loss of diversity on the Y chromosome is once again apparent (Table 1).
-
-%Although our sampling of each \textit{R.
-hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a secondarily-derived $XY_{1}Y_{2}$ sub-clade.
-As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the Florida sub-clade inferred here likely originated through hybridization between a female from the SC clade harboring the X-autosome fusion, and a male from the XY Texas race.
-
-%\begin{table}[htb]
-%\centering
-%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
-%\begin{tabular}{ccccccccc}
-%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
-%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
-%\midrule
-%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
-%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
-%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
-%\addlinespace
-
-%\bottomrule
-%\end{tabular}
-%\end{table}
-
-%Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
-Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}.
-Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race.
-If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can drive the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
-However, this would require extensive recombination suppression between the fused and unfused X chromosomes, and it is also possible that sex-specific demographic history has driven these patterns.
-It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
-
-In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 50 fold reduction compared to the mean $\theta_{Aut}$.
-We next consider several possible models - neutral and selective - that might explain this reduction.
-
-\subsection*{Female biased sex ratio and high variance in male offspring number}
-
-The occurrence of female-biased sex ratios in \textit{R.
-hastatulus} is expected to lower Y diversity through lowering male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome.
-In addition, the $N_{e}_{Y}$ might be further reduced due to high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
-hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there is reason to believe that there could be strong competition among males.
-
-%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
+\subsection*{Female biased sex ratio and variance in male offspring number}
 
@@
 
-In common with most flowering plants we do not have marker-based estimates of the variance in male reproductive success in \textit{R.
-hastatulus}.
-However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could jointly explain the level of Y/A diversity that we observed (see Methods).
-Conditioning on estimates of sex ratio bias in \textit{R.
-hastatulus}, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
-This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), and remains significant even if we consider an upper bound estimate obtained from our likelihood based estimated of the confidence interval.
-This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.
-
-Assuming that there is extensive variance in male reproductive success (\X), the predicted ratios of Y/A diversity (\X supp) were also significantly higher than our estimates for the empirically estimated sex ratio (\X table) .
-We also find, however, that purely neutral models in which the sex ratio was highly female-biased (\X), with level of variance in male reproductive success on the order of (\X), predicted a Y/A ratio that could not be rejected (\X).
-However, because a highly females biased sex ratio is expected to increase the X/A ratio as well, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure).
-Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity.
-
-\subsection*{Background Selection and Selective Sweeps}
-
-This approach was implemented rather than using analytical predictions of background selection for a few reasons.
-First,...
-because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
-
-discussion points to make:
-- Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-
-BGS predictions:
--"Traditionally, the term ‘‘background selection’’ is used to refer both to the general effects of purifying selection on linked neutral diversity as well as to the limiting behavior that emerges when Ns-->inf.
-"
-- "existing theory struggles to predict genetic diversity when many sites experience selection at the same time, which limits our ability to interpret variation in DNA sequence data.". In particular,
-
-- "For non-crossover regions, both with and without gene conversion, there is a rapid initial decline of B with L, but B levels off at a value of  0.015 for > 640 000 sites.
-B values predicted by the BGS model decrease log-linearly with L"
-
-Very relevant remarks From Good et al.:
-
-- "Here, we have shown that simple behavior emerges in the limit of widespread interference.
-When fitness variation is composed of many individual mutations, the magnitudes and signs of their fitness effects are relatively unimportant.
-Instead, molecular evolution is controlled by the variance in fitness within the population over some effectively asexual segment of the genome" " In other words, we cannot conclude that interference is negligible just because ‘‘Nes’’, as inferred from data, is larger than one."
-
-- estimates of "Nes" ignore linkage by fiat under the assumption that sites evolve independently.
-But these estimates become unreliable precisely when small- and intermediate-effect mutations are most common
-
-- "Individual fitness effects may play a central role in single-site models, but we have shown that global properties like the variance in fitness and the corresponding linkage scale are more relevant for predicting evolution in interfering populations.
-"
-
-- "We have provided further evidence that even weak purifying selection, when aggregated over a sufficiently large number of sites, can generate strong deviations from neutrality. ""
-
-- "Apart from an overall reduction in polymorphism, the most prominent features of this frequency spectrum include an excess of rare alleles "
-
-- "silent site diversity decays as p=p0*1=Ns, while the shape of the site-frequency spectrum, Qn(i), becomes independent of all underlying parameters."
-
-- " we showed that the reduction in silent site diversity on this chromosome (p=p0*7) is consistent with the parameters Ns<30, NU<300, and NR<0, which fall in the middle of the interference selection regime"
-
-- Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
-
-- iven that a large number of sites in the genome likely experience weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
-
-Classical population genetics fails to account for int
-erference between linked mutations, which grows increasingly severe as the density of selected polymorphisms increases.
-
-...there is no general analytical treatment describing the dynamics of interference among many alleles under selection and drift, with variable degrees of selection and linkage.
-Thus, forward computer simulations are still fundamental to the study of realistic situations.
-- dfe for the y is probably still fairly similar to that for the x -- shared genes
-
-- human Y: genes on y are not on X
-
-- 45331 = synsites in observed data after filtering
-  - summary stats phased texas autospy
-
-- doesnt matter about coding/non-coding in non-recombining regions? all of those neutral sites are linked
-
-%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
-
-%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
-
-%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
-
-%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
-
-%The agreement is reasonably good given the uncertainties involved.
-
-%silent site diversity for 22 genes on the D. miranda neo-Y is
-%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
-
-% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
-So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
-%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
-
-%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
-
-%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
-
-%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
-
-%The agreement is reasonably good given the uncertainties involved.
-
-%silent site diversity for 22 genes on the D. miranda neo-Y is
-%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
-
-% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
-So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
-
-%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}
-
-%A predicted consequence of a reduction in Ne is an accelerated rate of ¢xation of slightly deleterious substitu- tions, and a reduction in the rate of ¢xation of slightly advantageous mutations (Kimura 1983).
-Thus we might expect an increased rate of amino-acid replacement rela- tive to silent substitutions on evolving Y chromosomes (provided that su¤cient time has elapsed since the origin of the sex-chromosome system), and more ¢xations of non-preferred versus preferred synonymous substitutions at silent sites in exons (Charlesworth 1996a).
-
-%It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
-Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
-However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
-hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
-
-% A further interesting aspect of the results from the S.
-latifolia Y chromosome is that they do not readily ¢t a recent selective sweep.
-Such an event would eliminate all diversity, so that existing SLY1 polymorphisms must have accumulated since the selective sweep and should thus mostly be at low frequencies (see table 1).
-However, all six SLY1 polymorphic sites found are non-singletons, and no signi¢cant deviations from neutrality are detectable by standard test statistics (Fu & Li 1993).
-However, this does not take into account the probable population subdivision in this species.
-Although a selective sweep eliminating variability throughout the worldwide population of this species seems unlikely, e¡ects of selective sweeps within local populations cannot be excluded.
-
-%because the lack of recombination means that all genes on the Yof any given species share a common genealogy, which limits the amount of information that can be obtained from data on within-species variability.
-
-%This raises the further issue of whether Y-chromosome degeneration is driven mainly by the accumulation of deleterious mutations on the Y or is due to processes turning down the expression of genes on an evolving Y.
-
-%We have focused on the ¢rst of these possibilities, but it might well not completely explain Y degeneration.
-If mutations of relatively minor ¢tness e¡ects, with Nes not much bigger than 1, become ¢xed more rapidly on a degenerating Ychromosome than on the X, selection pres- sure would develop to reduce the transcription of Y genes, provided that this was compensated for in the hetero- gametic sex by an increased activity of their counterparts on the X.
+\subsection*{Purifying selection}
 
 \section*{Conclusions}
-
-%Our results indicate that highly reduced levels of variability in regions of the Drosophila genome that lack crossing over can largely be explained by the effects of selection on nonsynonymous sites subject to purifying selection of the magnitude indicated by recent work.
-It seems likely that similar effects will occur in other genomes or genomic regions with low rates of genetic recombination, such as many bacterial genomes, mitochondrial genomes and true Y or W chromosomes that are not completely degenerate, as well as asexual or highly self-fertilizing species.
-
-The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8).
-The extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the ....
-a large genomic regions that lack recombination.
 

### 2016-11-13 manuscript/Ydiversity_nov12.tex
minor edits, start on results

@@
 \subsection*{Simulations of purifying selection}
-To study the effects of purifying selection on expected levels of Y-chromosome diversity, we used a similar approach to \citep{Wilsonsayres2014} and conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
+To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
 We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism.
@@
 To match our sample size and the number of synonymous sites sample from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and simulated sequences contained 45,331bp of linked neutral sequence from which we calculated diversity.
-To examine the expected reduction in diversity relative to neutrality ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), and to estimate this parameter from our data, we ran simulations over a range of values of $L$ (up to a maximum of $5x10^{6}$) (Figure 3A).
-For each simulation, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}_{simulated}$, matched our empirical estimate, $\pi_{s}_{observed}$ (Figure 3B).
+To examine the expected reduction in diversity ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), and to estimate this parameter from our data, we ran simulations over a range of values of $L$ (up to a maximum of $5x10^{6}$) (Figure 3A).
+For each such simulation, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}_{simulated}$, was not significantly different from our empirical estimate, $\pi_{s}_{observed}$ (Figure 3B).
 
 \section*{Results and Discussion}
-
 \subsection*{Extensive loss of Y-chromosome diversity}
+%Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
+hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model.
+In particular, neutral diversity on Y-linked genes $\theta_{Y}$ is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome, and \textasciitilde 2.1\% of the value of autosomal genes.
+Talking 1/4 of mean autosomal diversity as the equilibrium expectation under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of \textasciitilde 93\% ($\pi/\pi_{0}$).
+By comparing X and Y diversity estimates to that estimated on autosomal genes, our results indicate that the difference in diversity between X and Y homologues is not due to an elevation of diversity on the X chromosome, but rather a Y-specific reduction.
+Our results to suggest a slight elevation in X-linked diversity relative to autosomes ($X/A=$), but this difference is not significant different from expectations under equal sex ratios (P), or the empirically-estimated ratio of 0.6 (P).
+Although our estimates of diversity have not beed normalized by divergence, previous estimates of the average synonymous substitution rate, weighted by the number of synonymous sites, between \textit{R.
+hastatulus} and the non-dioecious outgroup \textit{R.
+bucephalophorus} revealed similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
+It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
 

### 2016-11-13 manuscript/manuscript.tex
minor edits, start on results

@@
 \subsection*{Extensive loss of Y-chromosome diversity}
-Our analysis reveals that diversity on the \textit{R.
-hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, ....which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001).
-We also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1).
-Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction.
-Conceivably, such low diversity on the Y could arise from a low Y-linked mutation rate, or a lower mutation rate in males compared to females.
-However, these possibilities are unlikely because the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are not significantly different \citep{hough2014}.
-Similarly, our estimates of weighted average synonymous substitution rate between \textit{R.
-hastatulus} and the outgroup \textit{R.
-bucephalophorus} reveal similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
+Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
+hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model.
+In particular, neutral diversity on Y-linked genes $\theta_{Y}$ is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome, and \textasciitilde 2.1\% of the value of autosomal genes.
+Talking 1/4 of mean autosomal diversity as the equilibrium expectation under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of \textasciitilde 93\% ($\pi/\pi_{0}$).
+By comparing X and Y diversity estimates to that estimated on autosomal genes, our results indicate that the difference in diversity between X and Y homologues is not due to an elevation of diversity on the X chromosome, but rather a Y-specific reduction.
+Our results to suggest a slight elevation in X-linked diversity relative to autosomes ($X/A=$), but this difference is not significant different from expectations under equal sex ratios (P), or the empirically-estimated ratio of 0.6 (P).
+Although our estimates of diversity have not beed normalized by divergence, previous estimates of the average synonymous substitution rate, weighted by the number of synonymous sites, between \textit{R.
+hastatulus} and the non-dioecious outgroup \textit{R.
+bucephalophorus} revealed similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
+It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
 

### 2016-11-15 manuscript/Ydiversity_nov12.tex
edits

--- 2016-11-13 manuscript/Ydiversity_nov12.tex
@@
 For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
-The independent evolution of these phenomena in such taxonomically distant species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
+The independent evolution of these phenomena in such taxonomically distant groups suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
 
@@
 In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations (genetic hitchhiking).
-Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome such that selection at a focal site may "interfere" with selection at sites with which it is linked.
-Originally developed by Hill & Robertson (1966), a large body of work has now shown that such "selective interference" can substantially reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
+Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome, such that selection at a focal site may "interfere" with selection at sites with which it is linked \citep{hill1966HReffect}.
+A large body of work has now shown that in theory such "selective interference" effects can substantially reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}.
 These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which background selection, genetic hitchhiking, or selective interference have affected chromosome-wide levels of diversity in natural populations, and the relative importance of these processes, is not well-understood.
@@
 For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is attributable to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic}).
+Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is due to a historical excess of breeding females compared to males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic}).
 
-Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of either sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes.
+Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes.
 The time scales over which these different effects are likely to be important is therefore not well understood.
 In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations suggest that levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether purifying selection might have such strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of plants \citep{charlesworth2015plant}).
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA) \citep{lahn1999four,ross2005dna}, is not immediately obvious whether purifying selection might have such strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last \textasciitilde 20 MYA in the case of plants \citep{charlesworth2015plant}).
 Y-linked diversity loss might on the one hand be expected to be lower in younger systems due to a shorter history of recombination suppression.
 On the other hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
-Moreover, work by \citep{KaiserCharlesworth} and \citep{good2014genetic} have provided evidence that even weak purifying selection, if acting non-independently across large number of sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases.
+Moreover, there is clear evidence from simulation studies that even weak purifying selection, if acting non-independently across large number of sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
 Given that a large number of sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is a good candidate model for exploring the evolutionary dynamics of young plant sex chromosomes.
@@
 hastatulus }(Polygonaceae).
-This species is a dioecious annual with heteromorphic X and Y chromosomes that evolved approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an intraspecific polymorphism in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ("chromosomal races") \citep{smith1963mechanism}.
+This species is a dioecious annual with heteromorphic X and Y chromosomes that evolved \textasciitilde 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
+\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphic in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ( so-called "chromosomal races") \citep{smith1963mechanism}.
 The $XY_{1}XY_{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone significant gene loss and functional deterioration \citep{hough2014}.
+Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone gene loss and functional deterioration \citep{hough2014}.
 Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, we focus only on the ancestral Y chromosome, which occurs in both sex chromosome races.
@@
 Of particular importance for our study, \textit{R.
-hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
+hastatulus} populations have also been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
 Female-biased sex ratios are not uncommon in dioecious plants with heteromorphic X and Y chromosomes \citep{field2013comparative,hough2013evolutionarily}, and their occurrence in \textit{R.
-hastatulus} provides an excellent opportunity to study both the demographic and selective factors contributing to sex-linked variability.
+hastatulus} provides an excellent opportunity to investigate both the demographic and selective factors contributing to sex-linked variability.
 
@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model of the expected $N_{e}$ of males and females when the variance in offpring number is either smaller or larger than Poisson \citep{kimura1964number, hedrick2011genetics}.
-As we were primarily interested in the possible influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which such a model might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring, and calculated the male $N_{e}$ as:
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model for the expected $N_{e}$ of males and females when the variance in offpring number is smaller or larger than the standardly assumed Poisson \citep{kimura1964number, hedrick2011genetics}.
+As we were primarily interested in the influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which this effect might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring numbers, and calculated male $N_{e}$ as:
 
@@
 where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
-Given expected values of $N_e_{m}$ for a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding expectation for the $N_{e}$ of autosomes and sex chromosomes at equilibrium were calculated as:
+Given expected values of $N_e_{m}$ for a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding neutral expectation for autosomal and sex-linked $N_{e}$ at equilibrium were calculated as:
 
 \begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}} \label{eq:Ne}
+N_{e_{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}} \label{eq:Ne}
 \end{equation}
@@
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}} \label{eq:NeX}
+N_{e_{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}} \label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e{Y}} = \frac{N_{m}}{2} \label{eq:NeY}
+N_{e_{Y}} = \frac{N_{m}}{2} \label{eq:NeY}
 \end{equation}
 
-following \citep{wright1931evolution}.
-Note that with equal sex ratios $N_{f} = N_{m}$, and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) assuming the estimated \textit{R.
-hastatulus} sex ratio of 0.6 \citep{pickup2013influence}.
+respectively, following \citep{wright1931evolution}.
+Note that with equal sex ratios $N_{f} = N_{m}$, and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ and in the standard model.
+Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our observed Y/A and X/A diversity ratios to model predictions (shown in Figure 2) assuming the empirically-estimated \textit{R.
+hastatulus} population sex ratio of 0.6 \citep{pickup2013influence}.
 
@@
 \subsection*{Simulations of purifying selection}
-To study the effects of purifying selection on expected levels of Y-chromosome diversity, we used a similar approach to \citep{Wilsonsayres2014} and conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
+To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
 We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism.
-We used this estimated gamma distribution to parameterize the simulations, initializing each with our estimated $\theta$ from autosomal genes, but appropriately adjusted to reflect the expected reduction in Y-linked $N_{e}$ for a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$.
-To match our sample size and the number of synonymous sites sample from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and simulated sequences contained 45,331bp of linked neutral sequence from which we calculated diversity.
-To examine the expected reduction in diversity relative to neutrality ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), and to estimate this parameter from our data, we ran simulations over a range of values of $L$ (up to a maximum of $5x10^{6}$) (Figure 3A).
-For each simulation, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}_{simulated}$, matched our empirical estimate, $\pi_{s}_{observed}$ (Figure 3B).
+We then used this estimated gamma distribution to parameterize the simulations, initializing each with our estimated $\theta$ from autosomal genes, but appropriately adjusted to reflect the expected reduction in $N_{e}_{Y}$ for a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$.
+To match our sample size and the number of synonymous sites sample from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and contained 45,331bp of linked neutral sequence from which we calculated silent site diversity.
+To examine the expected reduction in diversity ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), we ran simulations over a range of values of $L$, up to a maximum of 5x$10^{6}$ (Figure 3A).
+To estimate $L$ from our data, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi__{simulated}$, was not significantly different from our empirical estimate, $\pi__{observed}$ (Figure 3B).
 
 \section*{Results and Discussion}
+\subsection*{Extensive loss of Y-chromosome diversity}
+Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
+hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model.
+In particular, neutral diversity on Y-linked genes $\theta_{Y}$ is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
+Talking 1/4 of the mean autosomal diversity as the equilibrium expectation under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of \textasciitilde 93\%.
+Note that by comparing X and Y diversity to autosomal diversity, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but rather a Y-specific reduction.
+Our results do suggest a slight elevation in X-linked diversity relative to the neutral prediction of X/A=0.75, but this difference is not significantly different from the expectation for the empirically-estimated ratio of 0.6.
+Although our estimates of diversity have not beed normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R.
+hastatulus} and the non-dioecious outgroup \textit{R.
+bucephalophorus} is not significantly different for sex-linked (0.2016) or autosomal genes (0.219) \citep{hough2014}.
+It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
 
-\subsection*{Extensive loss of Y-chromosome diversity}
+\subsection*{Female biased sex ratios and variance in male offspring number}
 
-\subsection*{Female biased sex ratio and variance in male offspring number}
-
-\begin{figure}[htbp]
-\centering
-\noindent
-\includegraphics[width=\linewidth]{figure2.jpg}
-\caption{new figure coming soon).
-}
-\label{fig:spectrum}
-\end{figure}
+%\begin{figure}[htbp]
+%\centering
+%\noindent
+%\includegraphics[width=\linewidth]{figure2.jpg}
+%\caption{new figure coming soon).
+%}
+%\label{fig:spectrum}
+%\end{figure}
 

### 2016-11-16 manuscript/Ydiversity_nov12.tex
minor edits

@@
 It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
+We next consider several possible models - neutral and selective - that might explain the loss of Y chromosome diversity.
 
-\subsection*{Female biased sex ratios and variance in male offspring number}
+\subsection*{female-biased sex ratios and variance in male offspring number}
+The occurrence of female-biased sex ratios in \textit{R.
+hastatulus} is expected to lower Y diversity through lowering male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome.
+In addition, the $N_{e}_{Y}$ might be further reduced due to high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
+hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there is reason to believe that there could be strong competition among males.
 

### 2016-11-19 manuscript/manuscript_nov18.tex
added stephen's changes in new document, varaince in rep success section updates

--- 2016-11-15 manuscript/Ydiversity_nov12.tex
@@
 
-\title{Nucleotide Diversity Loss on a Plant Y Chromosome Following Recent Recombination Suppression}
+\title{Hill-Robertson Interference Causes Reduced Genetic Diversity on a Young Plant Y-chromosome Following Recent Recombination Suppression}
 
@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic diversity.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect patterns of genetic diversity.
 On Y chromosomes with suppressed recombination, selection is expected to eliminate neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity.
-Here, we investigate the factors affecting levels of polymorphism within the genome during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
-Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
+However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity from natural populations.
+Here, we investigate the factors affecting levels of polymorphism during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
+Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of 93\% compared to the neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Instead, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations.
-Our results are in agreement with theory on "interference selection", and provide evidence that the effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
+Instead, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that diversity loss on the Y can be explained by purifying selection acting over a large number ($\geq$ 800 Kb) of genetically-linked sites.
+Our results are in agreement with theory on "interference selection", and provide empirical evidence that this effect can substantially reduce neutral diversity on a non-recombining chromosome.
 Given the recent origin of \textit{R.
-hastatulus} sex chromosomes (\textasciitilde 15MYA), our results imply that Y chromosome degeneration in the early stages may be largely driven by interference rather than positive selection for gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes (15MYA), our results imply that Y-chromosome degeneration in the early stages may be largely driven by such interference effects rather than by positive selection for gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 
-In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects positive of purifying selection need to be distinguished from these neutral processes.
-Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
+In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experienced by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects positive of purifying selection need to be distinguished from these neutral processes.
+Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of neutral diversity on X and Y, relative to autosomes.
 For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
@@
 However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA) \citep{lahn1999four,ross2005dna}, is not immediately obvious whether purifying selection might have such strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last \textasciitilde 20 MYA in the case of plants \citep{charlesworth2015plant}).
-Y-linked diversity loss might on the one hand be expected to be lower in younger systems due to a shorter history of recombination suppression.
+Y-linked diversity loss might on the one hand be expected to be lower in younger if there has been widespread gene silencing, as discovered on the \textit{Drosophil miranda} neo-Y chromosome \citep{zhou2012chromosome}.
 On the other hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
 Moreover, there is clear evidence from simulation studies that even weak purifying selection, if acting non-independently across large number of sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
-Given that a large number of sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is a good candidate model for exploring the evolutionary dynamics of young plant sex chromosomes.
+Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is a good candidate model for exploring the evolutionary dynamics of young plant sex chromosomes.
 
@@
 Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone gene loss and functional deterioration \citep{hough2014}.
-Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, we focus only on the ancestral Y chromosome, which occurs in both sex chromosome races.
+Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, we focus only on the XY system with the ancestral chromosome configuration,  as it is more likely to be at equilibrium than the neo-Y chromosome system that experienced the fusion event.
 
@@
 \subsection*{Phasing X and Y alleles}
-To estimate polymorphism for X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
+To estimate polymorphism for X and Y sequences separately in males, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
 In previous work, phasing alleles on \textit{R.
@@
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
-Finally, we filtered out SNPs with coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels.
-This filtering procedure resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes.
+Finally, we filtered out SNPs with coverage < 60, and those within a distance of 10bp or less from indels.
+This filtering procedure resulted in alignments of X and Y sequences for 372 sex-linked genes.
 
@@
 To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level polymorphism, obtained from the genetic cross, was zero.
-We identified 7 \% of sex-linked genes that either had  phasing errors of this kind genotyping errors.
+We identified 7\% of sex-linked genes that either had phasing errors genotyping errors.
 This corresponds to a SNP error rate estimate of 1.7 x 10-4.
 Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of the X/A ratio.
-However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we further filtered genes in which we found evidence for false-positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
+However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we conducted a further filtering step for Y-linked SNPs and identified false positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
 This final filtering was conducted by manually checking each individual putative polymorphism on the Y chromosome using IGV \citep{robinson2011integrative}.
@@
 \subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
-For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
+For each locus in our analysis, we calculated Watterson's (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
 To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2).
@@
 Our results do suggest a slight elevation in X-linked diversity relative to the neutral prediction of X/A=0.75, but this difference is not significantly different from the expectation for the empirically-estimated ratio of 0.6.
-Although our estimates of diversity have not beed normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R.
+Although our estimates of diversity have not been normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R.
 hastatulus} and the non-dioecious outgroup \textit{R.
-bucephalophorus} is not significantly different for sex-linked (0.2016) or autosomal genes (0.219) \citep{hough2014}.
+bucephalophorus} is not significantly different for sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
+Furthermore, we found no evidence for significant differences in substitution rate between Y and X chromosomes \citep{hough2014}.
 It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
 
+%\begin{table}[htb]
+%\centering
+%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
+%\begin{tabular}{ccccccccc}
+%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
+%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
+%\midrule
+%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
+%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
+%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
+%\addlinespace
+
+%\bottomrule
+%\end{tabular}
+%\end{table}
+
 \subsection*{Female biased sex ratios and variance in male offspring number}
+The occurrence of female-biased sex ratios in \textit{R.
+hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome.
+In addition, male $N_{e}$ can be further reduced by high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
+hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
+Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may be strong competition among males to fertilize females.
+Such competition should increase levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity (Charlesworth 1996b), while simultaneously reducing Y-linked diversity.
+Thus, to test whether high variance in male reproductive success could explain our observed Y/A and X/A ratios, we tested the fit of our data to predictions from a model of variance in male reproductive success \citep{kimura1964number}, using the estimated population sex ratio of 0.6 \citep{pickup2013influence}.
+As shown in Figure 2, our analysis indicated that, when evaluated at a sex ratio of 0.6, values for the variance in male reproductive success, $V_{k}_{m}$, that were large enough (range) to explain our observed reduction in Y/A diversity also predicted X/A diversity levels that were significantly greater than our observed values (range, P-value; Table 2).
+This indicates that the combined effects of a female-biased population sex ratio and extensive variance in male reproductive success is insufficient explain our observed  reduction in Y-chromosome diversity (Figure 2; Table 2).
 
-%\begin{figure}[htbp]
+%\begin{table}[htb]
 %\centering
-%\noindent
-%\includegraphics[width=\linewidth]{figure2.jpg}
-%\caption{new figure coming soon).
-%}
-%\label{fig:spectrum}
-%\end{figure}
+%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
+%\begin{tabular}{ccccccccc}
+%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
+%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
+%\midrule
+%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
+%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
+%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
+%\addlinespace
+
+%\bottomrule
+%\end{tabular}
+%\end{table}
+
+\begin{figure}[htbp]
+\centering
+\noindent
+\includegraphics[width=\linewidth]{figure2.jpg}
+\caption{new figure coming soon).
+}
+\label{fig:spectrum}
+\end{figure}
 
 \subsection*{Purifying selection}
+
+- Basic results; diversity loss consistent with purifying selection
+- HRI vs BGS; bgs over predicts diversity loss; good reasons to think we are in interference space. Include calculations from good et al; Figure 3 (sims); Figure 4 (bgs model -maybe put in supp)
+- Can we exclude sweeps empirically? Fitting frequency spectrum to model?; tajimas D?
+- Caveats; demography (e.g. changes in Ne) are not well known but could affect the metrics above; NeS assumes independence and can't be trusted; parameter uncertainties etc.
+- Implications for sex chromosome evolution in plants
 

### 2016-11-19 manuscript/manuscript_nov18.tex
results edits

@@
 Thus, to test whether high variance in male reproductive success could explain our observed Y/A and X/A ratios, we tested the fit of our data to predictions from a model of variance in male reproductive success \citep{kimura1964number}, using the estimated population sex ratio of 0.6 \citep{pickup2013influence}.
-As shown in Figure 2, our analysis indicated that, when evaluated at a sex ratio of 0.6, values for the variance in male reproductive success, $V_{k}_{m}$, that were large enough (range) to explain our observed reduction in Y/A diversity also predicted X/A diversity levels that were significantly greater than our observed values (range, P-value; Table 2).
+In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R.
+hastatulus}.
+However, as shown in Figure 2, we find that, when evaluated at a sex ratio of 0.6, values for the variance in male reproductive success, $V_{k}_{m}$, that were large enough (range) to explain our observed reduction in Y/A diversity also predicted X/A diversity levels that were significantly greater than our observed values (range, P-value; Table 2).
 This indicates that the combined effects of a female-biased population sex ratio and extensive variance in male reproductive success is insufficient explain our observed  reduction in Y-chromosome diversity (Figure 2; Table 2).
@@
 - Basic results; diversity loss consistent with purifying selection
+
 - HRI vs BGS; bgs over predicts diversity loss; good reasons to think we are in interference space. Include calculations from good et al; Figure 3 (sims); Figure 4 (bgs model -maybe put in supp)
+
+Given that a large number of sites in the genome likely experience weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
+Because there is no general analytical treatment describing the dynamics of such interference among a large number of linked sites under selection and drift, simulations are fundamental to investigating realistic situations.
+We therefore....
+
+...Individual fitness effects may play a central role in single-site models, but we have shown that global properties like the variance in fitness and the corresponding linkage scale are more relevant for predicting evolution in interfering populations.
+" - estimates of "Nes" ignore linkage by fiat under the assumption that sites evolve independently.
+But these estimates become unreliable precisely when small- and intermediate-effect mutations are most common - "Here, we have shown that simple behavior emerges in the limit of widespread interference.
+When fitness variation is composed of many individual mutations, the magnitudes and signs of their fitness effects are relatively unimportant.
+Instead, molecular evolution is controlled by the variance in fitness within the population over some effectively asexual segment of the genome" " In other words, we cannot conclude that interference is negligible just because ‘‘Nes’’, as inferred from data, is larger than one."
+
+- Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
+
+\subsection*{Can selective sweeps of beneficial mutations explain Y-diversity loss?}
 - Can we exclude sweeps empirically? Fitting frequency spectrum to model?; tajimas D?
@@
 
+I guess to my mind a sweep explanation is a priori implausible for the following reasons: 1) a species wide sweep, for example, would wipe out diversity entirely, so that the existing polymorphisms must have accumulated since the sweep, and that seems unlikely if none of our Y SNPs are singletons?
+2.
+If recombination is suppressed within the region that we are working with (which it must be, because the SNPs segregate as fully sex-linked), then the occurrence of chromosome-wide linkage in that nonrecombining region will also cause beneficial mutations to have a very low probability of fixation.
+The relevant linkage scale here is the entire length of the non-recombining region, L.
+I've been doing a bunch of these HRI simulations for the last couple of days, and just as an example, in a simple model with the following parameters:
+
+mutation rate: 1e-6
+recombination rate: 0
+generations: 1000 generations after equilibrium is reached (6000 generations total)
+L= 100000 (100kb)
+N=1000 individuals
+s =0.05 (beneficial)
+
+the reduction in the probability of fixation of a beneficial mutation arising from linkage among 100,000 sites in a nonrecombining chromosome, is 98!!!
+(the reduction is relative to a model of independence among  selected sites).
+That's huge (and it decreases with larger L).
+So at least on a first pass, the probability that a selected sweep is the explanation here seems very low.
+
+\subsubsection*{Implications for the evolution of young Y chromosomes}
+Right.
+On the other hand you can imagine that widespread silencing could happen precisely because of the problem of interference selection.
+For example, interference causes selection to be ineffective, so that Y chromosomes accumulate deleterious mutations and do not fix beneficial ones.
+This could then generate selection to down-regulate or silence such "unfit" Y-linked genes, as in an Orr & Kim type of model.
+So yeah, I don't think that finding many selected sites would necessarily preclude a model of widespread gene silencing.
+It could just mean that we haven't gotten there yet (which would still be interesting).
+
 \section*{Conclusions}
+
+This is why I think HRI is so important for the Y.
+Not only does it result in a significant reduction in diversity when selection acts on linked deleterious mutations, but it also substantially reduces the probability of fixation of beneficial mutations.
+I grant that the probability of fixation of a beneficial mutation, and therefore the probability that a sweep explains our data, is not zero, but given the probable size of L on the R hastatulus Y chromosome, I don't think the claim that "our data is just as easily explained by sweeps" is defensible.
+If it is, then we haven't really done anything valuable here.
 

### 2016-11-22 manuscript/manuscript_nov18.tex
table 1
edits

--- 2016-11-19 manuscript/manuscript_nov18.tex
@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model for the expected $N_{e}$ of males and females when the variance in offpring number is smaller or larger than the standardly assumed Poisson \citep{kimura1964number, hedrick2011genetics}.
-As we were primarily interested in the influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which this effect might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring numbers, and calculated male $N_{e}$ as:
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model for the expected $N_{e}$ of males and females when the variance in offspring number is smaller or larger than the standardly assumed Poisson \citep{kimura1964number, hedrick2011genetics}.
+Because we were primarily interested in the influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which this effect might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring numbers, and calculated male $N_{e}$ as:
 
@@
 where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
-Given expected values of $N_e_{m}$ for a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding neutral expectation for autosomal and sex-linked $N_{e}$ at equilibrium were calculated as:
+Using this model to generate expected values for the male effective population size, $N_{e}_{m}$, given values of $V_{k_{m}} > 2$ (Figure 1), the corresponding neutral expectations for the $N_{e}$ of autosomal and sex-linked genes at equilibrium were then calculated as:
 
@@
 respectively, following \citep{wright1931evolution}.
-Note that with equal sex ratios $N_{f} = N_{m}$, and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ and in the standard model.
-Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our observed Y/A and X/A diversity ratios to model predictions (shown in Figure 2) assuming the empirically-estimated \textit{R.
-hastatulus} population sex ratio of 0.6 \citep{pickup2013influence}.
+Note that with equal sex ratios ($N_{f} = N_{m}$) and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model.
+Figure 1 shows neutral predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our observed Y/A and X/A diversity ratios to model predictions for a sex ratio of 0.6 \citep{pickup2013influence}.
 
@@
 Thus, to test whether high variance in male reproductive success could explain our observed Y/A and X/A ratios, we tested the fit of our data to predictions from a model of variance in male reproductive success \citep{kimura1964number}, using the estimated population sex ratio of 0.6 \citep{pickup2013influence}.
-As shown in Figure 2, our analysis indicated that, when evaluated at a sex ratio of 0.6, values for the variance in male reproductive success, $V_{k}_{m}$, that were large enough (range) to explain our observed reduction in Y/A diversity also predicted X/A diversity levels that were significantly greater than our observed values (range, P-value; Table 2).
+In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R.
+hastatulus}.
+However, as shown in Figure 2, we find that, when evaluated at a sex ratio of 0.6, values for the variance in male reproductive success, $V_{k}_{m}$, that were large enough (range) to explain our observed reduction in Y/A diversity also predicted X/A diversity levels that were significantly greater than our observed values (range, P-value; Table 2).
 This indicates that the combined effects of a female-biased population sex ratio and extensive variance in male reproductive success is insufficient explain our observed  reduction in Y-chromosome diversity (Figure 2; Table 2).
 
-%\begin{table}[htb]
-%\centering
-%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
-%\begin{tabular}{ccccccccc}
-%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
-%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
-%\midrule
-%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
-%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
-%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
-%\addlinespace
+\begin{table}[htb]
+\centering
+\caption{Observed neutral diversity on sex chromosomes and autosomes in \textit{R. hastatulus}, and expected values for a population sex ratio $N_{f}/(N_{f}+N_{m})=0.6$.}
+\begin{tabular}{ccccccccc}
+\textbf{} & \multicolumn{2}{l}{\textbf{Observed}} & \multicolumn{3}{l}{\textbf{Expected}} \\
 
-%\bottomrule
-%\end{tabular}
-%\end{table}
+chr & $\theta$ & $/\theta_{A}$ & $\theta$ & $/\theta_{A}$ & P-value  \\
+\midrule
+A & 0.006 & 1 & 0.006 & 1 & 0.006 \\
+X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0019 \\
+Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ \\
+\addlinespace
+
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \includegraphics[width=\linewidth]{figure2.jpg}
-\caption{new figure coming soon).
+\caption{Observed and predicted X/A and Y/A ratios of neutral diversity.
+Because female-biased population sex ratios ($N_{f} > N_{m}$) and high variance in male reproductive success ($V_{k}_{m} > 2$) both reduce male effective population size ($N_{e}_{m}$), the expected nucleotide diversity ratios are shown for increasing values of the female-to-male $N_{e}$ ratio, $N_{e}_{f}/N_{e}_{m}$.
+Black bars correspond to predictions under the standard neutral model in which the population sex ratio is $N_{f}/(N_{f}+N_{m})=0.5$ and there is a Poisson distribution of offspring number ($V_{k}_{m} = V_{k}_{f} = \overline{k}=2$).
+Observed estimates of X/A and Y/A are given as the average $\theta$ across genes, weighted by the number of synonymous sites per gene.
+Confidence intervals were calculated by bootstrapping (20000 replicates) using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011), and using maximum likelihood for the Y-chromosome (Methods).
 }
@@
 
-- Basic results; diversity loss consistent with purifying selection
-- HRI vs BGS; bgs over predicts diversity loss; good reasons to think we are in interference space. Include calculations from good et al; Figure 3 (sims); Figure 4 (bgs model -maybe put in supp)
+%- Basic results; diversity loss consistent with purifying selection
+
+%- HRI vs BGS; bgs over predicts diversity loss; good reasons to think we are in interference space. Include calculations from good et al; Figure 3 (sims); Figure 4 (bgs model -maybe put in supp)
+
+%Given that a large number of sites in the genome likely experience weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
+Because there is no general analytical treatment describing the dynamics of such interference among a large number of linked sites under selection and drift, simulations are fundamental to investigating realistic situations.
+We therefore....
+
+%...Individual fitness effects may play a central role in single-site models, but we have shown that global properties like the variance in fitness and the corresponding linkage scale are more relevant for predicting evolution in interfering populations.
+" - estimates of "Nes" ignore linkage by fiat under the assumption that sites evolve independently.
+But these estimates become unreliable precisely when small- and intermediate-effect mutations are most common - "Here, we have shown that simple behavior emerges in the limit of widespread interference.
+When fitness variation is composed of many individual mutations, the magnitudes and signs of their fitness effects are relatively unimportant.
+Instead, molecular evolution is controlled by the variance in fitness within the population over some effectively asexual segment of the genome" " In other words, we cannot conclude that interference is negligible just because ‘‘Nes’’, as inferred from data, is larger than one."
+
+%- Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
+
+\subsection*{Can selective sweeps of beneficial mutations explain Y-diversity loss?}
 - Can we exclude sweeps empirically? Fitting frequency spectrum to model?; tajimas D?
@@
 
+%I guess to my mind a sweep explanation is a priori implausible for the following reasons: 1) a species wide sweep, for example, would wipe out diversity entirely, so that the existing polymorphisms must have accumulated since the sweep, and that seems unlikely if none of our Y SNPs are singletons?
+2.
+If recombination is suppressed within the region that we are working with (which it must be, because the SNPs segregate as fully sex-linked), then the occurrence of chromosome-wide linkage in that nonrecombining region will also cause beneficial mutations to have a very low probability of fixation.
+The relevant linkage scale here is the entire length of the non-recombining region, L.
+I've been doing a bunch of these HRI simulations for the last couple of days, and just as an example, in a simple model with the following parameters:
+
+%mutation rate: 1e-6
+%recombination rate: 0
+%generations: 1000 generations after equilibrium is reached (6000 generations total)
+%L= 100000 (100kb)
+%N=1000 individuals
+%s =0.05 (beneficial)
+
+%the reduction in the probability of fixation of a beneficial mutation arising from linkage among 100,000 sites in a nonrecombining chromosome, is 98!!!
+(the reduction is relative to a model of independence among  selected sites).
+That's huge (and it decreases with larger L).
+So at least on a first pass, the probability that a selected sweep is the explanation here seems very low.
+
+\subsubsection*{Implications for the evolution of young Y chromosomes}
+%Right.
+On the other hand you can imagine that widespread silencing could happen precisely because of the problem of interference selection.
+For example, interference causes selection to be ineffective, so that Y chromosomes accumulate deleterious mutations and do not fix beneficial ones.
+This could then generate selection to down-regulate or silence such "unfit" Y-linked genes, as in an Orr & Kim type of model.
+So yeah, I don't think that finding many selected sites would necessarily preclude a model of widespread gene silencing.
+It could just mean that we haven't gotten there yet (which would still be interesting).
+
 \section*{Conclusions}
+
+%This is why I think HRI is so important for the Y.
+Not only does it result in a significant reduction in diversity when selection acts on linked deleterious mutations, but it also substantially reduces the probability of fixation of beneficial mutations.
+I grant that the probability of fixation of a beneficial mutation, and therefore the probability that a sweep explains our data, is not zero, but given the probable size of L on the R hastatulus Y chromosome, I don't think the claim that "our data is just as easily explained by sweeps" is defensible.
+If it is, then we haven't really done anything valuable here.
 

### 2016-12-04 code/R/Nov25_2016.Rmd
major edits to ms, figures, etc.

---
output: html_notebook
---
#### import polymorphurama data
```{r}
#workingdir
setwd("/Users/josh/Github/XYdiversity/temp/workspace")

#sex chromosomes
txx <- read.csv("txx.csv")
txy <- read.csv("txy.csv")

#autosomes
tm_all <- read.csv("tm_all_new.csv")
tf_all <- read.csv("tf_all_new.csv")

```

#### calculate the average weighted theta value for each (synsites*theta_syntot); calculate weighted average
```{r}
#new columns=SynSitestot*theta_syntot (for each cell)
#TX_X
txx$weighted_theta <- txx$SynSitestot * txx$theta_syntot            
txx_ave_weighted_theta=(sum(txx$weighted_theta))/(sum(txx$SynSitestot))
#TX_Y
txy$weighted_theta <- txy$SynSitestot * txy$theta_syntot            
txy_ave_weighted_theta=(sum(txy$weighted_theta))/(sum(txy$SynSitestot))
```

#### plot of X and Y weighted mean theta
```{r}
barplot(c(txx_ave_weighted_theta,
          txy_ave_weighted_theta),
          ylab=expression(paste("Nucleotide Diversity (average weighted ", theta, " )")),
          names.arg=c("TX_X","TX_Y"),
          ylim=c(0,0.005))
```

```{r}

```

#### Calculate X/A diversity ratios

```{r}
#import autosomal genes
#females
tf_all$weighted_theta <- tf_all$SynSitestot * tf_all$theta_syntot            
tf_all_ave_weighted_theta=(sum(tf_all$weighted_theta))/(sum(tf_all$SynSitestot))
#males
tm_all$weighted_theta <- tm_all$SynSitestot * tm_all$theta_syntot            
tm_all_ave_weighted_theta=(sum(tm_all$weighted_theta))/(sum(tm_all$SynSitestot))

#X/A
tx_x_auto=txx_ave_weighted_theta/tf_all_ave_weighted_theta
#Y/A 
tx_y_auto=txy_ave_weighted_theta/tf_all_ave_weighted_theta

```

#### Neutral expectations for an equal sex ratio, poisson offspring number (Kimura and Crow model)

```{r}
#set numbers of males and females
vm=2 #variance in female rep success
vf=2 #variance in female rep success
kf=2 #mean number of offspring of females
km=2 #mean number of offspring of males
Nf=100000 #numner of females
Nm=100000 #number of males
kbar=2 # total mean offspring number

# males and female Ne
Nem=(Nm*km-1)/(km-1+(vm/kbar)) #effect of male reproductive variance on female Ne
Nef=(Nf*kf-1)/(kf-1+(vf/kbar)) #effect of female reproductive variance on female Ne

# chromosome Ne's
NeA=(4*Nem*Nef)/(Nem+Nef)
NeX=(9*Nem*Nef)/((4*Nem)+(2*Nef))
NeY=Nem/2

YA_poisson=NeY/NeA
XA_poisson=NeX/NeA

```

#### Diversity predictions for a range of Nef/(Nem+Nef) ratios
```{r}
# expected X/A and Y/A curves as functions of the sex ratio:
curve((9*(-1 + x))/(8*(1 - x)*(-2 + x)), from=0, to=1, ylim=c(0.002,1.12),
      ylab="X/A", xlab="Nef/(Nem+Nef)")

curve(1/(8*x), from=0, to=1, add=T)

#X/A
XA_0.5=((9*Nef*0.5*(0.5 - 1))/(2*(0.5 - 2)))/(4*Nef*(1 - 0.5)*0.5); XA_0.5
XA_0.6=((9*Nef*0.6*(0.6 - 1))/(2*(0.6 - 2)))/(4*Nef*(1 - 0.6)*0.6); XA_0.6
XA_0.7=((9*Nef*0.7*(0.7 - 1))/(2*(0.7 - 2)))/(4*Nef*(1 - 0.7)*0.7); XA_0.7
XA_0.8=((9*Nef*0.8*(0.8 - 1))/(2*(0.8 - 2)))/(4*Nef*(1 - 0.8)*0.8); XA_0.8
XA_0.9=((9*Nef*0.9*(0.9 - 1))/(2*(0.9 - 2)))/(4*Nef*(1 - 0.9)*0.9); XA_0.9

#Y/A
YA_0.5=1/(8*0.5); YA_0.5
YA_0.6=1/(8*0.6); YA_0.6
YA_0.7=1/(8*0.7); YA_0.7
YA_0.8=1/(8*0.8); YA_0.8
YA_0.9=1/(8*0.9); YA_0.9

curve(((9*(-1 + x))/(8*(1 - x)*(-2 + x)))/0.75, from=0.5, to=1, ylim=c(0,1.5), lwd=2, 
      ylab="observed/expected", xlab="sex ratio")
curve((1/(8*x))/0.25, from=0.5, to=1, lwd=2, add=T)
abline(lty=2,h=1.13)
abline(lty=2,h=0.0004)

```

So, theoretically, the maximum Y/A reduction relative to neutrality is 0.5. We have a reduction of 0.00156 (+/- ).

```{r}

plot(c(tx_x_auto/0.75,XA_0.6/0.75, xa_given_y/0.75),
        ylim=c(0,1.5),
        ylab="X/A relative to neutral model", main="X/A for a sex ratio of 0.6", 
        xlab="expected                                                      observed")
curve((1/(8*x))/0.25, from=0.5, to=1, lwd=2, add=T)
abline(lty=2,h=1.13)
abline(lty=2,h=0.0004)

```

#### Plot empirical, simulated, and theoretically predicted diversity ratios for a range of Ne sex ratios, with CI's
```{r}

xa_given_y=0.0003 # y theta * 3
ya_given_x=0.00156 #x theta * 1/3
XA=c(tx_x_auto, XA_0.5, XA_0.6, XA_0.7, XA_0.8, XA_0.9)
YA=c(tx_y_auto, YA_0.5, YA_0.6, YA_0.7, YA_0.8, YA_0.9)
ChrRatios=cbind(XA,YA)

barplot(as.matrix(ChrRatios), beside=TRUE, col=c("darkgreen", "black", "gray22","gray73","gray87","gray93","darkgreen","black", "gray22","gray73","gray87", "gray93"),
        ylab = expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
        ylim=c(0,1.14))
legend("topright", c("observed", "Nf/(Nf+Nm) = 0.5", "Nf/(Nf+Nm) = 0.6", "Nf/(Nf+Nm) = 0.7", "Nf/(Nf+Nm) = 0.8", "Nf/(Nf+Nm) = 0.9"), fill=c("darkgreen", "black", "gray22","gray73","gray87", "gray93"), box.lty=0, cex=0.7)

# empirical XA confidence intervals
segments(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75)
arrows(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75,angle = 90, code = 3, length = 0.04)

# simulated XA confidence intervals
#segments(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75)
#arrows(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75,angle = 90, code = 3, length = 0.04)

# empirical YA confidence intervals
segments(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75)
arrows(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)

# simulated YA confidence intervals
#segments(x0=10.5,y0=0.01,x1=10.5,y1=0.04,lwd=0.75)
#arrows(x0=10.5,y0=0.02,x1=10.5,y1=0.07,lwd=0.75,angle = 90, code = 3, length = 0.04)

```

```{r}
#0.25*(exp(-(0.02/((2*0.002)+0)))
bgs_exp=0.00168449

barplot(c(tx_y_auto,ya_given_x,bgs_exp), ylim=c(0,0.04), col="gray52", ylab="Y/A",
        names.arg=c("observed", "predicted from X", "predicted by bgs"), cex.names=0.75)
segments(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75)
arrows(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)

```

### 2016-12-04 manuscript/manuscript_Dec3.tex
major edits to ms, figures, etc.

\documentclass[9pt,twocolumn,twoside]{gsajnl}
% Use the documentclass option 'lineno' to view line numbers
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{adjustbox}
\usepackage{geometry}
\articletype{gos} % article type

\title{Hill-Robertson Interference Causes Reduced Genetic Diversity on a Young Plant Y-chromosome}

\author[$\ast$,$\dagger$,1]{Josh Hough}
\author[$\dagger$]{Wei Wang}
\author[$\dagger$]{Spencer C.H. Barrett}
\author[$\dagger$]{Stephen I. Wright}

\affil[$\ast$]{Department of Plant Sciences, University of California, Davis}
\affil[$\dagger$]{Department of Ecology and Evolutionary Biology, University of Toronto}

%\affil[$\dagger$]{Author two affiliation}
%\affil[$\ddagger$]{Author three affiliation}
%\affil[$\S$]{Author four affiliation}
%\affil[$\ast\ast$]{Author five affiliation}

\keywords{Sex Chromosome Evolution; Nucleotide Diversity; Recombination; Deleterious Mutations}

\runningtitle{Plant sex chromosome evolution} % For use in the footer

\correspondingauthor{Josh Hough}

\begin{abstract}
X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect patterns of genetic diversity. On Y chromosomes with suppressed recombination, natural selection is expected to eliminate neutral variation and lower the $N_{e}$ of Y compared to X chromosomes or autosomes. However, non-selective factors including female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity from natural populations. Here, we investigate the factors affecting levels of polymorphism during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes. Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of 99.2\% compared to the neutral expectation. We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success. Instead, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that diversity loss on the Y can be explained by purifying selection acting over a large number ($\geq$ 800 Kb) of genetically-linked sites. Our results are in agreement with theory on "interference selection", and provide evidence that this effect can substantially reduce nucleotide diversity on a young Y chromosome. Given the relatively recent origin of \textit{R. hastatulus} sex chromosomes, our results imply that Y-chromosome degeneration in the early stages may be largely driven by such interference effects rather than by positive selection for gene silencing followed by neutral genetic drift.
\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\marginmark
\firstpagefootnote
\correspondingauthoraffiliation{jhough@ucdavis.edu}
\vspace{-11pt}

\section*{Introduction}

\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plants and animals \citep{westergaard1958,ohno1967,bull1983,charlesworth1991,charlesworth2015plant}. Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes. In both groups, for example, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}. In some species, such genetic deterioration of the Y chromosome has also led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}. The independent evolution of these phenomena in such taxonomically distant groups as mammals and flowering plants suggests that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoretical and experimental biologists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.

One fundamental difference between X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population. Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas the  $N_{e}$ for genes on the X chromosome should be 3/4 that of autosomal genes (assuming an equal number of reproducing females and males). The lowered $N_{e}$ of the Y chromosome implies that the equilibrium level of neutral polymorphism - proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ - should be lower for Y-linked genes than for their X-linked counterparts. In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citep{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citep{smith1974hitch}). Furthermore, the build-up of genetic associations among selected mutations (high fitness covariance) on the Y means that selection will act non-independently across the chromosome, such that selection at any given focal site "interferes" with selection at the sites with which it is linked \citep{hill1966HReffect}. A large body of work has now shown that in theory such "selective interference" effects can substantially reduce both the efficacy of selection and the equilibrium level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}, with the magnitude of the effect depending strongly on the number and density of sites experiencing selection . These arguments all suggest that large non-recombining Y chromosomes should harbor a lower amount of neutral variability than predicted based on the number of Y chromosomes in a population (1/4 that of autosomes), and highlight that estimating the number of selected sites is a key test for determining the likely mechanism involved in driving diversity loss on a nonrecombining Y chromosome. However, the relative importance of neutral and selective factors for reducing chromosome-wide levels of diversity is not well-understood.

In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experienced by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects on diversity of positive of purifying selection need to be distinguished from these neutral processes. Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of silent site variability on X and Y chromosome, relative to values on autosomes. For example, high variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio compared to the neutral expectation of 1/4, but it also increases the X/A ratio. Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is due to a historical excess of breeding females compared to males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic}).

Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes. The time scales over which these different effects are likely to be important are therefore not well understood. In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations suggest that levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}. However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA) \citep{lahn1999four,ross2005dna}, it is not clear whether purifying selection might have similarly strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last \textasciitilde 20 MYA in the case of plants \citep{charlesworth2015plant}).

On the one hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}. Moreover, theory suggests that even weak purifying selection, if acting across a large number of genetically-linked sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}. Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is \textit{a priori} likely to have strong effects on young plant sex chromosomes. On the other hand, if there has been widespread gene silencing during the early stages of Y degeneration, as found on the \textit{Drosophila miranda} neo-Y chromosome \citep{zhou2012chromosome}, then Y-linked diversity loss might be low in younger sex chromosome systems. In particular, if gene silencing occurs early during Y chromosome degeneration, few sites are expected to be experiencing purifying selection, with degeneration primarily driven by neutral genetic drift rather than inefficient selection. Distinguishing the relative roles of selective interference and neutral drift during Y chromosome evolution is a major open issue.

To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R. hastatulus }(Polygonaceae). This species is a dioecious annual with heteromorphic X and Y chromosomes that originated \textasciitilde 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}. \textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphic in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ( so-called "chromosomal races") \citep{smith1963mechanism}. The $XY_{1}XY_{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}. Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone gene loss and functional deterioration \citep{hough2014}. Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, and to ensure that our diversity estimates are not biased by non-equilibrium conditions or other confounding factors arising from the X-autosome fusion in the $XY_{1}XY_{2}$ system, we focus only on the XY system.

Of particular importance for our study, \textit{R. hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}. Female-biased sex ratios are not uncommon in dioecious plants with heteromorphic X and Y chromosomes \citep{field2013comparative,hough2013evolutionarily}, and their occurrence in \textit{R. hastatulus} provides an opportunity to test both the demographic and selective contributions to changes in sex-linked variability during sex chromosome evolution.

\section*{Materials and Methods}
\subsection*{Population Samples and Sex-Linked Genes}
We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 12 population samples (1 male and 1 female from each of 6 populations). Samples were collected in 2010 from throughout the native range of \textit{R. hastatulus} (locations in Table S1), and plants were grown in the glasshouse at the University of Toronto from seeds collected from open-pollinated females. We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich). Isolation of mRNA and cDNA synthesis was conducted according to standard Illumina RNAseq procedures, and sequencing was conducted on two Illumina HiSeq lanes with 150-bp end reads at the Genome Quebec Innovation Center. Reads from these samples were mapped to the \textit{R. hastatulus} reference transcriptome \citep{hough2014} using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}. We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20. The population samples analyzed here were previously reported in Hough et al. (2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no. SRP041588. Here, to consider sex linked genes that were identified in all of our sequenced population samples, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).

\subsection*{Autosomal Genes}
In evaluating evidence for nucleotide diversity differences between X and Y chromosomes, it is important to distinguish between reduced Y-linked diversity, and the possibility that X-linked diversity is elevated above the level predicted from a neutral model. To do this, we normalized our sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below). Because the criteria for ascertaining autosomal loci in Hough et al. (2014) were based on identifying four segregating SNPs per locus, and since this set of genes is likely to be higher in diversity than the average autosomal gene, here we instead used the larger set of all non-sex linked (putatively autosomal) genes as our autosomal reference. We filtered genes in this set to remove any genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria. In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage (Beaudry et al., in prep) data (to filter out duplicates or genes with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females. Finally, we removed genes with fewer than 100 synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems. This filtering resulted in a final set of 12,356 autosomal genes.

\subsection*{Phasing X and Y alleles}
To estimate polymorphism for X and Y sequences separately in males, it is necessary to infer the phase of SNPs in sex-linked transcripts in males. In previous work, phasing alleles on \textit{R. hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross. Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome. Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences within each haplotype block (i.e., sites where all females were homozygous, and all males were heterozygous). Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output. In particular, this was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched. SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data. Finally, we filtered out SNPs with quality scores < 60, and those within a distance of 10bp or less from indels. This filtering procedure resulted in alignments of X and Y sequences for 372 sex-linked genes.

We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in Hough et al. (2014). To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level polymorphism, obtained from the genetic cross, was zero. We identified 7 \% of sex-linked genes that either had phasing errors genotyping errors. This corresponds to a SNP error rate estimate of 1.7 x 10-4. Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of the X/A ratio. However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we conducted a further filtering step for Y-linked SNPs and identified false positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression. This final filtering was conducted by manually checking each individual putative polymorphism on the Y chromosome using IGV \citep{robinson2011integrative}.

\subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2). We obtained 95\% confidence intervals for X/A and Y/A ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes.

Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the Y/A estimate may be wider than implied by bootstrapping. To address this, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1). Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate was not feasible, as this would require maximizing the likelihood estimate of the mutation rate for each locus independently. However, previous analysis of divergence data does not suggest important chromosomal differences in synonymous divergence contributing to diversity heterogeneity (see Results). Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci. Our model thus had two parameters: $\theta_{autosomal}$ and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes. We varied both parameters and evaluated the likelihood for \textit{f} from 0.001 to 1, and $\theta_{autosomal}$ per base pair from 0.001 to 0.01. In particular, we evaluated the likelihood of $\theta_{autosomal}$ and \textit{f}, given the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), the number of segregating sites on the Y chromosome ($S_{y}$), and the total number of sites per locus ($L_{i}$ and $L_{y}$). The likelihood of ($\theta_{autosomal}$, f) is therefore given by:

\begin{equation}
L \propto  P(S_{Y}, L_{Y} | f, \theta_{a})\prod_{i=1}^{n}P(S_{i}, L_{i} | \theta_{a})  \label{eq:L}
\end{equation}

\subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to neutral predictions for a range of values for the ratio of effective population sizes of females and males. Because we were primarily interested in determining the parameter space within which a reduced male effective population size could result in Y/A and X/A diversity ratios consistent with our data, we tested the fit of our data to predictions across the full range of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ (0.1 to 0.9; we excluded all-female or all-male cases). For autosomal and sex-linked effective population sizes given by:

\begin{equation}
N_{e_{A}} = \frac{4N_{e}_{m}N_{e}_{f}}{N_{e}_{m}+N_{e}_{f}} \label{eq:NeA}
\end{equation}

\begin{equation}
N_{e_{X}} = \frac{9N_{e}_{m}N_{e}_{f}}{4N_{e}_{m}+2N_{e}_{f}} \label{eq:NeX}
\end{equation}

\begin{equation}
N_{e_{Y}} = \frac{N_{e}_{m}}{2} \label{eq:NeY}
\end{equation}

\citep{wright1931evolution}, the corresponding expected X/A and Y/A ratios an be expressed as:

\begin{equation}
\frac{N_{e_{X}}}{N_{e_{A}}} = \frac{9 (-1 + r)}{8 (1 - r) (-2 + r))} \label{eq:XA}
\end{equation}

and
\begin{equation}
\frac{N_{e_{Y}}}{N_{e_{A}}} = \frac{1}{8r} \label{eq:YA}
\end{equation}

respectively, where $r = N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.  Note that when $N_{e}_{f} = N_{e}_{m}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}. With a female-biased ratio, however, the expected X/A ratio can become greater than 1, approaching 1.125 in the limit as r\rightarrow 1 \citep{caballero1995} (Figure 1).

\begin{figure}[htbp]
\centering
\noindent
\includegraphics[width=\linewidth]{figure1.jpg}
\caption{\textbf{A}. Expected effective population sizes, relative to $N$,  for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y) as a function of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$. \textbf{B}. The corresponding X/A and Y/A ratios of diversity predicted at equilibrium (where $\theta=4N_{e}\mu$), assuming equal neutral mutation rates for sex-linked and autosomal genes.
}
\end{figure}

\subsection*{Simulations of purifying selection}
To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}. We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms. We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta$ from autosomal genes, adjusted to reflect the expected neutral reduction in $N_{e}_{Y}$ for a sex ratio of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.6$. To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331bp of linked neutral sequence from which we calculated silent site diversity. To examine the expected reduction in diversity as a function of the number of selected sites ($L_{selected}$), we ran simulations over a range of values of $L_{selected}$, up to a maximum of $5x$10^{6}$ (Figure 3A). To estimate $L_{selected}$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{simulated}$, was less than 0.0001 from our empirical estimate, $\pi_{observed}$ (Figure 3B).

\section*{Results and Discussion}

\subsection*{Extensive loss of Y-chromosome diversity}
Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R. hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model. In particular, neutral diversity for Y-linked genes is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome. Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of 99.2\% . Note that by normalizing X and Y diversity estimates by autosomal diversity levels, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but a Y-specific reduction. Interestingly, our results also suggest an elevation of X-linked diversity ($X/A=0.85$) relative to the neutral prediction ($X/A=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2). Given the empirically-estimated sex ratio of 0.6 in \textit{R. hastatulus}\citep{pickup2013influence}, however, the elevation we observed is not unexpected; the neutrally-predicted X/A ratio for a sex ratio of 0.6 is $\approx 0.8$ (Figure 2; Equation 4). Although our estimates of diversity have not been normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R. hastatulus} and the non-dioecious outgroup \textit{R. bucephalophorus} is not significantly different for sex-linked (0.2016) and autosomal genes (0.219), and we found no evidence for significant differences in substitution rate between Y and X chromosomes \citep{hough2014}. It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.

\begin{table}[htb]
\centering
\caption{Observed and neutrally-expected silent site diversity on sex chromosomes and autosomes in \textit{R. hastatulus}}
\begin{tabular}{ccccccccc}
\textbf{} & \multicolumn{2}{l}{\textbf{Observed}} & \multicolumn{3}{l}{\textbf{Expected}} \\

chr & $\theta$ & $/\theta_{A}$ & $\theta$ & $/\theta_{A}$  \\
\midrule
A & 0.006 & 1 & N/A & 1 \\
X & 0.0047 & 0.85 & 0.0045 & 0.75 \\
Y & $\num{e-4}$ & 0.002 & 0.0015 & 0.25 \\
\addlinespace

\bottomrule
\end{tabular}
\end{table}

\subsection*{Female biased sex ratios and variance in male offspring number}
The occurrence of female-biased sex ratios in \textit{R. hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and thus a reduction the $N_{e}$ of the Y chromosome. Male $N_{e}$ could be further reduced by high variance in male reproductive success, which is not unusual in annual wind-pollinated plants such as \textit{R. hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production. Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may indeed be strong competition among males to fertilize females. Such competition should increase levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity. In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R. hastatulus}. We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values for $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.

As shown in Figure 2, the expected reduction in the Y/A ratio across the full range of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ is substantially higher than our observed reduction. Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.9$, the expectation for the Y/A ratio is $\approx 0.14$, which is substantially higher than our observed estimate ($Y/A= 0.002$) (Table 1; Figure 2; Figure S3). Moreover, such large reductions in $N_{e}_{m}$ also predict levels of X/A diversity that are not significantly larger than what we observed (Figure 2). Our results therefore indicate that, although reduced male $N_{e}$ arising from sex-biased demography is expected to contribute to reduced Y chromosome polymorphism, it is insufficient to explain the Y/A reduction we observed.

\begin{figure}[htbp]
\centering
\noindent
\includegraphics[width=\linewidth]{figure2.jpg}
\caption{Observed and predicted X/A and Y/A ratios of neutral diversity. Predictions are shown for increasing values of $N_{e}_{f}/N_{e}_{m}$. Observed estimates of X/A and Y/A were calculated as the average $\theta$ across genes, weighted by the number of synonymous sites per gene. Confidence intervals were calculated by bootstrapping (20000 replicates) using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011), and using maximum likelihood for the Y-chromosome (see Methods).
}
\label{fig:spectrum}
\end{figure}

\subsection*{Purifying selection}
Selection against strongly deleterious mutations is expected be an important factor reducing genetic variability in regions with low recombination \citep{charlesworth1993effect, charlesworth1996background}. The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for the extensive diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and Drosophila \citep{mcallister1999, charlesworth1996CB}. Under background selection theory, the predicted loss of diversity is modeled as modeled as a reduction in the equilibrium $N_{e}$ such that $\pi \approx 4f_{0}N_{e}\mu$, where $f_{0}=e^{-\frac{U}{sh+r}}$, $U$ is the mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate. This yields the expected value of $\pi$ relative to that in the absence of BGS ($\pi_{0}$): $E[\pi]= \pi_{0}e^{\frac{-U}{2sh+r}}$, where , $\pi_{0}=\theta=4N_{e}\mu$. Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}{m}/2$), the expected reduction in diversity is obtained in this haploid model by replacing $N_{e}$ above with number of breeding males, $N_{e}_{m}$. Assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes because of the female biased sex ratio \citep{pickup2013influence} and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations on the X chromosome, the corresponding expected value is $E[\pi]= 6x10^{-5}$ for a mutation rate of $10^{-8}$ per nucleotide per generation. Although there is considerable uncertainty in these parameters, the calculation yields a value of diversity that is 67\% lower that what we observed (Table 1).

Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time, because the behavior of selected alleles interferes with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}. As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding this selective interference in realistic situations. We therefore conducted forward simulations of purifying selection using our estimated parameters for the distribution of fitness effects of deleterious mutations, and tested whether interference selection could result in a level of Y diversity at neutral sites similar to the level we observed. We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was a remarkably good fit to our data (Figure 3). In addition, and consistent with previous studies, our simulations also suggest a saturating loss of diversity with an increasing number of selected sites \citep{KaiserCharlesworth}.

Although we initialized our simulations with selection parameters estimated from the distribution of fitness effects of deleterious mutations, recent work has shown that, unlike with BGS, evolutionary dynamics in the “interference selection regime” are insensitive to differences in strength of selection, with expected diversity loss instead determined primarily by the population fitness variance over the non-recombining region, which is in turn expected to increase with the number of linked loci under selection \citep{good2014genetic}. Estimating the number of selected loci is therefore key to determining whether diversity on the Y chromosome may be affected by interference selection. Our maximum likelihood estimates of the number of selected sites, $L$, from our simulations, suggest that the number of sites subject to selection required to explain the diversity loss from the \textit{R. hastatulus} Y chromosome is ($\approx 800 kb$) (Figure 3).

\begin{figure}[t!]
\centering
\noindent
\includegraphics[width=\linewidth]{figure3.jpg}
\caption{Simulation results of the level of neutral diversity $pi_{syn}$ vs. the number of selected sites (top panel) with the dashed line corresponding to our point estimate for the level of neutral diversity on the Y chromosome, and the relative likelihood curve for the selected sites (bottom panel), with the dashed line corresponding to our approximate 95\% credibility interval.
}
\label{fig:spectrum}
\end{figure}

Although our analyses of purifying selection models are consistent with our observed reduction in Y-linked diversity, it is also possible that selective sweeps have contributed. We have not formally excluded this possibility. However, if patterns of molecular evolution on the Y are driven primarily by interference selection, as our results suggest, then a further implication is that beneficial mutations should suffer a reduction in their probability of fixation, limiting the scope for sweeps to drive down diversity. Nonetheless, there is abundant evidence that “soft sweeps” are important determinants of patterns of nucleotide diversity \citep{messer2013population}, and it is possible that they have contributed to the extremely low levels of Y-linked variability that we observed.

\section*{Conclusions}
The non-recombining region of a Y chromosome produces evolutionary dynamics that are similar to a haploid asexual population, with the lack of recombination resulting in build up of genetic associations among selected sites, a loss in the efficiency of selection, and an increase in population fitness variance \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}. Our study of neutral diversity levels on the relatively young \textit{R. hastatulus} sex chromosomes provides clear evidence for an extensive loss of neutral diversity on the Y chromosome. Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss on this chromosome, forward population genetic simulations suggest that purifying selection acting over a large number of selected sites results in an expected level of Y-linked diversity that is consistent with what we observed. Our results are consistent with recent theory on interference selection, and suggest that interference is an important force in the evolution of  young Y chromosomes. In particular, rather than widespread silencing of Y-linked genes and their subsequent degeneration through increased neutral drift, our study suggests that when a large number of linked sites are subject to purifying selection, Y-chromosome degeneration in the early stages may be largely driven by the effects of selective interference.
\section*{Acknowledgments}

\bibliography{bibliography}
\end{document}

### 2016-12-04 code/R/bgs.R

@@ -1,6 +1,6 @@
 print(20*math.exp(-0.02/(0.1+0.005)))
 
-curve(0.2083333*(exp(-(0.02/((2*x)+0)))), from=0, to=1,
+curve(0.25*(exp(-(0.02/((2*x)+0)))), from=0, to=1,
       main="Expected Y/A diversity under BGS",
       xlab="selection coefficient (s)",
       ylab = expression(paste("E[", pi,"]")), las=1,
@@ -25,3 +25,13 @@ text(0.03,0.21,expression(paste(, pi[0],)))
 
 (0.2083333*(exp(-(0.02/((2*0.1)+0))))
 )
+
+
+theta
+Ne
+u
+h
+r
+U
+
+pi_not=theta

### 2016-12-04 code/R/nov29.R

@@ -1,0 +1,34 @@
+barplot(as.matrix(ChrRatios), beside=TRUE, col=c("darkgreen", "black", "gray22","gray73","gray87","gray93","darkgreen","black", "gray22","gray73","gray87", "gray93"),
+        ylab = expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
+        ylim=c(0,1.2))
+legend("topright", c("observed", "Nf/(Nf+Nm) = 0.5", "Nf/(Nf+Nm) = 0.6", "Nf/(Nf+Nm) = 0.7", "Nf/(Nf+Nm) = 0.8", "Nf/(Nf+Nm) = 0.9"), fill=c("darkgreen", "black", "gray22","gray73","gray87", "gray93"), box.lty=0, cex=0.7)
+
+# empirical XA confidence intervals
+segments(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75)
+arrows(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75,angle = 90, code = 3, length = 0.04)
+
+# simulated XA confidence intervals
+#segments(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75)
+#arrows(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75,angle = 90, code = 3, length = 0.04)
+
+# empirical YA confidence intervals
+segments(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75)
+arrows(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)
+
+
+188266
+
+empirical_sfs=c(187850/188266, 222/188266, 73/188266, 32/188266, 20/188266, 45/188266, 24/188266, 0, 0, 0, 0, 0, 0)
+sfs=c(0.0756799, 0.0115273, 0.00639306, 0.00517519, 0.00541744)
+options(scipen=0)
+barplot(sfs, ylim=c(0,0.1), ylab="Relative SFS (Qn(i))", xlab="Derived allele frequency", main="coarse grained predictions")
+
+barplot(empirical_sfs, ylim=c(0,0.1))
+
+Original parameters: 127.67 1374 0
+Coarse-grained parameters: 171.123752462 1025.09778728 0
+Calculating time spectrum for parameters Ns=171.124, NU=1025.1, n=6
[…]

### 2016-12-04 code/course_grained.py

@@ -1,0 +1,23 @@
+import coarse_grained_theory
+import recombination_theory
+
+# Asexual population
+
+Ns = 127.67
+NU = 1374
+NR = 0
+n = 6
+
+# for rumex: Ns=127.67, NU=1374 s=6 nruns=10000
+
+# Calculate coarse-grained parameters
+Nseff,NUeff,NUneff = coarse_grained_theory.calculate_effective_params(Ns,NU)
+
+print "Original parameters:", Ns,NU,NR
+print "Coarse-grained parameters:", Nseff,NUeff,0
+
+
+# Calculate the shape of the SFS directly
+Q = coarse_grained_theory.calculate_Q(Ns,NU,n)
+
+print "Qn(i):", Q

### 2016-12-19 code/R/Nov25_2016.Rmd
organization; new edits up to Methods

@@
 ```{r}
-
+xa_given_y=0.0003 # y theta * 3
+ya_given_x=0.00156 #x theta * 1/3
 plot(c(tx_x_auto/0.75,XA_0.6/0.75, xa_given_y/0.75),
@@
 
-xa_given_y=0.0003 # y theta * 3
-ya_given_x=0.00156 #x theta * 1/3
 XA=c(tx_x_auto, XA_0.5, XA_0.6, XA_0.7, XA_0.8, XA_0.9)
@@
 #0.25*(exp(-(0.02/((2*0.002)+0)))
-bgs_exp=0.00168449
+#bgs_exp=0.00168449
 
-barplot(c(tx_y_auto,ya_given_x,bgs_exp), ylim=c(0,0.04), col="gray52", ylab="Y/A",
-        names.arg=c("observed", "predicted from X", "predicted by bgs"), cex.names=0.75)
-segments(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75)
-arrows(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)
+BGS_YA=0.0054
+SEXRATIO_YA=YA_0.6
+OBS_YA=0.018
+
+#barplot(c(OBS_YA/YA_0.6,BGS_YA/YA_0.6), ylim=c(0,0.2), col="gray52", ylab="Y/A ratio relative to the neutral prediction",
+#        names.arg=c("observed", "BGS theory prediction"), cex.names=0.75)
+#segments(x0=0.7,y0=0.009/YA_0.6,x1=0.7,y1=0.03/YA_0.6,lwd=0.75)
+#arrows(x0=0.7,y0=0.009/YA_0.6,x1=0.7,y1=0.035/YA_0.6,lwd=0.75,angle = 90, code = 3, length = 0.04)
+
+par(las=1)
+barplot(c(OBS_YA/YA_0.6,BGS_YA/YA_0.6), ylim=c(0,0.2), col="gray52",ylab=expression(paste(pi/pi[0])),
+        names.arg=c("observed", "BGS theory"), cex.names=0.75)
+segments(x0=0.7,y0=0.009/YA_0.6,x1=0.7,y1=0.035/YA_0.6,lwd=0.75)
+arrows(x0=0.7,y0=0.009/YA_0.6,x1=0.7,y1=0.035/YA_0.6,lwd=0.75,angle = 90, code = 3, length = 0.04)
 

### 2016-12-19 manuscript/manuscript_Dec19.tex
organization; new edits up to Methods

--- 2016-12-04 manuscript/manuscript_Dec3.tex
@@
 X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect patterns of genetic diversity.
-On Y chromosomes with suppressed recombination, natural selection is expected to eliminate neutral variation and lower the $N_{e}$ of Y compared to X chromosomes or autosomes.
-However, non-selective factors including female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity from natural populations.
-Here, we investigate the factors affecting levels of polymorphism during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
-Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of 99.2\% compared to the neutral expectation.
-We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Instead, using forward simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that diversity loss on the Y can be explained by purifying selection acting over a large number ($\geq$ 800 Kb) of genetically-linked sites.
-Our results are in agreement with theory on "interference selection", and provide evidence that this effect can substantially reduce nucleotide diversity on a young Y chromosome.
+On Y chromosomes with suppressed recombination, natural selection is expected to eliminate linked neutral variation and lower the $N_{e}$ of Y compared to X chromosomes or autosomes.
+However, non-selective factors including female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity in natural populations.
+Here, we investigate the processes affecting levels of polymorphism during sex chromosome evolution in the dioecious plant \textit{Rumex hastatulus} (Polygonaceae).
+Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their X-linked homologues, corresponding to a chromosome-wide reduction of 93\% compared to the standard neutral expectation.
+We demonstrate that the magnitude of this diversity loss is inconsistent with reduced male $N_{e}$ caused by neutral processes.
+Instead, using forward simulations and estimates of the fitness effects of deleterious mutations, we show that Y chromosome diversity loss can be explained by purifying selection on a large number ($\geq$ 800 kb) of genetically-linked sites.
+Our results are in agreement with theory on "interference selection", and provide evidence that this can considerably reduce levels of polymorphism on a young plant Y chromosome.
 Given the relatively recent origin of \textit{R.
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-In both groups, for example, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}.
-In some species, such genetic deterioration of the Y chromosome has also led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{bachtrog2013NRG,hough2014,bergero2015}, and in some species, such genetic degeneration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
 The independent evolution of these phenomena in such taxonomically distant groups as mammals and flowering plants suggests that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoretical and experimental biologists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
@@
 One fundamental difference between X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas the  $N_{e}$ for genes on the X chromosome should be 3/4 that of autosomal genes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that the equilibrium level of neutral polymorphism - proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ - should be lower for Y-linked genes than for their X-linked counterparts.
-In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citep{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citep{smith1974hitch}).
-Furthermore, the build-up of genetic associations among selected mutations (high fitness covariance) on the Y means that selection will act non-independently across the chromosome, such that selection at any given focal site "interferes" with selection at the sites with which it is linked \citep{hill1966HReffect}.
-A large body of work has now shown that in theory such "selective interference" effects can substantially reduce both the efficacy of selection and the equilibrium level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}, with the magnitude of the effect depending strongly on the number and density of sites experiencing selection .
-These arguments all suggest that large non-recombining Y chromosomes should harbor a lower amount of neutral variability than predicted based on the number of Y chromosomes in a population (1/4 that of autosomes), and highlight that estimating the number of selected sites is a key test for determining the likely mechanism involved in driving diversity loss on a nonrecombining Y chromosome.
-However, the relative importance of neutral and selective factors for reducing chromosome-wide levels of diversity is not well-understood.
+Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, and 1/3 that of X-linked genes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that the level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citealt{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citealt{smith1974hitch}).
+Furthermore, the build-up of genetic associations among selected mutations (fitness covariance) on the Y means that selection will act non-independently across the chromosome, such that selection at a focal site may "interfere" with selection at linked sites \citep{hill1966HReffect}.
+A large body of work has now shown that such "selective interference" effects are expected to reduce both the efficacy of selection and the equilibrium level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}, with the magnitude of the effect depending on the number and density of sites experiencing selection.
+These arguments all suggest that large non-recombining Y chromosomes with many linked selected sites should harbor a lower amount of neutral variability than predicted based on the number of Y chromosomes in a population (1/4 that of autosomes).
+However, the relative importance of neutral and selective factors for reducing chromosome-wide levels of diversity is not well understood.
 
-In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experienced by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects on diversity of positive of purifying selection need to be distinguished from these neutral processes.
-Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of silent site variability on X and Y chromosome, relative to values on autosomes.
-For example, high variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio compared to the neutral expectation of 1/4, but it also increases the X/A ratio.
-Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is due to a historical excess of breeding females compared to males \citep{hammer2008sex} (but see \citep{bustamante2009,hammer2010,cotter2016genetic}).
+In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected to reduce the $N_{e}$ experienced by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection need to be distinguished from these neutral processes.
+Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can be obtained by comparing levels of silent site variability on X and Y chromosome, relative to values on autosomes.
+For example, variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio relative to the neutral expectation (1/4), but also causes an increases the X/A ratio.
+Based on such comparisons, studies in humans, for example, have suggested that the observation of an inflated X/A ratio can be explained by a historical excess of breeding females compared to males (\citealt{hammer2008sex}; but see \citealt{bustamante2009,hammer2010,cotter2016genetic}).
 
-Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes.
-The time scales over which these different effects are likely to be important are therefore not well understood.
-In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations suggest that levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA) \citep{lahn1999four,ross2005dna}, it is not clear whether purifying selection might have similarly strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last \textasciitilde 20 MYA in the case of plants \citep{charlesworth2015plant}).
+Despite widespread interest in determining the evolutionary factors that affect neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about how sex ratio variation or selective processes have affected levels of diversity on more recently evolved sex chromosomes.
+The time scales over which these processes are likely to be important are therefore not well understood.
+In humans, estimates of silent site diversity on the Y are considerably lower than predicted under neutral evolutionary models, and simulations suggest that observed levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago \citep{lahn1999four,ross2005dna}, it is not clear whether purifying selection might have had a similarly strong effect on Y chromosomes that evolved \textit{do novo} from autosomes much more recently (e.g., within the last \textasciitilde 20 MYA in the case of dioecious plants; \citealt{charlesworth2015plant}).
 
-On the one hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
-Moreover, theory suggests that even weak purifying selection, if acting across a large number of genetically-linked sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
-Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is \textit{a priori} likely to have strong effects on young plant sex chromosomes.
-On the other hand, if there has been widespread gene silencing during the early stages of Y degeneration, as found on the \textit{Drosophila miranda} neo-Y chromosome \citep{zhou2012chromosome}, then Y-linked diversity loss might be low in younger sex chromosome systems.
-In particular, if gene silencing occurs early during Y chromosome degeneration, few sites are expected to be experiencing purifying selection, with degeneration primarily driven by neutral genetic drift rather than inefficient selection.
-Distinguishing the relative roles of selective interference and neutral drift during Y chromosome evolution is a major open issue.
+Simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
+Moreover, theory suggests that even weak purifying selection, when aggregated over a large number of genetically-linked sites, can generate strong deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
+Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is \textit{a priori} likely to be a strong force affecting the evolution of young plant sex chromosomes.
+However, if there has been widespread gene silencing during the early stages of Y-chromosome evolution, as in \textit{Drosophila miranda}, for example \citep{zhou2012chromosome}, then diversity loss might be expected to be lower on younger Y chromosomes.
+That is, if Y-linked gene silencing occurs early during sex chromosome evolution, then few sites may be under selection, and Y chromosome degeneration may be driven primarily by neutral genetic drift rather than inefficient selection.
+Understanding the relative importance of selection and neutral drift during Y chromosome evolution is therefore a major open issue.
 
-To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R.
-hastatulus }(Polygonaceae).
+To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{Rumex hastatulus }(Polygonaceae).
 This species is a dioecious annual with heteromorphic X and Y chromosomes that originated \textasciitilde 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphic in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ( so-called "chromosomal races") \citep{smith1963mechanism}.
-The $XY_{1}XY_{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone gene loss and functional deterioration \citep{hough2014}.
-Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, and to ensure that our diversity estimates are not biased by non-equilibrium conditions or other confounding factors arising from the X-autosome fusion in the $XY_{1}XY_{2}$ system, we focus only on the XY system.
+\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphism in sex chromosome system, in which both XY and $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ male genotypes occur in geographically distinct populations (so-called "chromosomal races"; \citealt{smith1963mechanism}.
+The derived $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the current XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
+Interestingly, despite the recent origin of sex chromosomes in both races, there is evidence that both the ancestral and neo-Y-chromosomes have undergone gene loss and functional deterioration \citep{hough2014}.
+Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, and to ensure that our diversity estimates are not biased by non-equilibrium conditions arising from the X-autosome fusion, we focus only on the XY system with the ancestral chromosome complement.
 
-Of particular importance for our study, \textit{R.
-hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Female-biased sex ratios are not uncommon in dioecious plants with heteromorphic X and Y chromosomes \citep{field2013comparative,hough2013evolutionarily}, and their occurrence in \textit{R.
-hastatulus} provides an opportunity to test both the demographic and selective contributions to changes in sex-linked variability during sex chromosome evolution.
+Of particular relevance to our study, \textit{R.
+hastatulus} populations have been found to exhibit female-biased reproductive sex ratios, with a mean sex ratio of $\approx 0.6$ \citep{pickup2013influence}.
+Indeed, the occurrence of female-biased sex ratios in dioecious flowering plants has been found to be more common in species with heteromorphic X and Y chromosomes than in those with monomorphic sex chromosomes \citep{field2013comparative,hough2013evolutionarily}.
+In \textit{R.
+hastatulus}, the availability of field measurements of sex ratio variation thus provides an opportunity to test both neutral and selective models of Y-chromosome diversity loss during sex chromosome evolution.
 
@@
 We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
-The population samples analyzed here were previously reported in Hough et al.
+The RNAseq data from these population samples were previously reported in Hough et al.
 (2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
 SRP041588.
-Here, to consider sex linked genes that were identified in all of our sequenced population samples, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).
+Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$ chromosome).
 
@@
 We filtered genes in this set to remove any genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
-In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage (Beaudry et al., in prep) data (to filter out duplicates or genes with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
-Finally, we removed genes with fewer than 100 synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
+In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (Beaudry et al., in prep), to filter out duplicates or genes with highly repetitive sequences, and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
+Finally, we removed genes with fewer than 100 filtered synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
 This filtering resulted in a final set of 12,356 autosomal genes.
@@
 (2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level polymorphism, obtained from the genetic cross, was zero.
-We identified 7 \% of sex-linked genes that either had phasing errors genotyping errors.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level of polymorphism, obtained from the genetic cross, was zero.
+We identified 7 \% of sex-linked genes that either had phasing errors or genotyping errors.
 This corresponds to a SNP error rate estimate of 1.7 x 10-4.
@@
 In particular, we evaluated the likelihood of $\theta_{autosomal}$ and \textit{f}, given the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), the number of segregating sites on the Y chromosome ($S_{y}$), and the total number of sites per locus ($L_{i}$ and $L_{y}$).
-The likelihood of ($\theta_{autosomal}$, f) is therefore given by:
+The likelihood of ($\theta_{autosomal}$, \textit{f}) is therefore given by:
 
@@
 Note that when $N_{e}_{f} = N_{e}_{m}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}.
-With a female-biased ratio, however, the expected X/A ratio can become greater than 1, approaching 1.125 in the limit as r\rightarrow 1 \citep{caballero1995} (Figure 1).
+With a female-biased ratio, however, the expected X/A ratio can become greater than 1, approaching 1.125 in the limit as $r\rightarrow 1$ \citep{caballero1995}.
 
@@
 \caption{\textbf{A}.
-Expected effective population sizes, relative to $N$,  for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y) as a function of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+Equilibrium effective population sizes, relative to $N$, for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y) as a function of the sex ratio, expressed as $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
 \textbf{B}.
-The corresponding X/A and Y/A ratios of diversity predicted at equilibrium (where $\theta=4N_{e}\mu$), assuming equal neutral mutation rates for sex-linked and autosomal genes.
+The corresponding X/A and Y/A ratios of nucleotide diversity predicted at equilibrium, assuming equal neutral mutation rates for sex-linked and autosomal genes, where $\theta=4N_{e}\mu$).
 }
@@
 To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
-We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms.
+We first estimated the distribution of fitness effects of deleterious amino acid mutations from our polymorphism data for X-linked genes using the method of \citet*{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms.
 We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta$ from autosomal genes, adjusted to reflect the expected neutral reduction in $N_{e}_{Y}$ for a sex ratio of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.6$.
 To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331bp of linked neutral sequence from which we calculated silent site diversity.
-To examine the expected reduction in diversity as a function of the number of selected sites ($L_{selected}$), we ran simulations over a range of values of $L_{selected}$, up to a maximum of $5x$10^{6}$ (Figure 3A).
+To examine the expected reduction in diversity as a function of the number of selected sites ($L_{selected}$), we ran simulations over a range of values of $L_{selected}$, up to a maximum of 5x$10^{6}$ (Figure 3A).
 To estimate $L_{selected}$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{simulated}$, was less than 0.0001 from our empirical estimate, $\pi_{observed}$ (Figure 3B).
@@
 In particular, neutral diversity for Y-linked genes is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
-Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of 99.2\% .
+Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of 93\% relative to the standard neutral model.
 Note that by normalizing X and Y diversity estimates by autosomal diversity levels, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but a Y-specific reduction.
-Interestingly, our results also suggest an elevation of X-linked diversity ($X/A=0.85$) relative to the neutral prediction ($X/A=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
+Interestingly, our results also suggest an elevation of X-linked diversity ($\textnormal{X/A}=0.85$) relative to the neutral prediction ($\textnormal{X/A}=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
 Given the empirically-estimated sex ratio of 0.6 in \textit{R.
@@
 \midrule
-A & 0.006 & 1 & N/A & 1 \\
-X & 0.0047 & 0.85 & 0.0045 & 0.75 \\
-Y & $\num{e-4}$ & 0.002 & 0.0015 & 0.25 \\
+A & 0.0055 & 1 & N/A & 1 \\
+X & 0.0047 & 0.85 & 0.0041 & 0.75 \\
+Y & $\num{e-4}$ & 0.018 & 0.0014 & 0.25 \\
 \addlinespace
@@
 
-\subsection*{Female biased sex ratios and variance in male offspring number}
+\subsection*{Female Biased Sex Ratios and Variance in Male Reproductive Success}
 The occurrence of female-biased sex ratios in \textit{R.
-hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and thus a reduction the $N_{e}$ of the Y chromosome.
+hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and thus a reduction in the $N_{e}$ of the Y chromosome.
 Male $N_{e}$ could be further reduced by high variance in male reproductive success, which is not unusual in annual wind-pollinated plants such as \textit{R.
 hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production.
-Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may indeed be strong competition among males to fertilize females.
-Such competition should increase levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity.
+Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may indeed be strong competition among males to fertilize ovules.
+Such competition should cause a proportional increase in levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity.
 In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R.
@@
 
-As shown in Figure 2, the expected reduction in the Y/A ratio across the full range of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ is substantially higher than our observed reduction.
-Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.9$, the expectation for the Y/A ratio is $\approx 0.14$, which is substantially higher than our observed estimate ($Y/A= 0.002$) (Table 1; Figure 2; Figure S3).
-Moreover, such large reductions in $N_{e}_{m}$ also predict levels of X/A diversity that are not significantly larger than what we observed (Figure 2).
+As shown in Figure 2, the expected reduction in the Y/A ratio as a function of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ is substantially higher than our observed Y/A diversity ratio.
+Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.9$, the expectation for the Y/A ratio is $\approx 0.14$, which is substantially higher than our observed estimate ($\textnormal{Y/A}=0.002$) (Table 1; Figure 2; Figure S3).
+Moreover, such large reductions in $N_{e}_{m}$ also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
 Our results therefore indicate that, although reduced male $N_{e}$ arising from sex-biased demography is expected to contribute to reduced Y chromosome polymorphism, it is insufficient to explain the Y/A reduction we observed.
@@
 \caption{Observed and predicted X/A and Y/A ratios of neutral diversity.
-Predictions are shown for increasing values of $N_{e}_{f}/N_{e}_{m}$.
+Predictions are shown for increasing values of the effective popuatlion size sex ratio, $N_{e}_{f}$/(N_{e}_{m} + $N_{e}_{f})$.
 Observed estimates of X/A and Y/A were calculated as the average $\theta$ across genes, weighted by the number of synonymous sites per gene.
-Confidence intervals were calculated by bootstrapping (20000 replicates) using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011), and using maximum likelihood for the Y-chromosome (see Methods).
+Confidence intervals were calculated by bootstrapping (20000 replicates) using the BCa method \citep{efron1987better} implemented in the Boot package in R \citep{canty2012boot}, and using maximum likelihood for the Y-chromosome (see Methods).
 }
-\label{fig:spectrum}
+\label{fig:ratios}
 \end{figure}
@@
 Selection against strongly deleterious mutations is expected be an important factor reducing genetic variability in regions with low recombination \citep{charlesworth1993effect, charlesworth1996background}.
-The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for the extensive diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and Drosophila \citep{mcallister1999, charlesworth1996CB}.
-Under background selection theory, the predicted loss of diversity is modeled as modeled as a reduction in the equilibrium $N_{e}$ such that $\pi \approx 4f_{0}N_{e}\mu$, where $f_{0}=e^{-\frac{U}{sh+r}}$, $U$ is the mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate.
-This yields the expected value of $\pi$ relative to that in the absence of BGS ($\pi_{0}$): $E[\pi]= \pi_{0}e^{\frac{-U}{2sh+r}}$, where , $\pi_{0}=\theta=4N_{e}\mu$.
-Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}{m}/2$), the expected reduction in diversity is obtained in this haploid model by replacing $N_{e}$ above with number of breeding males, $N_{e}_{m}$.
-Assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes because of the female biased sex ratio \citep{pickup2013influence} and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations on the X chromosome, the corresponding expected value is $E[\pi]= 6x10^{-5}$ for a mutation rate of $10^{-8}$ per nucleotide per generation.
-Although there is considerable uncertainty in these parameters, the calculation yields a value of diversity that is 67\% lower that what we observed (Table 1).
+The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and \textit{Drosophila} \citep{mcallister1999, charlesworth1996CB}.
+Under background selection theory, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi$ $\approx 4f_{0}N_{e}\mu$, where $f_{0}$=$e^{-\frac{U}{2sh+r}}$, $U$ is the diploid mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate \citep{hudson1995deleterious}.
+This yields the expected value of $\pi$: $E[\pi]=\pi_{0}e^{\frac{-U}{2sh+r}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS $\pi_{0}=\theta=4N_{e}\mu$.
+Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}_{m}/2$), the expected reduction in diversity is obtained in this model by replacing $N_{e}$ with number of breeding males, $N_{e}_{m}$.
+Here, assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
+hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 3x$10^{-5}$ for a mutation rate of 7x$10^{-9}$ per nucleotide per generation.
+Although there is considerable uncertainty in these parameters, the calculation yields an expected value of diversity that is 70\% lower that what we observed (Figure 4).
+
+\begin{figure}[t!]
+\centering
+\noindent
+\includegraphics[width=\linewidth]{figure3.jpg}
+\caption{Results from forward simulations of purifying selection.
+Top: neutral diversity $\pi_{syn}$ as a function of the number of selected sites.
+The dashed line corresponds to our estimate of diversity on the Y chromosome ($\theta_{Y}$).
+Bottom: the relative likelihood curve for the number of selected sites, with the dashed line reflecting the approximate 95\% credibility interval.
+}
+\label{fig:selectedsites}
+\end{figure}
 
@@
 We therefore conducted forward simulations of purifying selection using our estimated parameters for the distribution of fitness effects of deleterious mutations, and tested whether interference selection could result in a level of Y diversity at neutral sites similar to the level we observed.
-We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was a remarkably good fit to our data (Figure 3).
+We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was consistent with our data (Figure 3).
 In addition, and consistent with previous studies, our simulations also suggest a saturating loss of diversity with an increasing number of selected sites \citep{KaiserCharlesworth}.
@@
 \noindent
-\includegraphics[width=\linewidth]{figure3.jpg}
-\caption{Simulation results of the level of neutral diversity $pi_{syn}$ vs.
-the number of selected sites (top panel) with the dashed line corresponding to our point estimate for the level of neutral diversity on the Y chromosome, and the relative likelihood curve for the selected sites (bottom panel), with the dashed line corresponding to our approximate 95\% credibility interval.
+\includegraphics[width=\linewidth]{figure4.jpg}
+\caption{Estimated levels of Y-linked nucleotide diversity compared to estimates obtained from simulations of purifying selection among linked selected sites (HRI simulations) and analytical predictions from background selection theory.
+Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0} =4N_{e}_{m}\mu$, and $N_{e}_{m}$ is the expected value of $N_{e}_{m}$ for a sex ratio of 0.6 (e.g., $N_{e}_{m} = 0.4$x$N_{e}_{autosomes}$), as estimated in \textit{R.
+hastatulus} populations.
+The credibility interval for our observed Y diversity was obtained by maximum likelihood (see methods), and error bars for HRI simulations correspond to 1.5x the interquartile range for $\pi_{s}$ estimates, approximately equal to the 95\% confidence interval for the median.
+\citep{chambers1983graphical}.
 }
-\label{fig:spectrum}
+\label{fig:ydiversity}
 \end{figure}
@@
 hastatulus} sex chromosomes provides clear evidence for an extensive loss of neutral diversity on the Y chromosome.
-Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss on this chromosome, forward population genetic simulations suggest that purifying selection acting over a large number of selected sites results in an expected level of Y-linked diversity that is consistent with what we observed.
+Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss, forward population genetic simulations suggest that purifying selection acting over a large number of selected sites results in an expected level of Y-linked diversity that is consistent with what we observed.
 Our results are consistent with recent theory on interference selection, and suggest that interference is an important force in the evolution of  young Y chromosomes.
 In particular, rather than widespread silencing of Y-linked genes and their subsequent degeneration through increased neutral drift, our study suggests that when a large number of linked sites are subject to purifying selection, Y-chromosome degeneration in the early stages may be largely driven by the effects of selective interference.
+
 \section*{Acknowledgments}
-
+We thank Felix Baudry for providing genomic coverage data for quality filtering.
+This research was funded by Discovery grants to SCHB and SIW) from the Natural Sciences and Engineering Research Council of Canada and by an OGS scholarship to JH.
 \bibliography{bibliography}

### 2016-12-19 code/R/bgs.R

@@ -1,4 +1,28 @@
-print(20*math.exp(-0.02/(0.1+0.005)))
+# Y/A
+s=0.02
+h=0.5
+hs=0.01
+mu=7*10^-9
+U=2*mu
+theta=4*Nem*mu
+
+
+curve(x*(0.5)*(exp(-(U/((2*s*h))))), from=0, to=1,
+      ylab = expression(paste("E[", pi,"]")), las=1,
+      #ylim=c(0,0.3),
+      #xlim=c(0,0.2),lwd=3)
+)
+#abline(h=0.2,lty=2)
+#abline(h=0.02,lty=1, col="red",lwd=3)
+#legend(x=0.12,y=0.3, cex=0.8, box.lty=0, c("Nf/(Nf+Nm) = 0.6","U = 0.02","h = 1","r = 0", expression(paste("", pi[0], paste(" = ",N[em]/2)))))
+
+text(0.03,0.029, cex=1, expression(paste(, theta,)))
+text(0.03,0.21,expression(paste(, pi[0],)))
+
+
+
+
+#print(20*math.exp(-0.02/(0.1+0.005)))
 
 curve(0.25*(exp(-(0.02/((2*x)+0)))), from=0, to=1,
       main="Expected Y/A diversity under BGS",

### 2016-12-19 code/R/nov29_2.R

@@ -32,3 +32,9 @@ Qn(i): [ 1.          0.12343715  0.05657223  0.04769479  0.05992975]
 
 plot(log(0.12343715/1), log(0.05657223/0.12343715), log(0.04769479/0.05657223), log(0.05992975/0.04769479))
 plot(-2.09202, -0.780214, -0.170696, 0.228351)
+
+
+barplot(c(tx_y_auto,ya_given_x,bgs_exp), ylim=c(0,0.04), col="gray52", ylab="Y/A",
+        names.arg=c("observed", "predicted from X", "predicted by bgs"), cex.names=0.75)
+segments(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75)
+arrows(x0=0.7,y0=0.01,x1=0.7,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)

### 2016-12-20 manuscript/manuscript_Dec19.tex
fixed figure 1, new edits up to results

@@
 hastatulus} reference transcriptome \citep{hough2014} using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}.
-We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
-The RNAseq data from these population samples were previously reported in Hough et al.
-(2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
+We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit variant calling software \citep{mckenna2010genome}, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
+The RNAseq data from these population samples were previously reported in \citet{hough2014}, where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
 SRP041588.
@@
 \subsection*{Phasing X and Y alleles}
-To estimate polymorphism for X and Y sequences separately in males, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
+To estimate polymorphism for X and Y sequences separately in males, it is necessary to infer the phase of SNPs in sex-linked transcripts.
 In previous work, phasing alleles on \textit{R.
 hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross.
-Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome.
+Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the polymorphic variants on each chromosome.
 Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences within each haplotype block (i.e., sites where all females were homozygous, and all males were heterozygous).
 Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output.
-In particular, this was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
+In particular, this was done by matching the phase of fixed X-Y differences with neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
@@
 
-We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in Hough et al.
-(2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level of polymorphism, obtained from the genetic cross, was zero.
-We identified 7 \% of sex-linked genes that either had phasing errors or genotyping errors.
-This corresponds to a SNP error rate estimate of 1.7 x 10-4.
+We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in \citet{hough2014}.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level of polymorphism, indicated by the genetic cross, was zero.
+We identified 7\% of sex-linked genes that either had phasing errors or genotyping errors.
+This corresponds to a SNP error rate estimate of 1.7 x $10^{-4}$.
 Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of the X/A ratio.
@@
 \subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
-For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
-To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2).
+For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the diploid effective population size, and $\mu$ is the per-base mutation rate per-generation, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
+To compare sex-linked and autosomal loci, we calculated the average value of $\theta$ for each chromosome type, weighted by the number of synonymous sites in each gene (\X Figure 2).
 We obtained 95\% confidence intervals for X/A and Y/A ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A on each iteration for 20000 replicates each.
@@
 Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the Y/A estimate may be wider than implied by bootstrapping.
-To address this, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+To address this, we also used a maximum likelihood approach to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate was not feasible, as this would require maximizing the likelihood estimate of the mutation rate for each locus independently.
@@
 Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
-Our model thus had two parameters: $\theta_{autosomal}$ and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes.
-We varied both parameters and evaluated the likelihood for \textit{f} from 0.001 to 1, and $\theta_{autosomal}$ per base pair from 0.001 to 0.01.
-In particular, we evaluated the likelihood of $\theta_{autosomal}$ and \textit{f}, given the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), the number of segregating sites on the Y chromosome ($S_{y}$), and the total number of sites per locus ($L_{i}$ and $L_{y}$).
-The likelihood of ($\theta_{autosomal}$, \textit{f}) is therefore given by:
+Our model thus had two parameters: $\theta_{a}$, our estimate of Watterson's theta for autosomal genes, and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes.
+We varied both parameters and evaluated the likelihood of \textit{f} from 0.001 to 1, and $\theta_{a}$ per base pair from 0.001 to 0.01.
+In particular, we evaluated the likelihood of $\theta_{a}$ and \textit{f}, given (1) the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), (2) the number of segregating sites on the Y chromosome ($S_{y}$), and (3) the total number of sites per locus ($L_{i}$ and $L_{y}$).
+The likelihood of ($\theta_{a}$, \textit{f}) is given by:
 
@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to neutral predictions for a range of values for the ratio of effective population sizes of females and males.
-Because we were primarily interested in determining the parameter space within which a reduced male effective population size could result in Y/A and X/A diversity ratios consistent with our data, we tested the fit of our data to predictions across the full range of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ (0.1 to 0.9; we excluded all-female or all-male cases).
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to the corresponding neutral predictions for a given effective population size "sex ratio", $r=N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+Because we were primarily interested in determining the parameter space within which a reduced male $N_{e}$ could result in Y/A and X/A diversity ratios consistent with our data, we tested the fit of our estimates to predictions across the full range of $r$ (0.1 to 0.9; we excluded all-female or all-male cases).
 For autosomal and sex-linked effective population sizes given by:
@@
 
-\citep{wright1931evolution}, the corresponding expected X/A and Y/A ratios an be expressed as:
+\citep{wright1931evolution}, the corresponding expected X/A and Y/A ratios can be expressed as:
 
@@
 
-respectively, where $r = N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+respectively.
 Note that when $N_{e}_{f} = N_{e}_{m}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}.
@@
 \caption{\textbf{A}.
-Equilibrium effective population sizes, relative to $N$, for genes on autosomes (A), X chromosomes (X), and Y chromosomes (Y) as a function of the sex ratio, expressed as $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+Equilibrium effective population sizes, relative to $N$, for genes on autosomes, X chromosomes, and Y chromosomes as a function of the sex ratio.
 \textbf{B}.
-The corresponding X/A and Y/A ratios of nucleotide diversity predicted at equilibrium, assuming equal neutral mutation rates for sex-linked and autosomal genes, where $\theta=4N_{e}\mu$).
+The corresponding X/A and Y/A ratios of nucleotide diversity predicted at equilibrium, assuming equal neutral mutation rates for sex-linked and autosomal genes.
 }
@@
 We first estimated the distribution of fitness effects of deleterious amino acid mutations from our polymorphism data for X-linked genes using the method of \citet*{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms.
-We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta$ from autosomal genes, adjusted to reflect the expected neutral reduction in $N_{e}_{Y}$ for a sex ratio of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.6$.
-To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331bp of linked neutral sequence from which we calculated silent site diversity.
-To examine the expected reduction in diversity as a function of the number of selected sites ($L_{selected}$), we ran simulations over a range of values of $L_{selected}$, up to a maximum of 5x$10^{6}$ (Figure 3A).
-To estimate $L_{selected}$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{simulated}$, was less than 0.0001 from our empirical estimate, $\pi_{observed}$ (Figure 3B).
+We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta_{a}$, adjusted to reflect the neutrally expected $N_{e_{Y}}$ for a sex ratio of $r=0.6$.
+To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331 bp of linked neutral sequence from which we calculated silent site diversity, $\pi_{s}$.
+To examine the expected reduction in diversity as a function of the number of selected sites ($L$), we ran simulations over a range of values of $L$, up to a maximum of 5x$10^{6}$ (Figure 3).
+To obtain an estimate of $L$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}$, was less than 0.0001 from our empirical estimate (Figure 3).
 

### 2016-12-20 manuscript/manuscript_Dec19.tex
final edits to ms, supp info edits

@@
 Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
-hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model.
+hastatulus} reveals a widespread loss of neutral diversity from the Y chromosome.
 In particular, neutral diversity for Y-linked genes is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
-Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of 93\% relative to the standard neutral model.
+Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality, this corresponds to a chromosome-wide reduction of 93\% relative to the standard neutral model \citep{wright1931evolution}.
 Note that by normalizing X and Y diversity estimates by autosomal diversity levels, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but a Y-specific reduction.
-Interestingly, our results also suggest an elevation of X-linked diversity ($\textnormal{X/A}=0.85$) relative to the neutral prediction ($\textnormal{X/A}=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
-Given the empirically-estimated sex ratio of 0.6 in \textit{R.
-hastatulus}\citep{pickup2013influence}, however, the elevation we observed is not unexpected; the neutrally-predicted X/A ratio for a sex ratio of 0.6 is $\approx 0.8$ (Figure 2; Equation 4).
+Interestingly, our results also suggest an elevation of X-linked diversity ($\textnormal{X/A}=0.85$) compared to the neutral prediction ($\textnormal{X/A}=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
+Given the empirically-estimated sex ratio of $r=0.6$ in \textit{R.
+hastatulus} \citep{pickup2013influence}, however, the X/A elevation we observed is not unexpected; the neutrally-predicted X/A ratio for a sex ratio of 0.6 is $\approx 0.8$ (Figure 2; Equation 4).
 Although our estimates of diversity have not been normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R.
@@
 hastatulus}.
-We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values for $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values the $N_{e}$ sex ratio, $r$.
 
-As shown in Figure 2, the expected reduction in the Y/A ratio as a function of $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$ is substantially higher than our observed Y/A diversity ratio.
-Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $N_{e}_{f}/(N_{e}_{f} + N_{e}_{m}) = 0.9$, the expectation for the Y/A ratio is $\approx 0.14$, which is substantially higher than our observed estimate ($\textnormal{Y/A}=0.002$) (Table 1; Figure 2; Figure S3).
-Moreover, such large reductions in $N_{e}_{m}$ also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
+As shown in Figure 2, the expected reduction in Y/A as a function of $r$ is substantially higher than our observed Y/A diversity ratio.
+Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $r=0.9$, the expectation for Y/A is $\approx 0.14$, which is substantially higher than our observed estimate ($\textnormal{Y/A}=0.002$) (Table 1; Figure 2).
+Moreover, such large reductions in $N_{e}_{m}$ would also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
 Our results therefore indicate that, although reduced male $N_{e}$ arising from sex-biased demography is expected to contribute to reduced Y chromosome polymorphism, it is insufficient to explain the Y/A reduction we observed.
@@
 Under background selection theory, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi$ $\approx 4f_{0}N_{e}\mu$, where $f_{0}$=$e^{-\frac{U}{2sh+r}}$, $U$ is the diploid mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate \citep{hudson1995deleterious}.
-This yields the expected value of $\pi$: $E[\pi]=\pi_{0}e^{\frac{-U}{2sh+r}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS $\pi_{0}=\theta=4N_{e}\mu$.
+This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{2sh+r}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS $\pi_{0}=\theta=4N_{e}\mu$.
 Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}_{m}/2$), the expected reduction in diversity is obtained in this model by replacing $N_{e}$ with number of breeding males, $N_{e}_{m}$.
 Here, assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
-hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 3x$10^{-5}$ for a mutation rate of 7x$10^{-9}$ per nucleotide per generation.
-Although there is considerable uncertainty in these parameters, the calculation yields an expected value of diversity that is 70\% lower that what we observed (Figure 4).
+hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 3x$10^{-5}$ for a haploid Y chromosome mutation rate of 7x$10^{-9}$ per nucleotide per generation.
+Although there is considerable uncertainty in these parameters, the calculation yields an expected value of diversity that is 70\% lower that what we observed (Table 1; Figure 2; Figure 4).
 
@@
 
-Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time, because the behavior of selected alleles interferes with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
-As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding this selective interference in realistic situations.
+Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time.
+This is because high linkage disequilibrium causes the behavior of selected alleles to interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
+As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding such selective interference in realistic situations.
 We therefore conducted forward simulations of purifying selection using our estimated parameters for the distribution of fitness effects of deleterious mutations, and tested whether interference selection could result in a level of Y diversity at neutral sites similar to the level we observed.
-We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was consistent with our data (Figure 3).
+We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was consistent with our data (Figure 3; Figure 4).
 In addition, and consistent with previous studies, our simulations also suggest a saturating loss of diversity with an increasing number of selected sites \citep{KaiserCharlesworth}.
@@
 Our maximum likelihood estimates of the number of selected sites, $L$, from our simulations, suggest that the number of sites subject to selection required to explain the diversity loss from the \textit{R.
-hastatulus} Y chromosome is ($\approx 800 kb$) (Figure 3).
+hastatulus} Y chromosome is $\approx 800$ kb (Figure 3).
 
@@
 \caption{Estimated levels of Y-linked nucleotide diversity compared to estimates obtained from simulations of purifying selection among linked selected sites (HRI simulations) and analytical predictions from background selection theory.
-Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0} =4N_{e}_{m}\mu$, and $N_{e}_{m}$ is the expected value of $N_{e}_{m}$ for a sex ratio of 0.6 (e.g., $N_{e}_{m} = 0.4$x$N_{e}_{autosomes}$), as estimated in \textit{R.
+Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0} =4N_{e}_{y}\mu$, and $N_{e}_{y}$ is the expected value of $N_{e}_{y}$ for a sex ratio of $r=0.6$, as estimated in \textit{R.
 hastatulus} populations.
@@
 hastatulus} sex chromosomes provides clear evidence for an extensive loss of neutral diversity on the Y chromosome.
-Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss, forward population genetic simulations suggest that purifying selection acting over a large number of selected sites results in an expected level of Y-linked diversity that is consistent with what we observed.
-Our results are consistent with recent theory on interference selection, and suggest that interference is an important force in the evolution of  young Y chromosomes.
-In particular, rather than widespread silencing of Y-linked genes and their subsequent degeneration through increased neutral drift, our study suggests that when a large number of linked sites are subject to purifying selection, Y-chromosome degeneration in the early stages may be largely driven by the effects of selective interference.
+Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss, forward population genetic simulations suggest that the extensive loss of diversity is most likely due to purifying selection.
+However, while standard background selection theory over-predicted the loss of diversity on the Y chromosome by $\approx$ 70\%, simulations of purifying selection acting over a large number of linked selected sites resulted in a level of Y-diversity that was consistent with what we observed.
+These results are consistent with theory on interference selection, and suggest that the effects of interference is likely an important force in the evolution of young Y chromosomes.
+Moreover, our results imply that when a large number of linked sites are subject to purifying selection, Y-chromosome degeneration in the early stages may largely be driven by the effects of interference rather primarily by widespread silencing of Y-linked genes and their subsequent degeneration through neutral drift.
 

### 2016-12-21 manuscript/manuscript.tex
file organization, final edits to ms

@@
 \usepackage{geometry}
+\usepackage{pdfpages}
+
 \articletype{gos} % article type
 
-\title{Nucleotide Diversity Loss on a Plant Y Chromosome Following Recent Recombination Suppression}
+\title{Hill-Robertson Interference Causes Reduced Genetic Diversity on a Young Plant Y-chromosome}
 
@@
 \begin{abstract}
-X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic diversity.
-On Y chromosomes with suppressed recombination, selection is expected to eliminate neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes.
-However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity.
-Here, we investigate the factors affecting levels of polymorphism within the genome during sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious plant with young sex chromosomes.
-Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corresponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation.
-We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success.
-Instead, using forward simulations, we show that the loss of diversity on the Y can be explained by interference among a large number ($\geq$ 800 Kb) of weakly selected mutations.
-Our results are in agreement with theory on "interference selection", and provide evidence that the effects of purifying selection over a large number of genetically-linked sites can substantially reduce neutral diversity.
-Given the recent origin of \textit{R.
-hastatulus} sex chromosomes (\textasciitilde 15MYA), our results imply that Y chromosome degeneration in the early stages may be largely driven by interference rather than positive selection for gene silencing followed by neutral genetic drift.
+X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect patterns of genetic diversity.
+On Y chromosomes with suppressed recombination, natural selection is expected to eliminate linked neutral variation and lower the $N_{e}$ of Y compared to X chromosomes or autosomes.
+However, non-selective factors including female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity in natural populations.
+Here, we investigate the processes affecting levels of polymorphism during sex chromosome evolution in the dioecious plant \textit{Rumex hastatulus} (Polygonaceae).
+Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their X-linked homologues, corresponding to a chromosome-wide reduction of 93\% compared to the standard neutral expectation.
+We demonstrate that the magnitude of this diversity loss is inconsistent with reduced male $N_{e}$ caused by neutral processes.
+Instead, using forward simulations and estimates of the fitness effects of deleterious mutations, we show that Y chromosome diversity loss can be explained by purifying selection on a large number ($\geq$ 800 kb) of genetically-linked sites.
+Our results are in agreement with theory on "interference selection", and provide evidence that this can considerably reduce levels of polymorphism on a young plant Y chromosome.
+Given the relatively recent origin of \textit{R.
+hastatulus} sex chromosomes, our results imply that Y-chromosome degeneration in the early stages may be largely driven by such interference effects rather than by positive selection for gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, the degeneration of the Y has led to the evolution dosage compensation of the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
-The independent evolution of these phenomena in such taxonomically distant species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
+In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{bachtrog2013NRG,hough2014,bergero2015}, and in some species, such genetic degeneration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+The independent evolution of these phenomena in such taxonomically distant groups as mammals and flowering plants suggests that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoretical and experimental biologists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
 
-One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
-Thus, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, whereas the  $N_{e}$ for genes on the X chromosome should be 3/4 that of autosomal genes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that that the equilibrium level of neutral polymorphism - proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$ - should be lower for Y-linked genes than for their X-linked counterparts.
-In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection) and selective sweeps of strongly beneficial mutations (genetic hitchhiking).
-Furthermore, the build-up of linkage disequilibrium between selected mutations on the Y means that selection will act non-independently across the chromosome such that selection at a focal site may "interfere" with selection at sites with which it is linked.
-Originally developed by Hill & Robertson (1966), a large body of work has now shown that such "selective interference" can substantially reduce both the efficacy of selection and the level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000}.
-These arguments all suggest that non-recombining Y chromosomes should harbor a lower amount of neutral genetic variability than predicted based on the number of Y chromosomes in a population, but the extent to which background selection, genetic hitchhiking, or selective interference have affected chromosome-wide levels of diversity in natural populations, and the relative importance of these processes, is not well-understood.
+One fundamental difference between X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population.
+Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, and 1/3 that of X-linked genes (assuming an equal number of reproducing females and males).
+The lowered $N_{e}$ of the Y chromosome implies that the level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citealt{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citealt{smith1974hitch}).
+Furthermore, the build-up of genetic associations among selected mutations (fitness covariance) on the Y means that selection will act non-independently across the chromosome, such that selection at a focal site may "interfere" with selection at linked sites \citep{hill1966HReffect}.
+A large body of work has now shown that such "selective interference" effects are expected to reduce both the efficacy of selection and the equilibrium level of neutral variability \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}, with the magnitude of the effect depending on the number and density of sites experiencing selection.
+These arguments all suggest that large non-recombining Y chromosomes with many linked selected sites should harbor a lower amount of neutral variability than predicted based on the number of Y chromosomes in a population (1/4 that of autosomes).
+However, the relative importance of neutral and selective factors for reducing chromosome-wide levels of diversity is not well understood.
 
-In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected reduce the $N_{e}$ experience by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects positive of purifying selection need to be distinguished from these neutral processes.
-Because variance in male reproductive success reduces both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, evidence for this can therefore be obtained by comparing levels of neutral diversity on X and Y chromosomes relative to values on autosomes.
-For example, high variance in male reproductive success causes a reduction in the Y/A diversity ratio, but an increase in the X/A ratio.
-Based on such comparisons, studies in humans, for example, have suggested that the inflated X/A ratio is attributable to a historical excess of breeding females over males \citep{hammer2008sex} (and see \citep{bustamante2009,hammer2010,cotter2016genetic}).
+In addition to reduced diversity arising from selection, in species with female-biased sex ratios or extensive male-male competition, high variance in male reproductive success is also expected to reduce the $N_{e}$ experienced by genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of selection need to be distinguished from these neutral processes.
+Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can be obtained by comparing levels of silent site variability on X and Y chromosome, relative to values on autosomes.
+For example, variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio relative to the neutral expectation (1/4), but also causes an increases the X/A ratio.
+Based on such comparisons, studies in humans, for example, have suggested that the observation of an inflated X/A ratio can be explained by a historical excess of breeding females compared to males (\citealt{hammer2008sex}; but see \citealt{bustamante2009,hammer2010,cotter2016genetic}).
 
-Despite widespread interest in determining the evolutionary factors affecting neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about the influence of either sex ratio variation or linked selection in determining levels of diversity on more recently evolved sex chromosomes.
-The time scales over which these different effects are likely to be important is therefore not well understood.
-In humans, estimates of Y-linked diversity are considerably lower than predicted under neutral models, and simulations suggest that levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
-However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago (MYA), is not immediately obvious whether purifying selection might have such strong effects on Y chromosomes that evolved \textit{do novo} from autosomes over much more recent evolutionary time (e.g., within the last 20 MYA in the case of plants \citep{charlesworth2015plant}).
-Y-linked diversity loss might on the one hand be expected to be lower in younger systems due to a shorter history of recombination suppression.
-On the other hand, simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
-Moreover, work by \citep{KaiserCharlesworth} and \citep{good2014genetic} have provided evidence that even weak purifying selection, if acting non-independently across large number of sites, can generate substantial deviations from neutrality, whereas classic background selection theory breaks down in such cases.
-Given that a large number of sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is a good candidate model for exploring the evolutionary dynamics of young plant sex chromosomes.
+Despite widespread interest in determining the evolutionary factors that affect neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about how sex ratio variation or selective processes have affected levels of diversity on more recently evolved sex chromosomes.
+The time scales over which these processes are likely to be important are therefore not well understood.
+In humans, estimates of silent site diversity on the Y are considerably lower than predicted under neutral evolutionary models, and simulations suggest that observed levels of diversity are consistent with strong purifying selection \citep{Wilsonsayres2014}.
+However, given that human sex chromosomes evolved from autosomes \textasciitilde 200 million years ago \citep{lahn1999four,ross2005dna}, it is not clear whether purifying selection might have had a similarly strong effect on Y chromosomes that evolved \textit{do novo} from autosomes much more recently (e.g., within the last \textasciitilde 20 MYA in the case of dioecious plants; \citealt{charlesworth2015plant}).
 
-To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R.
-hastatulus }(Polygonaceae).
-This species is a dioecious annual with heteromorphic X and Y chromosomes that evolved approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an intraspecific polymorphism in sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct populations ("chromosomal races") \citep{smith1963mechanism}.
-The $XY_{1}XY_{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
-Despite the recent origin of sex chromosomes in both races, there is evidence that both Y's have undergone significant gene loss and functional deterioration \citep{hough2014}.
-Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, we focus only on the ancestral Y chromosome, which occurs in both sex chromosome races.
+Simulations of strong selection models (background selection and genetic hitchhiking) suggest that these processes may have the greatest effects during the earliest stages of sex chromosome evolution, before the Y has lost many of its genes \citep{bachtrog2008temporal}.
+Moreover, theory suggests that even weak purifying selection, when aggregated over a large number of genetically-linked sites, can generate strong deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
+Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is \textit{a priori} likely to be a strong force affecting the evolution of young plant sex chromosomes.
+However, if there has been widespread gene silencing during the early stages of Y-chromosome evolution, as in \textit{Drosophila miranda}, for example \citep{zhou2012chromosome}, then diversity loss might be expected to be lower on younger Y chromosomes.
+That is, if Y-linked gene silencing occurs early during sex chromosome evolution, then few sites may be under selection, and Y chromosome degeneration may be driven primarily by neutral genetic drift rather than inefficient selection.
+Understanding the relative importance of selection and neutral drift during Y chromosome evolution is therefore a major open issue.
 
-Of particular importance for our study, \textit{R.
-hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}.
-Female-biased sex ratios are not uncommon in dioecious plants with heteromorphic X and Y chromosomes \citep{field2013comparative,hough2013evolutionarily}, and their occurrence in \textit{R.
-hastatulus} provides an excellent opportunity to study both the demographic and selective factors contributing to sex-linked variability.
+To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution, we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{Rumex hastatulus }(Polygonaceae).
+This species is a dioecious annual with heteromorphic X and Y chromosomes that originated \textasciitilde 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
+\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphism in sex chromosome system, in which both XY and $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ male genotypes occur in geographically distinct populations (so-called "chromosomal races"; \citealt{smith1963mechanism}.
+The derived $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the current XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
+Interestingly, despite the recent origin of sex chromosomes in both races, there is evidence that both the ancestral and neo-Y-chromosomes have undergone gene loss and functional deterioration \citep{hough2014}.
+Here, to simplify our comparison of polymorphism levels on X, Y, and autosomes, and to ensure that our diversity estimates are not biased by non-equilibrium conditions arising from the X-autosome fusion, we focus only on the XY system with the ancestral chromosome complement.
+
+Of particular relevance to our study, \textit{R.
+hastatulus} populations have been found to exhibit female-biased reproductive sex ratios, with a mean sex ratio of $\approx 0.6$ \citep{pickup2013influence}.
+Indeed, the occurrence of female-biased sex ratios in dioecious flowering plants has been found to be more common in species with heteromorphic X and Y chromosomes than in those with monomorphic sex chromosomes \citep{field2013comparative,hough2013evolutionarily}.
+In \textit{R.
+hastatulus}, the availability of field measurements of sex ratio variation thus provides an opportunity to test both neutral and selective models of Y-chromosome diversity loss during sex chromosome evolution.
 
@@
 \subsection*{Population Samples and Sex-Linked Genes}
-We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 12 males and 12 females (1 male and 1 female from each of 6 populations).
+We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 12 population samples (1 male and 1 female from each of 6 populations).
 Samples were collected in 2010 from throughout the native range of \textit{R.
@@
 hastatulus} reference transcriptome \citep{hough2014} using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}.
-We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
-The population samples analyzed here were previously reported in Hough et al.
-(2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
+We used Picard tools (http://picard.sourceforge.net) to process mapping alignments for the Genome Analysis Toolkit variant calling software \citep{mckenna2010genome}, and subsequently removed genes with low coverage (<10x) and low Phred Quality Scores <20.
+The RNAseq data from these population samples were previously reported in \citet{hough2014}, where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis, and raw sequences are available from the GenBank Short Read Archive under accession no.
 SRP041588.
-Here, to consider sex linked genes that were identified in all of our sequenced population samples, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).
+Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$ chromosome).
 
@@
 We filtered genes in this set to remove any genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
-In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicates or genes with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
-Finally, we removed genes with fewer than 100 synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
+In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes where SNPs showed fixed heterozygosity in males and fixed homozygosity in females, (iii) genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (Beaudry et al., in prep), to filter out duplicates or genes with highly repetitive sequences, and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females.
+Finally, we removed genes with fewer than 100 filtered synonymous sites to avoid biasing our results toward genes that may have been particularly short due to assembly problems.
 This filtering resulted in a final set of 12,356 autosomal genes.
@@
 \subsection*{Phasing X and Y alleles}
-To estimate polymorphism for X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males.
+To estimate polymorphism for X and Y sequences separately in males, it is necessary to infer the phase of SNPs in sex-linked transcripts.
 In previous work, phasing alleles on \textit{R.
 hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross.
-Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome.
+Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the polymorphic variants on each chromosome.
 Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences within each haplotype block (i.e., sites where all females were homozygous, and all males were heterozygous).
 Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output.
-In particular, this was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
+In particular, this was done by matching the phase of fixed X-Y differences with neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, then the variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched.
 SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data.
-Finally, we filtered out SNPs with coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels.
-This filtering procedure resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes.
+Finally, we filtered out SNPs with quality scores < 60, and those within a distance of 10bp or less from indels.
+This filtering procedure resulted in alignments of X and Y sequences for 372 sex-linked genes.
 
-We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in Hough et al.
-(2014).
-To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level polymorphism, obtained from the genetic cross, was zero.
-We identified 7 \% of sex-linked genes that either had  phasing errors of this kind genotyping errors.
-This corresponds to a SNP error rate estimate of 1.7 x 10-4.
+We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in \citet{hough2014}.
+To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the true level of polymorphism, indicated by the genetic cross, was zero.
+We identified 7\% of sex-linked genes that either had phasing errors or genotyping errors.
+This corresponds to a SNP error rate estimate of 1.7 x $10^{-4}$.
 Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of the X/A ratio.
-However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we further filtered genes in which we found evidence for false-positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
+However, because this rate is high relative to the expected level of polymorphism on the Y chromosome, we conducted a further filtering step for Y-linked SNPs and identified false positive SNP calls arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression.
 This final filtering was conducted by manually checking each individual putative polymorphism on the Y chromosome using IGV \citep{robinson2011integrative}.
@@
 \subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
-For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
-To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2).
+For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the diploid effective population size, and $\mu$ is the per-base mutation rate per-generation, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
+To compare sex-linked and autosomal loci, we calculated the average value of $\theta$ for each chromosome type, weighted by the number of synonymous sites in each gene (\X Figure 2).
 We obtained 95\% confidence intervals for X/A and Y/A ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A on each iteration for 20000 replicates each.
 Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes.
-Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the estimate Y/A may be wider than implied by bootstrapping results.
-To address this, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+
+Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the Y/A estimate may be wider than implied by bootstrapping.
+To address this, we also used a maximum likelihood approach to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate was not feasible, as this would require maximizing the likelihood estimate of the mutation rate for each locus independently.
+However, previous analysis of divergence data does not suggest important chromosomal differences in synonymous divergence contributing to diversity heterogeneity (see Results).
 Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
-Our model thus had two parameters: $\theta_{autosomal}$ and Y/A.
-We varied both parameters and evaluated the likelihood for Y/A from 0.001 to 1, and $\theta_{autosomal}$ from 0.001 to 0.01.
+Our model thus had two parameters: $\theta_{a}$, our estimate of Watterson's theta for autosomal genes, and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes.
+We varied both parameters and evaluated the likelihood of \textit{f} from 0.001 to 1, and $\theta_{a}$ per base pair from 0.001 to 0.01.
+In particular, we evaluated the likelihood of $\theta_{a}$ and \textit{f}, given (1) the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), (2) the number of segregating sites on the Y chromosome ($S_{y}$), and (3) the total number of sites per locus ($L_{i}$ and $L_{y}$).
+The likelihood of ($\theta_{a}$, \textit{f}) is given by:
+
+\begin{equation}
+L \propto  P(S_{Y}, L_{Y} | f, \theta_{a})\prod_{i=1}^{n}P(S_{i}, L_{i} | \theta_{a})  \label{eq:L}
+\end{equation}
 
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, we compared our estimated levels diversity to predictions from a model of the expected $N_{e}$ of males and females when the variance in offpring number is either smaller or larger than Poisson \citep{kimura1964number, hedrick2011genetics}.
-As we were primarily interested in the possible influence of variance in male reproductive success on Y-linked $N_{e}$, and the parameter space within which such a model might generate Y/A and X/A ratios consistent with our data, we ignored non-poission variance in female offspring, and calculated the male $N_{e}$ as:
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to the corresponding neutral predictions for a given effective population size "sex ratio", $r=N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+Because we were primarily interested in determining the parameter space within which a reduced male $N_{e}$ could result in Y/A and X/A diversity ratios consistent with our data, we tested the fit of our estimates to predictions across the full range of $r$ (0.1 to 0.9; we excluded all-female or all-male cases).
+For autosomal and sex-linked effective population sizes given by:
 
 \begin{equation}
-N_{e_{m}}=\frac{N_{m}k_{m}-1}{k_{m}-1+\frac{V_{k_{m}}}{\overline{k}}} \label{eq:Vk}
-\end{equation}
-
-where $k_{m}$ and $V_{k_{m}}$ are the mean and variance in male offspring number, respectively, in a population of $N_{m}$ reproducing males, and $\overline{k}$ is the total mean number of progeny \citep{kimura1964number}.
-Given expected values of $N_e_{m}$ for a range of values for $V_{k_{m}}$ and $N_{m}$, the corresponding expectation for the $N_{e}$ of autosomes and sex chromosomes at equilibrium were calculated as:
-
-\begin{equation}
-N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}} \label{eq:Ne}
+N_{e_{A}} = \frac{4N_{e}_{m}N_{e}_{f}}{N_{e}_{m}+N_{e}_{f}} \label{eq:NeA}
 \end{equation}
@@
 \begin{equation}
-N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}} \label{eq:NeX}
+N_{e_{X}} = \frac{9N_{e}_{m}N_{e}_{f}}{4N_{e}_{m}+2N_{e}_{f}} \label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e{Y}} = \frac{N_{m}}{2} \label{eq:NeY}
+N_{e_{Y}} = \frac{N_{e}_{m}}{2} \label{eq:NeY}
 \end{equation}
 
-following \citep{wright1931evolution}.
-Note that with equal sex ratios $N_{f} = N_{m}$, and Poisson-distributed offspring numbers ($V_{k} = \overline{k}$), then $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$.
-Figure 1 shows these predictions as a function of the sex ratio for a range values of $V_{k_{m}}$, and we tested the fit of our empirical diversity ratios to these predictions (shown in Figure 2) assuming the estimated \textit{R.
-hastatulus} sex ratio of 0.6 \citep{pickup2013influence}.
+\citep{wright1931evolution}, the corresponding expected X/A and Y/A ratios can be expressed as:
+
+\begin{equation}
+\frac{N_{e_{X}}}{N_{e_{A}}} = \frac{9 (-1 + r)}{8 (1 - r) (-2 + r))} \label{eq:XA}
+\end{equation}
+
+and
+\begin{equation}
+\frac{N_{e_{Y}}}{N_{e_{A}}} = \frac{1}{8r} \label{eq:YA}
+\end{equation}
+
+respectively.
+Note that when $N_{e}_{f} = N_{e}_{m}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}.
+With a female-biased ratio, however, the expected X/A ratio can become greater than 1, approaching 1.125 in the limit as $r\rightarrow 1$ \citep{caballero1995}.
 
@@
 \includegraphics[width=\linewidth]{figure1.jpg}
-\caption{The relation between effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and the X chromosome (\textbf{B}), and corresponding normalized X/A and Y/A ratios (\textbf{C}).
-The sex ratio is shown as the proportion of females, $N_{f}/(N_{f}+N_{m})$, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$.
-Dotted curves show equilibrium predictions when both males and females produce Poisson-distributed offspring numbers and ($V_{k}$=$\overline{k}=2$).
-Solid curves correspond to increasing values for $V_{k_{m}}$, the variance in male offspring number, ranging from Poisson/2 (top solid curve in panels A and B) to 3*Poisson (bottom most curve).
-For (\textbf{C}) we assumed equal neutral mutation rates among genes and calculated the expected neutral diversity as $\theta=4N_{e}\mu$.
+\caption{\textbf{A}.
+Equilibrium effective population sizes, relative to $N$, for genes on autosomes, X chromosomes, and Y chromosomes as a function of the sex ratio.
+\textbf{B}.
+The corresponding X/A and Y/A ratios of nucleotide diversity predicted at equilibrium, assuming equal neutral mutation rates for sex-linked and autosomal genes.
 }
-\label{fig:spectrum}
 \end{figure}
@@
 \subsection*{Simulations of purifying selection}
-To study the effects of purifying selection on expected levels of Y-chromosome diversity, we used a similar approach to \citep{Wilsonsayres2014} and conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
-We first estimated the distribution of fitness effects of deleterious mutations from our polymorphism data for X-linked genes using the method of \citep{keightley2007joint}, which fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism.
-We used this estimated gamma distribution to parameterize the simulations, initializing each with our estimated $\theta$ from autosomal genes, but appropriately adjusted to reflect the expected reduction in Y-linked $N_{e}$ for a sex ratio of $N_{f}/(N_{f}+N_{m})=0.6$.
-To match our sample size and the number of synonymous sites sample from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and simulated sequences contained 45,331bp of linked neutral sequence from which we calculated diversity.
-To examine the expected reduction in diversity relative to neutrality ($\pi/\pi_{0}$) as a function of the number of selected sites ($L$), and to estimate this parameter from our data, we ran simulations over a range of values of $L$ (up to a maximum of $5x10^{6}$) (Figure 3A).
-For each simulation, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}_{simulated}$, matched our empirical estimate, $\pi_{s}_{observed}$ (Figure 3B).
+To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
+We first estimated the distribution of fitness effects of deleterious amino acid mutations from our polymorphism data for X-linked genes using the method of \citet*{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms.
+We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta_{a}$, adjusted to reflect the neutrally expected $N_{e_{Y}}$ for a sex ratio of $r=0.6$.
+To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331 bp of linked neutral sequence from which we calculated silent site diversity, $\pi_{s}$.
+To examine the expected reduction in diversity as a function of the number of selected sites ($L$), we ran simulations over a range of values of $L$, up to a maximum of 5x$10^{6}$ (Figure 3).
+To obtain an estimate of $L$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}$, was less than 0.0001 from our empirical estimate (Figure 3).
 
@@
 Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
-hastatulus} reveals that neutral diversity on the Y chromosome is substantially lower than expected from a standard neutral model.
-In particular, neutral diversity on Y-linked genes $\theta_{Y}$ is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome, and \textasciitilde 2.1\% of the value of autosomal genes.
-Talking 1/4 of mean autosomal diversity as the equilibrium expectation under neutrality \citep{wright1931evolution}, this corresponds to a chromosome-wide reduction of \textasciitilde 93\% ($\pi/\pi_{0}$).
-By comparing X and Y diversity estimates to that estimated on autosomal genes, our results indicate that the difference in diversity between X and Y homologues is not due to an elevation of diversity on the X chromosome, but rather a Y-specific reduction.
-Our results to suggest a slight elevation in X-linked diversity relative to autosomes ($X/A=$), but this difference is not significant different from expectations under equal sex ratios (P), or the empirically-estimated ratio of 0.6 (P).
-Although our estimates of diversity have not beed normalized by divergence, previous estimates of the average synonymous substitution rate, weighted by the number of synonymous sites, between \textit{R.
+hastatulus} reveals a widespread loss of neutral diversity from the Y chromosome.
+In particular, neutral diversity for Y-linked genes is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
+Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality, this corresponds to a chromosome-wide reduction of 93\% relative to the standard neutral model \citep{wright1931evolution}.
+Note that by normalizing X and Y diversity estimates by autosomal diversity levels, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but a Y-specific reduction.
+Interestingly, our results also suggest an elevation of X-linked diversity ($\textnormal{X/A}=0.85$) compared to the neutral prediction ($\textnormal{X/A}=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
+Given the empirically-estimated sex ratio of $r=0.6$ in \textit{R.
+hastatulus} \citep{pickup2013influence}, however, the X/A elevation we observed is not unexpected; the neutrally-predicted X/A ratio for a sex ratio of 0.6 is $\approx 0.8$ (Figure 2; Equation 4).
+Although our estimates of diversity have not been normalized by divergence, previous work has shown that the average synonymous substitution rate, between \textit{R.
 hastatulus} and the non-dioecious outgroup \textit{R.
-bucephalophorus} revealed similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.
+bucephalophorus} is not significantly different for sex-linked (0.2016) and autosomal genes (0.219), and we found no evidence for significant differences in substitution rate between Y and X chromosomes \citep{hough2014}.
 It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
 
-%After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig).
-Surprisingly, this analysis revealed that the Y chromosome from this race appears to be paraphyletic, with XYY samples from Florida being more closely related to the TX race than either is to samples from South Carolina (suppfig).
-Once we examine diversity for these two subclades separately, a strong loss of diversity on the Y chromosome is once again apparent (Table 1).
+\begin{table}[htb]
+\centering
+\caption{Observed and neutrally-expected silent site diversity on sex chromosomes and autosomes in \textit{R. hastatulus}}
+\begin{tabular}{ccccccccc}
+\textbf{} & \multicolumn{2}{l}{\textbf{Observed}} & \multicolumn{3}{l}{\textbf{Expected}} \\
 
-%Although our sampling of each \textit{R.
-hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a secondarily-derived $XY_{1}Y_{2}$ sub-clade.
-As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the Florida sub-clade inferred here likely originated through hybridization between a female from the SC clade harboring the X-autosome fusion, and a male from the XY Texas race.
+chr & $\theta$ & $/\theta_{A}$ & $\theta$ & $/\theta_{A}$  \\
+\midrule
+A & 0.0055 & 1 & N/A & 1 \\
+X & 0.0047 & 0.85 & 0.0041 & 0.75 \\
+Y & $\num{e-4}$ & 0.018 & 0.0014 & 0.25 \\
+\addlinespace
 
-%\begin{table}[htb]
-%\centering
-%\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
-%\begin{tabular}{ccccccccc}
-%\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
-%chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
-%\midrule
-%A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
-%X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
-%Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
-%\addlinespace
+\bottomrule
+\end{tabular}
+\end{table}
 
-%\bottomrule
-%\end{tabular}
-%\end{table}
+\subsection*{Female Biased Sex Ratios and Variance in Male Reproductive Success}
+The occurrence of female-biased sex ratios in \textit{R.
+hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and thus a reduction in the $N_{e}$ of the Y chromosome.
+Male $N_{e}$ could be further reduced by high variance in male reproductive success, which is not unusual in annual wind-pollinated plants such as \textit{R.
+hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production.
+Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may indeed be strong competition among males to fertilize ovules.
+Such competition should cause a proportional increase in levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity.
+In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R.
+hastatulus}.
+We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values the $N_{e}$ sex ratio, $r$.
 
-%Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2).
-Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}.
-Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race.
-If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can drive the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades.
-However, this would require extensive recombination suppression between the fused and unfused X chromosomes, and it is also possible that sex-specific demographic history has driven these patterns.
-It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.
-
-In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 50 fold reduction compared to the mean $\theta_{Aut}$.
-We next consider several possible models - neutral and selective - that might explain this reduction.
-
-\subsection*{Female biased sex ratio and high variance in male offspring number}
-
-The occurrence of female-biased sex ratios in \textit{R.
-hastatulus} is expected to lower Y diversity through lowering male $N_{e}$ and therefore the $N_{e}$ of the Y chromosome.
-In addition, the $N_{e}_{Y}$ might be further reduced due to high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R.
-hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977).
-Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there is reason to believe that there could be strong competition among males.
-
-%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).
+As shown in Figure 2, the expected reduction in Y/A as a function of $r$ is substantially higher than our observed Y/A diversity ratio.
+Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $r=0.9$, the expectation for Y/A is $\approx 0.14$, which is substantially higher than our observed estimate ($\textnormal{Y/A}=0.002$) (Table 1; Figure 2).
+Moreover, such large reductions in $N_{e}_{m}$ would also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
+Our results therefore indicate that, although reduced male $N_{e}$ arising from sex-biased demography is expected to contribute to reduced Y chromosome polymorphism, it is insufficient to explain the Y/A reduction we observed.
 
@@
 \includegraphics[width=\linewidth]{figure2.jpg}
-\caption{new figure coming soon).
+\caption{Observed and predicted X/A and Y/A ratios of neutral diversity.
+Predictions are shown for increasing values of the effective popuatlion size sex ratio, $N_{e}_{f}$/(N_{e}_{m} + $N_{e}_{f})$.
+Observed estimates of X/A and Y/A were calculated as the average $\theta$ across genes, weighted by the number of synonymous sites per gene.
+Confidence intervals were calculated by bootstrapping (20000 replicates) using the BCa method \citep{efron1987better} implemented in the Boot package in R \citep{canty2012boot}, and using maximum likelihood for the Y-chromosome (see Methods).
 }
-\label{fig:spectrum}
+\label{fig:ratios}
 \end{figure}
 
-In common with most flowering plants we do not have marker-based estimates of the variance in male reproductive success in \textit{R.
-hastatulus}.
-However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could jointly explain the level of Y/A diversity that we observed (see Methods).
-Conditioning on estimates of sex ratio bias in \textit{R.
-hastatulus}, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3).
-This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), and remains significant even if we consider an upper bound estimate obtained from our likelihood based estimated of the confidence interval.
-This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.
+\subsection*{Purifying selection}
+Selection against strongly deleterious mutations is expected be an important factor reducing genetic variability in regions with low recombination \citep{charlesworth1993effect, charlesworth1996background}.
+The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and \textit{Drosophila} \citep{mcallister1999, charlesworth1996CB}.
+Under background selection theory, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi$ $\approx 4f_{0}N_{e}\mu$, where $f_{0}$=$e^{-\frac{U}{2sh+r}}$, $U$ is the diploid mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate \citep{hudson1995deleterious}.
+This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{2sh+r}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS $\pi_{0}=\theta=4N_{e}\mu$.
+Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}_{m}/2$), the expected reduction in diversity is obtained in this model by replacing $N_{e}$ with number of breeding males, $N_{e}_{m}$.
+Here, assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
+hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 3x$10^{-5}$ for a haploid Y chromosome mutation rate of 7x$10^{-9}$ per nucleotide per generation.
+Although there is considerable uncertainty in these parameters, the calculation yields an expected value of diversity that is 70\% lower that what we observed (Table 1; Figure 2; Figure 4).
 
-Assuming that there is extensive variance in male reproductive success (\X), the predicted ratios of Y/A diversity (\X supp) were also significantly higher than our estimates for the empirically estimated sex ratio (\X table) .
-We also find, however, that purely neutral models in which the sex ratio was highly female-biased (\X), with level of variance in male reproductive success on the order of (\X), predicted a Y/A ratio that could not be rejected (\X).
-However, because a highly females biased sex ratio is expected to increase the X/A ratio as well, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure).
-Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity.
+\begin{figure}[t!]
+\centering
+\noindent
+\includegraphics[width=\linewidth]{figure3.jpg}
+\caption{Results from forward simulations of purifying selection.
+Top: estimated silent site diversity $\pi$ relative to neutrality ($\pi_{0}$) as a function of the number of selected sites.
+The dashed line corresponds to our estimate of diversity on the Y chromosome ($\theta_{Y}$).
+Bottom: the relative likelihood curve for the number of selected sites, with the dashed line reflecting the approximate 95\% credibility interval.
+}
+\label{fig:selectedsites}
+\end{figure}
 
-\subsection*{Background Selection and Selective Sweeps}
+Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time.
+This is because high linkage disequilibrium causes the behavior of selected alleles to interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
+As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding such selective interference in realistic situations.
+We therefore conducted forward simulations of purifying selection using our estimated parameters for the distribution of fitness effects of deleterious mutations, and tested whether interference selection could result in a level of Y diversity at neutral sites similar to the level we observed.
+We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was consistent with our data (Figure 3; Figure 4).
+In addition, and consistent with previous studies, our simulations also suggest a saturating loss of diversity with an increasing number of selected sites \citep{KaiserCharlesworth}.
 
-This approach was implemented rather than using analytical predictions of background selection for a few reasons.
-First,...
-because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.
+Although we initialized our simulations with selection parameters estimated from the distribution of fitness effects of deleterious mutations, recent work has shown that, unlike with BGS, evolutionary dynamics in the “interference selection regime” are insensitive to differences in strength of selection, with expected diversity loss instead determined primarily by the population fitness variance over the non-recombining region, which is in turn expected to increase with the number of linked loci under selection \citep{good2014genetic}.
+Estimating the number of selected loci is therefore key to determining whether diversity on the Y chromosome may be affected by interference selection.
+Our maximum likelihood estimates of the number of selected sites, $L$, from our simulations, suggest that the number of sites subject to selection required to explain the diversity loss from the \textit{R.
+hastatulus} Y chromosome is $\approx 800$ kb (Figure 3).
 
-discussion points to make:
-- Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
+\begin{figure}[t!]
+\centering
+\noindent
+\includegraphics[width=\linewidth]{figure4.jpg}
+\caption{Estimated levels of Y-linked nucleotide diversity compared to estimates obtained from simulations of purifying selection among linked selected sites (HRI simulations) and analytical predictions from background selection theory.
+Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0} =4N_{e}_{Y}\mu$, and $N_{e}_{Y}$ is the expected value of $N_{e}_{Y}$ for a sex ratio of $r=0.6$, as estimated in \textit{R.
+hastatulus} populations.
+The credibility interval for our observed Y diversity was obtained by maximum likelihood (see methods), and error bars for HRI simulations correspond to 1.5x the interquartile range for $\pi_{s}$ estimates, approximately equal to the 95\% confidence interval for the median.
+\citep{chambers1983graphical}.
+}
+\label{fig:ydiversity}
+\end{figure}
 
-BGS predictions:
--"Traditionally, the term ‘‘background selection’’ is used to refer both to the general effects of purifying selection on linked neutral diversity as well as to the limiting behavior that emerges when Ns-->inf.
-"
-- "existing theory struggles to predict genetic diversity when many sites experience selection at the same time, which limits our ability to interpret variation in DNA sequence data.". In particular,
-
-- "For non-crossover regions, both with and without gene conversion, there is a rapid initial decline of B with L, but B levels off at a value of  0.015 for > 640 000 sites.
-B values predicted by the BGS model decrease log-linearly with L"
-
-Very relevant remarks From Good et al.:
-
-- "Here, we have shown that simple behavior emerges in the limit of widespread interference.
-When fitness variation is composed of many individual mutations, the magnitudes and signs of their fitness effects are relatively unimportant.
-Instead, molecular evolution is controlled by the variance in fitness within the population over some effectively asexual segment of the genome" " In other words, we cannot conclude that interference is negligible just because ‘‘Nes’’, as inferred from data, is larger than one."
-
-- estimates of "Nes" ignore linkage by fiat under the assumption that sites evolve independently.
-But these estimates become unreliable precisely when small- and intermediate-effect mutations are most common
-
-- "Individual fitness effects may play a central role in single-site models, but we have shown that global properties like the variance in fitness and the corresponding linkage scale are more relevant for predicting evolution in interfering populations.
-"
-
-- "We have provided further evidence that even weak purifying selection, when aggregated over a sufficiently large number of sites, can generate strong deviations from neutrality. ""
-
-- "Apart from an overall reduction in polymorphism, the most prominent features of this frequency spectrum include an excess of rare alleles "
-
-- "silent site diversity decays as p=p0*1=Ns, while the shape of the site-frequency spectrum, Qn(i), becomes independent of all underlying parameters."
-
-- " we showed that the reduction in silent site diversity on this chromosome (p=p0*7) is consistent with the parameters Ns<30, NU<300, and NR<0, which fall in the middle of the interference selection regime"
-
-- Indeed, simulations suggest that the effects of HRI on 100,000 linked sites can reduce polymorphism levels by \textasciitilde 30 \% compared to neutral expectations \citep{mcvean2000}, suggesting that HRI is likely important for the evolution of Y chromosomes.
-
-- iven that a large number of sites in the genome likely experience weak purifying selection (\X refs), the "interference" effect on Y chromosomes with suppressed recombination might alone cause a substantial diversity loss.
-
-Classical population genetics fails to account for int
-erference between linked mutations, which grows increasingly severe as the density of selected polymorphisms increases.
-
-...there is no general analytical treatment describing the dynamics of interference among many alleles under selection and drift, with variable degrees of selection and linkage.
-Thus, forward computer simulations are still fundamental to the study of realistic situations.
-- dfe for the y is probably still fairly similar to that for the x -- shared genes
-
-- human Y: genes on y are not on X
-
-- 45331 = synsites in observed data after filtering
-  - summary stats phased texas autospy
-
-- doesnt matter about coding/non-coding in non-recombining regions? all of those neutral sites are linked
-
-%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
-
-%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
-
-%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
-
-%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
-
-%The agreement is reasonably good given the uncertainties involved.
-
-%silent site diversity for 22 genes on the D. miranda neo-Y is
-%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
-
-% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
-So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
-%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].
-
-%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations
-
-%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.
-
-%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%
-
-%The agreement is reasonably good given the uncertainties involved.
-
-%silent site diversity for 22 genes on the D. miranda neo-Y is
-%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome
-
-% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection.
-So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
-
-%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.
-Given the recent origin of \textit{R.
-hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}
-
-%A predicted consequence of a reduction in Ne is an accelerated rate of ¢xation of slightly deleterious substitu- tions, and a reduction in the rate of ¢xation of slightly advantageous mutations (Kimura 1983).
-Thus we might expect an increased rate of amino-acid replacement rela- tive to silent substitutions on evolving Y chromosomes (provided that su¤cient time has elapsed since the origin of the sex-chromosome system), and more ¢xations of non-preferred versus preferred synonymous substitutions at silent sites in exons (Charlesworth 1996a).
-
-%It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems.
-Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity.
-However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R.
-hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.
-
-% A further interesting aspect of the results from the S.
-latifolia Y chromosome is that they do not readily ¢t a recent selective sweep.
-Such an event would eliminate all diversity, so that existing SLY1 polymorphisms must have accumulated since the selective sweep and should thus mostly be at low frequencies (see table 1).
-However, all six SLY1 polymorphic sites found are non-singletons, and no signi¢cant deviations from neutrality are detectable by standard test statistics (Fu & Li 1993).
-However, this does not take into account the probable population subdivision in this species.
-Although a selective sweep eliminating variability throughout the worldwide population of this species seems unlikely, e¡ects of selective sweeps within local populations cannot be excluded.
-
-%because the lack of recombination means that all genes on the Yof any given species share a common genealogy, which limits the amount of information that can be obtained from data on within-species variability.
-
-%This raises the further issue of whether Y-chromosome degeneration is driven mainly by the accumulation of deleterious mutations on the Y or is due to processes turning down the expression of genes on an evolving Y.
-
-%We have focused on the ¢rst of these possibilities, but it might well not completely explain Y degeneration.
-If mutations of relatively minor ¢tness e¡ects, with Nes not much bigger than 1, become ¢xed more rapidly on a degenerating Ychromosome than on the X, selection pres- sure would develop to reduce the transcription of Y genes, provided that this was compensated for in the hetero- gametic sex by an increased activity of their counterparts on the X.
+Although our analyses of purifying selection models are consistent with our observed reduction in Y-linked diversity, it is also possible that selective sweeps have contributed.
+We have not formally excluded this possibility.
+However, if patterns of molecular evolution on the Y are driven primarily by interference selection, as our results suggest, then a further implication is that beneficial mutations should suffer a reduction in their probability of fixation, limiting the scope for sweeps to drive down diversity.
+Nonetheless, there is abundant evidence that “soft sweeps” are important determinants of patterns of nucleotide diversity \citep{messer2013population}, and it is possible that they have contributed to the extremely low levels of Y-linked variability that we observed.
 
 \section*{Conclusions}
-
-%Our results indicate that highly reduced levels of variability in regions of the Drosophila genome that lack crossing over can largely be explained by the effects of selection on nonsynonymous sites subject to purifying selection of the magnitude indicated by recent work.
-It seems likely that similar effects will occur in other genomes or genomic regions with low rates of genetic recombination, such as many bacterial genomes, mitochondrial genomes and true Y or W chromosomes that are not completely degenerate, as well as asexual or highly self-fertilizing species.
-
-The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8).
-The extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the ....
-a large genomic regions that lack recombination.
+The non-recombining region of a Y chromosome produces evolutionary dynamics that are similar to a haploid asexual population, with the lack of recombination resulting in build up of genetic associations among selected sites, a loss in the efficiency of selection, and an increase in population fitness variance \citep{fisher1930genetical, muller1964relation, hill1966HReffect, mcvean2000,KaiserCharlesworth,good2014genetic}.
+Our study of neutral diversity levels on the relatively young \textit{R.
+hastatulus} sex chromosomes provides clear evidence for an extensive loss of neutral diversity on the Y chromosome.
+Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss, forward population genetic simulations suggest that the extensive loss of diversity is most likely due to purifying selection.
+However, while standard background selection theory over-predicted the loss of diversity on the Y chromosome by $\approx$ 70\%, simulations of purifying selection acting over a large number of linked selected sites resulted in a level of Y-diversity that was consistent with what we observed.
+These results are consistent with theory on interference selection, and suggest that the effects of interference is likely an important force in the evolution of young Y chromosomes.
+Moreover, our results imply that when a large number of linked sites are subject to purifying selection, Y-chromosome degeneration in the early stages may largely be driven by the effects of interference rather primarily by widespread silencing of Y-linked genes and their subsequent degeneration through neutral drift.
 
 \section*{Acknowledgments}
+We thank Felix Baudry for providing genomic coverage data for quality filtering.
+This research was funded by Discovery grants to SCHB and SIW) from the Natural Sciences and Engineering Research Council of Canada and by an OGS scholarship to JH.
 
 \bibliography{bibliography}
+
+\include{}
+\includepdf[pages={1-}]{SI.pdf}
+
 \end{document}

### 2016-12-21 code/R/Figure1.R

@@ -1,97 +1,0 @@
-Nf=1000 #number of females
-Nm=1000 #number of males
-N=Nf+Nm #total number of individuals
-#x=Nf/(Nf+Nm) #proportion of males
-
-op <- par(mfrow = c(1, 3))
-m=0
-f=0
-
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m))))/N, from=0, to=1, lty=2,
-      xlab=expression(italic('N'[f]/'N'[m]+'N'[f])),
-      ylab=expression(italic('N'[e]/N)), 
-      main="Autosomes", ylim=c(0,1), xlim=c(0,1))
-
-mh=0.5
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+mh))))/N, from=0, to=1, add=T)
-
-m1=1
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m1))))/N, from=0, to=1, add=T)
-
-m2=2
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m2))))/N, from=0, to=1, add=T)
-
-m3=3
-curve(((4*N*(1-x)*x*N)/(((1-x)*N*(1+f))+(x*N*(1+m3))))/N, from=0, to=1, add=T)
-
-#for sex-linked genes
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m))))/N, from=0, to=1, lty=2,
-      xlab=expression(italic('N'[f]/'N'[f]+'N'[m])), 
-      ylab=expression(italic('N'[e]/N)), 
-      main="X chromosomes", ylim=c(0,1),xlim=c(0,1))
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+mh))))/N, from=0, to=1, add=T)
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m1))))/N, from=0, to=1, add=T)
-
-curve(((9*N*(1-x)*N*(x))/((4*N*(1-x)*(1+f))+(2*N*(x)*(1+m2))))/N, from=0, to=1, add=T)
-
[…]

### 2016-12-21 code/R/TOP.R

@@ -1,144 +1,0 @@
-vm <- c(0:100)
-for (val in vm) {
-  Nf=100000 #numner of females
-  Nm=100000 #number of males
-  N=Nf+Nf
-  vf=0 #variance in female rep success
-  kf=2 #mean number of offspring of females
-  km=2 #mean number of offspring of males
-  Nef=((Nf*(kf-1))/((kf-1)+(vf/kf))) #effect of female reproductive variance on female Ne
-  Nem=((Nm*(km-1))/((km-1)+(vm/km))) #effect of male reproductive variance on male Ne
-  Ne=Nem+Nef
-  Ne_06=4*Ne*(1-0.6)*0.6
-  Ne_07=4*Ne*(1-0.7)*0.7
-  Ne_08=4*Ne*(1-0.8)*0.8
-  Ne_09=4*Ne*(1-0.9)*0.9
-  Ne_ratio=Nef/Nem
-  A=(4*Nef*Nem)/(Nef+Nem) #autosome Ne 
-  X=(9*Nef*Nem)/(((2*Nef))+4*Nem) #X-chromosome Ne
-  Y=Nem/2
-  XA=X/A
-  YA=Y/A
-}
-d <- data.frame(vm,Nem,Nef,Ne_ratio,A,X,Y,XA,YA, Ne, Ne_06,Ne_07,Ne_08,Ne_09)
-
-plot(vm,(Nem/2)/N, type="l", xlim = c(0,100))
-plot(vm,(Ne)/N, type="l", xlim = c(0,20))
-plot(vm,YA,type="l", xlim = c(0,100))
-plot(vm,XA,type="l", xlim = c(0,100))
-
-
-require(ggplot2)
-ggplot(d, aes(vm)) +                   
-  geom_line(aes(y=(Ne/4)/N), colour="red") +
-  geom_line(aes(y=(Ne_06/4)/N), colour="blue") +
-  geom_line(aes(y=(Ne_07/4)/N), colour="green") +
-  geom_line(aes(y=(Ne_08/4)/N), colour="red") +
-  geom_line(aes(y=(Ne_09/4)/N), colour="brown")
-  
-Nf=100000 #numner of females
-Nm=100000 #number of males
-N=Nf+Nf
-ne_A_poisson=(4*N*(1-r)*r)/N
-ne_X_poisson=(9*N*r*(r-1)/(2*(r-2)))/N
-  
[…]

### 2016-12-21 code/R/bgs.R

@@ -1,61 +1,0 @@
-# Y/A
-s=0.02
-h=0.5
-hs=0.01
-mu=7*10^-9
-U=2*mu
-theta=4*Nem*mu
-
-
-curve(x*(0.5)*(exp(-(U/((2*s*h))))), from=0, to=1,
-      ylab = expression(paste("E[", pi,"]")), las=1,
-      #ylim=c(0,0.3),
-      #xlim=c(0,0.2),lwd=3)
-)
-#abline(h=0.2,lty=2)
-#abline(h=0.02,lty=1, col="red",lwd=3)
-#legend(x=0.12,y=0.3, cex=0.8, box.lty=0, c("Nf/(Nf+Nm) = 0.6","U = 0.02","h = 1","r = 0", expression(paste("", pi[0], paste(" = ",N[em]/2)))))
-
-text(0.03,0.029, cex=1, expression(paste(, theta,)))
-text(0.03,0.21,expression(paste(, pi[0],)))
-
-
-
-
-#print(20*math.exp(-0.02/(0.1+0.005)))
-
-curve(0.25*(exp(-(0.02/((2*x)+0)))), from=0, to=1,
-      main="Expected Y/A diversity under BGS",
-      xlab="selection coefficient (s)",
-      ylab = expression(paste("E[", pi,"]")), las=1,
-      ylim=c(0,0.3),
-      xlim=c(0,0.2),lwd=3)
-abline(h=0.2,lty=2)
-abline(h=0.02,lty=1, col="red",lwd=3)
-legend(x=0.12,y=0.3, cex=0.8, box.lty=0, c("Nf/(Nf+Nm) = 0.6","U = 0.02","h = 1","r = 0", expression(paste("", pi[0], paste(" = ",N[em]/2)))))
-
-text(0.03,0.029, cex=1, expression(paste(, theta,)))
-text(0.03,0.21,expression(paste(, pi[0],)))
-
-0.2083333*(exp(-0.02/((2*0.02)+0)))
-
-
-
-((9*N*0.6*(0.6-1)/(2*(0.6-2))))/(4*N*(1-0.6)*0.6)
-
-
-((N*(1-0.6)/2)/(4*N*(1-0.6)*0.6)
-)
-
-(0.2083333*(exp(-(0.02/((2*0.1)+0))))
-)
-
-
-theta
-Ne
-u
-h
-r
-U
-
-pi_not=theta

### 2016-12-21 code/R/crap.R

@@ -1,15 +1,0 @@
-kf=2 #mean number of offspring of females
-km=2 #mean number of offspring of males
-vm=2
-
-#effect of male reproductive variance on male Ne
-curve((Nm*(km-1))/((km-1)+(x/km))/N, from = 0, to = 1)
-
-#1:1
-mvalue=0.25
-mhvalue=0.125
-m1value=0.124875
-m2value=0.125
-m3value=0.125
-
-barplot(c(mvalue,mhvalue,0.14375,0.141525,0.14))

### 2016-12-21 code/R/nov29.R

@@ -1,34 +1,0 @@
-barplot(as.matrix(ChrRatios), beside=TRUE, col=c("darkgreen", "black", "gray22","gray73","gray87","gray93","darkgreen","black", "gray22","gray73","gray87", "gray93"),
-        ylab = expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-        ylim=c(0,1.2))
-legend("topright", c("observed", "Nf/(Nf+Nm) = 0.5", "Nf/(Nf+Nm) = 0.6", "Nf/(Nf+Nm) = 0.7", "Nf/(Nf+Nm) = 0.8", "Nf/(Nf+Nm) = 0.9"), fill=c("darkgreen", "black", "gray22","gray73","gray87", "gray93"), box.lty=0, cex=0.7)
-
-# empirical XA confidence intervals
-segments(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75)
-arrows(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-# simulated XA confidence intervals
-#segments(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75)
-#arrows(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-# empirical YA confidence intervals
-segments(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75)
-arrows(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-
-188266
-
-empirical_sfs=c(187850/188266, 222/188266, 73/188266, 32/188266, 20/188266, 45/188266, 24/188266, 0, 0, 0, 0, 0, 0)
-sfs=c(0.0756799, 0.0115273, 0.00639306, 0.00517519, 0.00541744)
-options(scipen=0)
-barplot(sfs, ylim=c(0,0.1), ylab="Relative SFS (Qn(i))", xlab="Derived allele frequency", main="coarse grained predictions")
-
-barplot(empirical_sfs, ylim=c(0,0.1))
-
-Original parameters: 127.67 1374 0
-Coarse-grained parameters: 171.123752462 1025.09778728 0
-Calculating time spectrum for parameters Ns=171.124, NU=1025.1, n=6
[…]

### 2016-12-21 code/R/nov29_2.R

@@ -1,40 +1,0 @@
-barplot(as.matrix(ChrRatios), beside=TRUE, col=c("darkgreen", "black", "gray22","gray73","gray87","gray93","darkgreen","black", "gray22","gray73","gray87", "gray93"),
-        ylab = expression(paste("Nucleotide Diversity ( ", theta, " ) ratio")),
-        ylim=c(0,1.2))
-legend("topright", c("observed", "Nf/(Nf+Nm) = 0.5", "Nf/(Nf+Nm) = 0.6", "Nf/(Nf+Nm) = 0.7", "Nf/(Nf+Nm) = 0.8", "Nf/(Nf+Nm) = 0.9"), fill=c("darkgreen", "black", "gray22","gray73","gray87", "gray93"), box.lty=0, cex=0.7)
-
-# empirical XA confidence intervals
-segments(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75)
-arrows(x0=1.5,y0=0.7,x1=1.5,y1=0.9,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-# simulated XA confidence intervals
-#segments(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75)
-#arrows(x0=2.5,y0=0.7,x1=2.5,y1=0.8,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-# empirical YA confidence intervals
-segments(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75)
-arrows(x0=8.5,y0=0.01,x1=8.5,y1=0.04,lwd=0.75,angle = 90, code = 3, length = 0.04)
-
-
-188266
-
-empirical_sfs=c(187850/188266, 222/188266, 73/188266, 32/188266, 20/188266, 45/188266, 24/188266, 0, 0, 0, 0, 0, 0)
-sfs=c(0.0756799, 0.0115273, 0.00639306, 0.00517519, 0.00541744)
-options(scipen=0)
-barplot(sfs, ylim=c(0,0.1), ylab="Relative SFS (Qn(i))", xlab="Derived allele frequency", main="coarse grained predictions")
-
-barplot(empirical_sfs, ylim=c(0,0.1))
-
-Original parameters: 127.67 1374 0
-Coarse-grained parameters: 171.123752462 1025.09778728 0
-Calculating time spectrum for parameters Ns=171.124, NU=1025.1, n=6
[…]

### 2016-12-21 code/SFS_code/convert_SFS_CODE.sh

@@ -1,20 +1,0 @@
-#!/bin/bash -l
-#SBATCH -D /home/jhough/projects/rumex/projects/Ydiversity
-#SBATCH -o /home/jhough/projects/rumex/projects/Ydiversity/logs/convert-stdout_%j.txt
-#SBATCH -e /home/jhough/projects/rumex/projects/Ydiversity/logs/convert_stderr_%j.txt
-#SBATCH -J convert_SFS
-#SBATCH --cpus-per-task=5
-
-set -e
-set -u
-set -o pipefail
-
-#see https://github.com/molpopgen/msstats
-#command structure: convertSFS CODE <input file> <option [args]> | msstats/src/msstats >out
-# used: - for ms style output; T 0 says to output only sysnonymous mutations
-
-convertSFS_CODE SFS.out -m T 0 | ./msstats/src/msstats > stats.txt
-
-# Then, using R,
-# x=read.table("stats",header=T)
-# mean(x$hsm01[x$pop==0])

### 2016-12-21 code/SFS_code/makeTask.pl

@@ -1,41 +1,0 @@
-#!/usr/bin/perl -w
-use strict;
-
-my $NSIMS = 1000;   #TOTAL NUMBER OF SIMULATIONS YOU WANT TO RUN
-my $NJOBS = 10;     #TOTAL NUMBER OF JOBS (PROCESSORS) YOU WANT TO USE
-my $FOLD = 0;       #START IN $FOLD+1
-my $THETA = 0.0015; #SET NON-VARYING PARAMETERS
-
-# SET PARAMETERS YOU WOULD LIKE TO VARY IN AN ARRAY AS FOLLOWS:
-my @RHO = (0, 0.001);
-
-my %SEEDS = ();
-#NOW LOOP OVER VALUES OF ALL PARAMETERS
-foreach my $r (@RHO){
-  #YOU CAN INCLUDE MULTIPLE foreach LOOPS IF VARYING MORE THAN 1 PARAMETER
-  $FOLD++;
-  unless(chdir $FOLD){
-    mkdir $FOLD;
-    chdir $FOLD;
-  }
-  print "$FOLD\n";
-
-  #DEFAULT OUTPUT FILE IS "tasklist"
-  open(OUT,">tasklist") or die "cannot write to tasklist\n";
-
-  # UPDATE DIRECTORY TO INDICATE THE LOCATION OF SFS_CODE AND UPDATE COMMAND LINE AS NECESSARY FOR YOUR SET OF SIMULATIONS
-  my $cmd = "time /DIRECTORY/sfs_code 1 ";
-  $cmd .= int($NSIMS/$NJOBS+0.5);
-  $cmd .= " -t $THETA -r $r ";
-  
-  for(my $i=0; $i<$NJOBS; $i++){
-    my $seed = int(rand(40000000));
-    if(exists($SEEDS{$seed})){ #ENSURE UNIQUE SEEDS FOR EACH TASK
-      redo;
-    }
-    print OUT "$cmd -s $seed\n";
-  }
-  
-  close(OUT);
-  chdir "..";
-}

### 2016-12-21 code/SFS_code/runTask.sh

@@ -1,10 +1,0 @@
-#!sh
-#$ -S /bin/bash
-#$ -t 1-10
-#$ -cwd
-TASKFILE=tasklist
-TASK=$(cat $TASKFILE | head -n $SGE_TASK_ID | tail -n 1)
-echo $TASK
-$TASK -o out_$SGE_TASK_ID.txt
-
-exit 0

### 2016-12-21 code/course_grained.py

@@ -1,23 +1,0 @@
-import coarse_grained_theory
-import recombination_theory
-
-# Asexual population
-
-Ns = 127.67
-NU = 1374
-NR = 0
-n = 6
-
-# for rumex: Ns=127.67, NU=1374 s=6 nruns=10000
-
-# Calculate coarse-grained parameters
-Nseff,NUeff,NUneff = coarse_grained_theory.calculate_effective_params(Ns,NU)
-
-print "Original parameters:", Ns,NU,NR
-print "Coarse-grained parameters:", Nseff,NUeff,0
-
-
-# Calculate the shape of the SFS directly
-Q = coarse_grained_theory.calculate_Q(Ns,NU,n)
-
-print "Qn(i):", Q

### 2017-01-04 manuscript/manuscript.tex
updates to manuscript and SI

@@
-\documentclass[9pt,twocolumn,twoside]{gsajnl}
+\documentclass[9pt,twocolumn,twoside,lineno]{gsajnl}
 % Use the documentclass option 'lineno' to view line numbers
@@
 
-\title{Hill-Robertson Interference Causes Reduced Genetic Diversity on a Young Plant Y-chromosome}
+\title{Hill-Robertson Interference Reduced Genetic Diversity on a Young Plant Y-chromosome}
 
@@
 
-\keywords{Sex Chromosome Evolution; Nucleotide Diversity; Recombination; Deleterious Mutations}
+\keywords{Nucleotide Diversity; Suppressed Recombination; Deleterious Mutations; Interference Selection}
 
-\runningtitle{Plant sex chromosome evolution} % For use in the footer
+\runningtitle{Sex chromosome evolution} % For use in the footer
 
@@
 On Y chromosomes with suppressed recombination, natural selection is expected to eliminate linked neutral variation and lower the $N_{e}$ of Y compared to X chromosomes or autosomes.
-However, non-selective factors including female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity in natural populations.
-Here, we investigate the processes affecting levels of polymorphism during sex chromosome evolution in the dioecious plant \textit{Rumex hastatulus} (Polygonaceae).
-Strikingly, we find that neutral diversity for genes on the Y is on average 2.1\% of the value for their X-linked homologues, corresponding to a chromosome-wide reduction of 93\% compared to the standard neutral expectation.
+However, female-biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y-diversity.
+Here, we investigate the factors affecting levels of polymorphism during sex chromosome evolution in the dioecious plant \textit{Rumex hastatulus} (Polygonaceae).
+Strikingly, we find that neutral diversity for genes on the Y chromosome is on average 2.1\% of the value for their X-linked homologues, corresponding to a chromosome-wide reduction of 93\% compared to the standard neutral expectation.
 We demonstrate that the magnitude of this diversity loss is inconsistent with reduced male $N_{e}$ caused by neutral processes.
 Instead, using forward simulations and estimates of the fitness effects of deleterious mutations, we show that Y chromosome diversity loss can be explained by purifying selection on a large number ($\geq$ 800 kb) of genetically-linked sites.
-Our results are in agreement with theory on "interference selection", and provide evidence that this can considerably reduce levels of polymorphism on a young plant Y chromosome.
+Our results are in agreement with theory on interference selection, and provide evidence that this can considerably reduce levels of polymorphism on a young plant Y chromosome.
 Given the relatively recent origin of \textit{R.
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{bachtrog2013NRG,hough2014,bergero2015}, and in some species, such genetic degeneration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and n some species, such genetic deterioration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
 The independent evolution of these phenomena in such taxonomically distant groups as mammals and flowering plants suggests that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoretical and experimental biologists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
@@
 Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, and 1/3 that of X-linked genes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that the level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+The lowered $N_{e}$ of the Y chromosome implies that level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
 In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citealt{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citealt{smith1974hitch}).
@@
 Moreover, theory suggests that even weak purifying selection, when aggregated over a large number of genetically-linked sites, can generate strong deviations from neutrality, whereas classic background selection theory breaks down in such cases \citep{mcvean2000, comeron2002population, KaiserCharlesworth, good2014genetic}.
-Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citep{good2014genetic} is \textit{a priori} likely to be a strong force affecting the evolution of young plant sex chromosomes.
-However, if there has been widespread gene silencing during the early stages of Y-chromosome evolution, as in \textit{Drosophila miranda}, for example \citep{zhou2012chromosome}, then diversity loss might be expected to be lower on younger Y chromosomes.
+Given that a large number of selected sites are likely to be in linkage disequilibrium on a recently evolved Y chromosome, such "interference selection" \textit{sensu} \citealt{good2014genetic} is \textit{a priori} likely to be a strong force affecting the evolution of young plant sex chromosomes.
+However, if there has been widespread gene silencing during the early stages of Y-chromosome evolution, as in \textit{Drosophila albomicans}, for example \citep{zhou2012chromosome}, then diversity loss might be expected to be lower on younger Y chromosomes.
 That is, if Y-linked gene silencing occurs early during sex chromosome evolution, then few sites may be under selection, and Y chromosome degeneration may be driven primarily by neutral genetic drift rather than inefficient selection.
@@
 This species is a dioecious annual with heteromorphic X and Y chromosomes that originated \textasciitilde 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making Y chromosomes in this species over 100 million years younger than the highly degenerated Y chromosomes in mammals \citep{lahn1999,ross2005dna}.
-\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphism in sex chromosome system, in which both XY and $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ male genotypes occur in geographically distinct populations (so-called "chromosomal races"; \citealt{smith1963mechanism}.
+\textit{Rumex hastatulus} has also received particular attention because of the occurrence of an interesting polymorphism in sex chromosome system, in which both XY and $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ male genotypes occur in geographically distinct populations (so-called "chromosomal races"; \citealt{smith1963mechanism}).
 The derived $\textnormal{XY}_\textnormal{1}\textnormal{Y}_\textnormal{2}$ sex chromosome system in this species is thought to have originated through an X-autosome fusion, with the current XY system maintaining the ancestral chromosome complement \citep{smith1964evolving}.
@@
 We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 12 population samples (1 male and 1 female from each of 6 populations).
-Samples were collected in 2010 from throughout the native range of \textit{R.
+Samples were collected in 2010 from throughout the native range of the Texas chromosome race \textit{R.
 hastatulus} (locations in Table S1), and plants were grown in the glasshouse at the University of Toronto from seeds collected from open-pollinated females.
@@
 SRP041588.
-Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$ chromosome).
+Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$chromosome).
 
@@
 To do this, we normalized our sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below).
-Because the criteria for ascertaining autosomal loci in Hough et al.
-(2014) were based on identifying four segregating SNPs per locus, and since this set of genes is likely to be higher in diversity than the average autosomal gene, here we instead used the larger set of all non-sex linked (putatively autosomal) genes as our autosomal reference.
+Because the criteria for ascertaining autosomal loci in \citet{hough2014} were based on identifying four segregating SNPs per locus, and since this set of genes is likely to be higher in diversity than the average autosomal gene, here we instead used the larger set of all non-sex linked (putatively autosomal) genes as our autosomal reference.
 We filtered genes in this set to remove any genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria.
@@
 For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the diploid effective population size, and $\mu$ is the per-base mutation rate per-generation, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}.
-To compare sex-linked and autosomal loci, we calculated the average value of $\theta$ for each chromosome type, weighted by the number of synonymous sites in each gene (\X Figure 2).
+To compare sex-linked and autosomal loci, we calculated the average value of $\theta$ for each chromosome type, weighted by the number of synonymous sites in each gene (Figure 2).
 We obtained 95\% confidence intervals for X/A and Y/A ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A on each iteration for 20000 replicates each.
@@
 Note that the lack of recombination on the Y chromosome implies that statistical assumptions about independence across loci are violated, suggesting that the true uncertainty in the Y/A estimate may be wider than implied by bootstrapping.
-To address this, we also used a maximum likelihood approach to independently estimate a credibility interval for the Y/A ratio (Figure S1).
+To address this, we also used a maximum likelihood approach \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio (Figure S1).
 Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate was not feasible, as this would require maximizing the likelihood estimate of the mutation rate for each locus independently.
@@
 Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci.
-Our model thus had two parameters: $\theta_{a}$, our estimate of Watterson's theta for autosomal genes, and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes.
-We varied both parameters and evaluated the likelihood of \textit{f} from 0.001 to 1, and $\theta_{a}$ per base pair from 0.001 to 0.01.
-In particular, we evaluated the likelihood of $\theta_{a}$ and \textit{f}, given (1) the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), (2) the number of segregating sites on the Y chromosome ($S_{y}$), and (3) the total number of sites per locus ($L_{i}$ and $L_{y}$).
-The likelihood of ($\theta_{a}$, \textit{f}) is given by:
+Our model thus had two parameters: $\theta_{aut}$, our estimate of Watterson's theta for autosomal genes, and \textit{f}, the ratio of effective population size of the Y chromosome to autosomes.
+We varied both parameters and evaluated the likelihood for \textit{f} from 0.001 to 1, and $\theta_{aut}$ per base pair from 0.001 to 0.01.
+In particular, we evaluated the likelihood of $\theta_{aut}$ and \textit{f}, given (1) the observed number of synonymous segregating sites for each \textit{i} of \textit{n} autosomal loci ($S_{i}$), (2) the number of segregating sites on the Y chromosome ($S_{y}$), and (3) the total number of sites per locus ($L_{i}$ and $L_{y}$).
+The likelihood of ($\theta_{aut}$, \textit{f}) is given by:
 
 \begin{equation}
-L \propto  P(S_{Y}, L_{Y} | f, \theta_{a})\prod_{i=1}^{n}P(S_{i}, L_{i} | \theta_{a})  \label{eq:L}
+  L \propto  P(S_{Y}, L_{Y} | f, \theta_{aut})\prod_{i=1}^{n}P(S_{i}, L_{i} | \theta_{aut}) \label{eq:L}
 \end{equation}
@@
 \subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
-To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to the corresponding neutral predictions for a given effective population size "sex ratio", $r=N_{e}_{f}/(N_{e}_{f} + N_{e}_{m})$.
+To test whether our estimated levels of diversity on X, Y and autosomal chromosomes could be explained by neutral processes, including female-biased population sex ratios or high variance in male reproductive success, we compared our normalized X/A and Y/A diversity estimates to the corresponding neutral predictions for a given effective population size "sex ratio", $r=N_{e_{f}}/(N_{e_{f}} + N_{e_{m}})$.
 Because we were primarily interested in determining the parameter space within which a reduced male $N_{e}$ could result in Y/A and X/A diversity ratios consistent with our data, we tested the fit of our estimates to predictions across the full range of $r$ (0.1 to 0.9; we excluded all-female or all-male cases).
@@
 \begin{equation}
-N_{e_{A}} = \frac{4N_{e}_{m}N_{e}_{f}}{N_{e}_{m}+N_{e}_{f}} \label{eq:NeA}
+N_{e_{A}} = \frac{4N_{e_{m}}N_{e_{f}}}{N_{e_{m}}+N_{e_{f}}} \label{eq:NeA}
 \end{equation}
@@
 \begin{equation}
-N_{e_{X}} = \frac{9N_{e}_{m}N_{e}_{f}}{4N_{e}_{m}+2N_{e}_{f}} \label{eq:NeX}
+N_{e_{X}} = \frac{9N_{e_{m}}N_{e_{f}}}{4N_{e_{m}}+2N_{e_{f}}} \label{eq:NeX}
 \end{equation}
@@
 \begin{equation}
-N_{e_{Y}} = \frac{N_{e}_{m}}{2} \label{eq:NeY}
+N_{e_{Y}} = \frac{N_{e_{m}}}{2} \label{eq:NeY}
 \end{equation}
@@
 respectively.
-Note that when $N_{e}_{f} = N_{e}_{m}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}.
+Note that when $N_{e_{f}} = N_{e_{m}}$ ($r=0.5$) then, $N_{e_{X}}/N_{e_{A}} = 0.75$, and $N_{e_{Y}}/N_{e_{A}} = 0.25$ as in the standard neutral model \citep{wright1931evolution}.
 With a female-biased ratio, however, the expected X/A ratio can become greater than 1, approaching 1.125 in the limit as $r\rightarrow 1$ \citep{caballero1995}.
@@
 To study the effects of purifying selection on expected levels of Y-chromosome diversity, we conducted forward-time simulations of haploid Y chromosomes using the software SFSCODE \citep{hernandez2008flexible}.
-We first estimated the distribution of fitness effects of deleterious amino acid mutations from our polymorphism data for X-linked genes using the method of \citet*{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphisms.
-We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta_{a}$, adjusted to reflect the neutrally expected $N_{e_{Y}}$ for a sex ratio of $r=0.6$.
+We first estimated the distribution of fitness effects of deleterious amino acid mutations from our polymorphism data for X-linked genes using the method of \citet*{keightley2007joint}, which fits a gamma distribution of negative selection coefficients to the observed frequency distribution of non-synonymous and synonymous polymorphisms.
+We then used this estimated gamma distribution to parameterize the simulations, initializing them with our estimated $\theta_{aut}$, adjusted to reflect the neutrally expected $N_{e_{Y}}$ for a sex ratio of $r = 0.6$.
 To match our sample size and the number of synonymous sites sampled from our data (see Supporting Information), the simulations sampled 6 haploid chromosomes, and the genome sequence contained 45,331 bp of linked neutral sequence from which we calculated silent site diversity, $\pi_{s}$.
 To examine the expected reduction in diversity as a function of the number of selected sites ($L$), we ran simulations over a range of values of $L$, up to a maximum of 5x$10^{6}$ (Figure 3).
-To obtain an estimate of $L$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}$, was less than 0.0001 from our empirical estimate (Figure 3).
+To obtain an estimate of $L$, we calculated the approximate likelihood of our observed data based on the proportion of simulations in which synonymous diversity, $\pi_{s}$, was less than 0.0001 from our empirical estimate (Figure 3B).
 
@@
 Note that by normalizing X and Y diversity estimates by autosomal diversity levels, our results indicate that the difference between X and Y homologues is not due to an elevation of diversity on the X chromosome, but a Y-specific reduction.
-Interestingly, our results also suggest an elevation of X-linked diversity ($\textnormal{X/A}=0.85$) compared to the neutral prediction ($\textnormal{X/A}=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
+Interestingly, our results also suggest an elevation of X-linked diversity ($X/A=0.85$) compared to the neutral prediction ($X/A=0.75$), though the confidence intervals on this estimate are wide, ranging from 0.74 to 0.95 (Figure 2).
 Given the empirically-estimated sex ratio of $r=0.6$ in \textit{R.
@@
 hastatulus} and the non-dioecious outgroup \textit{R.
-bucephalophorus} is not significantly different for sex-linked (0.2016) and autosomal genes (0.219), and we found no evidence for significant differences in substitution rate between Y and X chromosomes \citep{hough2014}.
+bucephalophorus} is comparable for sex-linked (0.2016) and autosomal genes (0.219), and we found no evidence for significant differences in substitution rate between Y and X chromosomes \citep{hough2014}.
 It is therefore unlikely that our results are caused by mutation rate differences between sex-linked and autosomal genes.
@@
 
-chr & $\theta$ & $/\theta_{A}$ & $\theta$ & $/\theta_{A}$  \\
+chr & $\theta$ & $/\theta_{aut}$ & $\theta$ & $/\theta_{aut}$  \\
 \midrule
@@
 hastatulus} is expected to lower Y diversity through a reduction in male $N_{e}$ and thus a reduction in the $N_{e}$ of the Y chromosome.
-Male $N_{e}$ could be further reduced by high variance in male reproductive success, which is not unusual in annual wind-pollinated plants such as \textit{R.
+Male $N_{e}$ could be further reduced by high variance in male reproductive success, which is expected in annual wind-pollinated plants such as \textit{R.
 hastatulus} that commonly exhibit extensive phenotypic plasticity in plant size and flower production.
 Given that male plants in this species produce large amounts of pollen, and female flowers are uniovulate, there may indeed be strong competition among males to fertilize ovules.
-Such competition should cause a proportional increase in levels of X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity.
+Such competition should cause a proportional increase in X-linked diversity to a level that is close to (or even higher than) levels of autosomal diversity \citep{caballero1995}, while simultaneously reducing Y-linked diversity.
 In common with most flowering plants, we do not have marker-based estimates of variance in male reproductive success in \textit{R.
 hastatulus}.
-We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values the $N_{e}$ sex ratio, $r$.
+We therefore tested whether an overall reduction in male $N_{e}$, arising either from high variance in male reproductive offspring number and/or a female biased population sex ratio, could explain our observed Y/A and X/A ratios, by comparing our data to neutral predictions across the full range of values for the $N_{e}$ sex ratio, $r$.
 
 As shown in Figure 2, the expected reduction in Y/A as a function of $r$ is substantially higher than our observed Y/A diversity ratio.
-Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $r=0.9$, the expectation for Y/A is $\approx 0.14$, which is substantially higher than our observed estimate ($\textnormal{Y/A}=0.002$) (Table 1; Figure 2).
-Moreover, such large reductions in $N_{e}_{m}$ would also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
+Indeed, the lower limit for the Y/A ratio in a neutral model is 1/8, such that even in the extreme case where $r=0.9$, the expectation for Y/A is $\approx 0.14$, which is substantially higher than our observed estimate ($Y/A= 0.018$) (Table 1; Figure 2).
+Moreover, such large reductions in $N_{e_{m}}$ would also predict levels of X/A diversity that are significantly larger than what we observed (Figure 2).
 Our results therefore indicate that, although reduced male $N_{e}$ arising from sex-biased demography is expected to contribute to reduced Y chromosome polymorphism, it is insufficient to explain the Y/A reduction we observed.
@@
 \caption{Observed and predicted X/A and Y/A ratios of neutral diversity.
-Predictions are shown for increasing values of the effective popuatlion size sex ratio, $N_{e}_{f}$/(N_{e}_{m} + $N_{e}_{f})$.
+Predictions are shown for increasing values of the effective population size sex ratio, $N_{e_{f}}/(N_{e_{m}} + N_{e_{f}})$.
 Observed estimates of X/A and Y/A were calculated as the average $\theta$ across genes, weighted by the number of synonymous sites per gene.
@@
 \subsection*{Purifying selection}
-Selection against strongly deleterious mutations is expected be an important factor reducing genetic variability in regions with low recombination \citep{charlesworth1993effect, charlesworth1996background}.
+Selection against strongly deleterious mutations is expected to be an important factor reducing genetic variability in regions with low recombination \citep{charlesworth1993effect, charlesworth1996background}.
 The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and \textit{Drosophila} \citep{mcallister1999, charlesworth1996CB}.
-Under background selection theory, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi$ $\approx 4f_{0}N_{e}\mu$, where $f_{0}$=$e^{-\frac{U}{2sh+r}}$, $U$ is the diploid mutation rate to strongly deleterious variants, $s$ and $h$ are selection and dominance coefficients, and $r$ is the recombination rate \citep{hudson1995deleterious}.
-This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{2sh+r}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS $\pi_{0}=\theta=4N_{e}\mu$.
-Because only reproducing males bearing Y chromosomes are relevant for determining the level of Y-linked diversity (i.e., $N_{e}_{Y} = N_{e}_{m}/2$), the expected reduction in diversity is obtained in this model by replacing $N_{e}$ with number of breeding males, $N_{e}_{m}$.
-Here, assuming that $N_{e}_{m}$ is 0.4 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
-hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 3x$10^{-5}$ for a haploid Y chromosome mutation rate of 7x$10^{-9}$ per nucleotide per generation.
+Under background selection theory for a Y chromosome, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi \approx \pi_{0}=(\frac{Y}{A})4N_{e}f_{0}\mu$, where $f_{0}=e^{-\frac{U}{sh}}$, $U$ is the mutation rate to strongly deleterious variants on the Y, and $s$ and $h$ are selection and dominance coefficients \citep{hudson1995deleterious}.
+This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{sh}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS, $\pi_{0}=(\frac{Y}{A})4N_{e}\mu$, where Y/A is the expected ratio of the effective population size of the Y relative to autosomes.
+Here, assuming that the proportion of males is 0.4, and therefore that Y/A is 0.21 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
+hastatulus} \citep{pickup2013influence}, and using $s$=0.003 as estimated from the distribution of fitness effects of deleterious mutations, the corresponding expected diversity on the Y chromosome under BGS is $E[\pi]$= 1.5x$10^{-5}$ for a haploid Y chromosome mutation rate of 7x$10^{-9}$ per nucleotide per generation and 1 MB of selected sites.
 Although there is considerable uncertainty in these parameters, the calculation yields an expected value of diversity that is 70\% lower that what we observed (Table 1; Figure 2; Figure 4).
@@
 Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time.
-This is because high linkage disequilibrium causes the behavior of selected alleles to interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
+This isbecause high linkage disequilibrium causes the behavior of selected alleles interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
 As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding such selective interference in realistic situations.
 We therefore conducted forward simulations of purifying selection using our estimated parameters for the distribution of fitness effects of deleterious mutations, and tested whether interference selection could result in a level of Y diversity at neutral sites similar to the level we observed.
-We found that models of purifying selection acting over a large number of linked selected sites resulted in an expected loss of diversity at neutral sites that was consistent with our data (Figure 3; Figure 4).
-In addition, and consistent with previous studies, our simulations also suggest a saturating loss of diversity with an increasing number of selected sites \citep{KaiserCharlesworth}.
+Our maximum likelihood estimates of the number of selected sites, $L$, from our simulations, suggest that the number of sites subject to selection required to explain the diversity loss from the \textit{R.
+hastatulus} Y chromosome is ($\approx 800 kb$) and may be as large as 2.5 MB (Figure 3).
+Thus, these results are consistent with the hypothesis that the early stages of Y degeneration are characterized by the persistence of a large number of sites subject to purifying selection.
 
-Although we initialized our simulations with selection parameters estimated from the distribution of fitness effects of deleterious mutations, recent work has shown that, unlike with BGS, evolutionary dynamics in the “interference selection regime” are insensitive to differences in strength of selection, with expected diversity loss instead determined primarily by the population fitness variance over the non-recombining region, which is in turn expected to increase with the number of linked loci under selection \citep{good2014genetic}.
-Estimating the number of selected loci is therefore key to determining whether diversity on the Y chromosome may be affected by interference selection.
-Our maximum likelihood estimates of the number of selected sites, $L$, from our simulations, suggest that the number of sites subject to selection required to explain the diversity loss from the \textit{R.
-hastatulus} Y chromosome is $\approx 800$ kb (Figure 3).
+While clearly approximate, we can ask how these estimates compare to how many selected sites we might expect to be present on the Y chromosome.
+Our previous estimates based on cytological measurements suggested that there may be roughly 5,600 genes on the X chromosome of the XY race \citep{hough2014}.
+We also estimated that approximately 28 percent of genes had degenerated, leaving us with an estimated 1568 genes remaining on the Y chromosome.
+Using the average number of codons in genes of \textit{Arabidopsis thaliana} (426) \citep{wortman2003annotation}, and assuming 2/3 of codons are under selection, this would imply 1.3 MB of selected sites, which is well within the range of our uncertainty (Figure 3).
+Thus, although our estimates are broad, these results are consistent with the retention of a large proportion of sites subject to purifying selection on the Y chromosome, as expected in the early stages of degeneration.
 
@@
 \caption{Estimated levels of Y-linked nucleotide diversity compared to estimates obtained from simulations of purifying selection among linked selected sites (HRI simulations) and analytical predictions from background selection theory.
-Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0} =4N_{e}_{Y}\mu$, and $N_{e}_{Y}$ is the expected value of $N_{e}_{Y}$ for a sex ratio of $r=0.6$, as estimated in \textit{R.
+Silent site diversity is reported relative to the neutral expectation ($\pi_{0}$), where $\pi_{0}=(\frac{Y}{A})4N_{e}\mu$, and Y/A is the expected ratio of the effective population size of the Y relative to autosomes for a sex ratio of $r=0.6$, as estimated in \textit{R.
 hastatulus} populations.
-The credibility interval for our observed Y diversity was obtained by maximum likelihood (see methods), and error bars for HRI simulations correspond to 1.5x the interquartile range for $\pi_{s}$ estimates, approximately equal to the 95\% confidence interval for the median.
-\citep{chambers1983graphical}.
+The credibility interval for our observed Y diversity was obtained by maximum likelihood (see methods), and error bars for HRI simulations correspond to 1.5x the interquartile range for $\pi_{s}$ estimates, approximately equal to the 95\% confidence interval for the median \citep{chambers1983graphical}.
 }
@@
 Whereas neutral models of sex-biased demography were unable to explain the magnitude of diversity loss, forward population genetic simulations suggest that the extensive loss of diversity is most likely due to purifying selection.
-However, while standard background selection theory over-predicted the loss of diversity on the Y chromosome by $\approx$ 70\%, simulations of purifying selection acting over a large number of linked selected sites resulted in a level of Y-diversity that was consistent with what we observed.
+However, while standard background selection theory over-predicted the loss of diversity on the Y chromosome by $\approx$ 70\%, simulations of purifying selection acting over a large number of linked selected sites resulted in a level of Y-diversity that was similar to what we observed.
 These results are consistent with theory on interference selection, and suggest that the effects of interference is likely an important force in the evolution of young Y chromosomes.
@@
 We thank Felix Baudry for providing genomic coverage data for quality filtering.
-This research was funded by Discovery grants to SCHB and SIW) from the Natural Sciences and Engineering Research Council of Canada and by an OGS scholarship to JH.
+This research was funded by Discovery grants to SCHB and SIW from the Natural Sciences and Engineering Research Council of Canada, and by an OGS scholarship to JH.
 
 \bibliography{bibliography}
-
-\include{}
-\includepdf[pages={1-}]{SI.pdf}
-
 \end{document}

### 2017-01-05 manuscript/manuscript.tex
final edits

@@
 Given the relatively recent origin of \textit{R.
-hastatulus} sex chromosomes, our results imply that Y-chromosome degeneration in the early stages may be largely driven by such interference effects rather than by positive selection for gene silencing followed by neutral genetic drift.
+hastatulus} sex chromosomes, our results imply that Y-chromosome degeneration in the early stages may be largely driven by such interference effects rather than by widespread gene silencing followed by neutral genetic drift.
 \end{abstract}
@@
 Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes.
-In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and n some species, such genetic deterioration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
+In both groups, the loss of recombination between X and Y chromosomes is associated with an accumulation of deleterious mutations and a gradual loss of functional genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, such genetic deterioration has also led to the evolution of dosage compensation on the X chromosome \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}.
 The independent evolution of these phenomena in such taxonomically distant groups as mammals and flowering plants suggests that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoretical and experimental biologists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
@@
 Therefore, genes on the Y chromosome are expected to experience an effective population size ($N_{e}$) that is 1/4 that of autosomal genes, and 1/3 that of X-linked genes (assuming an equal number of reproducing females and males).
-The lowered $N_{e}$ of the Y chromosome implies that level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
+The lowered $N_{e}$ of the Y chromosome implies that the level of neutral polymorphism maintained at equilibrium, proportional to the product of $N_{e}$ and the neutral mutation rate, $\mu$, should be lower for Y-linked genes than for their X-linked counterparts.
 In the absence of recombination, genes on the Y chromosome are also expected to be in strong linkage disequilibrium, making them vulnerable to diversity loss due to selection against strongly deleterious mutations (background selection; \citealt{charlesworth1993effect}) and selective sweeps of strongly beneficial mutations (genetic hitchhiking; \citealt{smith1974hitch}).
@@
 Because variance in male reproductive success differentially affects the $N_{e}$ of X, Y, and autosomal chromosomes \citep{kimura1964number,nomura2002effective}, evidence for this can be obtained by comparing levels of silent site variability on X and Y chromosome, relative to values on autosomes.
-For example, variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio relative to the neutral expectation (1/4), but also causes an increases the X/A ratio.
+For example, variance in male reproductive success reduces male $N_{e}$, which reduces the Y/A diversity ratio relative to the neutral expectation (1/4), but also causes an increase in the X/A ratio.
 Based on such comparisons, studies in humans, for example, have suggested that the observation of an inflated X/A ratio can be explained by a historical excess of breeding females compared to males (\citealt{hammer2008sex}; but see \citealt{bustamante2009,hammer2010,cotter2016genetic}).
 
-Despite widespread interest in determining the evolutionary factors that affect neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we know very little about how sex ratio variation or selective processes have affected levels of diversity on more recently evolved sex chromosomes.
+Despite widespread interest in determining the evolutionary factors that affect neutral diversity on sex chromosomes \citep{ellegren2011,bachtrog2013NRG}, we still know very little about how sex ratio variation or selective processes have affected levels of diversity on more recently evolved sex chromosomes.
 The time scales over which these processes are likely to be important are therefore not well understood.
@@
 SRP041588.
-Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$chromosome).
+Here, to focus on sex-linked genes likely to be relatively older and closer to equilibrium, we focused on the previously described set of 460 X/Y genes for which a Y homolog was found in both races (i.e., X/Y genes where the Y copy was inferred to be on the $\textnormal{Y}_\textnormal{1}$ chromosome).
 
@@
 \caption{\textbf{A}.
-Equilibrium effective population sizes, relative to $N$, for genes on autosomes, X chromosomes, and Y chromosomes as a function of the sex ratio.
+Normalized effective population sizes at equilibrium for genes on autosomes, X chromosomes, and Y chromosomes as a function of the sex ratio.
 \textbf{B}.
@@
 \subsection*{Extensive loss of Y-chromosome diversity}
-Our analysis of polymorphism levels across the genome in the dioecious plant \textit{R.
-hastatulus} reveals a widespread loss of neutral diversity from the Y chromosome.
-In particular, neutral diversity for Y-linked genes is on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
+Our analysis of polymorphism levels across the genome of the dioecious plant \textit{R.
+hastatulus} revealed a widespread loss of neutral diversity from the Y chromosome.
+In particular, neutral diversity for Y-linked genes was on average \textasciitilde 2.1\% of the value for their homologues on the X chromosome.
 Taking 1/4 of the mean autosomal diversity as the equilibrium expectation for $\theta_{Y}$ under neutrality, this corresponds to a chromosome-wide reduction of 93\% relative to the standard neutral model \citep{wright1931evolution}.
@@
 \midrule
-A & 0.0055 & 1 & N/A & 1 \\
+A & 0.0055 & 1 & 0.0055 & 1 \\
 X & 0.0047 & 0.85 & 0.0041 & 0.75 \\
@@
 The effects of such “background selection” (BGS) have been well studied theoretically \citep{charlesworth1997effects, nordborg1996effect, kim2000joint} and the theory has been used to explain patterns of genetic diversity across genomes \citep{comeron2014background} and plays a central role in explanations for diversity loss on Y chromosomes in mammals \citep{Wilsonsayres2014} and \textit{Drosophila} \citep{mcallister1999, charlesworth1996CB}.
-Under background selection theory for a Y chromosome, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi \approx \pi_{0}=(\frac{Y}{A})4N_{e}f_{0}\mu$, where $f_{0}=e^{-\frac{U}{sh}}$, $U$ is the mutation rate to strongly deleterious variants on the Y, and $s$ and $h$ are selection and dominance coefficients \citep{hudson1995deleterious}.
-This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{sh}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS, $\pi_{0}=(\frac{Y}{A})4N_{e}\mu$, where Y/A is the expected ratio of the effective population size of the Y relative to autosomes.
+Under background selection theory for a Y chromosome, the predicted loss of diversity is modeled as a reduction in the equilibrium $N_{e}$ such that $\pi \approx (\frac{Y}{A})4N_{e}f_{0}\mu$, where $f_{0}=e^{-\frac{U}{sh}}$, $U$ is the mutation rate to strongly deleterious variants on the Y, and $s$ and $h$ are selection and dominance coefficients \citep{hudson1995deleterious}.
+This yields the expected value of $\pi$ under the model: $E[\pi]=\pi_{0}e^{\frac{-U}{sh}}$, where $\pi_{0}$ is the expected diversity in the absence of BGS, $\pi_{0}=(\frac{Y}{A})4N_{e}\mu$, and Y/A is the expected ratio of the effective population size of the Y relative to autosomes.
 Here, assuming that the proportion of males is 0.4, and therefore that Y/A is 0.21 of the total $N_{e}$ estimated from autosomes as a result of the female biased sex ratio in \textit{R.
@@
 Importantly, BGS assumes independence among sites, and the theory breaks down when many linked sites are subject to selection at the same time.
-This isbecause high linkage disequilibrium causes the behavior of selected alleles interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
+This is because high linkage disequilibrium causes the behavior of selected alleles interfere with the action of selection at linked sites \citep{good2014genetic,KaiserCharlesworth}.
 As there are no analytical formulae for predicting the outcome of selection on many linked alleles experiencing selection and drift, however, simulations are fundamental to understanding such selective interference in realistic situations.
